---
title: "model_testing"
output: html_document
---

```{r}
#load libraries/packages
source("packages.R")

#load function 
source("df_years()_Function.R")
merged_testing <- df_years(yr1 = 2015) 
```


# FD Subgroup Analysis: SDG ~ SPI + subgroup(income group)

## Models - subgroups by income class 

```{r}
# Overall model
fd_overall <- lm(fd_sdg ~ fd_spi + log_gdppc, 
                data = fd_fe_data %>% 
                  filter(!is.na(fd_sdg) & !is.na(fd_spi)))
summary(fd_overall)

# High income countries
fd_high <- lm(fd_sdg ~ fd_spi + gdppc,
             data = fd_fe_data %>% 
               filter(!is.na(fd_sdg) & !is.na(fd_spi) & 
                     income_level == "High Income Countries"))
summary(fd_high)

# Upper-middle income countries
fd_upper <- lm(fd_sdg ~ fd_spi + log_gdppc,
              data = fd_fe_data %>% 
                filter(!is.na(fd_sdg) & !is.na(fd_spi) & 
                      income_level == "Upper-Middle Income Countries"))
summary(fd_upper)

# Lower-middle income countries
fd_lower <- lm(fd_sdg ~ fd_spi + log_gdppc,
              data = fd_fe_data %>% 
                filter(!is.na(fd_sdg) & !is.na(fd_spi) & 
                      income_level == "Lower-Middle Income Countries"))
summary(fd_lower)

# Low income countries
fd_low <- lm(fd_sdg ~ fd_spi + log_gdppc,
            data = fd_fe_data %>% 
              filter(!is.na(fd_sdg) & !is.na(fd_spi) & 
                    income_level == "Low Income Countries"))
summary(fd_low)
```


## FD Comparison Table: SDG ~ SPI

```{r}
# Extract coefficients and statistics from each model
fd_comparison <- data.frame(
  model = c("Overall", "High Income", "Upper-Middle Income", 
           "Lower-Middle Income", "Low Income"),
  coefficient = c(
    coef(fd_overall)["fd_spi"],
    coef(fd_high)["fd_spi"],
    coef(fd_upper)["fd_spi"],
    coef(fd_lower)["fd_spi"],
    coef(fd_low)["fd_spi"]
  ),
  intercept = c(
    coef(fd_overall)["(Intercept)"],
    coef(fd_high)["(Intercept)"],
    coef(fd_upper)["(Intercept)"],
    coef(fd_lower)["(Intercept)"],
    coef(fd_low)["(Intercept)"]
  ),
  t_value = c(
    summary(fd_overall)$coefficients["fd_spi", "t value"],
    summary(fd_high)$coefficients["fd_spi", "t value"],
    summary(fd_upper)$coefficients["fd_spi", "t value"],
    summary(fd_lower)$coefficients["fd_spi", "t value"],
    summary(fd_low)$coefficients["fd_spi", "t value"]
  ),
  p_value = c(
    summary(fd_overall)$coefficients["fd_spi", "Pr(>|t|)"],
    summary(fd_high)$coefficients["fd_spi", "Pr(>|t|)"],
    summary(fd_upper)$coefficients["fd_spi", "Pr(>|t|)"],
    summary(fd_lower)$coefficients["fd_spi", "Pr(>|t|)"],
    summary(fd_low)$coefficients["fd_spi", "Pr(>|t|)"]
  ),
  r_squared = c(
    summary(fd_overall)$r.squared,
    summary(fd_high)$r.squared,
    summary(fd_upper)$r.squared,
    summary(fd_lower)$r.squared,
    summary(fd_low)$r.squared
  ),
  adj_r_squared = c(
    summary(fd_overall)$adj.r.squared,
    summary(fd_high)$adj.r.squared,
    summary(fd_upper)$adj.r.squared,
    summary(fd_lower)$adj.r.squared,
    summary(fd_low)$adj.r.squared
  ),
  n_obs = c(
    nobs(fd_overall),
    nobs(fd_high),
    nobs(fd_upper),
    nobs(fd_lower),
    nobs(fd_low)
  )
)

# Format the numbers for better display
fd_comparison <- fd_comparison %>%
  mutate(
    coefficient = round(coefficient, 3),
    intercept = round(intercept, 3),
    t_value = round(t_value, 3),
    p_value = round(p_value, 4),
    r_squared = round(r_squared, 3),
    adj_r_squared = round(adj_r_squared, 3)
  )

# Print the comparison table
print(fd_comparison)

#export fd_comparison 
#write.csv(fd_comparison, file = "fd_comparison.csv", row.names=F) 
```

**RESULTS:**

The overall model shows a positive and statistically significant relationship (coefficient = 0.015, p = 0.0033) between improvements in statistical capacity and improvements in SDG performance. Although this confirms my hypothesis that better statistical systems contribute to better development outcomes, the intercept indicates that when SPI remains constant, SDG scores progress by 0.286 anyway.

The income group breakdown reveals fascinating variations:

1.  **High Income Countries** show a negative coefficient (-0.012) that is not statistically significant (p = 0.2537). This suggests that for developed economies, improvements in statistical capacity don't necessarily translate to SDG improvements, possibly because they already have well-established statistical systems and other factors drive their SDG performance.

2.  **Upper-Middle Income Countries** show a stronger positive effect (0.021) that is statistically significant (p = 0.0163). This indicates that for these emerging economies, investments in statistical capacity yield meaningful SDG improvements.

3.  **Lower-Middle Income Countries** demonstrate an even stronger relationship (0.025) with high statistical significance (p = 0.0088). This is the most significant relationship among all income groups.

4.  **Low Income Countries** show the strongest coefficient (0.026), suggesting the largest potential impact, though with slightly lower statistical significance (p = 0.0459) than lower-middle income countries.

The R-squared values are relatively low across all models (between 0.004 and 0.023), indicating that changes in statistical capacity explain only a small portion of the variation in SDG changes. This is common in first-difference models and suggests that many other factors also influence SDG progress.

1.  The relationship between statistical capacity and sustainable development follows an inverted U-shaped pattern across the development spectrum - minimal impact for high-income countries, strongest for middle-income countries, and slightly less strong (but still significant) for low-income countries.

2.  Statistical capacity improvements appear most beneficial for developing and emerging economies, suggesting that international efforts to strengthen statistical systems should prioritize these countries.

3.  The consistently positive intercepts (ranging from 0.224 to 0.364) indicate that SDG scores tend to improve over time regardless of changes in statistical capacity, though the rate of improvement varies by income level.


## FD Visual Analysis: SDG ~ SPI + Income Groups 

**Error-bar**

```{r}
# Create dataframe for plotting 
plot_data <- fd_comparison %>%
  select(model, coefficient, p_value) %>%
  mutate(significant = p_value < 0.05)

# Create Horizontal error-bar graph 
fd_errorbar <- ggplot(plot_data, aes(x = coefficient, y = model, color = significant)) +
  geom_point(size = 3) +
  geom_errorbarh(aes(xmin = 0, xmax = coefficient), height = 0.2) + 
  geom_vline(xintercept = 0, linetype = "dashed") +
  scale_color_manual(values = c("gray", "blue")) +
  labs(title = "Effect of SPI Change on SDG Change by Income Group",
       x = "Coefficient (Effect Size)",
       y = "") +
  theme_bw() +
  theme(legend.position = "none")

fd_errorbar

#export fd_errorbar
# ggsave("fd_errorbar.png", width = 10, height = 6)
```

**Scatter plot of FDs**

```{r}
fd_scatter <- ggplot(fd_fe_data %>% 
         filter(!is.na(fd_sdg) & !is.na(fd_spi)), 
       aes(x = fd_spi, y = fd_sdg, color = income_level)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", se = FALSE) +
  facet_wrap(~income_level) +
  labs(title = "Relationship between Changes in SPI and SDG by Income Group",
       x = "Change in Statistical Performance (SPI)",
       y = "Change in Sustainable Development (SDG)") +
  theme_bw()

fd_scatter

#export fd_scatter
# ggsave("iCloud_Drive/Documents/GitHub/QMSS_Thesis_Sanchez", width = 10, height = 6)
```




# Multiple Regression: Xs = p1-5

```{r}
# y & x vars + controls
sdg_overall_score <- merged_testing$sdg_overall 
p_1 <- merged_testing$p1_use #insert x var 2
p_2 <- merged_testing$p2_services #insert x var 3
p_3 <- merged_testing$p3_products #insert x var 3
p_4 <- merged_testing$p4_sources #insert x var 3
p_5 <- merged_testing$p5_infra #insert x var 3
gdppc <- merged_testing$gdppc

df_spi_p1to5 <- data.frame(p_1, p_2, p_3, p_4, p_5, gdppc)

#OLS lm & full summary table 
lm2 <- lm(sdg_overall_score ~ p_1 + p_2 + p_3 + p_4 + p_5 + log(gdppc), data=df_spi_p1to5)
summary(lm2)

#plotting residuals
plot(lm2)
#plotting scatterplot 
plot(df$y_var, df$x1_var)
abline(lm2)
```

# PCA: principle component analysis -- combine to single index (FOR FUTURE)
```{r}

# Standardize both measures
#spi_z <- scale(spi_comp)
#sci_z <- scale(sci_overall)

# Create composite (simple average)
#stat_capacity_index <- (spi_z + sci_z)/2

#extract common variance 
#pca_result <- prcomp(cbind(spi_comp, sci_overall), scale = TRUE)
#stat_capacity_pc1 <- pca_result$x[,1]  # First principal component
```


# R Stan, GLM, Markove  
```{r}
library(rstanarm)

gml_testing <- stan_glm(sdg_overall ~ spi_comp + regime_type + regime_type*spi_comp + log(gdppc), data = merged_testing)

# glm_testing
print(gml_testing)
summary(gml_testing)
```


# Two-Stage Least Squares (2SLS) with Fixed Effects
```{r}
library(plm)
library(lmtest)

# First stage: DI → SPI
fe_stage1 <- plm(
  spi_comp ~ lag(di_score, 1) + lag(spi_comp, 1) + log_gdppc + factor(year),
  data = fd_fe_data,
  index = c("country_code", "year"),
  model = "within",
  effect = "twoways"
)

# Second stage: SPI → SDG (with DI as control)
fe_stage2 <- plm(
  sdg_overall ~ spi_comp + lag(di_score, 1) + lag(sdg_overall, 1) + log_gdppc + factor(year),
  data = your_data,
  index = c("country", "year"),
  model = "within",
  effect = "twoways"
)

# Report with HAC standard errors (Newey-West)
results_stage1 <- coeftest(fe_stage1, vcov = NeweyWest(fe_stage1, lag = 1))
results_stage2 <- coeftest(fe_stage2, vcov = NeweyWest(fe_stage2, lag = 1))

# Extract indirect effect (DI → SPI → SDG)
indirect_effect <- coef(fe_stage1)["lag(di_score, 1)"] * coef(fe_stage2)["spi_comp"]
```


# Qualtitative variable testing 
```{r}
# Recode to dummy variables
merged_2015 <- merged_2015 %>% 
  # split income_level into dummy variables using case_when()
  mutate(income_level_recoded = case_when(
    income_level == "L" ~ 1,
    income_level == "LM" ~ 2,
    income_level == "UM" ~ 3,
    income_level == "H" ~ 4,
    TRUE ~ NA_real_  # Handle any other cases
  )) %>% 
  mutate(income_level_recoded = as.factor(income_level_recoded))

# as a regular categorical variable
eq_testing <- lm(sdg_overall ~ spi_comp + income_level_recoded, data = merged_2015)
summary(eq_testing)

# as a categorical factor(variable)
eq_testing_fcat <- lm(sdg_overall ~ spi_comp + factor(income_level_recoded), data = merged_2015)
summary(eq_testing_fcat) # LITERALLY SAME EVERYTHING AS ABOVE

#as a categorical variable with dummies 
merged_2015 <- merged_2015 %>% 
  # Create dummy variables for income_level_recoded
  mutate(income_level_recoded_1 = ifelse(income_level_recoded == "1", 1, 0),
         income_level_recoded_2 = ifelse(income_level_recoded == "2", 1, 0),
         income_level_recoded_3 = ifelse(income_level_recoded == "3", 1, 0),
         income_level_recoded_4 = ifelse(income_level_recoded == "4", 1, 0)) %>% 
  # Convert to factors
  mutate(income_level_recoded_1 = as.factor(income_level_recoded_1),
         income_level_recoded_2 = as.factor(income_level_recoded_2),
         income_level_recoded_3 = as.factor(income_level_recoded_3),
         income_level_recoded_4 = as.factor(income_level_recoded_4))

# testing with dummies
eq_testing_dummies <- lm(sdg_overall ~ spi_comp + income_level_recoded_2 + income_level_recoded_3 + income_level_recoded_4, data = merged_2015)
summary(eq_testing_dummies)

# CONCLUSION: EVERY VARIATION OF THE QUALITATIVE VARIABLE FORMAT (WHETHER factor(var) OR Cat1 + Cat2 +...) IS THE SAME! ALL STATISTICS, MODEL OR VARIABLE WISE, ARE EXACTLY IDENTICAL!
```

```{r}
eq_testing <- plm(sdg_overall ~ spi_comp + income_level_recoded, 
                  model = "pooling", 
                  index = c("country_code", "year"),
                  data = merged_2015)
summary(eq_testing)


```


# Non-Linear Functional Form Testing 
```{r}
merged_2015 <- merged_2015 %>% mutate(regime_type_2 = as.factor(regime_type_2), 
                                      regime_type_4 = as.factor(regime_type_4))

# Quadratic and log for gdppc
testing_quadratic_gdp <- lm(sdg_overall ~ spi_comp + di_score + log_gdppc + I(log_gdppc^2) + factor(year), data = merged_2015)
summary(testing_quadratic_gdp, vcov = vcovHC(testing_quadratic_gdp, cluster = "group", type = "HC1"))

# Adding centered predictors 
testing_centered <- lm(sdg_overall ~ cen_spi_comp + I(cen_spi_comp^2) + cen_di_score + I(cen_di_score^2) + cen_log_gdppc + I(cen_log_gdppc^2) + factor(year), data = merged_2015)
summary(testing_centered, vcov = vcovHC(testing_centered, cluster = "group", type = "HC1"))

# Adding interaction term: cen_spi_comp x income_level
testing_inter_spi_x_income <- lm(sdg_overall ~ cen_spi_comp*income_level + cen_di_score + cen_log_gdppc + I(cen_log_gdppc^2) + factor(year), data = merged_2015)
summary(testing_inter_spi_x_income, vcov = vcovHC(testing_inter_spi_x_income, cluster = "group", type = "HC1"))

# Adding interaction term: cen_spi_comp x regime_type

# Check for multicollinearity
vif(testing_quadratic_gdp)
vif(testing_centered)
vif(testing_inter_spi_x_income)
```
- As a control, income level takes on the same slope as that of the overall model
- 


but the interaction term shows that the slope is __steeper__ for low-income countries and __less steep__ for high-income countries. This suggests that the relationship between SPI and SDG is stronger in low-income countries, which aligns with the findings from the subgroup analysis.

# Cross-Validation for Model Testing
```{r}
library(caret)

# Set up k-fold cross-validation (k=10 is commonly used)
ctrl <- trainControl(
  method = "cv",           # Cross-validation
  number = 10,             # Number of folds (k=10)
  savePredictions = "final" # Save predictions for further analysis
)

# Fit your model with cross-validation
set.seed(123)  # For reproducibility
cv_model <- train(
  sdg_overall ~ cen_spi_comp + I(cen_spi_comp^2) + 
                cen_di_score + I(cen_di_score^2) + 
                cen_log_gdppc + I(cen_log_gdppc^2) + 
                factor(year),
  data = merged_2015,
  method = "lm", # Linear model
  trControl = ctrl
)

# View cross-validation results
print(cv_model)

```

