---
title: "model_testing"
output: html_document
---

```{r}
#load libraries/packages
source("packages.R")

#load function 
source("df_years()_Function.R")
merged_testing <- df_years(yr1 = 2015) 
```


# FD Subgroup Analysis: SDG ~ SPI + subgroup(income group)

## Models - subgroups by income class 

```{r}
# Overall model
fd_overall <- lm(fd_sdg ~ fd_spi + log_gdppc + population, 
                data = fd_fe_data %>% 
                  filter(!is.na(fd_sdg) & !is.na(fd_spi)))
summary(fd_overall)

# High income countries
fd_high <- lm(fd_sdg ~ fd_spi + gdppc + population, 
             data = fd_fe_data %>% 
               filter(!is.na(fd_sdg) & !is.na(fd_spi) & 
                     income_level == "High Income Countries"))
summary(fd_high)

# Upper-middle income countries
fd_upper <- lm(fd_sdg ~ fd_spi + log_gdppc + population, 
              data = fd_fe_data %>% 
                filter(!is.na(fd_sdg) & !is.na(fd_spi) & 
                      income_level == "Upper-Middle Income Countries"))
summary(fd_upper)

# Lower-middle income countries
fd_lower <- lm(fd_sdg ~ fd_spi + log_gdppc + population, 
              data = fd_fe_data %>% 
                filter(!is.na(fd_sdg) & !is.na(fd_spi) & 
                      income_level == "Lower-Middle Income Countries"))
summary(fd_lower)

# Low income countries
fd_low <- lm(fd_sdg ~ fd_spi + log_gdppc + population, 
            data = fd_fe_data %>% 
              filter(!is.na(fd_sdg) & !is.na(fd_spi) & 
                    income_level == "Low Income Countries"))
summary(fd_low)
```


## FD Comparison Table: SDG ~ SPI

```{r}
# Extract coefficients and statistics from each model
fd_comparison <- data.frame(
  model = c("Overall", "High Income", "Upper-Middle Income", 
           "Lower-Middle Income", "Low Income"),
  coefficient = c(
    coef(fd_overall)["fd_spi"],
    coef(fd_high)["fd_spi"],
    coef(fd_upper)["fd_spi"],
    coef(fd_lower)["fd_spi"],
    coef(fd_low)["fd_spi"]
  ),
  intercept = c(
    coef(fd_overall)["(Intercept)"],
    coef(fd_high)["(Intercept)"],
    coef(fd_upper)["(Intercept)"],
    coef(fd_lower)["(Intercept)"],
    coef(fd_low)["(Intercept)"]
  ),
  t_value = c(
    summary(fd_overall)$coefficients["fd_spi", "t value"],
    summary(fd_high)$coefficients["fd_spi", "t value"],
    summary(fd_upper)$coefficients["fd_spi", "t value"],
    summary(fd_lower)$coefficients["fd_spi", "t value"],
    summary(fd_low)$coefficients["fd_spi", "t value"]
  ),
  p_value = c(
    summary(fd_overall)$coefficients["fd_spi", "Pr(>|t|)"],
    summary(fd_high)$coefficients["fd_spi", "Pr(>|t|)"],
    summary(fd_upper)$coefficients["fd_spi", "Pr(>|t|)"],
    summary(fd_lower)$coefficients["fd_spi", "Pr(>|t|)"],
    summary(fd_low)$coefficients["fd_spi", "Pr(>|t|)"]
  ),
  r_squared = c(
    summary(fd_overall)$r.squared,
    summary(fd_high)$r.squared,
    summary(fd_upper)$r.squared,
    summary(fd_lower)$r.squared,
    summary(fd_low)$r.squared
  ),
  adj_r_squared = c(
    summary(fd_overall)$adj.r.squared,
    summary(fd_high)$adj.r.squared,
    summary(fd_upper)$adj.r.squared,
    summary(fd_lower)$adj.r.squared,
    summary(fd_low)$adj.r.squared
  ),
  n_obs = c(
    nobs(fd_overall),
    nobs(fd_high),
    nobs(fd_upper),
    nobs(fd_lower),
    nobs(fd_low)
  )
)

# Format the numbers for better display
fd_comparison <- fd_comparison %>%
  mutate(
    coefficient = round(coefficient, 3),
    intercept = round(intercept, 3),
    t_value = round(t_value, 3),
    p_value = round(p_value, 4),
    r_squared = round(r_squared, 3),
    adj_r_squared = round(adj_r_squared, 3)
  )

# Print the comparison table
print(fd_comparison)

#export fd_comparison 
#write.csv(fd_comparison, file = "fd_comparison.csv", row.names=F) 
```

**RESULTS:**

The overall model shows a positive and statistically significant relationship (coefficient = 0.015, p = 0.0033) between improvements in statistical capacity and improvements in SDG performance. Although this confirms my hypothesis that better statistical systems contribute to better development outcomes, the intercept indicates that when SPI remains constant, SDG scores progress by 0.286 anyway.

The income group breakdown reveals fascinating variations:

1.  **High Income Countries** show a negative coefficient (-0.012) that is not statistically significant (p = 0.2537). This suggests that for developed economies, improvements in statistical capacity don't necessarily translate to SDG improvements, possibly because they already have well-established statistical systems and other factors drive their SDG performance.

2.  **Upper-Middle Income Countries** show a stronger positive effect (0.021) that is statistically significant (p = 0.0163). This indicates that for these emerging economies, investments in statistical capacity yield meaningful SDG improvements.

3.  **Lower-Middle Income Countries** demonstrate an even stronger relationship (0.025) with high statistical significance (p = 0.0088). This is the most significant relationship among all income groups.

4.  **Low Income Countries** show the strongest coefficient (0.026), suggesting the largest potential impact, though with slightly lower statistical significance (p = 0.0459) than lower-middle income countries.

The R-squared values are relatively low across all models (between 0.004 and 0.023), indicating that changes in statistical capacity explain only a small portion of the variation in SDG changes. This is common in first-difference models and suggests that many other factors also influence SDG progress.

1.  The relationship between statistical capacity and sustainable development follows an inverted U-shaped pattern across the development spectrum - minimal impact for high-income countries, strongest for middle-income countries, and slightly less strong (but still significant) for low-income countries.

2.  Statistical capacity improvements appear most beneficial for developing and emerging economies, suggesting that international efforts to strengthen statistical systems should prioritize these countries.

3.  The consistently positive intercepts (ranging from 0.224 to 0.364) indicate that SDG scores tend to improve over time regardless of changes in statistical capacity, though the rate of improvement varies by income level.

## FD Visual Analysis: SDG ~ SPI + Income Groups 

**Error-bar**

```{r}
# Create dataframe for plotting 
plot_data <- fd_comparison %>%
  select(model, coefficient, p_value) %>%
  mutate(significant = p_value < 0.05)

# Create Horizontal error-bar graph 
fd_errorbar <- ggplot(plot_data, aes(x = coefficient, y = model, color = significant)) +
  geom_point(size = 3) +
  geom_errorbarh(aes(xmin = 0, xmax = coefficient), height = 0.2) + 
  geom_vline(xintercept = 0, linetype = "dashed") +
  scale_color_manual(values = c("gray", "blue")) +
  labs(title = "Effect of SPI Change on SDG Change by Income Group",
       x = "Coefficient (Effect Size)",
       y = "") +
  theme_bw() +
  theme(legend.position = "none")

fd_errorbar

#export fd_errorbar
# ggsave("fd_errorbar.png", width = 10, height = 6)
```

**Scatter plot of FDs**

```{r}
fd_scatter <- ggplot(fd_fe_data %>% 
         filter(!is.na(fd_sdg) & !is.na(fd_spi)), 
       aes(x = fd_spi, y = fd_sdg, color = income_level)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", se = FALSE) +
  facet_wrap(~income_level) +
  labs(title = "Relationship between Changes in SPI and SDG by Income Group",
       x = "Change in Statistical Performance (SPI)",
       y = "Change in Sustainable Development (SDG)") +
  theme_bw()

fd_scatter

#export fd_scatter
# ggsave("iCloud_Drive/Documents/GitHub/QMSS_Thesis_Sanchez", width = 10, height = 6)
```





# Backsliding FE Model 
```{r}
#NICEEEE THIS WORKS!!
regressed <- fd_fe_data %>%
  # Filter to post-2015 first (if you want to check events only in this period)
  filter(year >= 2015) %>%
  # Group by country to check for any event
  group_by(country_code) %>%
  # Flag countries with at least one aut_ep == 1
  mutate(has_aut_ep = any(aut_ep == 1, na.rm = TRUE)) %>%
  ungroup() %>%
  # Keep only countries with at least one event
  filter(has_aut_ep) %>%
  # Remove the helper column
  select(-has_aut_ep)

# panel df - all backsliding 
panel_regressed <- pdata.frame(
  regressed, 
  index = c("country_code", "year"))  # Replace 'year' with your time variable

# FE Model - ALL
fe_testing_all <- plm(spi_comp ~ di_score + log_gdppc + population + factor(year), 
                  data = panel_regressed,
                  model = "within")
summary(fe_testing_all)
coeftest(fe_testing_all, vcov = vcovHC(fe_testing_all, cluster = "group", type = "HC1"))

# Single Country: Hungary 
hun_regressed <- regressed %>% 
  filter(country_code == "HUN")

# panel df - Hungary 
panel_data_hun <- pdata.frame(
  hun_regressed, 
  index = c("country_code", "year")  # Replace 'year' with your time variable
)

# FE model - Hungary 
fe_testing <- plm(spi_comp ~ di_score + log_gdppc + population + factor(year), 
                  data = panel_data_hun,
                  model = "within")
summary(fe_testing)
coeftest(fe_testing, vcov = vcovHC(fe_testing, cluster = "group", type = "HC1"))
```

```{r}
# subset of countries that experienced atleast 1 democratization event 
democratized <- merged %>% 
  group_by(country_code) %>%
  filter(dem_ep == 1) %>%
  ungroup()
```


# OLS: including regime variable [which is best?]
```{r}
# 1. regime_type_2 (0 or 1) control
lm1_reg2 <- lm(sdg_overall ~ spi_comp + regime_type_2 + log(gdppc) + population,
                        data = merged_testing)
summary(lm1_reg2)

# 2. lib_dem (continuous 0-1) control
lm2_regcont <- lm(sdg_overall ~ spi_comp + lib_dem + log(gdppc) + population,
                        data = merged_testing)
summary(lm2_regcont)

# 3. regime_type_2 (0 or 1) interaction 
lm3_reg2_int <- lm(sdg_overall ~ spi_comp + spi_comp*regime_type_2 + log(gdppc) + population,
                        data = merged_testing)
summary(lm3_reg2_int)

# 4. lib_dem (continuous 0-1) interaction
lm4_libd_int <- lm(sdg_overall ~ spi_comp + spi_comp*lib_dem + log(gdppc) + population,
                        data = merged_testing)
summary(lm4_libd_int)

# 5. regime_type_2 (0 or 1) interaction + control
lm5_reg2_int_cont <- lm(sdg_overall ~ spi_comp + regime_type_2 + spi_comp*regime_type_2 + log(gdppc) + population,
                        data = merged_testing)

summary(lm5_reg2_int_cont)

# 6. lib_dem (continuous 0-1) interaction + control 
lm6_libd_int_cont <- lm(sdg_overall ~ spi_comp + lib_dem + spi_comp*lib_dem + log(gdppc) + population,
                        data = merged_testing)
summary(lm6_libd_int_cont)

# just interaction?
lm7_libd_int_cont <- lm(sdg_overall ~ spi_comp + spi_comp*lib_dem + log(gdppc) + population,
                        data = merged_testing)
summary(lm7_libd_int)
```
- Interactions are not statistically significant (lib_dem or )



## Multiple Regression: Xs = p1-5

```{r}
# y & x vars + controls
sdg_overall_score <- merged_testing$sdg_overall 
p_1 <- merged_testing$p1_use #insert x var 2
p_2 <- merged_testing$p2_services #insert x var 3
p_3 <- merged_testing$p3_products #insert x var 3
p_4 <- merged_testing$p4_sources #insert x var 3
p_5 <- merged_testing$p5_infra #insert x var 3
gdppc <- merged_testing$gdppc

df_spi_p1to5 <- data.frame(p_1, p_2, p_3, p_4, p_5, gdppc)

#OLS lm & full summary table 
lm2 <- lm(sdg_overall_score ~ p_1 + p_2 + p_3 + p_4 + p_5 + log(gdppc), data=df_spi_p1to5)
summary(lm2)

#plotting residuals
plot(lm2)
#plotting scatterplot 
plot(df$y_var, df$x1_var)
abline(lm2)
```

## Combine to single index: principle component analysis (FOR FUTURE)
```{r}

# Standardize both measures
#spi_z <- scale(spi_comp)
#sci_z <- scale(sci_overall)

# Create composite (simple average)
#stat_capacity_index <- (spi_z + sci_z)/2

#extract common variance 
#pca_result <- prcomp(cbind(spi_comp, sci_overall), scale = TRUE)
#stat_capacity_pc1 <- pca_result$x[,1]  # First principal component
```


## R Stan, GLM, Markove  
```{r}
library(rstanarm)

gml_testing <- stan_glm(sdg_overall ~ spi_comp + regime_type + regime_type*spi_comp + log(gdppc) + population, data = merged_testing)

# glm_testing
print(gml_testing)
summary(gml_testing)
```

#First difference
```{r}
fd_model <- plm(
  spi_comp ~ di_score + log(gdppc) + population + year_fct + gini,
  data = merged_testing,
  index = c("country_code", "year"),
  model = "fd"
)

summary(fd_model)

# Check for serial correlation
lmtest::dwtest(fd_model)

```


## Fixed Effects Model
```{r}
library(plm)
fe_model <- plm(sdg_overall ~ spi_comp + regime_type + gdppc + population,
                data = merged_testing, 
                index = c("country_code", "year"), 
                model = "within")

summary(fe_model)
```


# Mediation Model: SDG ~ SPI(DI)
```{r}
# Model SPI’s effect on SDG outcomes with regime change as mediator

library(mediation)

med_model <- mediate(
  model.m = lm(spi_comp ~ di_score + log(gdppc) + population + gini, data = fd_fe_data),
  model.y = lm(sdg_overall ~ spi_comp + di_score + log(gdppc) + population + gini, data = fd_fe_data),
  treat = "aut_ep",
  mediator = "spi_comp"
)

summary(med_model)
```


# Significance of factor(year) Dummy Variable 

## factor(year) for OLS: SDG ~ SPI
```{r}
#excludes year dummy 
ols_spi <- lm(y_var_sdg ~ x_var_spi + log_gdppc + population + di_score + gini, data = merged)
summary(ols_spi)

#includes year dummy 
ols_spi_fyr <- lm(y_var_sdg ~ x_var_spi + log_gdppc + population + di_score + gini + factor(year), data = merged)
summary(ols_spi_fyr)

# Joint significance
anova(ols_spi, ols_spi_fyr)
```

## factor(year) for OLS: SDG ~ SCI
```{r}
#excludes year dummy 
ols_sci <- lm(y_var_sdg ~ x_var_sci + log_gdppc + population + di_score + gini, data = merged)
summary(ols_sci)

#includes year dummy 
ols_sci_fyr <- lm(y_var_sdg ~ x_var_sci + log_gdppc + population + di_score + gini + factor(year), data = merged)
summary(ols_sci_fyr)

# Joint significance
anova(ols_sci, ols_sci_fyr)
```

## factor(year) for OLS: SDG ~ SPI + SCI 
```{r}
#excludes year dummy 
ols_multiple <- lm(y_var_sdg ~ x_var_spi + x_var_sci + log_gdppc + population + di_score + gini, data = merged)
summary(ols_multiple)

#includes year dummy 
ols_multiple_fyr <- lm(y_var_sdg ~ x_var_spi + x_var_sci + log_gdppc + population + di_score + gini + factor(year), data = merged)
summary(ols_multiple_fyr)

# Joint significance
anova(ols_multiple, ols_multiple_fyr)
```


## VIF: Check for multicollinearity for models 
```{r}
# mainly concerned abt log_gdppc & factor(time)

#SPI VIF 
car::vif(ols_spi_fyr)  

#SCI VIF
car::vif(ols_sci_fyr) 

#SPI + SCI VIF
car::vif(ols_multiple_fyr) 
```

