---
title: "model_testing"
output: html_document
---

```{r}
#load libraries/packages
source("packages.R")

#load function 
source("df_years()_Function.R")
merged_testing <- df_years(yr1 = 2015) 
```


# FD Subgroup Analysis: SDG ~ SPI + subgroup(income group)

## Models - subgroups by income class 

```{r}
# Overall model
fd_overall <- lm(fd_sdg ~ fd_spi + log_gdppc, 
                data = fd_fe_data %>% 
                  filter(!is.na(fd_sdg) & !is.na(fd_spi)))
summary(fd_overall)

# High income countries
fd_high <- lm(fd_sdg ~ fd_spi + gdppc,
             data = fd_fe_data %>% 
               filter(!is.na(fd_sdg) & !is.na(fd_spi) & 
                     income_level == "High Income Countries"))
summary(fd_high)

# Upper-middle income countries
fd_upper <- lm(fd_sdg ~ fd_spi + log_gdppc,
              data = fd_fe_data %>% 
                filter(!is.na(fd_sdg) & !is.na(fd_spi) & 
                      income_level == "Upper-Middle Income Countries"))
summary(fd_upper)

# Lower-middle income countries
fd_lower <- lm(fd_sdg ~ fd_spi + log_gdppc,
              data = fd_fe_data %>% 
                filter(!is.na(fd_sdg) & !is.na(fd_spi) & 
                      income_level == "Lower-Middle Income Countries"))
summary(fd_lower)

# Low income countries
fd_low <- lm(fd_sdg ~ fd_spi + log_gdppc,
            data = fd_fe_data %>% 
              filter(!is.na(fd_sdg) & !is.na(fd_spi) & 
                    income_level == "Low Income Countries"))
summary(fd_low)
```


## FD Comparison Table: SDG ~ SPI

```{r}
# Extract coefficients and statistics from each model
fd_comparison <- data.frame(
  model = c("Overall", "High Income", "Upper-Middle Income", 
           "Lower-Middle Income", "Low Income"),
  coefficient = c(
    coef(fd_overall)["fd_spi"],
    coef(fd_high)["fd_spi"],
    coef(fd_upper)["fd_spi"],
    coef(fd_lower)["fd_spi"],
    coef(fd_low)["fd_spi"]
  ),
  intercept = c(
    coef(fd_overall)["(Intercept)"],
    coef(fd_high)["(Intercept)"],
    coef(fd_upper)["(Intercept)"],
    coef(fd_lower)["(Intercept)"],
    coef(fd_low)["(Intercept)"]
  ),
  t_value = c(
    summary(fd_overall)$coefficients["fd_spi", "t value"],
    summary(fd_high)$coefficients["fd_spi", "t value"],
    summary(fd_upper)$coefficients["fd_spi", "t value"],
    summary(fd_lower)$coefficients["fd_spi", "t value"],
    summary(fd_low)$coefficients["fd_spi", "t value"]
  ),
  p_value = c(
    summary(fd_overall)$coefficients["fd_spi", "Pr(>|t|)"],
    summary(fd_high)$coefficients["fd_spi", "Pr(>|t|)"],
    summary(fd_upper)$coefficients["fd_spi", "Pr(>|t|)"],
    summary(fd_lower)$coefficients["fd_spi", "Pr(>|t|)"],
    summary(fd_low)$coefficients["fd_spi", "Pr(>|t|)"]
  ),
  r_squared = c(
    summary(fd_overall)$r.squared,
    summary(fd_high)$r.squared,
    summary(fd_upper)$r.squared,
    summary(fd_lower)$r.squared,
    summary(fd_low)$r.squared
  ),
  adj_r_squared = c(
    summary(fd_overall)$adj.r.squared,
    summary(fd_high)$adj.r.squared,
    summary(fd_upper)$adj.r.squared,
    summary(fd_lower)$adj.r.squared,
    summary(fd_low)$adj.r.squared
  ),
  n_obs = c(
    nobs(fd_overall),
    nobs(fd_high),
    nobs(fd_upper),
    nobs(fd_lower),
    nobs(fd_low)
  )
)

# Format the numbers for better display
fd_comparison <- fd_comparison %>%
  mutate(
    coefficient = round(coefficient, 3),
    intercept = round(intercept, 3),
    t_value = round(t_value, 3),
    p_value = round(p_value, 4),
    r_squared = round(r_squared, 3),
    adj_r_squared = round(adj_r_squared, 3)
  )

# Print the comparison table
print(fd_comparison)

#export fd_comparison 
#write.csv(fd_comparison, file = "fd_comparison.csv", row.names=F) 
```

**RESULTS:**

The overall model shows a positive and statistically significant relationship (coefficient = 0.015, p = 0.0033) between improvements in statistical capacity and improvements in SDG performance. Although this confirms my hypothesis that better statistical systems contribute to better development outcomes, the intercept indicates that when SPI remains constant, SDG scores progress by 0.286 anyway.

The income group breakdown reveals fascinating variations:

1.  **High Income Countries** show a negative coefficient (-0.012) that is not statistically significant (p = 0.2537). This suggests that for developed economies, improvements in statistical capacity don't necessarily translate to SDG improvements, possibly because they already have well-established statistical systems and other factors drive their SDG performance.

2.  **Upper-Middle Income Countries** show a stronger positive effect (0.021) that is statistically significant (p = 0.0163). This indicates that for these emerging economies, investments in statistical capacity yield meaningful SDG improvements.

3.  **Lower-Middle Income Countries** demonstrate an even stronger relationship (0.025) with high statistical significance (p = 0.0088). This is the most significant relationship among all income groups.

4.  **Low Income Countries** show the strongest coefficient (0.026), suggesting the largest potential impact, though with slightly lower statistical significance (p = 0.0459) than lower-middle income countries.

The R-squared values are relatively low across all models (between 0.004 and 0.023), indicating that changes in statistical capacity explain only a small portion of the variation in SDG changes. This is common in first-difference models and suggests that many other factors also influence SDG progress.

1.  The relationship between statistical capacity and sustainable development follows an inverted U-shaped pattern across the development spectrum - minimal impact for high-income countries, strongest for middle-income countries, and slightly less strong (but still significant) for low-income countries.

2.  Statistical capacity improvements appear most beneficial for developing and emerging economies, suggesting that international efforts to strengthen statistical systems should prioritize these countries.

3.  The consistently positive intercepts (ranging from 0.224 to 0.364) indicate that SDG scores tend to improve over time regardless of changes in statistical capacity, though the rate of improvement varies by income level.


## FD Visual Analysis: SDG ~ SPI + Income Groups 

**Error-bar**

```{r}
# Create dataframe for plotting 
plot_data <- fd_comparison %>%
  select(model, coefficient, p_value) %>%
  mutate(significant = p_value < 0.05)

# Create Horizontal error-bar graph 
fd_errorbar <- ggplot(plot_data, aes(x = coefficient, y = model, color = significant)) +
  geom_point(size = 3) +
  geom_errorbarh(aes(xmin = 0, xmax = coefficient), height = 0.2) + 
  geom_vline(xintercept = 0, linetype = "dashed") +
  scale_color_manual(values = c("gray", "blue")) +
  labs(title = "Effect of SPI Change on SDG Change by Income Group",
       x = "Coefficient (Effect Size)",
       y = "") +
  theme_bw() +
  theme(legend.position = "none")

fd_errorbar

#export fd_errorbar
# ggsave("fd_errorbar.png", width = 10, height = 6)
```

**Scatter plot of FDs**

```{r}
fd_scatter <- ggplot(fd_fe_data %>% 
         filter(!is.na(fd_sdg) & !is.na(fd_spi)), 
       aes(x = fd_spi, y = fd_sdg, color = income_level)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", se = FALSE) +
  facet_wrap(~income_level) +
  labs(title = "Relationship between Changes in SPI and SDG by Income Group",
       x = "Change in Statistical Performance (SPI)",
       y = "Change in Sustainable Development (SDG)") +
  theme_bw()

fd_scatter

#export fd_scatter
# ggsave("iCloud_Drive/Documents/GitHub/QMSS_Thesis_Sanchez", width = 10, height = 6)
```







# Multiple Regression: Xs = p1-5

```{r}
# y & x vars + controls
sdg_overall_score <- merged_testing$sdg_overall 
p_1 <- merged_testing$p1_use #insert x var 2
p_2 <- merged_testing$p2_services #insert x var 3
p_3 <- merged_testing$p3_products #insert x var 3
p_4 <- merged_testing$p4_sources #insert x var 3
p_5 <- merged_testing$p5_infra #insert x var 3
gdppc <- merged_testing$gdppc

df_spi_p1to5 <- data.frame(p_1, p_2, p_3, p_4, p_5, gdppc)

#OLS lm & full summary table 
lm2 <- lm(sdg_overall_score ~ p_1 + p_2 + p_3 + p_4 + p_5 + log(gdppc), data=df_spi_p1to5)
summary(lm2)

#plotting residuals
plot(lm2)
#plotting scatterplot 
plot(df$y_var, df$x1_var)
abline(lm2)
```

# PCA: principle component analysis -- combine to single index (FOR FUTURE)
```{r}

# Standardize both measures
#spi_z <- scale(spi_comp)
#sci_z <- scale(sci_overall)

# Create composite (simple average)
#stat_capacity_index <- (spi_z + sci_z)/2

#extract common variance 
#pca_result <- prcomp(cbind(spi_comp, sci_overall), scale = TRUE)
#stat_capacity_pc1 <- pca_result$x[,1]  # First principal component
```


# R Stan, GLM, Markove  
```{r}
library(rstanarm)

gml_testing <- stan_glm(sdg_overall ~ spi_comp + regime_type + regime_type*spi_comp + log(gdppc), data = merged_testing)

# glm_testing
print(gml_testing)
summary(gml_testing)
```


# Two-Stage Least Squares (2SLS) with Fixed Effects
```{r}
library(plm)
library(lmtest)

# First stage: DI → SPI
fe_stage1 <- plm(
  spi_comp ~ lag(di_score, 1) + lag(spi_comp, 1) + log_gdppc + factor(year),
  data = fd_fe_data,
  index = c("country_code", "year"),
  model = "within",
  effect = "twoways"
)

# Second stage: SPI → SDG (with DI as control)
fe_stage2 <- plm(
  sdg_overall ~ spi_comp + lag(di_score, 1) + lag(sdg_overall, 1) + log_gdppc + factor(year),
  data = your_data,
  index = c("country", "year"),
  model = "within",
  effect = "twoways"
)

# Report with HAC standard errors (Newey-West)
results_stage1 <- coeftest(fe_stage1, vcov = NeweyWest(fe_stage1, lag = 1))
results_stage2 <- coeftest(fe_stage2, vcov = NeweyWest(fe_stage2, lag = 1))

# Extract indirect effect (DI → SPI → SDG)
indirect_effect <- coef(fe_stage1)["lag(di_score, 1)"] * coef(fe_stage2)["spi_comp"]
```

# Non-Linear Functional Form Testing 

## Quadratic Terms [structure]
```{r}

# linear model - no quadratic terms
ols_spi_reg <- plm(formula = sdg_overall ~ spi_comp + di_score + log_gdppc + factor(year),
               model = "pooling", 
               index = c("country_code", "year"),
               data = merged)
summary(ols_spi_quad, vcov = vcovHC(ols_spi_quad, cluster = "group", type = "HC1"))

# quadratic model
ols_spi_quad <- plm(formula = sdg_overall ~ spi_comp + I(spi_comp^2) + di_score + I(di_score^2) + log_gdppc + I(log_gdppc^2) + factor(year),
               model = "pooling", 
               index = c("country_code", "year"),
               data = merged)
summary(ols_spi_quad, vcov = vcovHC(ols_spi_quad, cluster = "group", type = "HC1"))

# Compare models
library(lmtest)
library(sandwich)
# Compare models using AIC and BIC
aic_ols <- AIC(ols_spi_reg)
bic_ols <- BIC(ols_spi_reg)
aic_quad <- AIC(ols_spi_quad)
bic_quad <- BIC(ols_spi_quad)
comparison <- data.frame(
  Model = c("OLS", "OLS with Quadratic Term"),
  AIC = c(aic_ols, aic_quad),
  BIC = c(bic_ols, bic_quad)
)

#comparing using adjusted R sq 
# Compare adjusted R-squared
adj_r_squared_ols <- summary(ols_spi_reg)$adj.r.squared
adj_r_squared_quad <- summary(ols_spi_quad)$adj.r.squared
comparison_adj_r2 <- data.frame(
  Model = c("OLS", "OLS with Quadratic Term"),
  Adjusted_R_squared = c(adj_r_squared_ols, adj_r_squared_quad)
)


# Simplify by dropping non-significant terms
model_reduced <- lm(y ~ x1 + I(x1^2) + x2 + controls, data = df)  

```



# Quadratic terms 
```{r}


# Create a sequence of SPI values for prediction
spi_seq <- seq(min(merged$spi_comp, na.rm = TRUE), max(merged$spi_comp, na.rm = TRUE), length.out = 100)
# Create a new data frame for predictions
pred_data <- data.frame(spi_comp = spi_seq, di_score = mean(merged$di_score, na.rm = TRUE), log_gdppc = mean(merged$log_gdppc, na.rm = TRUE), year = factor("2020"))
# Predict SDG values using both models
pred_spi <- predict(ols_spi, newdata = pred_data)
pred_spi_quad <- predict(ols_spi_quad, newdata = pred_data)
# Combine predictions into a data frame
pred_df <- data.frame(spi_comp = spi_seq, 
                       sdg_overall_spi = pred_spi, 
                       sdg_overall_spi_quad = pred_spi_quad)
# Plot the results
ggplot(pred_df, aes(x = spi_comp)) +
  geom_line(aes(y = sdg_overall_spi, color = "Linear Model"), size = 1) +
  geom_line(aes(y = sdg_overall_spi_quad, color = "Quadratic Model"), size = 1) +
  labs(title = "Effect of SPI on SDG with and without Quadratic Term",
       x = "SPI Composite Score",
       y = "SDG Overall Score") +
  scale_color_manual(values = c("blue", "red")) +
  theme_minimal() +
  theme(legend.title = element_blank())
# Export the plot
# ggsave("spi_sdg_quad_plot.png", width = 10, height = 6)
# Export the results
# write.csv(pred_df, "spi_sdg_predictions.csv", row.names = FALSE)
# Export the models
# saveRDS(ols_spi, "ols_spi_model.rds")
# saveRDS(ols_spi_quad, "ols_spi_quad_model.rds")
# Load the models
# ols_spi <- readRDS("ols_spi_model.rds")
# ols_spi_quad <- readRDS("ols_spi_quad_model.rds")
# Export the comparison table
# comparison_table <- data.frame(
#   Model = c("OLS", "OLS with Quadratic Term"),
#   R_squared = c(summary(ols_spi)$r.squared, summary(ols_spi_quad)$r.squared),
#   Adj_R_squared = c(summary(ols_spi)$adj.r.squared, summary(ols_spi_quad)$adj.r.squared),
#   AIC = c(AIC(ols_spi), AIC(ols_spi_quad)),
#   BIC = c(BIC(ols_spi), BIC(ols_spi_quad))
# )


