---
title: "model_testing"
output: html_document
---

```{r}
#load libraries/packages
source("packages.R")

#load function 
source("df_years()_Function.R")
merged_testing <- df_years(yr1 = 2015) 
```


# OLS: including regime variable [which is best?]
```{r}
# 1. regime_type_2 (0 or 1) control
lm1_reg2 <- lm(sdg_overall ~ spi_comp + regime_type_2 + log(gdppc) + population,
                        data = merged_testing)
summary(lm1_reg2)

# 2. lib_dem (continuous 0-1) control
lm2_regcont <- lm(sdg_overall ~ spi_comp + lib_dem + log(gdppc) + population,
                        data = merged_testing)
summary(lm2_regcont)

# 3. regime_type_2 (0 or 1) interaction 
lm3_reg2_int <- lm(sdg_overall ~ spi_comp + spi_comp*regime_type_2 + log(gdppc) + population,
                        data = merged_testing)
summary(lm3_reg2_int)

# 4. lib_dem (continuous 0-1) interaction
lm4_libd_int <- lm(sdg_overall ~ spi_comp + spi_comp*lib_dem + log(gdppc) + population,
                        data = merged_testing)
summary(lm4_libd_int)

# 5. regime_type_2 (0 or 1) interaction + control
lm5_reg2_int_cont <- lm(sdg_overall ~ spi_comp + regime_type_2 + spi_comp*regime_type_2 + log(gdppc) + population,
                        data = merged_testing)

summary(lm5_reg2_int_cont)

# 6. lib_dem (continuous 0-1) interaction + control 
lm6_libd_int_cont <- lm(sdg_overall ~ spi_comp + lib_dem + spi_comp*lib_dem + log(gdppc) + population,
                        data = merged_testing)
summary(lm6_libd_int_cont)

# just interaction?
lm7_libd_int_cont <- lm(sdg_overall ~ spi_comp + spi_comp*lib_dem + log(gdppc) + population,
                        data = merged_testing)
summary(lm7_libd_int)
```
- Interactions are not statistically significant (lib_dem or )



## Multiple Regression: Xs = p1-5

```{r}
# y & x vars + controls
sdg_overall_score <- merged_testing$sdg_overall 
p_1 <- merged_testing$p1_use #insert x var 2
p_2 <- merged_testing$p2_services #insert x var 3
p_3 <- merged_testing$p3_products #insert x var 3
p_4 <- merged_testing$p4_sources #insert x var 3
p_5 <- merged_testing$p5_infra #insert x var 3
gdppc <- merged_testing$gdppc

df_spi_p1to5 <- data.frame(p_1, p_2, p_3, p_4, p_5, gdppc)

#OLS lm & full summary table 
lm2 <- lm(sdg_overall_score ~ p_1 + p_2 + p_3 + p_4 + p_5 + log(gdppc), data=df_spi_p1to5)
summary(lm2)

#plotting residuals
plot(lm2)
#plotting scatterplot 
plot(df$y_var, df$x1_var)
abline(lm2)
```

# OLS: spi ~ democracy index [DR. DI GENNERO's MODEL]
```{r}
ols_spi_di <- lm(spi_comp ~ di_score + log(gdppc) + population + year_fct + gini, data = merged_testing)
summary(ols_spi_di)

ggplot(merged_testing, aes(di_score, spi_comp))+
  geom_point()+
  geom_smooth(method=lm) #makes the line into a linear model

#correlation - variance 
spi_di_cor <- cor(merged_testing$spi_comp, merged_testing$di_score, use = "complete.obs")
spi_di_cor

#r-squared
spi_di_cor_2 <- cor(merged_testing$spi_comp, merged_testing$di_score, use = "complete.obs")^2
spi_di_cor_2
```


#First difference
```{r}
fd_model <- plm(
  spi_comp ~ di_score + log(gdppc) + population + year_fct + gini,
  data = merged_testing,
  index = c("country_code", "year"),
  model = "fd"
)

summary(fd_model)

# Check for serial correlation
lmtest::dwtest(fd_model)

```

**PART 1: Create lagged variables and first differences**

```{r}
# Create a simplified data frame with essential variables
fd_data_testing <- merged_testing %>%
  dplyr::select(country_name.x, country_code, year, sdg_overall, spi_comp, di_score, income_level, gini)

# PART 1: Create lagged variables and first differences
fd_data_testing <- fd_data_testing %>%
  arrange(country_code, year) %>% 
  group_by(country_code) %>%
  mutate(
    # Create lagged variables with explicit dplyr reference
    sdg_lag = dplyr::lag(sdg_overall, n = 1),
    spi_lag = dplyr::lag(spi_comp, n = 1),
    di_lag = dplyr::lag(di_score, n = 1),
    
    # Calculate first differences
    fd_sdg = sdg_overall - sdg_lag,
    fd_spi = spi_comp - spi_lag,
    fd_di = di_score - di_lag
  ) %>%
  ungroup()

# To verify the lag creation works correctly
#head(fd_data_testing %>% dplyr::select(country_code, year, sdg_overall, sdg_lag, fd_sdg, spi_comp, spi_lag, fd_spi) %>% arrange(country_code, year), 10)
```

**PART 2: Run FD models for each income group**

```{r}
# Overall model
fd_overall <- lm(fd_sdg ~ fd_spi, 
                data = fd_data_testing %>% 
                  filter(!is.na(fd_sdg) & !is.na(fd_spi)))
summary(fd_overall)

# High income countries
fd_high <- lm(fd_sdg ~ fd_spi, 
             data = fd_data_testing %>% 
               filter(!is.na(fd_sdg) & !is.na(fd_spi) & 
                     income_level == "High Income Countries"))
summary(fd_high)

# Upper-middle income countries
fd_upper <- lm(fd_sdg ~ fd_spi, 
              data = fd_data_testing %>% 
                filter(!is.na(fd_sdg) & !is.na(fd_spi) & 
                      income_level == "Upper-Middle Income Countries"))
summary(fd_upper)

# Lower-middle income countries
fd_lower <- lm(fd_sdg ~ fd_spi, 
              data = fd_data_testing %>% 
                filter(!is.na(fd_sdg) & !is.na(fd_spi) & 
                      income_level == "Lower-Middle Income Countries"))
summary(fd_lower)

# Low income countries
fd_low <- lm(fd_sdg ~ fd_spi, 
            data = fd_data_testing %>% 
              filter(!is.na(fd_sdg) & !is.na(fd_spi) & 
                    income_level == "Low Income Countries"))
summary(fd_low)
```
- No need for time variable


## R Stan, GLM, Markove  
```{r}
library(rstanarm)

gml_testing <- stan_glm(sdg_overall ~ spi_comp + regime_type + regime_type*spi_comp + log(gdppc) + population, data = merged_testing)

# glm_testing
print(gml_testing)
summary(gml_testing)
```

## Fixed Effects Model
```{r}
library(plm)
fe_model <- plm(sdg_overall ~ spi_comp + regime_type + gdppc + population,
                data = merged_testing, 
                index = c("country_code", "year"), 
                model = "within")

summary(fe_model)
```


## Testing Index Reliability (Cronbach's Alpha)

```{r}

```

