---
title: "backsliding_countries"
output: html_document
---

# preparation 
```{r}
setwd("~/Documents/GitHub/QMSS_Thesis_Sanchez")

#load libraries 
source("packages.R")

#load data 
event_data_all <- read_csv("data/Main CSV Outputs/merged_final_df.csv")

#selecting vars
event_data <- event_data_all %>% 
  dplyr::select(country_code, year, sdg_overall, spi_comp, di_score, aut_ep, dem_ep, regch_event, regime_type_2, regime_type_4, elect_dem, lib_dem, part_dem, delib_dem, egal_dem, academ_free, income_level, income_level_lab, log_gdppc, population) %>% 
  arrange(country_code, year) %>%  # Critical for correct lagging
  filter(year >= 2016)
```

# country_events dataframe: identifies/summarizes transition events per country
```{r}
#counting number of event-years: dem_ep & aut_ep
country_events <- event_data %>%
  group_by(country_code) %>%
  summarize(
    aut_ep_duration = sum(aut_ep == 1, na.rm = TRUE),
    dem_ep_duration = sum(dem_ep == 1, na.rm = TRUE), 
  )

# creating condition dummy variables
country_events <- country_events %>%
  mutate(
    has_aut_ep = (aut_ep_duration > 0) * 1, # experienced atleast 1 aut episode 
    has_dem_ep = (dem_ep_duration > 0) * 1, # experienced atleast 1 dem episode 
    either_event = (aut_ep_duration > 0 | dem_ep_duration > 0) * 1, # experience either kind of event
    neither_event = (aut_ep_duration == 0 & dem_ep_duration == 0) * 1 # experienced neither event 
  )

# labels for sorting 
country_events <- country_events %>%
  mutate(event_pattern = case_when(
    has_aut_ep == 1 & has_dem_ep == 1 ~ "Both Events", #must go first to label both events 
    has_aut_ep == 1 ~ "Backsliding Only",
    has_dem_ep == 1 ~ "Democratizing Only",
    neither_event == 1 ~ "Neither Event",
  ))

# Creating summary statistics 
summary_counts <- country_events %>%
  summarize(
    num_countries_with_aut_events = sum(has_aut_ep),
    num_countries_with_dem_events = sum(has_dem_ep),
    num_countries_with_events = sum(either_event),
    num_countries_without_events = sum(neither_event), 
    total_countries = n()
  )

# frequency
print(summary_counts)

# proportions 
print(summary_counts / 167)

# frequency table of exclusive event patterns
table(country_events$event_pattern)

# proportion table of exclusive event patterns
prop.table(country_events$event_pattern) # CANNOT DO BC event_patterns IS A STRING <-  FIX LATERRR
```
**Frequency**
167: countries total

Within the time period of 2015-2023:  

60 countries experienced at least 1 episode of autocratization   
  *40 countries experienced ONLY autocratization*  
35 countries experienced at least 1 episode of democratization  
  *15 countries experienced ONLY democratization*  
75 countries experienced at least 1 episode of either autocratization or democratization  
  *20 countries experienced BOTH democratization and autocratization*  
92 countries experienced neither kind of episode, autocratization or democratization  

**Proportions**
- The proportion of countries that have experienced at least 1 episode of autocratization in the time period between 2015 and 2023 is 60/167 = 0.3592814 (35.93%)
- The proportion of countries that have experienced at least 1 episode of democratization in the time period between 2015 and 2023 is 35/167 = 0.2095808 (20.96%)
- The proportion of countries that have experienced either kind of transition episode (democratization or autocratization) in the time period between 2015 and 2023 is 75/167 =  0.4491018 (44.91%)
- The proportion of countries that have NOT experienced any kind of transition episode the time period between 2015 and 2023 is 92/167 = 0.5508982 (55.09%)

# merging country_events df with event_data
```{r}
# Adding country_events data to event_data
event_data <- event_data %>%
  left_join(country_events, by = 'country_code')
  #filter(has_aut_ep == TRUE | has_dem_ep == TRUE) #FILTERS TO ONLY CHANGING REGIMES

# subgroups: regressing and democratizing coutries 
aut_data <- event_data %>% filter(has_aut_ep == 1) ## experienced at least 1 backsliding event 
dem_data <- event_data %>% filter(has_dem_ep == 1) ## experienced at least 1 democratizing event 
```

# complete regime change
```{r}
# Frequency/proportion table: instances of backsliding, stability, democratization from 2004-2024
table(event_data$regch_event) # frequency
prop.table(table(event_data$regch_event)) # proportions 

#Frequency of countries that experienced democratization 2015-2024
hist(dem_data$year[dem_data$regch_event==1], breaks = 20)

#Frequency of countries that experienced autocratization 2015-2024
hist(aut_data$year[aut_data$regch_event==-1], breaks = 20)

#Frequency of countries that experienced no change 2015-2024 [FIX HERE AND BELOW]
hist(merged$year[merged$regch_event==0], breaks = 20)
```




# Backsliding FD Models 
```{r}
###### SPI ~ DI (M ~ X) ######
fd_spi_di_aut <- plm(
  formula = spi_comp ~ di_score + log_gdppc,
  index = c("country_code", "year"),
  data = aut_data, 
  model = "fd" #FD
) 
# Country-level custard Robust SEs
summary(fd_spi_di_aut, vcov = vcovHC(fd_spi_di_aut, cluster = "group", type = "HC1"))

###### SDG ~ SPI (Y ~ M) ######
fd_sdg_spi_aut <- plm(
  formula = sdg_overall ~ spi_comp + di_score + log_gdppc,
  index = c("country_code", "year"),
  data = aut_data, 
  model = "fd"
)
# Country-level custard Robust SEs
summary(fd_sdg_spi_aut, vcov = vcovHC(fd_sdg_spi_aut, cluster = "group", type = "HC1"))
```
On average, when backsliding countries experience a change in Democracy score, is there a change in statistical capacity (SPI)? On average, when backsliding countries experience a change in statistical capacity (SPI) do they similalry experience a change in SDG score? 

# Backsliding FE Models
```{r}
###### SPI ~ DI (M ~ X) ######
fe_spi_di_aut <- plm(
  formula = spi_comp ~ di_score + log_gdppc + factor(year),
  index = c("country_code", "year"),
  data = aut_data,
  model = "within" 
)
summary(fe_spi_di_aut, vcov = vcovHC(fe_spi_di_aut, cluster = "group", type = "HC1")) # Robust SEs

###### SDG ~ SPI (Y ~ M) ######
fe_sdg_spi_aut <- plm(
  sdg_overall ~ spi_comp + di_score + log_gdppc + factor(year),
  index = c("country_code", "year"),
  data = aut_data,
  model = "within" 
)
summary(fe_sdg_spi_aut, vcov = vcovHC(fe_sdg_spi_aut, cluster = "group", type = "HC1")) # Robust SEs
```

# Single Country: Hungary 
```{r}
#filter for hungary 
hun_regressed <- aut_data %>% 
  filter(country_code == "HUN")

# FE model - Hungary 
fe_testing <- plm(
  formula = spi_comp ~ di_score + log_gdppc,
  data = aut_data, sub = (country_code == 'HUN'),
  index = c("country_code", "year"),
  model = "within")
summary(fe_testing, vcov = vcovHC(fe_testing, type = "HC3"))
```

#testing for autocorrelation 
```{r}
lmtest::bgtest(fe_testing)
lmtest::bgtest(fe_testing, order = 1)
```


```{r}
# subset of countries that experienced atleast 1 democratization event 
democratized <- merged %>% 
  group_by(country_code) %>%
  filter(dem_ep == 1) %>%
  ungroup()
```
