---
title: "Component 1: Multiple OLS, Interactions, Subgroups"
author: "Sevastian Sanchez"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  word_document: default
  html_document: default
---

# All Countries, Preliminary Analysis (SPI x SDGs)

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  tidy = TRUE,
  tidy.opts = list(width.cutoff = 60)  # Adjust width as needed
)
```

#FIRST: Libraries, Directory & Data 

```{r include=FALSE}
# set working directory 
setwd("~/Documents/GitHub/QMSS_Thesis_Sanchez")

#load libraries 
source("packages.R")
```

#SECOND: Run function in r-script: df_years_function.R   
*[ADJUST TIME OR SKIP AND LOADING DATA FROM DIRECTORY]*

```{r eval=FALSE, include=FALSE}
#df_years2.0():[Updated Version] used for extracting all available country data (based on country_codes) from specified range of years

#load function: all available data within year range 
source("df_years2.0_Function.R")
#specify start & end year
#merged_full <- df_years2.0(2004, 2023) 

#saving full version as CSV output 
#write.table(merged_full, file = 'merged_full_df.csv', row.names=F, sep = ",")

# Older Version, df_years() - Archived:
#source("df_years()_Function.R")
#merged2 <- df_years(yr1=2004) %>% 
#  select(country_name.x, country_code, year, year_fct, sdg_overall, spi_comp, sci_overall, di_score, regime_type_2, #regime_type_4, regch_event, income_level, income_level_lab, gini, log_gdppc, population)
```

# THIRD: Load and Refine Data   
*[SKIP IF LOADING FROM DIRECTORY]*
```{r eval=FALSE, include=FALSE}
# load merged_full csv [SKIP IF YOU CHANGED YEARS ABOVE]
#merged_full <- read_csv('data/Main CSV Outputs/merged_full_final_df.csv')

# FILTERING FOR ALL COUNTRIES WITH ATLEAST 1 SDG VALUE 
sdg_cols <- paste0("goal", 1:17)

# Identify & remove countries where all SDG columns are NA across all years
countries_all_sdg_missing <- merged_full %>%
  group_by(country_code) %>%
  summarise(all_missing = all(if_all(all_of(sdg_cols), is.na))) %>%
  filter(all_missing) %>%
  pull(country_code)
merged_cleaned <- merged_full %>% # Removing all rows for those countries
  filter(!country_code %in% countries_all_sdg_missing)

#saving merged CSV output 
#write.table(merged_cleaned, file = 'merged_cleaned.csv', row.names=F, sep = ",")
```

# FOURTH: Load cleaned 'merged' Dataset  
*[ADJUST VARIABLES OR SKIP IF LOADING FROM CSV]*
```{r eval=FALSE, include=FALSE}
#Load final cleaned 'merged' Dataset & selecting variables 
merged_final <- read_csv("data/Main CSV Outputs/merged_cleaned.csv") %>% 
  dplyr::select(country_name, country_code, year, year_fct, sdg_overall, spi_comp, sci_overall, di_score, regime_type_2, regime_type_4, regch_event, aut_ep, dem_ep, income_level, income_level_lab, gini, log_gdppc, population, p1_use, p2_services, p3_products, p4_sources, p5_infra, goal1, goal2, goal3, goal4, goal5, goal6, goal7, goal8, goal9, goal10, goal11, goal12, goal13, goal14, goal15, goal16, goal17)

#writing final merged csv 
#write.table(merged_final, file = 'merged_final_df.csv', row.names=F, sep = ",")
```

# FINALLY: LOAD FINAL MERGED CSV 
```{r}
#load final merged df 
merged <- read_csv("data/Main CSV Outputs/merged_final_df.csv")
```

# COMPONENT 1: COMPARING SPI & SCI X VARIABLES

**Aggregated SPI & SDG Scores**  
*H0: Null, there is no relationship*  
*H1: there is a statistically significant relationship between overall SPI and SDG composite scores*

```{r}
#correlation coefficients (r-squared), WITHOUT control variables 

#x-var 1 = spi
correlation_sdg_spi <- cor(merged$sdg_overall, merged$spi_comp, use = "complete.obs")^2

#x-var 2 = sci
correlation_sdg_sci <- cor(merged$sdg_overall, merged$sci_overall, use = "complete.obs")^2 

#x-var 3 = di
correlation_sdg_di <- cor(merged$sdg_overall, merged$di_score, use = "complete.obs")^2 

# pasting result 
string_corcoef <- "Correlation coefficient:" 
paste(string_corcoef, correlation_sdg_spi, "(SPI)", correlation_sdg_sci, "(SCI)", correlation_sdg_di, "(DI)") 
```

Correlation coefficient/R-sq (SPI): 0.616037202309322  
Correlation coefficient/R-sq (SCI): 0.410940651230861
Correlation coefficient/R-sq (DI): 0.452968121616442

## NAIVE OLS: Comparing SPI & SCI w/o controls 
Finding estimated impact of variables on SDG status prior to adding controls or robust SEs
```{r}
# 2. OLS for SPI and SDG - Overall 
ols_spi_naive <- lm(sdg_overall ~ spi_comp, data = merged)
summary(ols_spi_naive)

# 2. OLS for SCI and SDG - Overall 
ols_sci_naive <- lm(sdg_overall ~ sci_overall, data = merged)
summary(ols_sci_naive)

# 3. Multiple Regression with both SPI and SCI
ols_multiple_naive <- lm(sdg_overall ~ spi_comp + sci_overall, data = merged)
summary(ols_multiple_naive)
```
ols_spi_naive: 0.47806 (p-value < 0.001)  
ols_sci_naive: 0.39081 (p-value < 0.001)  
ols_multiple_naive: spi: 0.28779 (p-value < 0.001); sci: 0.15311 (p-value < 0.001)

The impact of SCI on SDG and SPI on SDG are statistically significant, in all models. SPI appears to have a greater impact on SDGs compared to that of SCI, regardless of the model. All of this is without controls. 

#Checking for Heteroskedasticity: residual plots 
```{r}
#residual plots 
plot(ols_spi_naive, which = 1)  # SPI model  
plot(ols_sci_naive, which = 1)  # SDG model 
plot(ols_multiple_naive, which = 1)  # SDG model controlled
```
U-shaped residuals detected, suggests non-linearity of x-variable terms. Additional tests reconfirm non-linearity (See Breusch-Pagan Test below). 

## TEST 1: Comparing SPI & SCI WITH controls AND Robust Standard Errors  
Applying controls and robust (Huber-White) standard errors  

*H0: Null, SCI model > SPI model*   
*H1: SPI model > SCI model*

```{r}
# 1. OLS for SPI and SDG - Overall 
ols_spi <- lm(sdg_overall ~ spi_comp + log_gdppc + population +  di_score + year_fct, data = merged)
summary(ols_spi)
coeftest(ols_spi, vcov = vcovHC(ols_spi, type = "HC1"))

# 2. OLS for SCI and SDG - Overall 
ols_sci <- lm(sdg_overall ~ sci_overall + log_gdppc + population + di_score + year_fct, data = merged)
summary(ols_sci)
coeftest(ols_sci, vcov = vcovHC(ols_sci, type = "HC1"))

# 3. Multiple Regression with both SPI and SCI
ols_multiple <- lm(sdg_overall ~ spi_comp + sci_overall + log_gdppc + population + di_score + year_fct, data = merged)
summary(ols_multiple)
coeftest(ols_multiple, vcov = vcovHC(ols_multiple, type = "HC1"))
```

# Creating dataframe combining summary and Robust SE results 
```{r}
# For model statistics extraction 
summary_spi <- summary(ols_spi)
summary_sci <- summary(ols_sci)
summary_multiple <- summary(ols_multiple)

# extracting robust SEs and coefficients (using coeftest)
rob_stats_spi <- coeftest(ols_spi, vcov = vcovHC(ols_spi, type = "HC1"))
rob_stats_sci <- coeftest(ols_sci, vcov = vcovHC(ols_sci, type = "HC1"))
rob_stats_multiple <- coeftest(ols_multiple, vcov = vcovHC(ols_multiple, type = "HC1"))

#SPI Statistics DF
spi_df <- data.frame(
  model = "M1: ols_spi",
  term = rownames(rob_stats_spi),
  estimate = rob_stats_spi[, 1],
  std.error = rob_stats_spi[, 2],
  t.statistic = rob_stats_spi[, 3],
  p.value = rob_stats_spi[, 4],
  residual.SE = summary_spi$sigma,
  r.squared = summary_spi$r.squared,
  adj.r.squared = summary_spi$adj.r.squared,
  row.names = NULL
)

#SCI Statistics DF
sci_df <- data.frame(
  model = "M2: ols_sci",
  term = rownames(rob_stats_sci),
  estimate = rob_stats_sci[, 1],
  std.error = rob_stats_sci[, 2],
  t.statistic = rob_stats_sci[, 3],
  p.value = rob_stats_sci[, 4],
  residual.SE = summary_sci$sigma,
  r.squared = summary_sci$r.squared,
  adj.r.squared = summary_sci$adj.r.squared,
  row.names = NULL
)

#Combined Mod Statistics DF
multiple_df <- data.frame(
  model = "M3: ols_multiple",
  term = rownames(rob_stats_multiple),
  estimate = rob_stats_multiple[, 1],
  std.error = rob_stats_multiple[, 2],
  t.statistic = rob_stats_multiple[, 3],
  p.value = rob_stats_multiple[, 4],
  residual.SE = summary_multiple$sigma,
  r.squared = summary_multiple$r.squared,
  adj.r.squared = summary_multiple$adj.r.squared,
  row.names = NULL
)

# Bind all together into one tidy dataframe
robust_mods_df <- bind_rows(spi_df, sci_df, multiple_df)

# Attributes under column names 
attr(robust_mods_df$std.error, "label") <- "Robust Std. Errors Adjusted"
attr(robust_mods_df$t.statistic, "label") <- "Robust Std. Errors Adjusted"
attr(robust_mods_df$p.value, "label") <- "Robust Std. Errors Adjusted"

#save to output_CSVs folder
write.csv(robust_mods_df, file = "output_CSVs/ols_mods_results.csv")

# View the result
print(robust_mods_df)
```

We reject the null hypothesis that there is no relationship between SPI and SDG composite scores. Additionally, we reject the null hypothesis that there is no relationship between SCI and SDG composite scores. Holding all else constant (log GDP per capita, democracy score and population), SPI and SCI exhibit positive moderate and statistically significant relationships with SDG status.

ols_spi: 0.28641 (p-value < 0.001)  
ols_sci: 0.23976 (p-value < 0.001)  
ols_multiple: spi: 0.11945 (p-value < 0.001); sci: 0.14876 (p-value < 0.001)

When compared in separate models, SPI has a greater impact on SDG status (0.28641) than SCI (0.23976). This suggests that a one-unit increase in SPI is associated with a larger improvement in SDG outcomes compared to a one-unit increase in SCI, holding all controls constant. 

Interestingly, the opposite holds true in a multiple regression model containing both SPI and SCI. SPI's impact on SDG status (0.11945) (net of SPI) is less than that of SCI's (0.14876) (net of SCI), holding all controls constant. When together, the coefficients represent the unique impact of each predictor variable (measures of statistical capacity) on SDG status, net of all other variables.

Model 1 (ols_spi) does not control for SCI and model 2 (ols_sci) does not control for spi -- this is okay. SPI is the predecessor of the SCI, sharing/data overlap, and so it is expected to have significant statistical correlation (multicollinearity). This is likely what explains the significant reduction of both coefficients as seen in model 3: 0.28641 to 0.11945 for SPI (58.29% decrease); and from 0.23976 to 0.14876 for SCI (37.95% decrease). Such indicates that they're both capturing much of the same underlying relationship with SDG status.

However, the fact that both SPI and SCI remain significant when included together (model 3) with a high adjusted R-sq (0.7614) suggests that they capture different dimensions of statistical capacity that independently contribute to SDG status.

### Checking for Multicolinearity: VIF of SPI & SCI 

```{r}
# Check correlation between SPI and SCI
cor(merged$spi_comp, merged$sci_overall, use = "complete.obs")

# Check VIF (Variance Inflation Factor) in Model 3
vif(ols_multiple)

#make into Datatable 
vif_vals <- vif(ols_multiple)   # returns a named vector
tidy_vif <- enframe(vif_vals, name = "term", value = "vif")
print(tidy_vif)
```

**colinearity:** The correlation between SCI and SPI is about 0.8277. When placed within the same model, SCI inflated the standard error of SPI from 0.01396 to 0.02671. SCI had a similar reaction from the SPI with its standard error increasing from 0.00969 to 0.02398.

**VIF:** Such multicollinearity is reflected by the VIF test which accounts for all x variables in the model instead of just the two measures of statistical capacity (SCI & SPI).

_VIF Results:_
  term          vif
1 spi_comp     4.29
2 sci_overall  3.79
3 log_gdppc    1.48
4 population   1.03
5 di_score     1.60
6 year_fct     1.29 (categorical)

Overall there reveals no severe multicollinearity (all GVIF < 5). There is moderate correlation between statistical capacity measures (spi_comp and sci_overall) with SPI moderately inflated by a factor of 4.29 and SCI inflated by a factor of  3.79. Nevertheless, it is acceptable to include both in the same model as doing so will not severely impact estimates with both factors less than 5.0. Even so, there are significant limitations in either model that warrant strong consideration, including sample size, and longitudinal suitability. All other variables show minimal multicollinearity concerns.

### Checking misspecification missing non-linear or omitted interactions
```{r}
#lm test package 
resettest(ols_spi, power = 2:3, type = "fitted")
resettest(ols_sci, power = 2:3, type = "fitted")
resettest(ols_multiple, power = 2:3, type = "fitted")
```
All three models show statistically significant evidence of misspecification. Given the high variablility of statistical capacity measures and control variables like GDP Per Capita and Total Population, misspecification here is likely to be the result of ommitted interaction terms or heteroskedasticity.

As pointed out by (AUTHOR) who maintained that ____. 
Thus, to test for the existence of non-linear omitted variables, or lack thereof, I deploy a Breusch-Pagan test for heteroskedasticity.

### Checking for Heteroskedasticity: Breusch-Pagan Test 
This validates the need for integrating robust standard errors in our models
```{r}
#Breusch-Pagan tests 
bptest(ols_spi)
bptest(ols_sci)
bptest(ols_multiple)

#make into objects 
bp_spi <- bptest(ols_spi)
bp_sci <- bptest(ols_sci)
bp_multiple <- bptest(ols_multiple)

# combine for data frame 
bp_tests <- list(
  ols_spi = bp_spi,
  ols_sci = bp_sci,
  ols_multiple = bp_multiple
)

# Tidy all tests and add a "model" column
df_bptests <- bp_tests %>% 
  map_df(~ tidy(.x), .id = "model")


#write.table(df_bptests, file = 'output_CSVs/df_bptests_heterosked.csv', row.names=F, sep = ",")
```
Model:         BP statistic   p-value
ols_spi           192.11       < 2.2e-16        
ols_sci           101.71       < 2.2e-16 
ols_multiple      36.54       < 2.169e-06

The Breusch-Pagan Test was applied to test to see whether residuals are constant across observations, which signals unaccounted non-linear relationships, especially with macro factors such as GDP Per Capita and Population in the models. This is important because Ordinary Least Squares models assume constant error variance. In such a complex world of diverse cultural and everchanging political structures across 200 countries, cross-national data, especially in development, is rarely ever linear. Accordingly, this test evaluates the extent of such non-linearity among specified predictors. 

As such, results indicate strong evidence of heteroskedasticity in all three models. The small p-values in all models indicates that the variance of residuals are not constant across observations in all three models. This reinforces the motivation behind applying robust standard errors, which have been integrated to all OLS models. Without Robust SEs, there is a risk of inflated t-statistics, leading to false significance and misinterpretation of results.

Despite the improvement from 192.11 (SPI) and 101.71 (SCI) to 36.54 (Both), there still remains statically significant heteroskedasticity in the combined model. Both statistical capacity measures create a better-specified model (ols_multiple), though not enough to eliminate heteroskedasticity entirely.

### Missing Data Structure & Interpretations 
**Systematic, non-random missing data pattern:** SPI has near complete country data coverage (165 out of 168 countries with an SDG score), but with a stubborn temporal limitation (2016-2023). On the other hand, SCI has longer temporal coverage (2004-2020), but lacks reporting on high-income countries focusing primarily on the developing world (123 out of 168 countries with an SDG score). 

In model 1 (SDG ~ SPI), democracy score (di_score) is not statistically significant (-0.0243, p=0.8673). However, in model 2 (SDG ~ SCI) democracy score is highly significant (0.5556, p < 0.001). In model 3, with both SCI and SPI, democracy score is marginally significant (0.0301, p=0.0484). This suggests that SCI's relationship with SDG outcomes may be closely linked to regime/democratic governance. Considering the non-random missing data structured previously mentioned, the difference in significance between the models for democracy score signals an even greater need to perform subgroup analysis of countries at different stages/levels of development. 

### AIC/BIC Checking Fit

```{r}
# Compare all three models with AIC
AIC(ols_spi, ols_sci, ols_multiple)

# Compare all three models with BIC
BIC(ols_spi, ols_sci, ols_multiple)

# INTO DATAFRAME 
aic_vals <- c(
  AIC(ols_spi),
  AIC(ols_sci),
  AIC(ols_multiple)
)

bic_vals <- c(
  BIC(ols_spi),
  BIC(ols_sci),
  BIC(ols_multiple)
)

# Model names
model_names <- c("ols_spi", "ols_sci", "ols_multiple")

# Combine into a dataframe
aic_bic_ols_results <- data.frame(
  model = model_names,
  AIC = aic_vals,
  BIC = bic_vals
)

print(aic_bic_ols_results)

# saving to output_CSVs
write.csv(aic_bic_ols_results, file = "output_CSVs/aic_bic_ols_results.csv")
```
**AIC/BIC Results**
AIC(ols_spi, ols_sci, ols_multiple)
             df      AIC
ols_spi       7 6482.761
ols_sci       7 9006.394
ols_multiple  8 3279.338

BIC(ols_spi, ols_sci, ols_multiple)
             df      BIC
ols_spi       7 6517.673
ols_sci       7 9043.652
ols_multiple  8 3314.060

**Adjusted R-squares**
ols_spi: 0.7725
ols_sci: 0.7383
ols_multiple: 0.7614

## Selecting Best Model
**Best fit:** ols_spi (Adj Rsq: 0.7725) (AIC/BIC: 6482.761, 6517.673) (n=1082)
Model 3 (ols_multiple) sacrifices a significant portion of its sample size in order to include both statistical capacity measures. While ols_multiple appears to outperform the other two models, the lower AIC/BIC partially reflects its smaller sample size, not necessarily a better model fit. Model 1 (ols_spi) provides better country coverage and sustains slightly higher explanatory power (0.7725 > 0.7614) than model 3.

**Selecting model:** This analysis is specifically focused on overall statistical capacity rather than comparisons of measures. Model 1 (ols_spi) reveals a better adjusted R-squares value than that of model 2 (Adj Rsq: 0.7725 > 0.7383) and model 3 (Adj Rsq: 0.7725 > 0.7614). Model 3, containing both SPI and SCI, has greater explanatory power than model 2, but as mentioned, has a much smaller sample size, which significantly impedes results. Models 1 and 2 have significantly more country-year data points (n=1082 and n=1513, respectively) for regression analysis compared to model 3 (n=566). With all else considered, this study employs the Statistical Performance Index (SPI) as the primary measure of statistical capacity. 

### Visual Analysis of Fit: SCI & SPI x SDG
```{r}
#define regression line colors
spi_line <- "steelblue4"
sci_line <- "darkgoldenrod"

# Creating scatterplot with both SPI and SCI on the same plot
Compare_fit <- ggplot(merged, aes(x = spi_comp, y = sdg_overall))+
  geom_smooth(aes(x = spi_comp, y = sdg_overall), 
              color = spi_line, 
              method = "lm", 
              linewidth = 0.75, 
              se = FALSE)+ # Regression line SPI
  geom_smooth(aes(x = sci_overall, y = sdg_overall), 
              color = sci_line, 
              method = "lm", 
              linewidth = 0.75, 
              se = FALSE)+ # Regression line for SCI
  geom_point(aes(color = "spi_comp"), alpha=0.50, size = 0.5)+ # Scatter plot for SPI
  geom_point(aes(x = sci_overall, y = sdg_overall, color = "sci_overall"), 
             alpha=0.5, size = 0.5)+ # Add SCI points w/different color
  scale_color_manual(values = c("spi_comp" = "steelblue1", 
                                "sci_overall" = "darkgoldenrod1")) +
  labs(title = "Comparing SPI & SCI Measures Against SDG Index",
       x = "SPI & SCI Scores (0-100)",
       y = "SDG Composite Scores (0-100)",
       color = "Statistical Capacity Measure") +  # Title for legend
  theme_bw() # Optional: adds a clean, black and white theme

Compare_fit

#make interactive
#ggplotly(Compare_fit)

# Save to specific folder
ggsave("figures/spi_sci_plot.png", Compare_fit, width = 10, height = 6)
```
The SPI regression line is expected to appear higher in terms of SDG Score compared the SCI model because SPI countries include higher-income countries. As previously mentioned, the SCI soley focuses on lower to upper-middle income countries (146 countries over 17 years). 

# INTERACTIONS AND SUBGROUP ANALYSIS

## TEST 2: Checking for Interactions [REDO RESULTS FOR ROBUST SEs]: 
- Is there a need for subgroup analysis, and if so, by what kind of group? 
- Options: GNI Classification (income_level), regime_type_2, regime_type_4, di_score
```{r}
merged_2015 <- merged %>% 
  filter(year > 2015) %>% 
  mutate(regime_type_2 = as.factor(regime_type_2), 
         regime_type_4 = as.factor(regime_type_4))

#interaction 1: does GNI Classification (income_level) affect the relationship between x (spi) & y (sdg)?
inc_lev_interaction <- lm(sdg_overall ~ spi_comp + spi_comp*income_level + log_gdppc + population + year_fct, data = merged_2015)
#summary(inc_lev_interaction)
coeftest(inc_lev_interaction, vcov = vcovHC(inc_lev_interaction, type = "HC1")) #Robust SE

#interaction 2: does regime_type_2 affect the relationship between x (spi) & y (sdg)?
reg_type2_interaction <- lm(sdg_overall ~ spi_comp + spi_comp*regime_type_2 + log_gdppc + population + year_fct,data = merged_2015)
#summary(reg_type2_interaction)
coeftest(reg_type2_interaction, vcov = vcovHC(reg_type2_interaction, type = "HC1")) #Robust SE

#interaction 3: does regime_type_4 affect the relationship between x (spi) & y (sdg)?
reg_type4_interaction <- lm(sdg_overall ~ spi_comp + spi_comp*regime_type_4 + log_gdppc + population + year_fct,
                        data = merged_2015)
#summary(reg_type4_interaction)
coeftest(reg_type4_interaction, vcov = vcovHC(reg_type4_interaction, type = "HC1")) #Robust SE

#interaction 4: does di affect the relationship between x (spi) & y (sdg)?
reg_type_di_interaction <- lm(sdg_overall ~ spi_comp + spi_comp*di_score + log_gdppc + population + year_fct,
                        data = merged_2015)
#summary(reg_type_di_interaction)
coeftest(reg_type_di_interaction, vcov = vcovHC(reg_type_di_interaction, type = "HC1")) #Robust SE
```
*Interaction 1: GNI Income Classification: Yes* there are statistically significant interactions found from GNI classifications that affects the relationship between spi and sdgs

*Interaction 2: Binary Regime Type: Yes* there are statistically significant interactions (mildly, p = 0.0566) found from regime type (autocracy vs democracy) that affects the relationship between spi and sdgs. 

*Interaction 3: Categorical Regime type (4 options): No* there are statistically significant interactions found depending on regime type (Closed autocracy, electoral autocracy, electoral democracy, liberal democracy) that affects the relationship between spi and sdgs. 

*Interaction 4: Continuous di_score [0-1] Regime type: No* there are no statistically significant interactions found from regime type (infinate between 0-10) that affects the relationship between spi and sdgs. 

## Interactions DF
```{r}
################ in a table #################
ct <- coeftest(reg_type_di_interaction, vcov = vcovHC(reg_type_di_interaction, type = "HC1"))

# Convert to tidy dataframe
ct_tidy <- tidy(ct)

# In stargazer 
stargazer(ct_tidy, type = "text", summary = FALSE, rownames = FALSE)

# In tidy table 
ct_tidy
```


## TEST 3: WB GNI Classifications: income_level ("H", "UM", "LM", "L") 
**Disaggregated/Grouped by Development Status:** Make 4 regression models and then put them all together in a table to compare the slopes and R-sq values.
```{r}
# 1. Overall model (all countries)
overall_lm <- lm(sdg_overall ~ spi_comp + di_score + log_gdppc + population + year_fct,
                        data = merged_2015)
#summary(overall_lm)
coeftest(overall_lm, vcov = vcovHC(overall_lm, type = "HC1")) #Robust SE

# 2. High income countries
high_inc_lm <- lm(sdg_overall ~ spi_comp + di_score + log_gdppc + population + year_fct,
                        data = merged_2015 %>% 
                          filter(income_level == "H"))
#summary(high_inc_lm)
coeftest(high_inc_lm, vcov = vcovHC(high_inc_lm, type = "HC1")) #Robust SE

# 3. Upper-middle income countries
upper_mid_lm <- lm(sdg_overall ~ spi_comp + di_score + log_gdppc + population + year_fct,
                         data = merged_2015 %>% 
                           filter(income_level == "UM"))
#summary(upper_mid_lm)
coeftest(upper_mid_lm, vcov = vcovHC(upper_mid_lm, type = "HC1")) #Robust SE

# 4. Lower-middle income countries
lower_mid_lm <- lm(sdg_overall ~ spi_comp + di_score + log_gdppc + population + year_fct,
                         data = merged_2015 %>% 
                           filter(income_level == "LM"))
#summary(lower_mid_lm)
coeftest(lower_mid_lm, vcov = vcovHC(lower_mid_lm, type = "HC1")) #Robust SE

# 5. Low income countries
low_inc_lm <- lm(sdg_overall ~ spi_comp + di_score + log_gdppc + population + year_fct,
                       data = merged_2015 %>% 
                         filter(income_level == "L"))
#summary(low_inc_lm)
coeftest(low_inc_lm, vcov = vcovHC(low_inc_lm, type = "HC1")) #Robust SE
```

# GNI Classifications DF
```{r}
# 1. Overall model
overall_ct <- coeftest(overall_lm, vcov = vcovHC(overall_lm, type = "HC1"))
overall_df <- data.frame(
  model = "overall_lm",
  term = rownames(overall_ct),
  estimate = overall_ct[, "Estimate"],
  std.error = overall_ct[, "Std. Error"],
  t.statistic = overall_ct[, "t value"],
  p.value = overall_ct[, "Pr(>|t|)"],
  n = nobs(overall_lm)
)

# 2. High income countries
high_inc_ct <- coeftest(high_inc_lm, vcov = vcovHC(high_inc_lm, type = "HC1"))
high_inc_df <- data.frame(
  model = "high_inc_lm",
  term = rownames(high_inc_ct),
  estimate = high_inc_ct[, "Estimate"],
  std.error = high_inc_ct[, "Std. Error"],
  t.statistic = high_inc_ct[, "t value"],
  p.value = high_inc_ct[, "Pr(>|t|)"],
  n = nobs(high_inc_lm)
)

# 3. Upper-middle income countries
upper_mid_ct <- coeftest(upper_mid_lm, vcov = vcovHC(upper_mid_lm, type = "HC1"))
upper_mid_df <- data.frame(
  model = "upper_mid_lm",
  term = rownames(upper_mid_ct),
  estimate = upper_mid_ct[, "Estimate"],
  std.error = upper_mid_ct[, "Std. Error"],
  t.statistic = upper_mid_ct[, "t value"],
  p.value = upper_mid_ct[, "Pr(>|t|)"],
  n = nobs(upper_mid_lm)
)

# 4. Lower-middle income countries
lower_mid_ct <- coeftest(lower_mid_lm, vcov = vcovHC(lower_mid_lm, type = "HC1"))
lower_mid_df <- data.frame(
  model = "lower_mid_lm",
  term = rownames(lower_mid_ct),
  estimate = lower_mid_ct[, "Estimate"],
  std.error = lower_mid_ct[, "Std. Error"],
  t.statistic = lower_mid_ct[, "t value"],
  p.value = lower_mid_ct[, "Pr(>|t|)"],
  n = nobs(lower_mid_lm)
)

# 5. Low income countries
low_inc_ct <- coeftest(low_inc_lm, vcov = vcovHC(low_inc_lm, type = "HC1"))
low_inc_df <- data.frame(
  model = "low_inc_lm",
  term = rownames(low_inc_ct),
  estimate = low_inc_ct[, "Estimate"],
  std.error = low_inc_ct[, "Std. Error"],
  t.statistic = low_inc_ct[, "t value"],
  p.value = low_inc_ct[, "Pr(>|t|)"],
  n = nobs(low_inc_lm)
)

# Combine all results
gni_classes_ols <- bind_rows(
  overall_df,
  high_inc_df,
  upper_mid_df,
  lower_mid_df,
  low_inc_df
)

attr(gni_classes_ols $std.error, "label") <- "Robust Std. Errors Adjusted"
attr(gni_classes_ols $t.statistic, "label") <- "Robust Std. Errors Adjusted"
attr(gni_classes_ols $p.value, "label") <- "Robust Std. Errors Adjusted"

gni_classes_ols

#interactive table for side access 
#datatable(gni_classes_ols, caption = "Regression Results, by GNI Country Classifications")

# write to directory 
write.csv(gni_classes_ols, "output_CSVs/gni_classes_ols.csv")
```

##Visualizing Slopes: plotting multiple regression - by subgroup
```{r}
viz_gni_class <- ggplot(data = merged_2015, aes(x = spi_comp, 
                                           y = sdg_overall, 
                                           color = income_level_lab)) +
  geom_point(alpha = 0.25, size = 0.75) + 
  # Overall regression line (black)
  geom_smooth(aes(group = 1), 
              method = "lm", 
              linewidth = 0.75, 
              se = FALSE, 
              color = "black") +
  # Group-specific regression lines
  geom_smooth(method = "lm", 
              linewidth = 0.65, 
              se = FALSE) +
  scale_color_manual(
    values = c("High Income Countries" = "#1D6A96", 
               "Upper-Middle Income Countries" = "#4CB5AE",
               "Lower-Middle Income Countries" = "#F3A738",
               "Low Income Countries" = "#C02942")
  ) +
  labs(title = "Relationship between SPI and SDG by World Bank Income Classification",
       x = "Statistical Performance Indicators (SPI)",
       y = "Sustainable Development Goals (SDG)",
       color = "Income Classification") +
  theme_bw() 

viz_gni_class
#ggplotly(viz_gni_class)

# Save to specific folder
ggsave("figures/gni_subgroups_ols.png", viz_gni_class, width = 10, height = 6)
```
[more results TBD]

## Coefficient & Interval Plot
```{r}
# New fd with SPI coefficients data
spi_plot_data <- data.frame(
  model = c("overall_lm", "high_inc_lm", "upper_mid_lm", "lower_mid_lm", "low_inc_lm"),
  estimate = c(0.286414727883271, 0.355563975838364, 0.174313752024353, 0.382217438772999, 0.169651614020134),
  std.error = c(0.0156271299809294, 0.0170293332032076, 0.0139294904502819, 0.0262707606446246, 0.0441323334102258)
)

# Calculate confidence intervals
spi_plot_data <- spi_plot_data %>%
  mutate(
    CI_lower = estimate - 1.96 * std.error,
    CI_upper = estimate + 1.96 * std.error
  )

# Set model order
model_order <- c("low_inc_lm", "lower_mid_lm", "upper_mid_lm", "high_inc_lm", "overall_lm")
spi_plot_data$model <- factor(spi_plot_data$model, levels = model_order)

# Create the coefficient plot
coef_inter_spi_plot <- ggplot(spi_plot_data, aes(x = estimate, y = model)) +
  geom_point(size = 3, color = "dodgerblue") +
  geom_errorbarh(aes(xmin = CI_lower, xmax = CI_upper), 
                 height = 0.2, 
                 color = "dodgerblue") +
  labs(
    title = "Coefficient Estimates with 95% Confidence Intervals for SPI by Income Group Models",
    x = "Coefficient Estimate for SPI",
    y = NULL
  ) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_x_continuous(limits = c(0, 0.45)) +
  scale_y_discrete(labels = c("low_inc_lm" = "Low Income", 
                              "lower_mid_lm" = "Lower-Middle Income",
                              "upper_mid_lm" = "Upper-Middle Income", 
                              "high_inc_lm" = "High Income",
                              "overall_lm" = "Overall"))

coef_inter_spi_plot

ggsave("figures/coef_inter_spi_plot.png", coef_inter_spi_plot, width = 9, height = 5)
```
