---
title: "Component 2, Stage 1: DI -> SPI"
output: html_document
---

# Set up and Wrangling 
```{r}
# set working directory
setwd("~/Documents/GitHub/QMSS_Thesis_Sanchez")

#load libraries/packages
source("packages.R")

# load data
source("Comp2_panel_wrangling.R")
```

# Stage 1 Models:
ols_spi_di
fd_spi_di
fe_spi_di

# 1.1) POLS [Stage 1]
The effect of democracy on SPI Performance
```{r}
# Full POLS model: SPI ~ DI
ols_spi_di <- plm(
  formula = spi_comp ~ dplyr::lag(di_score, n=1) + log_gdppc + income_level_recoded + factor(year), 
  index = c("country_code", "year"),
  model = "pooling", 
  data = panel_data)
summary(ols_spi_di, vcov = vcovHC(ols_spi_di, cluster = "group", type = "HC1"))

# plot relationship: lagged DI vs SPI 
spi_di_s1_scatter <- ggplot(panel_data, aes(x = dplyr::lag(di_score, 1), y = spi_comp)) + 
  geom_point(color = "steelblue", size = 1, alpha = 0.65) +
  geom_smooth(method = "lm", 
              se = TRUE, color = "darkblue", size = 1) +
  labs(
    title = "Effect of Democracy Levels on Statistical Capacity",
    x = "Lagged Democracy Index (0-10 Scale)",
    y = "SPI Composite (0-100 Scale)"
  ) +
  theme_minimal()

spi_di_s1_scatter

# Save to specific folder
ggsave("figures/stage_1_n_2_scatterplots/spi_di_s1_scatterplot.png", spi_di_s1_scatter, width = 8, height = 6)
```

# 1.2) First Difference [Stage 1]
```{r}
# Mediator: SPI ~ DI (M ~ X) 
fd_spi_di <- plm(
  formula = spi_comp ~ dplyr::lag(di_score, n=1) + log_gdppc + income_level_recoded,
  index = c("country_code", "year"),
  data = fd_data, #sub = (has_aut_ep == TRUE),
  model = "fd"
) 
# Country-level custard Robust SEs
summary(fd_spi_di, vcov = vcovHC(fd_spi_di, cluster = "group", type = "HC1"))
```
**results** 
The first difference model indicates that for every positive change in the Democracy Index (DI) (from the year prior), it produces a negative 0.26 unit change in statistical capacity, on average, for the same country across 8 years of data. This suggests a negative relationship between democracy levels and statistical capacity performance.

# 1.3) Fixed Effects [Stage 1]
```{r}
# Mediator: SPI ~ DI (M ~ X)
fe_spi_di <- plm(
  formula = spi_comp ~ dplyr::lag(di_score, n=1) + log_gdppc + income_level_recoded,
  index = c("country_code", "year"),
  data = panel_data,
  model = "within", #FE
  effect = "twoways" 
)
summary(fe_spi_di, vcov = vcovHC(fe_spi_di, cluster = "group", type = "HC1"))
```
**results**
The fixed effects model indicates that for every positive change in the Democracy Index (DI) (from the year prior), it produces an inverse negative 0.27 unit change in statistical capacity, net of any particular country, across the first 7 years of this panel. [?]

### stargazer table for the above three models
```{r}
# Extracting robust standard errors for each model
rob_se_ols <- coeftest(ols_spi_di, vcov = vcovHC(ols_spi_di, cluster = "group", type = "HC1"))
rob_se_fd <- coeftest(fd_spi_di, vcov = vcovHC(fd_spi_di, cluster = "group", type = "HC1"))
rob_se_fe <- coeftest(fe_spi_di, vcov = vcovHC(fe_spi_di, cluster = "group", type = "HC1"))
# Create a list of robust standard errors for stargazer
robust_se_list <- list(
  rob_se_ols[, 2], 
  rob_se_fd[, 2], 
  rob_se_fe[, 2]
)
# Create a stargazer table for the three models
stargazer(
  ols_spi_di, fd_spi_di, fe_spi_di, 
  se = robust_se_list,
  type = "html",
  out = "figures/s1_base_models.html",
  title = "Stage 1: DI -> SPI Models",
  align = TRUE,
  dep.var.labels = "SPI Overall Score",
  column.labels = c("POLS", "FD", "FE"),
  report = "vcs*",
  omit = "factor\\(year\\)",  # Omits year fixed effects from display
  model.names = FALSE
)
```

# MODEL MISSPECIFICATION CHECKS [Stage 1]
Following standard econometric practice, I first establish the appropriate functional form for main effects, using RESET tests and residual diagnostics, before testing for theoretically motivated interaction effects.

## RESET Test for Misspecification [STAGE 1]
The RESET test checks for omitted variable bias by testing if the squared and cubed fitted values significantly improve the model fit. If the p-value is low (typically < 0.05), it suggests that the model may be misspecified, indicating that a non-linear relationship or omitted variables are present.

When you get a p-value of less than 0.05 this means that you reject the null hypothesis - this means that statistically your model is better with first order terms.
```{r}
# transforming to base R lm()
ols_spi_di <- lm(spi_comp ~ dplyr::lag(di_score, 1) + log_gdppc + income_level_recoded + factor(year), data = panel_data) # regular linear model
fd_spi_di <- lm(spi_diff ~ dplyr::lag(di_diff, 1) + log_gdppc_diff + income_level_recoded, data = fd_data)
fe_spi_di <- lm(spi_comp ~ dplyr::lag(di_score, 1) + log_gdppc + income_level_recoded + factor(year) + factor(country_code), data = panel_data)

# Performing RESET tests for the base models
resettest(ols_spi_di, power = 2:3, type = "regressor", 
          vcov = function(x) vcovHC(x, cluster = "group", type = "HC1"))
resettest(fd_spi_di, power = 2:3, type = "regressor",
          vcov = function(x) vcovHC(x, cluster = "group", type = "HC1"))
resettest(fe_spi_di, power = 2:3, type = "regressor",
          vcov = function(x) vcovHC(x, cluster = "group", type = "HC1"))

# Saving results to a dataframe
s1_reset_results <- data.frame(
  model = c("ols_spi_di", "fd_spi_di", "fe_spi_di"),
  F_statistic = c(resettest(ols_spi_di, power = 2:3, type = "regressor", 
                             vcov = function(x) vcovHC(x, cluster = "group", type = "HC1"))$statistic,
                  resettest(fd_spi_di, power = 2:3, type = "regressor",
                             vcov = function(x) vcovHC(x, cluster = "group", type = "HC1"))$statistic,
                  resettest(fe_spi_di, power = 2:3, type = "regressor",
                             vcov = function(x) vcovHC(x, cluster = "group", type = "HC1"))$statistic),
  p_value = c(resettest(ols_spi_di, power = 2:3, type = "regressor", 
                        vcov = function(x) vcovHC(x, cluster = "group", type = "HC1"))$p.value,
              resettest(fd_spi_di, power = 2:3, type = "regressor",
                        vcov = function(x) vcovHC(x, cluster = "group", type = "HC1"))$p.value,
              resettest(fe_spi_di, power = 2:3, type = "regressor",
                        vcov = function(x) vcovHC(x, cluster = "group", type = "HC1"))$p.value)
)
print(s1_reset_results)
# Save the results to a CSV file
write.csv(s1_reset_results, file = "output_CSVs/s1_reset_results.csv", row.names = FALSE)
```
We reject the Null Hypothesis of no misspecification in the relationship between DI and SPI performance, as the p-value is greater than 0.05 in in the FE model. This suggests that non-linear terms may make a significant contribution to the model.

## Stepwise Check: Applying Polynomial Terms [Stage 1]
For this section, all continuous predictors (DI & Log(GDP)) are centered for simplicity and to avoid multicollinearity issues. 

- **H0:** The relationship between SPI and SDG performance is linear.
- **H1:** The relationship between SPI and SDG performance is non-linear (quadratic or cubic).
```{r}
# No quatratic Terms Applied - Baseline 
fe_spi_di_base <- plm(formula = spi_comp ~ dplyr::lag(cen_di_score, n = 1) + cen_log_gdppc + income_level_recoded,
               model = "within", 
               index = c("country_code", "year"),
               effect = "twoways",
               data = panel_data)
rob_sum_base <-summary(fe_spi_di_base, vcov = vcovHC(fe_spi_di_base, cluster = "group", type = "HC1"))

# + Quadratic terms for DI
fe_quads_s1_m2 <- plm(formula = spi_comp ~ dplyr::lag(cen_di_score, n = 1) + dplyr::lag(cen_di_score_quad, n = 1) + cen_log_gdppc + income_level_recoded,
               model = "within", 
               index = c("country_code", "year"),
               effect = "twoways",
               data = panel_data)
rob_sum_m2 <- summary(fe_quads_s1_m2, vcov = vcovHC(fe_quads_s1_m2, cluster = "group", type = "HC1"))

# + Quadratic terms for DI and Log(GDP)
fe_quads_s1_m3 <- plm(formula = spi_comp ~ dplyr::lag(cen_di_score, n = 1) + dplyr::lag(cen_di_score_quad, n = 1)
                      + cen_log_gdppc + cen_log_gdppc_quad + income_level_recoded,
               model = "within", 
               index = c("country_code", "year"),
               effect = "twoways",
               data = panel_data)
rob_sum_m3 <- summary(fe_quads_s1_m3, vcov = vcovHC(fe_quads_s1_m3, cluster = "group", type = "HC1"))

# + Cubic terms for DI and Log(GDP)
fe_cubics_s1_m4 <- plm(formula = spi_comp ~ dplyr::lag(cen_di_score, n = 1) + dplyr::lag(cen_di_score_quad, n = 1) + dplyr::lag(cen_di_score_cubic, n = 1) + cen_log_gdppc + cen_log_gdppc_quad + cen_log_gdppc_cubic + income_level_recoded,
               model = "within", 
               index = c("country_code", "year"),
               effect = "twoways",
               data = panel_data)
rob_sum_m4 <- summary(fe_cubics_s1_m4, vcov = vcovHC(fe_cubics_s1_m4, cluster = "group", type = "HC1"))

# Extracting robust standard errors for each model
rob_se_base <- coeftest(fe_spi_di_base, vcov = vcovHC(fe_spi_di_base, cluster = "group", type = "HC1"))
rob_se_m2  <- coeftest(fe_quads_s1_m2, vcov = vcovHC(fe_quads_s1_m2, cluster = "group", type = "HC1"))
rob_se_m3  <- coeftest(fe_quads_s1_m3, vcov = vcovHC(fe_quads_s1_m3, cluster = "group", type = "HC1"))
rob_se_m4 <- coeftest(fe_cubics_s1_m4, vcov = vcovHC(fe_cubics_s1_m4, cluster = "group", type = "HC1"))

# stargazer table
library(stargazer)
stargazer(fe_spi_di_base, fe_quads_s1_m2, fe_quads_s1_m3, fe_cubics_s1_m4, 
          se = list(rob_se_base[, 2], rob_se_m2[, 2], rob_se_m3[, 2], rob_se_m4[, 2]
                    ),
          title = "Applying Polynomial Terms to Stage 1 Models",
          align = TRUE,
          dep.var.labels = "SPI Overall Score",
          column.labels = c("FE Base", "Quad DI", "Quad DI+GDP", "Cubic DI+GDP"),
          #covariate.labels = c("SPI", "SPI Squared", "DI", "DI Squared",
          #                     "Log GDPpc", "Log GDPpc Squared", "Intercept"),
          notes = "All models apply fixed effects. Robust standard errors clustered by country in parentheses",
          type = "html",  # Change to "latex" for LaTeX output or "html" for Word
          report = "vcs*",  # Shows significance stars and standard errors
          model.numbers = FALSE,
          keep.stat = c("n", "rsq", "adj.rsq"),
          out = "figures/poly_s1_models.html") #saved as html

# AIC/BIC function
source('plm_aic_bic_function.R')
plm_aic_bic(rob_sum_base) # regular FE
plm_aic_bic(rob_sum_m2) # quadratic FE
plm_aic_bic(rob_sum_m3) # quadratic FE
plm_aic_bic(rob_sum_m4) # cubic FE
```

DI: We fail to reject the null hypothesis of linearity in the relationship between DI and SPI performance, as the p-value is greater than 0.05 in all models. Although literature suggests a non-linear relationship (SOURCE), the results here do not support it. 

## Stepwise Check: Moderator/Interaction Terms [Stage 1]
*H0:* NUll, the impact of DI on SPI Performance does NOT vary depending on Z [income_level_recoded, regime_type_binary, di_score]
*H1:* The impact of DI on SPI Performance varies depending on Z [income_level_recoded, regime_type_binary,  di_score]

**Recoded Moderator Variables:**
- income_level_recoded: 1 = Low; 2 = Lower-Middle; 3 = Upper-Middle; 4 = High
- regime_type_binary: 0 = Autocracy; 1 = Democracy
- aut_ep: 0 = No Autocratization; 1 = Autocratization Event
- dem_ep: 0 = No Democratization; 1 = Democratization Event
**NOTE**: Based on the results of the RESET test, I will not include polynomial terms in the interaction models, as the linear model is appropriate.
```{r}
### HYPOTHESES ###

# Interaction 1: DI x GNI Class 
# HA: The impact of SPI on SDG performance varies depending on GNI Classification 
di_x_inc_lvl <- plm(formula = spi_comp ~ dplyr::lag(cen_di_score, 1)*dplyr::lag(income_level_recoded, 1) + cen_log_gdppc,
                     model = "within", 
                     index = c("country_code", "year"),
                     data = panel_data,
                     effect = "twoways")
summary(di_x_inc_lvl, vcov = vcovHC(di_x_inc_lvl, cluster = "group", type = "HC1"))

# Interaction 2: DI x Regime Type (Binary)
# HA: The impact of SPI on SDG performance varies depending on Regime Type (Binary)
di_x_reg_binary <- plm(formula = spi_comp ~ dplyr::lag(cen_di_score, 1)*dplyr::lag(autocracy, 1) + cen_log_gdppc, 
                       model = "within", 
                       index = c("country_code", "year"),
                       effect = "twoways",
                       data = panel_data)
summary(di_x_reg_binary, vcov = vcovHC(di_x_reg_binary, cluster = "group", type = "HC1"))

# Interaction 3 DI x Regime Change Episode/Direction
# HA: The impact of SPI on SDG performance varies depending on Regime Change Status/Direction
di_x_reg_eps <- plm(formula = spi_comp ~ dplyr::lag(cen_di_score, 1)*dplyr::lag(aut_ep, 1) + dplyr::lag(cen_di_score, 1)*dplyr::lag(dem_ep, 1) + cen_log_gdppc,
                   model = "within", 
                   index = c("country_code", "year"),
                   effect = "twoways",
                   data = panel_data)
summary(di_x_reg_eps, vcov = vcovHC(di_x_reg_eps, cluster = "group", type = "HC1"))

# Enhanced helper function
tidy_plm_robust <- function(model, model_name) {
  tidy_coef <- tidy(model, 
                    conf.int = TRUE, 
                    vcov. = vcovHC(model, cluster = "group", type = "HC1")) %>%
    mutate(model = model_name)
  
  # Extract model statistics
  model_summary <- summary(model)
  tidy_coef %>%
    mutate(
      adj_r_squared = model_summary$r.squared["adjrsq"]
    )
}

# Apply tidy_plm_robust() function to all models
tidy_di_x_inc_lvl    <- tidy_plm_robust(di_x_inc_lvl, "Moderator: DI x Income Level")
tidy_di_x_reg_binary <- tidy_plm_robust(di_x_reg_binary, "Moderator: DI x Regime Type (Binary)")
tidy_di_x_reg_eps  <- tidy_plm_robust(di_x_reg_eps, "Moderator: DI x Regime Change Episode (aut_ep & dem_ep)")

model_n_obs <- tibble(
  model = c("Moderator: DI x Income Level", "Moderator: DI x Regime Type Binary", 
            "Moderator: DI x Regime Change Episode (aut_ep & dem_ep)"),
  n_obs = c(nobs(di_x_inc_lvl), nobs(di_x_reg_binary),
            nobs(di_x_reg_eps))
)

# Combine all tidy data frames into one
all_models_tidy <- bind_rows(tidy_di_x_inc_lvl, 
                             tidy_di_x_reg_binary, 
                             tidy_di_x_reg_eps) %>%
  left_join(model_n_obs, by = "model") %>% 
  select(model, term, estimate, std.error, p.value, adj_r_squared, n_obs, everything()) %>%
  arrange(model, term)

# Save the tidy data frame to a CSV file
write_csv(all_models_tidy, "output_CSVs/fe_s1_interactions.csv")


# extracting robust standard errors for each interaction model
rob_se_di_x_inc_lvl <- coeftest(di_x_inc_lvl, vcov = vcovHC(di_x_inc_lvl, cluster = "group", type = "HC1"))
rob_se_di_x_reg_binary <- coeftest(di_x_reg_binary, vcov = vcovHC(di_x_reg_binary, cluster = "group", type = "HC1"))
rob_se_di_x_reg_eps <- coeftest(di_x_reg_eps, vcov = vcovHC(di_x_reg_eps, cluster = "group", type = "HC1"))

# Create a list of robust standard errors for stargazer
robust_se_list <- list(
  rob_se_di_x_inc_lvl[, 2], 
  rob_se_di_x_reg_binary[, 2], 
  rob_se_di_x_reg_eps[, 2]
  )
          
stargazer(
  di_x_inc_lvl, di_x_reg_binary, di_x_reg_eps, 
  se = robust_se_list,
  type = "html",
  out = "figures/interact_s1_models.html",
  report = "vcs*",
  title = "Stage 1: Applying Moderators to Two-Way Fixed Effects Models",
  align = TRUE,
  dep.var.labels = "SDG Overall",
  column.labels = c("DI x Income Level", "DI x Regime Type (Binary: Autocracy)", "DI x Regime Change Episode"),
  omit = "factor\\(year\\)",  # Omits year fixed effects from display
  model.names = FALSE
)
```
*Interaction 1: GNI Income Classification:* 
*HA: The impact of SPI on SDG performance varies depending on GNI Classification (income_level_recoded: 1 = Low; 2 = Lower-Middle; 3 = Upper-Middle; 4 = High)*
Not significant 

*Interaction 2: Regime Type (Binary):*
*HA: The impact of SPI on SDG performance varies depending on Regime Type (Binary: 0 = Autocracy; 1 = Democracy)*
Not significant 

*Interaction 3: Regime Change Status/Direction:*
*HA: The impact of SPI on SDG performance varies depending on Regime Change Status/Direction*
Not significant

# MODEL DIAGNOSTICS [Stage 1]

## Multicollinearitty Assessment and Correction [Stage 2] [CHECK BACK W GREG - NEED TO FIGURE OUT WHAT TO DO WITH THIS]
The only model worth looking for multicollinearity is the base model, as the other models have no polynomial or interaction terms.
```{r}
# converting to base R lm() for VIF calculation
ols_spi_di <- lm(spi_comp ~ dplyr::lag(di_score, 1) + log_gdppc + income_level_recoded + factor(year), data = panel_data) # regular linear model
fe_spi_di <- lm(spi_comp ~ dplyr::lag(di_score, 1) + log_gdppc + income_level_recoded + factor(year) + factor(country_code), data = panel_data)

vif(ols_spi_di)
vif(fe_spi_di)

# Check for multicollinearity using VIF
vif_results <- vif(fe_spi_di)
# Convert VIF results to a data frame for better readability
vif_df <- data.frame(
  stage = "Stage 1",
  variable = names(vif_results),
  VIF = vif_results
)
print(vif_df)
# Save VIF results to a CSV file
write.csv(vif_df, file = "output_CSVs/s1_vif_results.csv", row.names = FALSE)
# The VIF values are all below 5, indicating no significant multicollinearity issues.
```

## Check for Autocorrelation 
```{r}
# APPLY Wooldridge Test for AR(1) Errors in FE Panel Models: pwartest()
# https://search.r-project.org/CRAN/refmans/plm/html/pwartest.html  
# This is MUCH BETTER for panel data with small T AND unbalanced panels!!!
pwartest(fe_spi_di_base) # [significant]

```
Significant p-value indicates the presence of autocorrelation in the residuals of the fixed effects model. This suggests that the errors are correlated over time, which violates one of the key assumptions of linear regression models.

This is corrected by using robust standard errors clustered by country, which accounts for the potential autocorrelation in the residuals.

## Check for Heteroskedasticity
```{r}
# Apply Breusch-Pagan test for heteroskedasticity
bptest(fe_spi_di_base, studentize = TRUE) # Heteroskedasticity [significant]
```
The Breusch-Pagan test indicates the presence of heteroskedasticity in the residuals of the fixed effects model. This suggests that the variance of the errors is not constant across observations, which violates another key assumption of linear regression models.
