---
title: "Component 2, Stage 2: SPI -> SDG"
output: html_document
---
# Stage 2: SDG ~ SPI (Mediator) + DI 

# Set up and Wrangling 
```{r}
# set working directory
setwd("~/Documents/GitHub/QMSS_Thesis_Sanchez")

#load libraries/packages
source("packages.R")

# load data
source("Comp2_panel_wrangling.R")
```

# 2.1) POLS SDG ~ SPI [Stage 2]
```{r}
ols_sdg_spi <- plm(formula = sdg_overall ~ dplyr::lag(spi_comp, 1) + dplyr::lag(di_score, 1) + log_gdppc + income_level_recoded + factor(year), 
               model = "pooling", 
               index = c("country_code", "year"),
               data = panel_data)
summary(ols_sdg_spi, vcov = vcovHC(ols_sdg_spi, cluster = "group", type = "HC1"))

# plot relationship: Lagged SPI vs SDG
sdg_spi_s2_scatter <- ggplot(panel_data, aes(x = dplyr::lag(spi_comp, 1), y = sdg_overall)) + 
  geom_point(color = "steelblue4", size = 1, alpha = 0.65) +
  geom_smooth(method = "lm", 
              se = TRUE, color = "darkgreen", size = 1) +
  labs(
    title = "Effect of Statistical Capacity on SDG Performance",
    x = "Lagged SPI Composite (0-100 Scale)",
    y = "SDG Composite (0-100 Scale)"
  ) +
  theme_minimal()

# Save to specific folder
ggsave("figures/stage_1_n_2_scatterplots/sdg_spi_s2_scatterplot.png", sdg_spi_s2_scatter, width = 8, height = 6)
```


# 2.2) First Difference [Stage 2]
```{r}
# Outcome: SDG ~ SPI (Y ~ M)
fd_sdg_spi <- plm(
  formula = sdg_overall ~ dplyr::lag(spi_comp, 1) + dplyr::lag(di_score, 1) + log_gdppc + income_level_recoded,
  index = c("country_code", "year"),
  data = fd_data, 
  model = "fd"
)
# Country-level custard Robust SEs
summary(fd_sdg_spi, vcov = vcovHC(fd_sdg_spi, cluster = "group", type = "HC1"))
```


# 2.3) Fixed Effects [Stage 2]
```{r}
# Outcome: SDG ~ SPI (Y ~ M)
fe_sdg_spi <- plm(
  sdg_overall ~ dplyr::lag(spi_comp, 1) + dplyr::lag(di_score, 1) + log_gdppc + income_level_recoded,
  index = c("country_code", "year"),
  data = panel_data,
  model = "within", #FE
  effect = "twoways"
)
summary(fe_sdg_spi, vcov = vcovHC(fe_sdg_spi, cluster = "group", type = "HC1")) # Robust SEs
```

# make a stargazer table of the above three models 
```{r}
# stargazer table for POLS, FD, and FE models
stargazer(ols_sdg_spi, fd_sdg_spi, fe_sdg_spi, 
          se = list(sqrt(diag(vcovHC(ols_sdg_spi, cluster = "group", type = "HC1"))),
                    sqrt(diag(vcovHC(fd_sdg_spi, cluster = "group", type = "HC1"))),
                    sqrt(diag(vcovHC(fe_sdg_spi, cluster = "group", type = "HC1")))),
          title = "Stage 2 Models: SDG ~ SPI",
          align = TRUE,
          dep.var.labels = "SDG Overall Score",
          column.labels = c("POLS", "FD", "FE"),
          covariate.labels = c("Lagged SPI Composite", "Lagged DI Score", "Log GDP per Capita", 
                               "Income Level (Recoded)", "Year Fixed Effects"),
          notes = "Robust standard errors clustered by country in parentheses",
          type = "html", # Change to "latex" for LaTeX output or "html" for Word
          report = "vcs*", # Shows significance stars and standard errors
          model.numbers = FALSE,
          keep.stat = c("n", "rsq", "adj.rsq"),
          omit = "factor\\(year\\)",  # Omits year fixed effects from display
          out = "figures/s2_base_models.html") #saved as html

```


# Model Selection: Adj. R^2 & AIC/BIC [Stage 2]
```{r}
# AIC/BIC function
source('plm_aic_bic_function.R')

# POLS
plm_aic_bic(ols_sdg_spi) 

# First difference 
plm_aic_bic(fd_sdg_spi)

# Fixed Effects 
plm_aic_bic(fe_sdg_spi) 
```

# MODEL MISSPECIFICATION CHECKS [Stage 2]

## RESET Test for Misspecification [STAGE 2]
```{r}
ols_sdg_spi <- lm(sdg_overall ~ dplyr::lag(spi_comp, 1) + dplyr::lag(di_score, 1) + log_gdppc + income_level_recoded + factor(year), data = panel_data) # regular linear model
fd_sdg_spi <- lm(sdg_overall ~ dplyr::lag(spi_comp, 1) + dplyr::lag(di_diff, 1) + log_gdppc_diff + income_level_recoded, data = fd_data)
fe_sdg_spi <- lm(sdg_overall ~ dplyr::lag(spi_comp, 1) + dplyr::lag(di_score, 1) + log_gdppc + income_level_recoded + factor(year) + factor(country_code), data = panel_data)

# Run RESET tests w/ Robust SEs [Stage 2]
resettest(ols_sdg_spi, power = 2:3, type = "regressor", 
          vcov = function(x) vcovHC(x, cluster = "group", type = "HC1"))
resettest(fd_sdg_spi, power = 2:3, type = "regressor",
          vcov = function(x) vcovHC(x, cluster = "group", type = "HC1"))
resettest(fe_sdg_spi, power = 2:3, type = "regressor",
          vcov = function(x) vcovHC(x, cluster = "group", type = "HC1"))

# Saving results to a dataframe
s2_reset_results <- data.frame(
  model = c("ols_sdg_spi", "fd_sdg_spi", "fe_sdg_spi"),
  F_statistic = c(resettest(ols_sdg_spi, power = 2:3, type = "regressor", 
                             vcov = function(x) vcovHC(x, cluster = "group", type = "HC1"))$statistic,
                  resettest(fd_sdg_spi, power = 2:3, type = "regressor",
                             vcov = function(x) vcovHC(x, cluster = "group", type = "HC1"))$statistic,
                  resettest(fe_sdg_spi, power = 2:3, type = "regressor",
                             vcov = function(x) vcovHC(x, cluster = "group", type = "HC1"))$statistic),
  p_value = c(resettest(ols_sdg_spi, power = 2:3, type = "regressor", 
                        vcov = function(x) vcovHC(x, cluster = "group", type = "HC1"))$p.value,
              resettest(fd_sdg_spi, power = 2:3, type = "regressor",
                        vcov = function(x) vcovHC(x, cluster = "group", type = "HC1"))$p.value,
              resettest(fe_sdg_spi, power = 2:3, type = "regressor",
                        vcov = function(x) vcovHC(x, cluster = "group", type = "HC1"))$p.value)
)
print(s2_reset_results)

# Save the results to a CSV file
write.csv(s2_reset_results, file = "output_CSVs/s2_reset_results.csv", row.names = FALSE)
```
Results from the reset test indicate that the null hypothesis of  is rejected for all models, suggesting potential non-linearity or omitted variable bias in the models.

## Stepwise Check: Applying Polynomial Terms [Stage 2]
For this section, all continuous predictors (DI, SPI, Log(GDP)) in non-linear models are centered to avoid multicollinearity issues. 

- **H0:** The relationship between SPI and SDG performance is linear.
- **H1:** The relationship between SPI and SDG performance is non-linear (quadratic or cubic).
```{r}
# No quadratic Terms Applied - Baseline [SIGNIFICANT DI]
fe_sdg_spi_base <- plm(formula = sdg_overall ~ dplyr::lag(cen_spi_comp, n = 1) + dplyr::lag(cen_di_score, n = 1) + cen_log_gdppc + income_level_recoded, 
  index = c("country_code", "year"),
  data = panel_data,
  model = "within", #FE
  effect = "twoways")
rob_sum_base <- summary(fe_sdg_spi_base, vcov = vcovHC(fe_sdg_spi_base, cluster = "group", type = "HC1"))

# + Quadratic term for SPI [SIGNIFICANT SPI QUAD]
fe_quads_s2_m2 <- plm(formula =
  sdg_overall ~ dplyr::lag(cen_spi_comp, n = 1) + dplyr::lag(cen_spi_comp_quad, n = 1) + dplyr::lag(cen_di_score, n = 1) + cen_log_gdppc + income_level_recoded,
  index = c("country_code", "year"),
  data = panel_data,
  model = "within", #FE
  effect = "twoways")
rob_sum_m2 <- summary(fe_quads_s2_m2, vcov = vcovHC(fe_quads_s2_m2, cluster = "group", type = "HC1"))

# + Quadratic terms for SPI and DI
fe_quads_s2_m3 <- plm(formula = sdg_overall ~ dplyr::lag(cen_spi_comp, n = 1) + dplyr::lag(cen_spi_comp_quad, n = 1) + dplyr::lag(cen_di_score, n = 1) + dplyr::lag(cen_di_score_quad, n = 1) + cen_log_gdppc + income_level_recoded,
               model = "within", 
               index = c("country_code", "year"),
               effect = "twoways",
               data = panel_data)
rob_sum_m3 <- summary(fe_quads_s2_m3, vcov = vcovHC(fe_quads_s2_m3, cluster = "group", type = "HC1"))

# + Cubic terms for SPI & DI 
fe_cubics_s2_m4 <- plm(formula = sdg_overall ~ dplyr::lag(cen_spi_comp, n = 1) + dplyr::lag(cen_spi_comp_quad, n = 1) + dplyr::lag(cen_spi_comp_cubic, n = 1) + dplyr::lag(cen_di_score, n = 1) + dplyr::lag(cen_di_score_quad, n = 1) + dplyr::lag(cen_di_score_cubic, n = 1) + cen_log_gdppc + income_level_recoded,
               model = "within", 
               index = c("country_code", "year"),
               effect = "twoways",
               data = panel_data)
rob_sum_m4 <- summary(fe_cubics_s2_m4, vcov = vcovHC(fe_cubics_s2_m4, cluster = "group", type = "HC1"))

# Extracting robust standard errors for each model
rob_se_base <- coeftest(fe_sdg_spi_base, vcov = vcovHC(fe_sdg_spi_base, cluster = "group", type = "HC1"))
rob_se_m2  <- coeftest(fe_quads_s2_m2, vcov = vcovHC(fe_quads_s2_m2, cluster = "group", type = "HC1"))
rob_se_m3  <- coeftest(fe_quads_s2_m3, vcov = vcovHC(fe_quads_s2_m3, cluster = "group", type = "HC1"))
rob_se_m4 <- coeftest(fe_cubics_s2_m4, vcov = vcovHC(fe_cubics_s2_m4, cluster = "group", type = "HC1"))

# stargazer table 
# [RERUN FULL CHUNK TO RESET NAMES FOR ACURATE RESULTS]
stargazer(fe_sdg_spi_base, fe_quads_s2_m2, fe_quads_s2_m3, fe_cubics_s2_m4, 
          se = list(rob_se_base[, 2], rob_se_m2[, 2], rob_se_m3[, 2], rob_se_m4[, 2]
                    ),
          title = "Applying Polynomial Terms to Stage 2 Models",
          align = TRUE,
          dep.var.labels = "SDG Overall Score",
          column.labels = c("FE Base", "Quad SPI", "Quad SPI+DI", "Cubic SPI+DI"),
          #covariate.labels = c("SPI", "SPI Squared", "DI", "DI Squared",
          #                     "Log GDPpc", "Intercept"),
          notes = "All models apply fixed effects. Robust standard errors clustered by country in parentheses",
          type = "html", # Change to "latex" for LaTeX output or "html" for Word
          report = "vcs*", # Shows significance stars and standard errors
          model.numbers = FALSE,
          keep.stat = c("n", "rsq", "adj.rsq"),
          out = "figures/poly_s2_models.html") #saved as html

# AIC/BIC for both models (sourcing function for plm models)
# [RERUN FULL CHUNK TO RESET NAMES FOR ACURATE RESULTS]
plm_aic_bic(rob_sum_base) # regular FE
plm_aic_bic(rob_sum_m2) # quadratic SPI FE
plm_aic_bic(rob_sum_m3) # quadratic SPI + DI FE
plm_aic_bic(rob_sum_m4) # cubic SPI + DI FE
```
**M1: standard FE**
AIC: 2324.805
BIC: 2447.637 #lowest BIC

**M2: quadratic SPI, FE**
AIC: 2307.741 #lowest AIC
BIC: 2451.045

**M3: quadratic SPI + DI, FE**
AIC: 2313.891
BIC: 2477.667

**M4: cubic SPI + DI, FE**
AIC: 2324.069
BIC: 2528.789

# Interaction Effects [Stage 2]
*H0:* NUll, the impact of SPI on SDG Performance does NOT vary depending on Z [income_level_recoded, regime_type_binary, regime_type_categ, di_score]
*H1:* The impact of SPI on SDG Performance varies depending on Z [income_level_recoded, regime_type_binary, regime_type_categ, di_score]

**Recoded Moderator Variables:**
- income_level_recoded: 1 = Low; 2 = Lower-Middle; 3 = Upper-Middle; 4 = High
- regime_type_binary: 0 = Autocracy; 1 = Democracy
- di_score: 0-1, continuous (Democracy Index)
- aut_ep: 0 = No Autocratization; 1 = Autocratization Event
- dem_ep: 0 = No Democratization; 1 = Democratization Event
```{r}
# Interaction 1: SPI x GNI Class 
# HA: The impact of SPI on SDG performance varies depending on GNI Classification 
spi_x_inc_lvl <- plm(formula = sdg_overall ~ cen_spi_comp*dplyr::lag(income_level_recoded, 1) + cen_di_score +
                       cen_log_gdppc + I(cen_log_gdppc^2),
                     model = "within", 
                     index = c("country_code", "year"),
                     data = panel_data,
                     effect = "twoways")
summary(spi_x_inc_lvl, vcov = vcovHC(spi_x_inc_lvl, cluster = "group", type = "HC1"))

# Interaction 2: SPI x DI
# HA: The impact of SPI on SDG performance varies depending on Democracy Score (cen_di_score: 0-1, continuous)
spi_x_di_score <- plm(formula = sdg_overall ~ cen_spi_comp*cen_di_score + cen_log_gdppc + I(cen_log_gdppc^2) + factor(income_level_recoded),
                      model = "within", 
                      index = c("country_code", "year"),
                      data = panel_data,
                      effect = "twoways")
summary(spi_x_di_score, vcov = vcovHC(spi_x_di_score, cluster = "group", type = "HC1"))

# Interaction 3: SPI x Regime Type (binary)
# HA: The impact of SPI on SDG performance varies depending on Regime Type (Binary: 0 = Autocracy; 1 = Democracy)
spi_x_reg_binary <- plm(formula = sdg_overall ~ cen_spi_comp*factor(regime_type_binary) + cen_di_score +
                          cen_log_gdppc + I(cen_log_gdppc^2) + factor(income_level_recoded),
                        model = "within", 
                        index = c("country_code", "year"),
                        data = panel_data,
                        effect = "twoways")
summary(spi_x_reg_binary, vcov = vcovHC(spi_x_reg_binary, cluster = "group", type = "HC1"))

# Interaction 4: SPI x Regime Change Event (Autocratization or Democratization)
# HA: The impact of SPI on SDG performance varies depending on Regime Change Status/Direction
spi_x_reg_eps <- plm(formula = sdg_overall ~ cen_spi_comp*aut_ep + cen_spi_comp*dem_ep +
                     cen_log_gdppc + I(cen_log_gdppc^2) + factor(income_level_recoded),
                   model = "within", 
                   index = c("country_code", "year"),
                   data = panel_data,
                   effect = "twoways")
summary(spi_x_reg_eps, vcov = vcovHC(spi_x_reg_eps, cluster = "group", type = "HC1"))


```


```{r}
# Final Model [TO BE UPDATED BASED ON STAGE 2 RESULTS]
final_model <- plm(formula = sdg_overall ~ cen_spi_comp*factor(income_level_recoded) + cen_spi_comp*cen_di_score + cen_log_gdppc + I(cen_log_gdppc^2), 
                   model = "within", 
                   index = c("country_code", "year"),
                   data = panel_data,
                   effect = "twoways")
summary(final_model, vcov = vcovHC(final_model, cluster = "group", type = "HC1"))
```




