---
title: 'Component 2: FE, FD, Granger Causality'
author: "Sevastian Sanchez"
date: "`r Sys.Date()`"
output:
  word_document: default
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  tidy = TRUE,
  tidy.opts = list(width.cutoff = 60)  # Adjust width as needed
)
```

# TEMPORAL MEDIATION ANALYSIS: First Difference, Fixed Effects, Lags, Checks
**Acronyms:**
SDG (sdg_overall): UN Sustainable Development Goals Index [0-100, 100 = optimal SDG performance, continuous]
SPI (spi_comp): Statistical Performance Indicators [0-100, 100 = full statistical capacity, continuous]
DI (di_score): Democracy Score [0-10, 10 = fully democratic, continuous]

**Main controls**
log_gdppc: Log of GDP Per Capita of given country (based on reported gdp/population from the world bank)

##setup: packages and data
```{r include=FALSE}
setwd("~/Documents/GitHub/QMSS_Thesis_Sanchez")

#load libraries 
source("packages.R")

#load function 
#source("df_years2.0_Function.R")

#load df_years() function: 2015-present
#fd_fe_data <- df_years2.0(2004, 2023)

#read merged_df csv
fd_fe_data <- read_csv("data/Main CSV Outputs/merged_final_df.csv")

#selecting vars
fd_fe_data <- fd_fe_data %>% 
  dplyr::select(country_code, year, sdg_overall, spi_comp, di_score, log_gdppc, income_level, aut_ep, dem_ep, regch_event, elect_dem, lib_dem, part_dem, delib_dem, egal_dem, academ_free) %>% 
  arrange(country_code, year) %>%  # Critical for correct lagging
  filter(year >= 2016)

# adding has_aut_ep and has_dem_ep for regressing regimes (experienced atleast 1 aut_ep/dem_ep)
fd_fe_data <- fd_fe_data %>%
  # Group by country to check for any event
  group_by(country_code) %>%
  mutate(has_aut_ep = any(aut_ep == 1, na.rm = TRUE)) %>%
  mutate(has_dem_ep = any(dem_ep == 1, na.rm = TRUE)) %>% 
  ungroup() 

# centralized variables for higher-order multicollinearity 
fd_fe_data <- fd_fe_data %>%
  mutate(
    cen_sdg_overall = sdg_overall - mean(sdg_overall, na.rm = TRUE),
    cen_spi_comp = spi_comp - mean(spi_comp, na.rm = TRUE),
    cen_di_score = di_score - mean(di_score, na.rm = TRUE),
    cen_log_gdppc = log_gdppc - mean(log_gdppc, na.rm = TRUE)
  )

# year-year lags for FD Models - DF
fd_data <- fd_fe_data %>%
  select(country_code, year, sdg_overall, di_score, spi_comp, log_gdppc) %>%
  filter(!is.na(di_score) & !is.na(spi_comp)) %>% 
  arrange(country_code, year) %>%
  group_by(country_code) %>%
  mutate(
    # first differences for selected variables
    sdg_diff = sdg_overall - dplyr::lag(sdg_overall, n=1),
    di_diff = di_score - dplyr::lag(di_score, n=1),
    spi_diff = spi_comp - dplyr::lag(spi_comp,n=1),
    log_gdppc_diff = log_gdppc - dplyr::lag(log_gdppc, n=1),
    # lagged first differences 
    di_diff_lag1 = dplyr::lag(di_diff, n=1),
    di_diff_lag2 = dplyr::lag(di_diff, n=2),
    spi_diff_lag1 = dplyr::lag(spi_diff, n=1),
    spi_diff_lag2 = dplyr::lag(spi_diff, n=2),
    log_gdppc_diff_lag1 = dplyr::lag(log_gdppc_diff, n=1),
    log_gdppc_diff_lag1 = dplyr::lag(log_gdppc_diff, n=2)
  ) %>%
  ungroup()
```

# Mediation analysis
To test if SPI mediates the relationship between regime type and SDG outcomes:  
Democratic backsliding → reduces SPI → slows SDG progress  

*H0: SPI DOES NOT mediate (indirectly effect) the relationship between regime type and SDG status* 
*H1: SPI mediates (indirectly effects) the relationship between regime type and SDG status* 

## TEST 1: POLS Mediation (Y ~ M ~ X) [See >> SPI~DI_Validation.Rmd]


# **Fixed Effects**
[description...]
- Smooths our averages, non-immediate changes 

## TEST 2: Fixed Effects Models (Y ~ M ~ X)
```{r}
###### Total Effect: SDG ~ DI (Y ~ X) ######
fe_sdg_di <- plm(
  formula = sdg_overall ~ di_score + log_gdppc + factor(year),
  index = c("country_code", "year"),
  data = fd_fe_data,
  model = "within" #FE
)
summary(fe_sdg_di, vcov = vcovHC(fe_sdg_di, cluster = "group", type = "HC1")) # Robust SEs

###### Mediator: SPI ~ DI (M ~ X) ######
fe_spi_di <- plm(
  formula = spi_comp ~ di_score + log_gdppc + factor(year),
  index = c("country_code", "year"),
  data = fd_fe_data,
  model = "within" #FE
)
summary(fe_spi_di, vcov = vcovHC(fe_spi_di, cluster = "group", type = "HC1")) # Robust SEs

###### Outcome: SDG ~ SPI (Y ~ M) ######
fe_sdg_spi <- plm(
  sdg_overall ~ spi_comp + di_score + log_gdppc + factor(year),
  index = c("country_code", "year"),
  data = fd_fe_data,
  model = "within" #FE
)
summary(fe_sdg_spi, vcov = vcovHC(fe_sdg_spi, cluster = "group", type = "HC1")) # Robust SEs

####### MEDIATION ANALYSIS (Y ~ (M) ~ X) ####### 
# converted to lm form: total effect
#fe_sdg_di_lm <- lm(
#  sdg_overall ~ di_score + log_gdppc + 
#    factor(year) + factor(country_code),
#  data = fd_fe_data
#)
# converted to lm form: mediation
#fe_spi_di_lm <- lm(
#  spi_comp ~ di_score + log_gdppc + 
#    factor(year) + factor(country_code),
#  data = fd_fe_data
#)
# converted to lm form: outcome 
#fe_sdg_spi_lm <- lm(
#  sdg_overall ~ spi_comp + di_score + log_gdppc + 
#    factor(year) + factor(country_code),
#  data = fd_fe_data
#)

#Mediation model 
#set.seed(124) #seed
#med_fe_results <- mediate(
#  fe_spi_di_lm, # mediator model
#  fe_sdg_spi_lm, # outcome model 
#  treat = "di_score",   # Treatment variable (regime change)
#  mediator = "spi_comp", # SPI mediator
#  robustSE = TRUE
#)
#summary(med_fe_results)

# Sensitivity Analysis
#sens_med_fe <- medsens(med_fe_results, rho.by = 0.1)
#plot(sens_med_fe)

#### pbg tests for autocorrelation ####
# total effect 
pbgtest(fe_sdg_di, order = 1) # AR(1)
pbgtest(fe_sdg_di, order = 2) # AR(2)
# mediator 
pbgtest(fe_spi_di, order = 1) # AR(1)
pbgtest(fe_spi_di, order = 2) # AR(2)
# outcome
pbgtest(fe_sdg_spi, order = 1) # AR(1)
pbgtest(fe_sdg_spi, order = 2) # AR(2)

# APPLY Wooldridge Test for AR(1) Errors in FE Panel Models: pwartest()
# https://search.r-project.org/CRAN/refmans/plm/html/pwartest.html 
# This is MUCH BETTER for panel data with small T AND unbalanced panels!!!
pwartest(fe_sdg_di) # AR(1) [significant]
pwartest(fe_spi_di) # AR(1) [significant]
pwartest(fe_sdg_spi) # AR(1) [significant]

```
FE Results: Both the total effect and mediator models are not statistically significant. The outcome model, however, the effect of the independent variable (statistical capacity measured in SPI) on the dependent variable (SDG performance), with all time-invariant factors constant, indicate that a one unit increase in statistical capacity is associated with a 0.0439 point increase in SDG progress. This relationship is statistically significant (p < 0.05).

Additionally, within democracy level (di_score) and SDG progress (sdg_overall) relationship (total effect model), all year coefficients (factored to measure within unit changes), incrementally increase starting from years 2017 to 2023. There is a similar upward trend found between statistical capacity (spi_comp) and democracy level (mediator model) and between SDG performance and statistical capacity (outcome model). All models exhibit extremely high statistical significance across all years (p < 0.001). 

Total Effect (SDG ~ DI): 0.103688 (p = 0.2746) [not significant]
  factor(2017): 0.432077 (p < 0.001) [significant]
  factor(2023): 2.179580 (p < 0.001) [significant]
Mediator (SPI ~ DI): -0.4976 (p = 0.3886) [not significant]
  factor(2017): 0.432077 (p < 0.001) [significant]
  factor(2023): 2.179580 (p < 0.001) [significant]
Outcome (SDG ~ SPI): 0.043900 (p = 0.003263**) [significant]
  factor(2017): 0.432077 (p < 0.001) [significant]
  factor(2023): 2.179580 (p < 0.001) [significant]

Given the total effect two-way fixed effects models are statistically significant, there is no evidence of a possible mediation pathway between democracy levels and SDG progress. Accordingly, there is no reason to assume that statistical capacity mediates this relationship. No mediation analysis is necessary. 

*Panel Breusch-Godfrey Test for Autocorrlation*
This study adopts the Breusch-Godfrey (BG) approach to detect *high order* autocorrelation in panal data, in addition to other forms of testing. The results identify significant autocorrelation found in all three models: total effect, mediator and outcome models. 

## TEST X, First Difference: SPI ~ DI & SDG ~ SPI [FOR ROBUSTNESS]
**First Difference (FD)** removes all time-invariant characteristics of each unit (like geography, culture, or baseline wealth). However, variables that do change over time, such as GDP per capita, should be controlled.
```{r}
#fd dataframe 
fd_data <- fd_fe_data %>%
  select(country_code, year, sdg_overall, di_score, spi_comp, log_gdppc) %>%
  filter(!is.na(di_score) & !is.na(spi_comp)) %>% 
  arrange(country_code, year) %>%
  group_by(country_code) %>%
  mutate(
    # first differences for selected variables
    sdg_diff = sdg_overall - dplyr::lag(sdg_overall, n=1),
    di_diff = di_score - dplyr::lag(di_score, n=1),
    spi_diff = spi_comp - dplyr::lag(spi_comp,n=1),
    log_gdppc_diff = log_gdppc - dplyr::lag(log_gdppc, n=1),
    # lagged first differences 
    di_diff_lag1 = dplyr::lag(di_diff, n=1),
    di_diff_lag2 = dplyr::lag(di_diff, n=2),
    spi_diff_lag1 = dplyr::lag(spi_diff, n=1),
    spi_diff_lag2 = dplyr::lag(spi_diff, n=2),
    log_gdppc_diff_lag1 = dplyr::lag(log_gdppc_diff, n=1),
    log_gdppc_diff_lag1 = dplyr::lag(log_gdppc_diff, n=2)
  ) %>%
  ungroup()

###### Total Effect: SDG ~ SPI (Y ~ X) ######
fd_spi_di <- plm(
  formula = sdg_overall ~ di_score + log_gdppc,
  index = c("country_code", "year"),
  data = fd_data, 
  model = "fd"
) 
# Country-level custard Robust SEs
summary(fd_spi_di, vcov = vcovHC(fd_spi_di, cluster = "group", type = "HC1"))

###### Mediator: SPI ~ DI (M ~ X) ######
fd_spi_di <- plm(
  formula = spi_comp ~ di_score + log_gdppc,
  index = c("country_code", "year"),
  data = fd_data, #sub = (has_aut_ep == TRUE),
  model = "fd"
) 
# Country-level custard Robust SEs
summary(fd_spi_di, vcov = vcovHC(fd_spi_di, cluster = "group", type = "HC1"))

###### Outcome: SDG ~ SPI (Y ~ M) ######
fd_sdg_spi <- plm(
  formula = sdg_overall ~ spi_comp + di_score + log_gdppc,
  index = c("country_code", "year"),
  data = fd_data, 
  model = "fd"
)
# Country-level custard Robust SEs
summary(fd_sdg_spi, vcov = vcovHC(fd_sdg_spi, cluster = "group", type = "HC1"))

###### MEDIATION ANALYSIS (Y ~ (M) ~ X) #######

# mediation 
fd_spi_di_lm <- lm(
  spi_diff ~ di_diff + log_gdppc_diff,
  data = fd_data
)
# outcome 
fd_sdg_spi_lm <- lm(
  sdg_diff ~ spi_diff + di_diff + log_gdppc_diff,
  data = fd_data
)

# mediation model 
set.seed(124) #seed 
med_results <- mediate(
  model.m = fd_spi_di_lm,  # Mediator model (DI → SPI)
  model.y = fd_sdg_spi_lm, # Outcome model (SPI → SDG, controlling for DI)
  treat = "di_diff",       # Treatment variable (first-differenced DI)
  mediator = "spi_diff",   # Mediator (first-differenced SPI)
  boot = TRUE,             # Use bootstrapping for SEs
  sims = 1000              # Number of bootstrap samples
)
summary(med_results)

# Sensitivity Analysis: after running mediation
sens_med_fd <- medsens(med_fd_results, rho.by = 0.1)
plot(sens_med_fd)

##### pbg tests for autocorrelation ####
pbgtest(fd_sdg_di, order = 1) # AR(1)
pbgtest(fd_sdg_di, order = 2) # AR(2)
# mediator 
pbgtest(fd_spi_di, order = 1) # AR(1)
pbgtest(fd_spi_di, order = 2) # AR(2)
# outcome
pbgtest(fd_sdg_spi, order = 1) # AR(1)
pbgtest(fd_sdg_spi, order = 2) # AR(2)
```
total effect: On average, when a country improves it's democracy score (di_score) by 1 point, that same country typically sees a 0.029029 increase in SDG progress, holding log GDP per capita constant. However, this model is not statistically significant, which contrasts the significant results of the total effect fixed effect's model (Y ~ X).

mediator: On average, when a country improves it's democracy score (di_score) by 1 point, that same country typically sees a -0.711012-point fall in statistical capacity (spi_comp), holding log GDP per capita constant. Unlike that of the fixed effects model, first difference estimates between SPI and Democracy Index scores are statistically significant. *[FACT CHECK]*

outcome: Similarly, on average, when a country experiences a 1 point increase in statistical capacity (spi_comp), that same country typically sees a 0.0162939-point increase in SDG score, holding log GDP per capita constant. Like the mediator model, the effect is statistically significant in the outcome model. 

Total Effect (SDG ~ DI): 0.022099 (p = 0.7925) [not significant]
Mediator (SPI ~ DI): -0.711012 (p = 0.033029*) [significant]
Outcome (SDG ~ SPI): 0.0162939 (p = 0.00658**) [significant]

Although there is no statistically significant relationship that suggests that changes in democracy scores have a *direct* affect on SDG performance, there is significant evidence of a potential *indirect* mediation pathway between democracy levels and SDG progress. The presence of a non-significant total effect does not preclude the existence of mediation, as indirect pathways (e.g., DI → SPI → SDG) may be obscured by countervailing direct effects or temporal heterogeneity in the mediation process. First-difference models, which isolate short-term within-country variation, are particularly adept at identifying immediate mediation mechanisms-such as rapid improvements in statistical capacity (SPI) following democratic reforms (DI) that static specifications may overlook due to aggregation over time.

[something here about incrementalism in democratic backsliding]

## TEST 3: Panel Granger Tests: DI & SPI
*[RQ: Do changes in statistical capacity (SPI) reflect a precondition or a symptom of autocratization?]*
Essentially tests whether a predictor variable in a prior year has an affect on a present day dependent variable. In this case, my study utilizes a granger test to both understand directionality of effects, assessing potential reverse causality of x and y variables, and determine the need to incorporate lagged predictors (DI and SPI variables).
```{r}
#balancing panel data 
bal_granger <- fd_fe_data %>%
  select(country_code, year, sdg_overall, spi_comp, di_score) %>% 
  group_by(country_code) %>%
  filter(!any(is.na(across(everything())))) %>%
  ungroup()

# validation 
table(bal_granger$country_code) # Number of rows per country
colSums(is.na(bal_granger)) # Number of NAs per column 

# Identify countries lost due to NA filtering
original_countries <- unique(fd_fe_data$country_code) #all data
filtered_countries <- unique(bal_granger$country_code) #filtered data
lost_countries <- setdiff(original_countries, filtered_countries) # Countries lost
# print lost countries 
print(lost_countries)

# SPI ~ DI (order = 1 and 2) 
spi_di_order1 <- pgrangertest(formula = spi_comp ~ di_score, order = 1, # [SIGNIFICANT]
            index = c("country_code", "year"), 
            test = "Zbar",  # Zbar needed for balanced panels 
            data = bal_granger)
spi_di_order2 <- pgrangertest(formula = spi_comp ~ di_score, order = 2, # [SIGNIFICANT]
            index = c("country_code", "year"), 
            test = "Zbar",
            data = bal_granger)

# DI ~ SPI (order = 1 and 2) 
di_spi_order1 <- pgrangertest(formula = di_score ~ spi_comp, order = 1, # [SIGNIFICANT]
            index = c("country_code", "year"), 
            test = "Zbar",
            data = bal_granger)
di_spi_order2 <- pgrangertest(formula = di_score ~ spi_comp, order = 2, # [SIGNIFICANT, over-fitting warning]
            index = c("country_code", "year"), 
            test = "Zbar",
            data = bal_granger)
spi_di_order1
spi_di_order2
di_spi_order1
di_spi_order2
```

spi_comp ~ di_score[order=1]: Zbar = 19.265, p-value < 2.2e-16 ***
spi_comp ~ di_score[order=2]: Zbar = 128135, p-value < 2.2e-16 ***
di_score ~ spi_comp[order=1]: Zbar = 15.559, p-value < 2.2e-16 ***
di_score ~ spi_comp[order=2]: Zbar = 6237.9, p-value < 2.2e-16 *** [significant-with warning]

As speculated/hypothesized, democracy scores (di_score) do help predict current SPI scores. That said, the same is true in the opposite direction; all panel granger tests results showed high statistical significance in either direction. this suggests a bi-directional relationship between democracy level and statistical capacity. 

## TEST 4: Granger Tests: SPI & SDG
```{r}
# SDG ~ SPI (order = 1 and 2)
sdg_spi_order1 <- pgrangertest(sdg_overall ~ spi_comp, order = 1, # [SIGNIFICANT]
             index = c("country_code", "year"), 
             test = "Zbar",
             data = bal_granger)
sdg_spi_order2 <- pgrangertest(sdg_overall ~ spi_comp, order = 2, # [SIGNIFICANT]
             index = c("country_code", "year"),
             test = "Zbar",
             data = bal_granger)

# SPI ~ SDG (order = 1 and 2)
spi_sdg_order1 <- pgrangertest(spi_comp ~ sdg_overall, order = 1, # [SIGNIFICANT]
             index = c("country_code", "year"), 
             test = "Zbar",
             data = bal_granger)
spi_sdg_order2 <- pgrangertest(spi_comp ~ sdg_overall, order = 2, # [SIGNIFICANT]
             index = c("country_code", "year"), 
             test = "Zbar",
             data = bal_granger)

sdg_spi_order1
sdg_spi_order2
spi_sdg_order1
spi_sdg_order2
```

SDG ~ SPI[order = 1]: Zbar = 24.285, p-value < 2.2e-16 ***
SDG ~ SPI[order = 2]: Zbar = 5824.2, p-value < 2.2e-16 ***
SPI ~ SDG[order = 1]: Zbar = 28.998, p-value < 2.2e-16 ***
SPI ~ SDG[order = 2]: Zbar = 3498.3, p-value < 2.2e-16 ***

Unexpectedly, the results of the panel granger causality tests also showed strong inication of bi-directionality between statistical capacity (SPI) and SDG performance. This highlights a potential feedback loop. 

It is possible that certain SDG goals/indicators could predict statistical capacity, such as SDG 17 (strong institutions) and/or SDG 4 (Quality Education) that could facilitate the growth of national statistical capacity. Deeper analysis of separate SDG score (disaggregated by individual score 1-17) impacts on predictor variables (in a similar methodological scheme) would provide interesting insights into this 'chicken or the egg' problem on a deeper level. 

## P Granger Test Dataframe 
```{r}
# loads function to extract and build df 
source("panel_granger_function.R")

# granger results dataframe 
pgranger_results <- extract_pgrang_results(spi_di_order1, spi_di_order2, di_spi_order1, di_spi_order2, sdg_spi_order1, sdg_spi_order2, spi_sdg_order1, spi_sdg_order2)

#save file
write.csv(pgranger_results, file = "output_CSVs/pgranger_results.csv")
```

# Temporal Dependent Lags
Considering results from the granger test, there is still evidence that democracy score changes preceed changes in SPI. This is in line with studies that sustain it can take some time for implementation of statistical capacity structures to produce results.

Combined with results from the first difference and fixed effects mediation models, it seems across all 167 countries, there is a greater potential for immediate and possibly short-lived effects on statistical capacity than gradual and sustained occurrences. That said, it is important to consider the propotion of countries that not only experienced an episode of regime trantion, but also those that have actually undergone complete regime transformation in either direction (i.e., from a democracy to autocracy or vice-versa). 

## TEST X, Fixed Effects + LAGS: SPI ~ DI & SDG ~ SPI
```{r eval=FALSE, include=FALSE}
###### SPI ~ DI (M ~ X) ######
fe_spi_di_lag <- plm(
  spi_comp ~ di_score + lag(di_score, 1) + lag(di_score, 2) + log_gdppc + factor(year),
  index = c("country_code", "year"),
  data = fd_fe_data,
  model = "within" 
)
summary(fe_spi_di_lag, vcov = vcovHC(fe_spi_di_lag, cluster = "group", type = "HC1")) # Robust SEs

###### SDG ~ SPI (Y ~ M) ######
fe_sdg_spi_lag <- plm(
  sdg_overall ~ spi_comp + lag(spi_comp, 1) + lag(spi_comp, 2) + di_score + lag(di_score, 1) + lag(di_score, 2) + log_gdppc + factor(year),
  index = c("country_code", "year"),
  data = fd_fe_data,
  model = "within" 
)
summary(fe_sdg_spi_lag, vcov = vcovHC(fe_sdg_spi_lag, cluster = "group", type = "HC1")) # Robust SEs

#### pbg tests for autocorrelation ####
# mediator 
pbgtest(fe_spi_di_lag, order = 1) # AR(1)
pbgtest(fe_spi_di_lag, order = 2) # AR(2)
# outcome
pbgtest(fe_sdg_spi_lag, order = 1) # AR(1)
pbgtest(fe_sdg_spi_lag, order = 2) # AR(2)
```


## TEST X, First Difference + LAGS: SPI ~ DI & SDG ~ SPI 
```{r}
###### SPI ~ DI (M ~ X) ######
fd_spi_di_lag <- plm(
  spi_comp ~ di_score + lag(di_score, 1) + lag(di_score, 2) + log_gdppc,
  index = c("country_code", "year"), # two-way
  data = fd_fe_data,
  model = "fd" 
)
summary(fd_spi_di_lag, vcov = vcovHC(fd_spi_di_lag, cluster = "group", type = "HC1")) # Robust SEs

###### SDG ~ SPI (Y ~ M) ######
fd_sdg_spi_lag <- plm(
  sdg_overall ~ spi_comp + lag(spi_comp, 1) + lag(spi_comp, 2) + di_score + lag(di_score, 1) + lag(di_score, 2) + log_gdppc,
  index = c("country_code", "year"),
  data = fd_fe_data,
  model = "fd" 
)
summary(fd_sdg_spi_lag, vcov = vcovHC(fd_sdg_spi_lag, cluster = "group", type = "HC1")) # Robust SEs

#### pbg tests for autocorrelation ####
# mediator 
pbgtest(fd_spi_di_lag, order = 1) # AR(1)
pbgtest(fd_spi_di_lag, order = 2) # AR(2)
# outcome
pbgtest(fd_sdg_spi_lag, order = 1) # AR(1)
pbgtest(fd_sdg_spi_lag, order = 2) # AR(2)
```
*fd_spi_di_lag*
di_score: -0.82917 (p = 0.01966) [significant]
plm::lag(di_score, 1): -0.64157 (p = 0.09281) [significant]

*fd_sdg_spi_lag*
spi_comp: 0.0111140 (p=0.08375) [kinda/not significant]
lag(spi_comp, 1): 0.0075453 (p=0.21093) [not significant]
di_score: 0.0655585 (p=0.43406) [not significant]
lag(di_score, 1): 0.1369567 (p=0.06269) [kinda/not significant]

## ROBUSTNESS CHECKS [SEE DISTRIBUTED_LAGS_TESTING]
```{r}
#AIC/BIC function
source('plm_aic_bic_function.R')

# fixed effects 
plm_aic_bic(fe_spi_di_lag)
plm_aic_bic(fe_sdg_spi_lag) #best AIC (by like 3 points only)
# first difference 
plm_aic_bic(fd_spi_di_lag)
plm_aic_bic(fd_sdg_spi_lag) #best BIC 
```

## First Difference Mediation Effect + Lags: connecting DI -> SDGs
Mediation analysis to formally test and quantify SPI's role in transmitting regime change immediate effects (first difference) to SDGs. This is how to quantify the connection between the Regime Change and the SDGs
```{r eval=FALSE, include=FALSE}
# mediator: SPI ~ DI 
fd_spi_di_lag_lm <- lm(
  spi_diff ~ di_diff + di_diff_lag1 + di_diff_lag2 + log_gdppc_diff,
  data = fd_data
)
summary(fd_spi_di_lag_lm, vcov = vcovHC(fd_spi_di_lag_lm, cluster = "group", type = "HC1")) # Robust SEs

# outcome: SDG ~ SPI 
fd_sdg_spi_lag_lm <- lm(
  sdg_diff ~ spi_diff + spi_diff_lag1 + spi_diff_lag2 + 
                      di_diff + di_diff_lag1 + di_diff_lag2 + 
                      log_gdppc_diff,
  data = fd_data
)
summary(fd_sdg_spi_lag_lm, vcov = vcovHC(fd_sdg_spi_lag_lm, cluster = "group", type = "HC1")) # Robust SEs

# Mediation model 1 [full mediation effect(?)]
set.seed(321)
med_fd_lag_results1 <- mediate(
  model.m = fd_spi_di_lag_lm,  # Mediator model (DI → SPI)
  model.y = fd_sdg_spi_lag_lm, # Outcome model (SPI → SDG, controlling for DI)
  treat = "di_diff",           # Treatment (democracy index) [current]
  mediator = "spi_diff", # Mediator (statistical capacity) [current]
  boot = TRUE,              
  sims = 1000         
)
summary(med_fd_lag_results1) 

# Mediation model 2 [no effect]
set.seed(361)
med_fd_lag_results2 <- mediate(
  model.m = fd_spi_di_lag_lm,  # Mediator model (DI → SPI)
  model.y = fd_sdg_spi_lag_lm, # Outcome model (SPI → SDG, controlling for DI)
  treat = "di_diff",           # Treatment (democracy index) [current]
  mediator = "spi_diff_lag1", # Mediator (statistical capacity) [1 period]
  boot = TRUE,              
  sims = 1000         
)
summary(med_fd_lag_results2) 

# Mediation model 3 [full mediation effect (?)]
set.seed(456)
med_fd_lag_results3 <- mediate(
  model.m = fd_spi_di_lag_lm, # mediator model
  model.y = fd_sdg_spi_lag_lm, # outcome model
  treat = "di_diff", # Treatment (regime change) [current]
  mediator = "spi_diff_lag2", # SPI mediator [2 periods]
  sims = 1000,        
  boot = TRUE
)
summary(med_fd_lag_results3)

# Mediation model 4 [direct effect & slight indirect effect]
set.seed(654)
med_fd_lag_results4 <- mediate(
  model.m = fd_spi_di_lag_lm, # mediator model
  model.y = fd_sdg_spi_lag_lm, # outcome model
  treat = "di_diff_lag1", # Treatment (regime change) [1 period]
  mediator = "spi_diff", # SPI mediator [current]
  sims = 1000,        
  boot = TRUE
)
summary(med_fd_lag_results4)

# Mediation model 5 [direct effect]
set.seed(654)
med_fd_lag_results5 <- mediate(
  model.m = fd_spi_di_lag_lm, # mediator model
  model.y = fd_sdg_spi_lag_lm, # outcome model
  treat = "di_diff_lag1", # Treatment (regime change) [1 period]
  mediator = "spi_diff_lag1", # SPI mediator [1 period]
  sims = 1000,        
  boot = TRUE
)
summary(med_fd_lag_results5)

# Mediation model 6 [direct effect, slight indirect effect]
set.seed(654)
med_fd_lag_results6 <- mediate(
  model.m = fd_spi_di_lag_lm, # mediator model
  model.y = fd_sdg_spi_lag_lm, # outcome model
  treat = "di_diff_lag1", # Treatment (regime change) [1 period]
  mediator = "spi_diff_lag2", # SPI mediator [2 period]
  sims = 1000,        
  boot = TRUE
)
summary(med_fd_lag_results6)

# Mediation model 7 [full mediation effect(?)]
set.seed(658)
med_fd_lag_results7 <- mediate(
  model.m = fd_spi_di_lag_lm, # mediator model
  model.y = fd_sdg_spi_lag_lm, # outcome model
  treat = "di_diff_lag2", # Treatment (regime change) [2 period]
  mediator = "spi_diff", # SPI mediator [current]
  sims = 1000,        
  boot = TRUE
)
summary(med_fd_lag_results7)

# Mediation model 8 [nothing]
set.seed(659)
med_fd_lag_results8 <- mediate(
  model.m = fd_spi_di_lag_lm, # mediator model
  model.y = fd_sdg_spi_lag_lm, # outcome model
  treat = "di_diff_lag2", # Treatment (regime change) [2 period]
  mediator = "spi_diff_lag1", # SPI mediator [1 period]
  sims = 1000,        
  boot = TRUE
)
summary(med_fd_lag_results8)

# Mediation model 9 [full mediation(?)]
set.seed(664)
med_fd_lag_results9 <- mediate(
  model.m = fd_spi_di_lag_lm, # mediator model
  model.y = fd_sdg_spi_lag_lm, # outcome model
  treat = "di_diff_lag2", # Treatment (regime change) [2 period]
  mediator = "spi_diff_lag2", # SPI mediator [2 periods]
  sims = 1000,        
  boot = TRUE
)
summary(med_fd_lag_results9)

##### Sensitivity analysis #####
# mediation model 1
sens_med1 <- medsens(med_fd_lag_results1, rho.by = 0.1) # [non-significant]
summary(sens_med1)
# mediation model 2
sens_med2 <- medsens(med_fd_lag_results2, rho.by = 0.1) 
summary(sens_med2)
# mediation model 3
sens_med3 <- medsens(med_fd_lag_results3, rho.by = 0.1) 
summary(sens_med3)
# mediation model 4
sens_med4 <- medsens(med_fd_lag_results4, rho.by = 0.1) 
summary(sens_med4)
# mediation model 5
sens_med5 <- medsens(med_fd_lag_results5, rho.by = 0.1) 
summary(sens_med5)
# mediation model 6
sens_med6 <- medsens(med_fd_lag_results6, rho.by = 0.1) 
summary(sens_med6)
# mediation model 7
sens_med7 <- medsens(med_fd_lag_results7, rho.by = 0.1) 
summary(sens_med7)
# mediation model 8
sens_med8 <- medsens(med_fd_lag_results8, rho.by = 0.1)
summary(sens_med8)
# mediation model 9
sens_med9 <- medsens(med_fd_lag_results9, rho.by = 0.1)
summary(sens_med9)
```