---
title: "Component 2: FE, FD, Granger Causality"
author: "Sevastian Sanchez"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  tidy = TRUE,
  tidy.opts = list(width.cutoff = 60)  # Adjust width as needed
)
```

# TEMPORAL MEDIATION ANALYSIS: First Difference, Fixed Effects, Lags, Checks
**Acronyms:**
SDG (sdg_overall): UN Sustainable Development Goals Index [0-100, 100 = optimal SDG performance, continuous]
SPI (spi_comp): Statistical Performance Indicators [0-100, 100 = full statistical capacity, continuous]
DI (di_score): Democracy Score [0-10, 10 = fully democratic, continuous]

**Main controls**
log_gdppc: Log of GDP Per Capita of given country (based on reported gdp/population from the world bank)

##setup: packages and data
```{r include=FALSE}
setwd("~/Documents/GitHub/QMSS_Thesis_Sanchez")

#load libraries 
source("packages.R")

#load function 
#source("df_years2.0_Function.R")

#load df_years() function: 2015-present
#fd_fe_data <- df_years2.0(2004, 2023)

#read merged_df csv
fd_fe_data <- read_csv("data/Main CSV Outputs/merged_final_df.csv")

#selecting vars
fd_fe_data <- fd_fe_data %>% 
  dplyr::select(country_code, year, sdg_overall, spi_comp, di_score, aut_ep, dem_ep, elect_dem, lib_dem, part_dem, delib_dem, egal_dem, academ_free, income_level, income_level_lab, log_gdppc, population) %>% 
  arrange(country_code, year) %>%  # Critical for correct lagging
  filter(year >= 2016)

# adding has_aut_ep and has_dem_ep for regressing regimes (experienced atleast 1 aut_ep/dem_ep)
fd_fe_data <- fd_fe_data %>%
  # Group by country to check for any event
  group_by(country_code) %>%
  mutate(has_aut_ep = any(aut_ep == 1, na.rm = TRUE)) %>%
  mutate(has_dem_ep = any(dem_ep == 1, na.rm = TRUE)) %>% 
  ungroup() 
```

# TEST 1: OLS Mediation analysis [REDO RESULTS FOR ROBUST SEs]
To test if SPI mediates the relationship between regime type and SDG outcomes:  
Democratic backsliding → reduces SPI → slows SDG progress  

*H0: SPI DOES NOT mediate (indirectly effect) the relationship between regime type and SDG status* 
*H1: SPI mediates (indirectly effects) the relationship between regime type and SDG status* 

# TEST X: The effect of democracy score on SDG (Y ~ X)
```{r}
#Total Effect: Check if regime type directly affects SDG scores (without SPI) + Controls & RobustSEs
sdg_di_ols <- plm(formula = sdg_overall ~ di_score + log_gdppc + factor(year), 
                        index = c("country_code", "year"),
                        model = "pooling", 
                        data = fd_fe_data)
summary(sdg_di_ols, vcov = vcovHC(sdg_di_ols, cluster = "group", type = "HC1"))

#plot relationship
plot_sdg_di <- ggplot(fd_fe_data, aes(x = di_score, y = sdg_overall)) + 
  geom_point(color = "#EE6A50", size = 1, alpha = 0.65) +
  geom_smooth(method = "lm", se = TRUE, color = "darkred", size = 1) +
  labs(
    title = "Effect of Democracy Score on SDG Status",
    x = "Democracy Score",
    y = "SDG Progress"
  ) +
  theme_minimal()

plot_sdg_di
```

## TEST X: OLS The effect of democracy score on SDG (Y ~ M ~ X)
- ACME (Average Causal Mediation Effect): SPI’s indirect effect.
- ADE (Average Direct Effect): Regime type’s direct effect, excluding SPI
```{r}
#OLS Mediator model: Check if regime type affects SPI (M)
med_spi_ols <- lm(spi_comp ~ di_score + log_gdppc + factor(year), data = fd_fe_data)
summary(med_spi_ols, vcov = vcovHC(med_spi_ols, cluster = "group", type = "HC1"))

#OLS Outcome model: Check if SPI affects SDG scores (Y ~ M) while controlling for regime type (X)
output_sdg_ols <- lm(sdg_overall ~ spi_comp + di_score + log_gdppc + factor(year), data = fd_fe_data)
summary(output_sdg_ols, vcov = vcovHC(output_sdg_ols, cluster = "group", type = "HC1"))

#seed to reproduce estimates 
set.seed(125)
#OLS Mediation test: Quantify how much of regime type’s effect on SDGs operates through SPI
med_model <- mediate(med_spi_ols, output_sdg_ols, 
                     treat = "di_score", 
                     mediator = "spi_comp", 
                     sims = 1000, 
                     robustSE = TRUE)
summary(med_model)

#Sensitivity Analysis: Tests how robust mediation findings are to potential unmeasured confounding variables
sens_med <- medsens(med_model, rho.by = 0.1)
summary(sens_med)
plot(sens_med)
```
*Results:*
ACEM: SPI’s indirect effect = 1.0285  units (p < 2e-16***)
ADE: Regime type’s direct effect, excluding SPI = 0.2027  (p = 0.064) [just over 0.05]
Total Effect: = 1.2312 units (p < 2e-16***)
Proportion Mediated: = 0.8347 or 83.47% of total units

`Proportion Mediated` = the percent difference between the total effect (SDG ~ DI) and indirect ACEM effect (+ SPI)

Interpretation: A 1-unit DI increase boosts SDG scores by 1.2312 total units, with 1.0285 units (83.47% of units) transmitted through SPI. The remaining 0.2027 units represents democracy score's effect (e.g., governance reforms unrelated to statistics) on SDG progress. 

Because the indirect effect of SPI on sdg_overall (ACME) is highly significant (p < 0.001), there is evidence to suggest that SPI mediates the regime-SDG relationship, based on the model. 

Because the ADE (the direct effect between di_score on sdg_overall) is not significant (or just barely), SPI has a complete/almost-complete effect on SDG status when SPI is accounted for, based on the model. 

*Robustness check: sensitivity analysis*
This tests how much unmeasured confounding would be needed to nullify the mediation effect. Based on the results, a rho = 0.5 indicates the mediation is moderately robust. there would need to be relatively strong unmeasured cofounders to nullify the results.

Any unmeasured confounder would need to explain 25% of the residual variance in both models to eliminate the mediation effect (R²_MR²_Y = 0.25). 
-----------------------------------------------------------------------------------------------------

# **First Difference & Fixed Effects**

## TEST X, First Difference: SPI ~ DI & SDG ~ SPI 
**First Difference (FD)** removes all time-invariant characteristics of each unit (like geography, culture, or baseline wealth). However, variables that do change over time, such as GDP per capita, should be controlled.
```{r}
###### SPI ~ DI (M ~ X) ######
fd_spi_di <- plm(
  formula = spi_comp ~ di_score + log_gdppc,
  index = c("country_code", "year"),
  data = fd_fe_data, #sub = (has_aut_ep == TRUE),
  model = "fd" #FD
) 
# Country-level custard Robust SEs
summary(fd_spi_di, vcov = vcovHC(fd_spi_di, cluster = "group", type = "HC1"))

###### SDG ~ SPI (Y ~ M) ######
fd_sdg_spi <- plm(
  formula = sdg_overall ~ spi_comp + di_score + log_gdppc,
  index = c("country_code", "year"),
  data = fd_fe_data, 
  model = "fd" #FD
)
# Country-level custard Robust SEs
summary(fd_sdg_spi, vcov = vcovHC(fd_sdg_spi, cluster = "group", type = "HC1"))

###### MEDIATION ANALYSIS (Y ~ (M) ~ X) ######
#seed 
set.seed(124)

med_fd_results <- mediate(
  fd_spi_di, # mediator model
  fd_sdg_spi, # outcome model 
  treat = "di_score",   # Treatment variable (regime change)
  mediator = "spi_comp", # SPI mediator
  robustSE = TRUE
)
summary(med_fd_results)

# Sensitivity Analysis: after running mediation
sens_med_fd <- medsens(med_fd_results, rho.by = 0.1)
plot(sens_med_fd)

#### pbg tests for autocorrelation ####
pbgtest(fd_spi_di, order = 1) # AR(1)
pbgtest(fd_sdg_spi, order = 2) # AR(2)
```

Results: 

FD Models: 
SPI ~ DI: -0.9927 (p = 0.0083**)
  Adj.R-squared: 0.0076734
SDG ~ SPI: 0.0158 (p = 0.0032**)
  Adj.R-squared:
On average, when a country improves its DI score by 1 point, that same country typically sees a -0.9927-point fall in SPI. 
On average, when a country improves its SPI score by 1 point, that same country typically sees a 0.0158-point fall in SDG score.  


## TEST X, Fixed Effects Models: SPI ~ DI & SDG ~ SPI  
**Fixed Effects (FE)** [description...]
```{r}
###### SPI ~ DI (M ~ X) ######
fe_spi_di <- plm(
  formula = spi_comp ~ di_score + log_gdppc + factor(year),
  index = c("country_code", "year"),
  data = fd_fe_data,
  model = "within" #FE
)
summary(fe_spi_di, vcov = vcovHC(fe_spi_di, cluster = "group", type = "HC1")) # Robust SEs

###### SDG ~ SPI (Y ~ M) ######
fe_sdg_spi <- plm(
  sdg_overall ~ spi_comp + di_score + log_gdppc + factor(year),
  index = c("country_code", "year"),
  data = fd_fe_data,
  model = "within" #FE
)
summary(fe_sdg_spi)
summary(fe_sdg_spi, vcov = vcovHC(fe_sdg_spi, cluster = "group", type = "HC1")) # Robust SEs

####### MEDIATION ANALYSIS (Y ~ (M) ~ X) ####### 
#seed
set.seed(124)

med_fe_results <- mediate(
  fe_spi_di, # mediator model
  fe_sdg_spi, # outcome model 
  treat = "di_score",   # Treatment variable (regime change)
  mediator = "spi_comp", # SPI mediator
  robustSE = TRUE
)
summary(med_fe_results)

# Sensitivity Analysis: after running mediation
sens_med_fe <- medsens(med_fe_results, rho.by = 0.1)
plot(sens_med_fe)

#### pbg tests for autocorrelation ####
pbgtest(fe_spi_di, order = 1) # AR(1)
pbgtest(fe_spi_di, order = 2) # AR(2)
```
Results:

SPI ~ DI: -0.4976 (p = 0.3886) [not significant]
  factor(year): 2.0536 (p < 2e-16***)
SDG ~ SPI: 0.0324 (p = 0.1247) [not significant]
  factor(year): 2.5292e-01 (p < 2.2e-16***)

SPI~DI is not statistically significant. This contrasts from previous first difference estimates that found both SPI~DI and SDG~SPI models statistically significant (p < 0.05). Interestingly, the factored-year variable (factor(year)) in both FE models are extremely statistically significant with increasing coefficient estimates over the years. This suggests underlying trends that statistical capacity is, on average, increasing based on the model. 

## TEST X: Granger Tests: DI & SPI
Essentially tests whether a predictor variable in a prior year has an affect on a present day dependent variable. In this case, my study utilizes a granger test to both understand directionality of effects, assessing potential reverse causality of x and y variables, and determine the need to incorporate lagged predictors (DI and SPI variables).
```{r}
bal_granger_data <- fd_fe_data %>% select(sdg_overall, spi_comp, di_score, year, country_code)
colSums(is.na(bal_granger_data))

# SPI ~ DI [SIGNIFICANT]
pgrangertest(formula = spi_comp ~ di_score, order = 1, 
            index = c("country_code", "year"), 
            test = "Zbar",  # Changed to Zbar? [according to perplexity]
            data = bal_granger_data)
pgrangertest(spi_comp ~ di_score, order = 2, 
            index = c("country_code", "year"), 
            data = bal_granger_data)

# DI ~ SPI [NOT]
pgrangertest(di_score ~ spi_comp, order = 1, 
            index = c("country_code", "year"), 
            data = fd_fe_data)
pgrangertest(di_score ~ spi_comp, order = 2, 
            index = c("country_code", "year"), 
            data = fd_fe_data)
```

spi_comp ~ di_score[order=1]: F = 16.97, p = 4.048e-05 (***)
- As speculated/hypothesized, democracy scores (di_score) DO help predict current SPI scores. 
- Although we recieve statistically significant results upon increasing the lag order, the F statistic drops sharply from 16.97 to 8.8021 and then continues falling gradually, as p-value increases ever so slightly (although still extremely below alpha 0.05). Accordingly, 1st orderd lagged di_score predictors are optimal and sufficient for the following Fixed Effects models. 

di_score ~ spi_comp[order=1]: F = 0.4086, p = 0.5228
- As expected, past SPI scores do NOT help predict current democracy scores (di_score)
- even incorporating more lags, the relationships remain not statistically significant 

**Bottom Line**
Changes in regime characteristics precede changes in statistical capacity, NOT vice versa???

## TEST X: Granger Tests: SPI & SDG
```{r}
# SDG ~ SPI [SIGNIFICANT?]
pgrangertest(sdg_overall ~ spi_comp, order = 1, 
             index = c("country_code", "year"), 
             data = fd_fe_data)
pgrangertest(sdg_overall ~ spi_comp, order = 2, 
             index = c("country_code", "year"),
             data = fd_fe_data)

# SPI ~ SDG [NOT?]
pgrangertest(spi_comp ~ sdg_overall, order = 1, 
             index = c("country_code", "year"), 
             data = fd_fe_data)
pgrangertest(spi_comp ~ sdg_overall, order = 2, 
             index = c("country_code", "year"), 
             data = fd_fe_data)
```
Results:

SDG~SPI [order = 1]: F = 2.5447, p = 0.1109

SPI~SDG [order = 1]: F = 8.7453, p = 0.00316 (**)

Unexpectedly, the results of the granger causality tests showed that it is SDG status that preceeds changes in statistical capacity, similar to DI score. It is possible that certain SDG goals/indicators could predict statistical capacity, such as SDG 17 (strong institutions) and/or SDG 4 (Quality Education) that could facilitate the growth of national statistical capacity. Deeper analysis of separate SDG score (disaggregated by individual score 1-17) impacts on predictor variables (in a similar methodological scheme) would provide interesting insights into this 'chicken or the egg' problem on a deeper level. This study reconsiders strategy due to suprising, but valid, evidence of reverse causality.  

# Temporal Dependent Lags
Considering results from the granger test, there is evidence that democracy score changes preceed changes in SPI (first). This is in line with studies that sustain it can take some time for implementation of statistical capacity structures to produce results.

Similarly, I suspect that regime transition has a delayed effect on statistical capacity, although I suspect it varies in time (do we notice a drop in SPI as a precondition or a symptom of autocratization)?

## Significant Autocorrelation: create lagged variables and first differences [MAY NOT NEED]

```{r eval=FALSE, include=FALSE}
# Create lagged variables and first differences
panel_data <- panel_data %>%
  arrange(country_code, year) %>% 
  group_by(country_code) %>%
  mutate(
    # 1 year lags
    sdg_lag_1 = dplyr::lag(sdg_overall, n = 1),
    spi_lag_1 = dplyr::lag(spi_comp, n = 1),
    di_lag_1 = dplyr::lag(di_score, n = 1),
    # 2-year lags 
    sdg_lag_2 = dplyr::lag(sdg_overall, n = 2),
    spi_lag_2 = dplyr::lag(spi_comp, n = 2),
    di_lag_2 = dplyr::lag(di_score, n = 2),
    
    # Calculate first differences
    fd_sdg_1 = sdg_overall - sdg_lag_1,
    fd_spi_1 = spi_comp - spi_lag_1,
    fd_di_1 = di_score - di_lag_1,
    fd_sdg_2 = sdg_overall - sdg_lag_2,
    fd_spi_2 = spi_comp - spi_lag_2,
    fd_di_2 = di_score - di_lag_2
  ) %>%
  ungroup()

# To verify the lag creation works correctly
#head(fd_data_testing %>% dplyr::select(country_code, year, sdg_overall, sdg_lag, fd_sdg, spi_comp, spi_lag, fd_spi) %>% arrange(country_code, year), 10)
```


## TEST X, First Difference + LAGS: SPI ~ DI & SDG ~ SPI 
```{r}
###### SPI ~ DI (M ~ X) ######
fd_spi_di_lag <- plm(
  spi_comp ~ di_score + lag(di_score, 1) + log_gdppc,
  index = c("country_code", "year"), # two-way
  data = fd_fe_data,
  model = "fd" 
)
summary(fd_spi_di_lag, vcov = vcovHC(fd_spi_di_lag, cluster = "group", type = "HC1")) # Robust SEs

###### SDG ~ SPI (Y ~ M) ######
fd_sdg_spi_lag <- plm(
  sdg_overall ~ spi_comp + lag(spi_comp, 1) + di_score + lag(di_score, 1) + log_gdppc,
  index = c("country_code", "year"),
  data = fd_fe_data,
  model = "fd" 
)
summary(fd_sdg_spi_lag, vcov = vcovHC(fd_sdg_spi_lag, cluster = "group", type = "HC1")) # Robust SEs
```
*fd_spi_di_lag*
di_score: -0.82917 (p = 0.01966) [significant]
plm::lag(di_score, 1): -0.64157 (p = 0.09281) [kinda/not significant]

*fd_sdg_spi_lag*
spi_comp: 0.0111140 (p=0.08375) [kinda/not significant]
lag(spi_comp, 1): 0.0075453 (p=0.21093) [not significant]
di_score: 0.0655585 (p=0.43406) [not significant]
lag(di_score, 1): 0.1369567 (p=0.06269) [kinda/not significant]

## TEST X, Fixed Effects + LAGS: SPI ~ DI & SDG ~ SPI
```{r eval=FALSE, include=FALSE}
###### SPI ~ DI (M ~ X) ######
fe_spi_di_lag <- plm(
  spi_comp ~ di_score + plm::lag(di_score, 1) + log_gdppc + factor(year),
  index = c("country_code", "year"),
  data = fd_fe_data,
  model = "within" 
)
summary(fe_spi_di_lag, vcov = vcovHC(fe_spi_di_lag, cluster = "group", type = "HC1")) # Robust SEs

###### SDG ~ SPI (Y ~ M) ######
fe_sdg_spi_lag <- plm(
  sdg_overall ~ spi_comp + plm::lag(spi_comp) + di_score + plm::lag(di_score, 1) + log_gdppc + factor(year),
  index = c("country_code", "year"),
  data = fd_fe_data,
  model = "within" 
)
summary(fe_sdg_spi_lag, vcov = vcovHC(fe_sdg_spi_lag, cluster = "group", type = "HC1")) # Robust SEs
```
*fe_spi_di_lag*
di_score: -0.31101 (p = 0.5556) [not significant]
plm::lag(di_score, 1): 0.14895 (p=0.7736) [not significant]

*fe_sdg_spi_lag*
spi_comp: 0.0242614 (p = 0.0324657) [significant]
plm::lag(spi_comp): 0.0239796 (p = 0.0324657) [significant]
di_score:  0.1844171 (p = 0.0110328) [significant]
plm::lag(di_score, 1): -0.0345650 (p = 0.7485891) [not significant]

## ROBUSTNESS CHECKS [SEE DISTRIBUTED_LAGS_TESTING]

### Breusch-Godfrey Autocorrelation Tests: AR(1) + AR(2)
```{r}
#mediator models 
pbgtest(fe_spi_di, order = 1) # AR(1)
pbgtest(fe_spi_di, order = 2) # AR(2)
```
This analysis does not test for autocorrelation by using the Durban Watson or ACF approach because they're designed for single time series models, not panel models. Instead, this study adopts the Breusch-Godfrey (BG) approach to detect *high order* autocorrelation in addition to other forms. 

## Mediation Effect: connecting Regime -> SDGs 
Mediation analysis to formally test and quantify SPI's role in transmitting regime change effects to SDGs. This is how to quantify the connection between the Regime Change and the SDGs

## Mediation Effect + Lags: connecting DI -> SDGs
```{r eval=FALSE, include=FALSE}
## CONFIRM THE MED MODELS BELOW ARE OKAY W PERPLEXITY 

# Mediation analysis - FD models 
med_fd_lag_results <- mediate(
  fd_spi_di_lag, # mediator model
  fd_sdg_spi_lag, # outcome model 
  treat = "di_score",   # Treatment variable (regime change)
  mediator = "spi_comp", # SPI mediator
  sims = 1000,        # Number of bootstrap simulations
  robustSE = TRUE     # Cluster SEs by country
)

summary(med_fd_lag_results)

# After running mediation
sens_med <- medsens(med_fd_lag_results, rho.by = 0.1)
plot(sens_med)

# Mediation analysis - FE models 
med_fe_lag_results <- mediate(
  fe_spi_di_lag, # mediator model
  fe_sdg_spi_lag, # outcome model 
  treat = "di_score",   # Treatment variable (regime change)
  mediator = "spi_comp", # SPI mediator
  sims = 1000,        # Number of bootstrap simulations
  robustSE = TRUE     # Cluster SEs by country
)

summary(med_fe_lag_results)

# After running mediation
sens_med <- medsens(med_fe_lag_results, rho.by = 0.1)
plot(sens_med)
```
