---
title: "Component 2, Stage 1: DI -> SPI"
output: html_document
---

# Set up and Wrangling 
```{r}
# set working directory
setwd("~/Documents/GitHub/QMSS_Thesis_Sanchez")

#load libraries/packages
source("packages.R")

# load data
source("Comp2_panel_wrangling.R")
```

# Stage 1 Models:
ols_spi_di
fd_spi_di
fe_spi_di
ols_spi_di_L1
fd_spi_di_L1
fe_spi_di_L1

# 1.1) POLS [Stage 1]: SPI ~ DI
The effect of democracy on SPI Performance
```{r}
# Contemporaneous Effect: SPI ~ DI
ols_spi_di <- plm(
  formula = spi_comp ~ di_score + log_gdppc + income_level_recoded + factor(year), 
  index = c("country_code", "year"),
  model = "pooling", 
  data = panel_data)
summary(ols_spi_di, vcov = vcovHC(ols_spi_di, cluster = "group", type = "HC1"))

# Adding Lag1: SPI ~ DI
ols_spi_di_L1 <- plm(
  formula = spi_comp ~ di_score + dplyr::lag(di_score, n=1) + log_gdppc + income_level_recoded + factor(year), 
  index = c("country_code", "year"),
  model = "pooling", 
  data = panel_data)
summary(ols_spi_di_L1, vcov = vcovHC(ols_spi_di_L1, cluster = "group", type = "HC1"))

# Adding Lag2: SPI ~ DI
ols_spi_di_L2 <- plm(
  formula = spi_comp ~ di_score + dplyr::lag(di_score, n=1) + dplyr::lag(di_score, n=2) + log_gdppc + income_level_recoded + factor(year), 
  index = c("country_code", "year"),
  model = "pooling", 
  data = panel_data)
summary(ols_spi_di_L2, vcov = vcovHC(ols_spi_di_L2, cluster = "group", type = "HC1"))
```

**results**

## POLS Summary Table
```{r}
# --- POLS MODELS TABLE ---

# Extracting robust standard errors for the two POLS models
rob_se_pols <- list(
  coeftest(ols_spi_di, vcov = vcovHC(ols_spi_di, cluster = "group", type = "HC1"))[, 2],
  coeftest(ols_spi_di_L1, vcov = vcovHC(ols_spi_di_L1, cluster = "group", type = "HC1"))[, 2],
  coeftest(ols_spi_di_L2, vcov = vcovHC(ols_spi_di_L2, cluster = "group", type = "HC1"))[, 2]
)

# Generate the stargazer table for POLS models
stargazer(
  ols_spi_di, ols_spi_di_L1, ols_spi_di_L2,
  se = rob_se_pols,
  type = "html",
  out = "figures/s1_pols_lags.html",
  title = "Stage I Models: Pooled OLS (Current and Lagged Effects)",
  align = TRUE,
  dep.var.labels = "SPI Overall Score",
  column.labels = c("Contemporaneous", "Lag 1", "Lag 2"),
  covariate.labels = c("DI Composite", "Lagged DI (t-1)", "Lagged DI (t-2)",  "Log(GDPper-capita)", "Lower-Middle Income (LMIC)", "Upper-Middle Income (UMIC)", "High Income (HIC)", "Intercept"),
  report = "vcs*",
  notes = "Robust standard errors clustered by country in parentheses",
  omit = "factor\\(year\\)",
  model.names = FALSE
)
```


# 1.2) First Difference [Stage 1]: SPI ~ DI
```{r}
# Contemporaneous Effect: SPI ~ DI
fd_spi_di <- plm(
  formula = spi_comp ~ di_score + log_gdppc + income_level_recoded,
  index = c("country_code", "year"),
  data = fd_data, #sub = (has_aut_ep == TRUE),
  model = "fd"
)
summary(fd_spi_di, vcov = vcovHC(fd_spi_di, cluster = "group", type = "HC1"))

# Adding Lag1: SPI ~ DI
fd_spi_di_L1 <- plm(
  formula = spi_comp ~ di_score + dplyr::lag(di_score, n=1) + log_gdppc + income_level_recoded,
  index = c("country_code", "year"),
  data = fd_data, #sub = (has_aut_ep == TRUE),
  model = "fd"
)
summary(fd_spi_di_L1, vcov = vcovHC(fd_spi_di_L1, cluster = "group", type = "HC1"))

# Adding Lag2: SPI ~ DI
fd_spi_di_L2 <- plm(
  formula = spi_comp ~ di_score + dplyr::lag(di_score, n=1) + dplyr::lag(di_score, n=2) + log_gdppc + income_level_recoded,
  index = c("country_code", "year"),
  data = fd_data, #sub = (has_aut_ep == TRUE),
  model = "fd"
)
summary(fd_spi_di_L2, vcov = vcovHC(fd_spi_di_L2, cluster = "group", type = "HC1"))
```
**results** 

## First Difference Summary Table
```{r}
# --- FD MODELS TABLE ---
# extracting robust standard errors for the three FD models
rob_se_fd <- list(
  coeftest(fd_spi_di, vcov = vcovHC(fd_spi_di, cluster = "group", type = "HC1"))[, 2],
  coeftest(fd_spi_di_L1, vcov = vcovHC(fd_spi_di_L1, cluster = "group", type = "HC1"))[, 2],
  coeftest(fd_spi_di_L2, vcov = vcovHC(fd_spi_di_L2, cluster = "group", type = "HC1"))[, 2]
)
# stargazer table for FD models
stargazer(
  fd_spi_di, fd_spi_di_L1, fd_spi_di_L2,
  se = rob_se_fd,
  type = "html",
  out = "figures/s1_fd_lags.html",
  title = "Stage I Models: First Difference (Current and Lagged Effects)",
  align = TRUE,
  dep.var.labels = "SPI Overall Score",
  column.labels = c("Contemporaneous", "Lag 1", "Lag 2"),
  covariate.labels = c("DI Composite", "Lagged DI (t-1)", "Lagged DI (t-2)",  "Log(GDPper-capita)", "Lower-Middle Income (LMIC)", "Upper-Middle Income (UMIC)", "High Income (HIC)", "Intercept"),
  report = "vcs*",
  notes = "Robust standard errors clustered by country in parentheses",
  omit = "factor\\(year\\)",
  model.names = FALSE
)
```

# 1.3) Fixed Effects [Stage 1]: SPI ~ DI
```{r}
# Contemporaneous Effect: SPI ~ DI 
fe_spi_di <- plm(
  formula = spi_comp ~ di_score + log_gdppc + income_level_recoded,
  index = c("country_code", "year"),
  data = panel_data,
  model = "within", #FE
  effect = "twoways" 
)
summary(fe_spi_di, vcov = vcovHC(fe_spi_di, cluster = "group", type = "HC1"))

# Adding Lag1: SPI ~ DI
fe_spi_di_L1 <- plm(
  formula = spi_comp ~ di_score + dplyr::lag(di_score, n=1) + log_gdppc + income_level_recoded,
  index = c("country_code", "year"),
  data = panel_data,
  model = "within", #FE
  effect = "twoways" 
)
summary(fe_spi_di_L1, vcov = vcovHC(fe_spi_di_L1, cluster = "group", type = "HC1"))

# Adding Lag2: SPI ~ DI
fe_spi_di_L2 <- plm(
  formula = spi_comp ~ di_score + dplyr::lag(di_score, n=1) + dplyr::lag(di_score, n=2) + log_gdppc + income_level_recoded,
  index = c("country_code", "year"),
  data = panel_data,
  model = "within", #FE
  effect = "twoways" 
)
summary(fe_spi_di_L2, vcov = vcovHC(fe_spi_di_L2, cluster = "group", type = "HC1"))

```
**results**

## Fixed Effects Summary Table
```{r}
# --- FE MODELS TABLE ---
# Extracting robust standard errors for the three FE models
rob_se_fe <- list(
  coeftest(fe_spi_di, vcov = vcovHC(fe_spi_di, cluster = "group", type = "HC1"))[, 2],
  coeftest(fe_spi_di_L1, vcov = vcovHC(fe_spi_di_L1, cluster = "group", type = "HC1"))[, 2],
  coeftest(fe_spi_di_L2, vcov = vcovHC(fe_spi_di_L2, cluster = "group", type = "HC1"))[, 2]
)
# stargazer table for FE models
stargazer(
  fe_spi_di, fe_spi_di_L1, fe_spi_di_L2,
  se = rob_se_fe,
  type = "html",
  out = "figures/s1_fe_lags.html",
  title = "Stage I Models: Fixed Effects (Current and Lagged Effects)",
  align = TRUE,
  dep.var.labels = "SPI Overall Score",
  column.labels = c("Contemporaneous", "Lag 1", "Lag 2"),
  covariate.labels = c("DI Composite", "Lagged DI (t-1)", "Lagged DI (t-2)",  "Log(GDPper-capita)", "Lower-Middle Income (LMIC)", "Upper-Middle Income (UMIC)", "High Income (HIC)", "Intercept"),
  report = "vcs*",
  notes = "Robust standard errors clustered by country in parentheses",
  omit = "factor\\(year\\)",
  model.names = FALSE
)
```

### stargazer table for only lag2 models
```{r}
# stargazer table for only lag2 models
stargazer(
  ols_spi_di_L2, fd_spi_di_L2, fe_spi_di_L2,
  se = list(
    coeftest(ols_spi_di_L2, vcov = vcovHC(ols_spi_di_L2, cluster = "group", type = "HC1"))[, 2],
    coeftest(fd_spi_di_L2, vcov = vcovHC(fd_spi_di_L2, cluster = "group", type = "HC1"))[, 2],
    coeftest(fe_spi_di_L2, vcov = vcovHC(fe_spi_di_L2, cluster = "group", type = "HC1"))[, 2]
  ),
  type = "html",
  title = "Stage I Models: Lagged Effects (Lag 2)",
  align = TRUE,
  dep.var.labels = "SPI Overall Score",
  column.labels = c("POLS", "FD", "FE"),
  covariate.labels = c("DI Composite", "Lagged DI (t-1)", "Lagged DI (t-2)",  "Log(GDPper-capita)", "Lower-Middle Income (LMIC)", "Upper-Middle Income (UMIC)", "High Income (HIC)", "Intercept"),
  report = "vcs*",
  notes = "Robust standard errors clustered by country in parentheses",
  omit = "factor\\(year\\)",
  model.names = FALSE,
  out = "figures/s1_lag2_mods.html"
)

```

# MODEL MISSPECIFICATION CHECKS [Stage 1]
Following standard econometric practice, I first establish the appropriate functional form for main effects, using RESET tests and residual diagnostics, before testing for theoretically motivated interaction effects.

## RESET Test for Misspecification [STAGE 1]
The RESET test checks for omitted variable bias by testing if the squared and cubed fitted values significantly improve the model fit. If the p-value is low (typically < 0.05), it suggests that the model may be misspecified, indicating that a non-linear relationship or omitted variables are present.

When you get a p-value of less than 0.05 this means that you reject the null hypothesis - this means that statistically your model is better with first order terms.
```{r}
# transforming to base R lm()
ols_spi_di <- lm(spi_comp ~ dplyr::lag(di_score, 1) + log_gdppc + income_level_recoded + factor(year), data = panel_data) # regular linear model
fd_spi_di <- lm(spi_diff ~ dplyr::lag(di_diff, 1) + log_gdppc_diff + income_level_recoded, data = fd_data)
fe_spi_di <- lm(spi_comp ~ dplyr::lag(di_score, 1) + log_gdppc + income_level_recoded + factor(year) + factor(country_code), data = panel_data)

# Performing RESET tests for the base models
resettest(ols_spi_di, power = 2:3, type = "regressor", 
          vcov = function(x) vcovHC(x, cluster = "group", type = "HC1"))
resettest(fd_spi_di, power = 2:3, type = "regressor",
          vcov = function(x) vcovHC(x, cluster = "group", type = "HC1"))
resettest(fe_spi_di, power = 2:3, type = "regressor",
          vcov = function(x) vcovHC(x, cluster = "group", type = "HC1"))

# Saving results to a dataframe
s1_reset_results <- data.frame(
  model = c("ols_spi_di", "fd_spi_di", "fe_spi_di"),
  F_statistic = c(resettest(ols_spi_di, power = 2:3, type = "regressor", 
                             vcov = function(x) vcovHC(x, cluster = "group", type = "HC1"))$statistic,
                  resettest(fd_spi_di, power = 2:3, type = "regressor",
                             vcov = function(x) vcovHC(x, cluster = "group", type = "HC1"))$statistic,
                  resettest(fe_spi_di, power = 2:3, type = "regressor",
                             vcov = function(x) vcovHC(x, cluster = "group", type = "HC1"))$statistic),
  p_value = c(resettest(ols_spi_di, power = 2:3, type = "regressor", 
                        vcov = function(x) vcovHC(x, cluster = "group", type = "HC1"))$p.value,
              resettest(fd_spi_di, power = 2:3, type = "regressor",
                        vcov = function(x) vcovHC(x, cluster = "group", type = "HC1"))$p.value,
              resettest(fe_spi_di, power = 2:3, type = "regressor",
                        vcov = function(x) vcovHC(x, cluster = "group", type = "HC1"))$p.value)
)
print(s1_reset_results)
# Save the results to a CSV file
write.csv(s1_reset_results, file = "output_CSVs/s1_reset_results.csv", row.names = FALSE)
```
We reject the Null Hypothesis of no misspecification in the relationship between DI and SPI performance, as the p-value is greater than 0.05 in in the FE model. This suggests that non-linear terms may make a significant contribution to the model.

## Stepwise Check: Applying Polynomial Terms [Stage 1]
For this section, all continuous predictors (DI & Log(GDP)) are centered for simplicity and to avoid multicollinearity issues. 

- **H0:** The relationship between SPI and SDG performance is linear.
- **H1:** The relationship between SPI and SDG performance is non-linear (quadratic or cubic).
```{r}
# No quadratic Terms Applied - Baseline 
fe_spi_di_base <- plm(formula = spi_comp ~ cen_di_score_lag1 + log_gdppc + income_level_recoded,
               model = "within", 
               index = c("country_code", "year"),
               effect = "twoways",
               data = panel_data)
rob_sum_base <-summary(fe_spi_di_base, vcov = vcovHC(fe_spi_di_base, cluster = "group", type = "HC1"))

# + Quadratic terms for DI
fe_quads_s1_m2 <- plm(formula = spi_comp ~ cen_di_score_lag1 + cen_di_score_lag1_sq + cen_log_gdppc + income_level_recoded,
               model = "within", 
               index = c("country_code", "year"),
               effect = "twoways",
               data = panel_data)
rob_sum_m2 <- summary(fe_quads_s1_m2, vcov = vcovHC(fe_quads_s1_m2, cluster = "group", type = "HC1"))

# + Quadratic terms for DI and Log(GDP)
fe_quads_s1_m3 <- plm(formula = spi_comp ~ cen_di_score_lag1 + cen_di_score_lag1_sq
                      + cen_log_gdppc + cen_log_gdppc_sq + income_level_recoded,
               model = "within", 
               index = c("country_code", "year"),
               effect = "twoways",
               data = panel_data)
rob_sum_m3 <- summary(fe_quads_s1_m3, vcov = vcovHC(fe_quads_s1_m3, cluster = "group", type = "HC1"))

# + Cubic terms for DI and Log(GDP)
fe_cubics_s1_m4 <- plm(formula = spi_comp ~ cen_di_score_lag1 + cen_di_score_lag1_sq + cen_di_score_lag1_cub + cen_log_gdppc + cen_log_gdppc_sq + cen_log_gdppc_cub + income_level_recoded,
               model = "within", 
               index = c("country_code", "year"),
               effect = "twoways",
               data = panel_data)
rob_sum_m4 <- summary(fe_cubics_s1_m4, vcov = vcovHC(fe_cubics_s1_m4, cluster = "group", type = "HC1"))

# Extracting robust standard errors for each model
rob_se_base <- coeftest(fe_spi_di_base, vcov = vcovHC(fe_spi_di_base, cluster = "group", type = "HC1"))
rob_se_m2  <- coeftest(fe_quads_s1_m2, vcov = vcovHC(fe_quads_s1_m2, cluster = "group", type = "HC1"))
rob_se_m3  <- coeftest(fe_quads_s1_m3, vcov = vcovHC(fe_quads_s1_m3, cluster = "group", type = "HC1"))
rob_se_m4 <- coeftest(fe_cubics_s1_m4, vcov = vcovHC(fe_cubics_s1_m4, cluster = "group", type = "HC1"))

# stargazer table
library(stargazer)
stargazer(fe_spi_di_base, fe_quads_s1_m2, fe_quads_s1_m3, fe_cubics_s1_m4, 
          se = list(rob_se_base[, 2], rob_se_m2[, 2], rob_se_m3[, 2], rob_se_m4[, 2]
                    ),
          title = "Applying Polynomial Terms to Stage 1 Models",
          align = TRUE,
          dep.var.labels = "SPI Overall Score",
          column.labels = c("FE Base", "Quad DI", "Quad DI+GDP", "Cubic DI+GDP"),
          #covariate.labels = c("SPI", "SPI Squared", "DI", "DI Squared",
          #                     "Log GDPpc", "Log GDPpc Squared", "Intercept"),
          notes = "All models apply fixed effects. Robust standard errors clustered by country in parentheses",
          type = "html",  # Change to "latex" for LaTeX output or "html" for Word
          report = "vcs*",  # Shows significance stars and standard errors
          model.numbers = FALSE,
          keep.stat = c("n", "rsq", "adj.rsq"),
          out = "figures/poly_s1_models.html") #saved as html

# AIC/BIC function
source('plm_aic_bic_function.R')
plm_aic_bic(rob_sum_base) # regular FE
plm_aic_bic(rob_sum_m2) # quadratic FE
plm_aic_bic(rob_sum_m3) # quadratic FE
plm_aic_bic(rob_sum_m4) # cubic FE
```

DI: We fail to reject the null hypothesis of linearity in the relationship between DI and SPI performance, as the p-value is greater than 0.05 in all models. Although literature suggests a non-linear relationship (SOURCE), the results here do not support it. 



# MODEL DIAGNOSTICS [Stage 1]

## Multicollinearity Assessment and Correction [Stage 2] [CHECK BACK W GREG - NEED TO FIGURE OUT WHAT TO DO WITH THIS]
The only model worth looking for multicollinearity is the base model, as the other models have no polynomial or interaction terms.
```{r}
# converting to base R lm() for VIF calculation
ols_spi_di <- lm(spi_comp ~ dplyr::lag(di_score, 1) + log_gdppc + income_level_recoded + factor(year), data = panel_data) # regular linear model
fe_spi_di <- lm(spi_comp ~ dplyr::lag(di_score, 1) + log_gdppc + income_level_recoded + factor(year) + factor(country_code), data = panel_data)

vif(ols_spi_di)
vif(fe_spi_di)

# Check for multicollinearity using VIF
vif_results <- vif(fe_spi_di)
# Convert VIF results to a data frame for better readability
vif_df <- data.frame(
  stage = "Stage 1",
  variable = names(vif_results),
  VIF = vif_results
)
print(vif_df)
# Save VIF results to a CSV file
write.csv(vif_df, file = "output_CSVs/s1_vif_results.csv", row.names = FALSE)
# The VIF values are all below 5, indicating no significant multicollinearity issues.
```

## Check for Autocorrelation 
```{r}
# APPLY Wooldridge Test for AR(1) Errors in FE Panel Models: pwartest()
# https://search.r-project.org/CRAN/refmans/plm/html/pwartest.html  
# This is MUCH BETTER for panel data with small T AND unbalanced panels!!!
pwartest(fe_spi_di_base) # [significant]

```
Significant p-value indicates the presence of autocorrelation in the residuals of the fixed effects model. This suggests that the errors are correlated over time, which violates one of the key assumptions of linear regression models.

This is corrected by using robust standard errors clustered by country, which accounts for the potential autocorrelation in the residuals.

## Check for Heteroskedasticity
```{r}
# Apply Breusch-Pagan test for heteroskedasticity
bptest(fe_spi_di_base, studentize = TRUE) # Heteroskedasticity [significant]
```
The Breusch-Pagan test indicates the presence of heteroskedasticity in the residuals of the fixed effects model. This suggests that the variance of the errors is not constant across observations, which violates another key assumption of linear regression models.
