---
title: "Comp2_residual_diagn"
author: "Sevastian Sanchez"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loading Data & Packages
```{r}
setwd("~/Documents/GitHub/QMSS_Thesis_Sanchez")

#load libraries 
source("packages.R")

#load function 
#source("df_years2.0_Function.R")

#load df_years() function: 2015-present
#fd_fe_data <- df_years2.0(2004, 2023)

#read merged_df csv
fd_fe_data <- read_csv("data/Main CSV Outputs/merged_final_df.csv")

#selecting vars
fd_fe_data <- fd_fe_data %>% 
  dplyr::select(country_code, year, sdg_overall, spi_comp, di_score, log_gdppc, income_level, aut_ep, dem_ep, regch_event, elect_dem, lib_dem, part_dem, delib_dem, egal_dem, academ_free) %>% 
  arrange(country_code, year) %>%  # Critical for correct lagging
  filter(year >= 2016)

# adding has_aut_ep and has_dem_ep for regressing regimes (experienced atleast 1 aut_ep/dem_ep)
fd_fe_data <- fd_fe_data %>%
  # Group by country to check for any event
  group_by(country_code) %>%
  mutate(has_aut_ep = any(aut_ep == 1, na.rm = TRUE)) %>%
  mutate(has_dem_ep = any(dem_ep == 1, na.rm = TRUE)) %>% 
  ungroup() 

# centralized variables for higher-order multicollinearity 
fd_fe_data <- fd_fe_data %>%
  mutate(
    cen_sdg_overall = sdg_overall - mean(sdg_overall, na.rm = TRUE),
    cen_spi_comp = spi_comp - mean(spi_comp, na.rm = TRUE),
    cen_di_score = di_score - mean(di_score, na.rm = TRUE),
    cen_log_gdppc = log_gdppc - mean(log_gdppc, na.rm = TRUE)
  )

# year-year lags for FD Models - DF
fd_data <- fd_fe_data %>%
  select(country_code, year, sdg_overall, di_score, spi_comp, log_gdppc) %>%
  filter(!is.na(di_score) & !is.na(spi_comp)) %>% 
  arrange(country_code, year) %>%
  group_by(country_code) %>%
  mutate(
    # first differences for selected variables
    sdg_diff = sdg_overall - dplyr::lag(sdg_overall, n=1),
    di_diff = di_score - dplyr::lag(di_score, n=1),
    spi_diff = spi_comp - dplyr::lag(spi_comp,n=1),
    log_gdppc_diff = log_gdppc - dplyr::lag(log_gdppc, n=1),
    # lagged first differences 
    di_diff_lag1 = dplyr::lag(di_diff, n=1),
    di_diff_lag2 = dplyr::lag(di_diff, n=2),
    spi_diff_lag1 = dplyr::lag(spi_diff, n=1),
    spi_diff_lag2 = dplyr::lag(spi_diff, n=2),
    log_gdppc_diff_lag1 = dplyr::lag(log_gdppc_diff, n=1),
    log_gdppc_diff_lag1 = dplyr::lag(log_gdppc_diff, n=2)
  ) %>%
  ungroup()
```


# POLS MEDIATION - RESIDUAL DIAGNOSTICS

## PLOS Residuals for mediator model: SPI ~ DI + Controls [STAGE 1]
```{r}
# Extract the data actually used in the model
ols_spi_di <- lm(spi_comp ~ di_score + log_gdppc + factor(year), data = fd_fe_data) # regular linear model
model_data1 <- model.frame(ols_spi_di)

# saving as png 
png("figures/ols1_residual_plots.png", width = 8, height = 7, units = "in", res = 300)
par(mfrow = c(2, 2))

# Residuals vs Fitted Values plot for ols_spi_di
plot(ols_spi_di, which = 1,
     main = "Residuals vs Fitted Model", caption = "ols_spi_di", pch = 1, cex = 0.65, col = "#595959")
abline(h = 0, col = "black", lty = 2, lwd = 1)
fitted_vals <- fitted(ols_spi_di)
lines(lowess(fitted_vals, resid(ols_spi_di)), col = "red", lwd =  1.5)

# Residuals vs di_score plot
plot(model_data1$di_score, resid(ols_spi_di),
     xlab = "di_score", ylab = "Residuals", main = "Residuals vs di_score", pch = 1, cex = 0.65, col = "darkred")
abline(h = 0, col = "black", lty = 2, lwd = 1)
lines(lowess(model_data1$di_score, resid(ols_spi_di)), col = "red", lwd = 1.5)

# Residuals vs log_gdppc plot
plot(model_data1$log_gdppc, resid(ols_spi_di),
     xlab = "log_gdppc", ylab = "Residuals", main = "Residuals vs log_gdppc", pch = 1, cex = 0.65, col = "darkgreen")
abline(h = 0, col = "black", lty = 2, lwd = 1)
lines(lowess(model_data1$log_gdppc, resid(ols_spi_di)), col = "red", lwd =  1.5)

dev.off() # to save 

par(mfrow = c(1, 1))
```


## POLS Residuals for base/outcome model: SDG ~ SPI + DI + Controls [STAGE 2]
```{r}
# Extract the data actually used in the base model
ols_base_lm <- lm(sdg_overall ~ spi_comp + di_score + log_gdppc + factor(year), data = fd_fe_data) # regular linear model
model_data2 <- model.frame(ols_base_lm)

# saving as png 
png("figures/ols2_residual_plots.png", width = 8, height = 7, units = "in", res = 300)
par(mfrow = c(2, 2))

# Residuals vs Fitted Values plot for ols_base_lm
plot(ols_base_lm, which = 1,
     main = "Residuals vs Fitted Model", caption = "ols_base", pch = 1, cex = 0.65, col = "#595959")
abline(h = 0, col = "black", lty = 2, lwd = 1)
fitted_vals <- fitted(ols_base_lm)
lines(lowess(fitted_vals, resid(ols_base_lm)), col = "red", lwd =  1.5)

# Residuals vs spi_comp plots
plot(model_data2$spi_comp, resid(ols_base_lm),
     xlab = "spi_comp", ylab = "Residuals", main = "Residuals vs spi_comp", pch = 1, cex = 0.65, col = "darkblue")
abline(h = 0, col = "black", lty = 2, lwd = 1)
lines(lowess(model_data2$spi_comp, resid(ols_base_lm)), col = "red", lwd =  1.5)

# Residuals vs di_score plot
plot(model_data2$di_score, resid(ols_base_lm),
     xlab = "di_score", ylab = "Residuals", main = "Residuals vs di_score", pch = 1, cex = 0.65, col = "darkred")
abline(h = 0, col = "black", lty = 2, lwd = 1)
lines(lowess(model_data2$di_score, resid(ols_base_lm)), col = "red", lwd = 1.5)

# Residuals vs log_gdppc plot
plot(model_data2$log_gdppc, resid(ols_base_lm),
     xlab = "log_gdppc", ylab = "Residuals", main = "Residuals vs log_gdppc", pch = 1, cex = 0.65, col = "darkgreen")
abline(h = 0, col = "black", lty = 2, lwd = 1)
lines(lowess(model_data2$log_gdppc, resid(ols_base_lm)), col = "red", lwd =  1.5)

dev.off() # to save 

par(mfrow = c(1, 1))
```

### Experimental - Quad term for log_gdppc in base model
```{r}
quad2_log_gdppc <- I(fd_fe_data$log_gdppc^2)

ols_base_lm_quad <- lm(sdg_overall ~ spi_comp + di_score + log_gdppc + quad2_log_gdppc + factor(year), data = fd_fe_data) # regular linear model
model_data3 <- model.frame(ols_base_lm_quad)

par(mfrow = c(2, 2))

# Residuals vs Fitted Values plot for ols_base_lm_quad
plot(ols_base_lm_quad, which = 1,
     main = "Residuals vs Fitted Model", caption = "ols_base", pch = 1, cex = 0.65, col = "#595959")
abline(h = 0, col = "black", lty = 2, lwd = 1)
fitted_vals <- fitted(ols_base_lm_quad)
lines(lowess(fitted_vals, resid(ols_base_lm_quad)), col = "red", lwd =  1.5)

# Residuals vs spi_comp plots
plot(model_data3$spi_comp, resid(ols_base_lm_quad),
     xlab = "spi_comp", ylab = "Residuals", main = "Residuals vs spi_comp", pch = 1, cex = 0.65, col = "darkblue")
abline(h = 0, col = "black", lty = 2, lwd = 1)
lines(lowess(model_data3$spi_comp, resid(ols_base_lm_quad)), col = "red", lwd =  1.5)

# Residuals vs di_score plot
plot(model_data3$di_score, resid(ols_base_lm_quad),
     xlab = "di_score", ylab = "Residuals", main = "Residuals vs di_score", pch = 1, cex = 0.65, col = "darkred")
abline(h = 0, col = "black", lty = 2, lwd = 1)
lines(lowess(model_data3$di_score, resid(ols_base_lm_quad)), col = "red", lwd = 1.5)

# Residuals vs log_gdppc plot
plot(model_data3$quad2_log_gdppc, resid(ols_base_lm_quad),
     xlab = "quad2_log_gdppc", ylab = "Residuals", main = "Residuals vs log_gdppc", pch = 1, cex = 0.65, col = "darkgreen")
abline(h = 0, col = "black", lty = 2, lwd = 1)
lines(lowess(model_data3$quad2_log_gdppc, resid(ols_base_lm_quad)), col = "red", lwd =  1.5)

par(mfrow = c(1, 1))
```


# Fixed Effects 

## FE Residuals for mediator model: SPI ~ DI + Controls [STAGE 1]
```{r}
# Extract the data actually used in the model
fe_spi_di <- lm(spi_comp ~ di_score + log_gdppc + factor(year) + factor(country_code), data = fd_fe_data)
model_data_fe1 <- model.frame(fe_spi_di)

# saving as png
png("figures/fe1_residual_plots.png", width = 8, height = 7, units = "in", res = 300)
par(mfrow = c(2, 2))

# Residuals vs Fitted Values plot for fe_spi_di
plot(fe_spi_di, which = 1,
     main = "FE Residuals vs Fitted Model", caption = "fe_spi_di [Stage 1]", pch = 1, cex = 0.65, col = "#595959")
abline(h = 0, col = "black", lty = 2, lwd = 1)
fitted_vals_fe <- fitted(fe_spi_di)
lines(lowess(fitted_vals_fe, resid(fe_spi_di)), col = "red", lwd =  1.5)

# Residuals vs di_score plot
plot(model_data_fe1$di_score, resid(fe_spi_di),
     xlab = "di_score", ylab = "Residuals", main = "FE Residuals vs di_score", pch = 1, cex = 0.65, col = "darkred")
abline(h = 0, col = "black", lty = 2, lwd = 1)
lines(lowess(model_data_fe1$di_score, resid(fe_spi_di)), col = "red", lwd = 1.5)

# Residuals vs log_gdppc plot
plot(model_data_fe1$log_gdppc, resid(fe_spi_di),
     xlab = "log_gdppc", ylab = "Residuals", main = "FE Residuals vs log_gdppc", pch = 1, cex = 0.65, col = "darkgreen")
abline(h = 0, col = "black", lty = 2, lwd = 1)
lines(lowess(model_data_fe1$log_gdppc, resid(fe_spi_di)), col = "red", lwd =  1.5)

dev.off() # to save
par(mfrow = c(1, 1))
```

## FE Residuals for base/outcome model: SDG ~ SPI + DI + Controls [STAGE 2]
```{r}
# Extract the data actually used in the base model
fe_base_lm <- lm(sdg_overall ~ spi_comp + di_score + log_gdppc + factor(year) + factor(country_code), data = fd_fe_data)
model_data_fe2 <- model.frame(fe_base_lm)

# saving as png
png("figures/fe2_residual_plots.png", width = 8, height = 7, units = "in", res = 300)
par(mfrow = c(2, 2))

# Residuals vs Fitted Values plot for fe_base_lm
plot(fe_base_lm, which = 1,
     main = "FE Residuals vs Fitted Model", caption = "fe_base_lm [Stage 2]", pch = 1, cex = 0.65, col = "#595959")
abline(h = 0, col = "black", lty = 2, lwd = 1)
fitted_vals_fe2 <- fitted(fe_base_lm)
lines(lowess(fitted_vals_fe2, resid(fe_base_lm)), col = "red", lwd =  1.5)

# Residuals vs spi_comp plots
plot(model_data_fe2$spi_comp, resid(fe_base_lm),
     xlab = "spi_comp", ylab = "Residuals", main = "FE Residuals vs spi_comp", pch = 1, cex = 0.65, col = "darkblue")
abline(h = 0, col = "black", lty = 2, lwd = 1)
lines(lowess(model_data_fe2$spi_comp, resid(fe_base_lm)), col = "red", lwd =  1.5)

# Residuals vs di_score plot
plot(model_data_fe2$di_score, resid(fe_base_lm),
     xlab = "di_score", ylab = "Residuals", main = "FE Residuals vs di_score", pch = 1, cex = 0.65, col = "darkred")
abline(h = 0, col = "black", lty = 2, lwd = 1)
lines(lowess(model_data_fe2$di_score, resid(fe_base_lm)), col = "red", lwd = 1.5)

# Residuals vs log_gdppc plot
plot(model_data_fe2$log_gdppc, resid(fe_base_lm),
     xlab = "log_gdppc", ylab = "Residuals", main = "FE Residuals vs log_gdppc", pch = 1, cex = 0.65, col = "darkgreen")
abline(h = 0, col = "black", lty = 2, lwd = 1)
lines(lowess(model_data_fe2$log_gdppc, resid(fe_base_lm)), col = "red", lwd =  1.5)

dev.off() # to save
par(mfrow = c(1, 1))
```
- No unsusual curvatures or patterns. 
- Homoscedasticity seems reasonable.
- No obvious outliers.
- Residuals appear roughly normally distributed.
- Overall, the residual diagnostics suggest that the linear regression assumptions are reasonably met for both models.

## First Differences Residuals for mediator model: SPI ~ DI + Controls [STAGE 1]
```{r}
# Extract the data actually used in the model
fd_spi_di <- lm(spi_diff ~ di_diff + log_gdppc_diff, data = fd_data)
model_data_fd1 <- model.frame(fd_spi_di)

# saving as png
png("figures/fd1_residual_plots.png", width = 8, height = 7, units = "in", res = 300)
par(mfrow = c(2, 2))

# Residuals vs Fitted Values plot for fd_spi_di
plot(fd_spi_di, which = 1,
     main = "FD Residuals vs Fitted Model", caption = "fd_spi_di [Stage 1]", pch = 1, cex = 0.65, col = "#595959")
abline(h = 0, col = "black", lty = 2, lwd = 1)
fitted_vals_fd <- fitted(fd_spi_di)
lines(lowess(fitted_vals_fd, resid(fd_spi_di)), col = "red", lwd =  1.5)

# Residuals vs di_diff plot
plot(model_data_fd1$di_diff, resid(fd_spi_di),
     xlab = "di_diff", ylab = "Residuals", main = "FD Residuals vs di_diff", pch = 1, cex = 0.65, col = "darkred")
abline(h = 0, col = "black", lty = 2, lwd = 1)
lines(lowess(model_data_fd1$di_diff, resid(fd_spi_di)), col = "red", lwd = 1.5)

# Residuals vs log_gdppc_diff plot
plot(model_data_fd1$log_gdppc_diff, resid(fd_spi_di),
     xlab = "log_gdppc_diff", ylab = "Residuals", main = "FD Residuals vs log_gdppc_diff", pch = 1, cex = 0.65, col = "darkgreen")
abline(h = 0, col = "black", lty = 2, lwd = 1)
lines(lowess(model_data_fd1$log_gdppc_diff, resid(fd_spi_di)), col = "red", lwd =  1.5)

dev.off() # to save
par(mfrow = c(1, 1))
```

## First Differences Residuals for base/outcome model: SDG ~ SPI + DI + Controls [STAGE 2]
```{r}
# Extract the data actually used in the base model
fd_base_lm <- lm(sdg_diff ~ spi_diff + di_diff + log_gdppc_diff, data = fd_data)
model_data_fd2 <- model.frame(fd_base_lm)

# saving as png
png("figures/fd2_residual_plots.png", width = 8, height = 7, units = "in", res = 300)
par(mfrow = c(2, 2))

# Residuals vs Fitted Values plot for fd_base_lm
plot(fd_base_lm, which = 1,
     main = "FD Residuals vs Fitted Model", caption = "fd_base_lm [Stage 2]", pch = 1, cex = 0.65, col = "#595959")
abline(h = 0, col = "black", lty = 2, lwd = 1)
fitted_vals_fd2 <- fitted(fd_base_lm)
lines(lowess(fitted_vals_fd2, resid(fd_base_lm)), col = "red", lwd =  1.5)

# Residuals vs spi_comp plots
plot(model_data_fd2$spi_diff, resid(fd_base_lm),
     xlab = "spi_diff", ylab = "Residuals", main = "FD Residuals vs spi_comp (Difference)", pch = 1, cex = 0.65, col = "darkblue")
abline(h = 0, col = "black", lty = 2, lwd = 1)
lines(lowess(model_data_fd2$spi_diff, resid(fd_base_lm)), col = "red", lwd =  1.5)

# Residuals vs di_score plot
plot(model_data_fd2$di_diff, resid(fd_base_lm),
     xlab = "di_diff", ylab = "Residuals", main = "FD Residuals vs di_score (Difference)", pch = 1, cex = 0.65, col = "darkred")
abline(h = 0, col = "black", lty = 2, lwd = 1)
lines(lowess(model_data_fd2$di_diff, resid(fd_base_lm)), col = "red", lwd = 1.5)

# Residuals vs log_gdppc_diff plot
plot(model_data_fd2$log_gdppc_diff, resid(fd_base_lm),
     xlab = "log_gdppc_diff", ylab = "Residuals", main = "FD Residuals vs log_gdppc (Difference)", pch = 1, cex = 0.65, col = "darkgreen")
abline(h = 0, col = "black", lty = 2, lwd = 1)
lines(lowess(model_data_fd2$log_gdppc_diff, resid(fd_base_lm)), col = "red", lwd =  1.5)

dev.off() # to save
par(mfrow = c(1, 1))
```



