---
title: "First Tests: OLS, Interactions, Subgroups"
author: "Sevastian Sanchez"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
  word_document: default
---

# All Countries, Preliminary Analysis (SPI x SDGs)

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  tidy.opts = list(width.cutoff = 60)
)
```

#FIRST: Libraries, Directory & Data 

```{r}
# set working directory 
setwd("~/Documents/GitHub/QMSS_Thesis_Sanchez")

#load libraries 
source("packages.R")
```

#SECOND: Run function in r-script: df_years_function.R   
*[ADJUST TIME OR SKIP AND LOADING DATA FROM DIRECTORY]*

```{r eval=FALSE, include=FALSE}
#df_years2.0():[Updated Version] used for extracting all available country data (based on country_codes) from specified range of years

#load function: all available data within year range 
source("df_years2.0_Function.R")

#specify start & end year
merged_full <- df_years2.0(2004, 2023) 

#saving full version as CSV output 
#write.table(merged_full, file = 'merged_full_df.csv', row.names=F, sep = ",")

# Older Version, df_years() - Archived:
#source("df_years()_Function.R")
#merged2 <- df_years(yr1=2004) %>% 
#  select(country_name.x, country_code, year, year_fct, sdg_overall, spi_comp, sci_overall, di_score, regime_type_2, #regime_type_4, regch_event, income_level, income_level_lab, gini, log_gdppc, population)
```

# THIRD: Load and Refine Data   
*[SKIP IF LOADING FROM DIRECTORY]*
```{r eval=FALSE, include=FALSE}
# load merged_full csv [SKIP IF YOU CHANGED YEARS ABOVE]
#merged_full <- read_csv('data/Main CSV Outputs/merged_full_final_df.csv')

# FILTERING FOR ALL COUNTRIES WITH ATLEAST 1 SDG VALUE 
sdg_cols <- paste0("goal", 1:17)

# Identify & remove countries where all SDG columns are NA across all years
countries_all_sdg_missing <- merged_full %>%
  group_by(country_code) %>%
  summarise(all_missing = all(if_all(all_of(sdg_cols), is.na))) %>%
  filter(all_missing) %>%
  pull(country_code)
merged_cleaned <- merged_full %>% # Removing all rows for those countries
  filter(!country_code %in% countries_all_sdg_missing)

#saving merged CSV output 
#write.table(merged_cleaned, file = 'merged_cleaned.csv', row.names=F, sep = ",")
```

# FOURTH: Load cleaned 'merged' Dataset  
*[ADJUST VARIABLES OR SKIP IF LOADING FROM CSV]*
```{r eval=FALSE, include=FALSE}
#Load final cleaned 'merged' Dataset & selecting variables 
merged_final <- read_csv("data/Main CSV Outputs/merged_cleaned.csv") %>% 
  dplyr::select(country_name, country_code, year, year_fct, sdg_overall, spi_comp, sci_overall, di_score, regime_type_2, regime_type_4, regch_event, aut_ep, dem_ep, income_level, income_level_lab, gini, log_gdppc, population, p1_use, p2_services, p3_products, p4_sources, p5_infra, goal1, goal2, goal3, goal4, goal5, goal6, goal7, goal8, goal9, goal10, goal11, goal12, goal13, goal14, goal15, goal16, goal17)

#writing final merged csv 
#write.table(merged_final, file = 'merged_final_df.csv', row.names=F, sep = ",")
```

# FINALLY: LOAD FINAL MERGED CSV 
```{r}
#load final merged df 
merged <- read_csv("data/Main CSV Outputs/merged_final_df.csv)
```

# NEW DF: counting the number of NAs (across all years) per country, by variables: SPI, SCI, SDG
```{r eval=FALSE, include=FALSE}
# List of columns to count NAs in
target_cols <- c("spi_comp", "sdg_overall", "sci_overall", "p1_use", "p2_services", 
                 "p3_products", "p4_sources", "p5_infra", paste0("goal", 1:17))

# Create new dataframe with NA counts per country
merged_na_counts <- merged %>%
  group_by(country_code) %>%
  summarise(across(
    .cols = all_of(target_cols),
    .fns = ~ sum(is.na(.x)),
    .names = "{.col}_na_count"
  )) %>%
  ungroup()

#saving merged_na_counts as CSV output 
#write.table(merged_na_counts, file = 'merged_na_counts.csv', row.names=F, sep = ",")

#load CSV 
merged_na_counts <- read_csv("data/Main CSV Outputs/merged_na_counts.csv")
```

## All countries 2019 data only
```{r eval=FALSE, include=FALSE}
#subset for only 2019 
merged_cs_2019 <- merged[merged$year==2019,]

#exporting cross-sectional 2019 data AS: merged_cs2019; will save to current working directory 
#write.table(merged_cs_2019, file = "merged_cs2019.csv", row.names=F, sep = ",")
```

# Descriptive stats & Missing Data: Full Dataset 
```{r}
# x & y variables 
df_sdg_statcap <- data.frame(merged$sdg_overall, merged$spi_comp, merged$sci_overall, merged$di_score)

#N total observations: country-years
colSums(!is.na(df_sdg_statcap))
# N total NAs
colSums(is.na(df_sdg_statcap))

# N countries per variable? 
sapply(c("sdg_overall", "di_score", "spi_comp", "sci_overall"), function(var) {
  length(unique(merged$country_code[!is.na(merged[[var]])]))
})

# N years per variable? 
sapply(c("sdg_overall", "di_score", "spi_comp", "sci_overall"), function(var) {
  length(unique(merged$year[!is.na(merged[[var]])]))
})
```
**Years 2004-2023:**

Number of observations per variable:
sdg_overall spi_comp sci_overall  di_score 
     3360      1300      2072      2560

Number of NAs per variable:
sdg_overall spi_comp sci_overall  di_score 
        0      2060      1288       800 
        
Number of countries per variable:
sdg_overall    di_score    spi_comp  sci_overall 
        168         160         165         123 

Number of years per variable:
sdg_overall    di_score    spi_comp sci_overall 
         20          16           8          17 
# Histograms 
```{r}
# SPI & SCI HISTOGRAMS: side-by-side layout
png("figures/spi_sci_hist.png", width = 9, height = 5, unit = "in", res = 300) #SPI & SCI Side By Side 
par(mfrow = c(1, 2))

# SPI histogram
hist(merged$spi_comp, main = "Statistical Performance Index (2016-2023)", 
     xlab = "Composite Score (0-100)", 
     ylab = "Frequency (country-years)")

# SCI histogram
hist(merged$sci_overall, main = "Statistical Capacity Index (2004-2020)", 
     xlab = "Composite Score (0-100)", 
     ylab = "Frequency (country-years)")

par(mfrow = c(1, 1)) #side by side 
dev.off()

# SDG histogram 
png("figures/sdg_hist.png", width = 5, height = 4, unit = "in", res = 300) #Saving to figures folder 
sdg_hist <- hist(merged$sdg_overall, main = "Sustainable Development Goals (2004-2023)", 
                 xlab = "Composite Score (0-100)", 
                 ylab = "Frequency (country-years)")
dev.off()
```

*SDG VARIABLE NOTE: While all member states have a country profile, only those with less than 20% missing data are ranked in the SDG Index, resulting in 167 countries being ranked in the 2024 report.*

## Box plots
```{r}
#SPI & SCI BOX PLOTS: SIDE BY SIDE 
png("figures/spi_sci_boxplots.png", width = 9, height = 5, units = "in", res = 300)
par(mfrow = c(1, 2))

# SPI Distribution by regime type 
boxplot(spi_comp ~ regime_type_2, data = merged,
        main = "SPI Distribution by Regime Type",
        xlab = "Regime Type",
        ylab = "Composite Score (1-100)",
        col = c("salmon", "skyblue"))

legend("bottomright",                      # Position of the legend
       legend = c("Democracy", "Autocracy"),  # Group names
       fill = c("skyblue", "salmon"),        # Colors matching the boxplot
       border = "black",                     # Box border color (optional)
       bty = "n")   


# SCI Distribution by regime type (only considers lower income / developing countries)
boxplot(sci_overall ~ regime_type_2, data = merged,
        main = "SCI Distribution by Regime Type",
        xlab = "Regime Type",
        ylab = "Composite Score (1-100)",
        col = c("salmon", "skyblue"))

legend("bottomright",                      # Position of the legend
       legend = c("Democracy", "Autocracy"),  # Group names
       fill = c("skyblue", "salmon"),        # Colors matching the boxplot
       border = "black",                     # Box border color (optional)
       bty = "n")   

par(mfrow = c(1, 1)) #side by side 
dev.off()

# SDG Box plot 
png("figures/sdg_boxplots.png", width = 5, height = 4, units = "in", res = 300)
boxplot(sdg_overall ~ regime_type_2, data = merged,
        main = "SDG Distribution by Regime Type",
        xlab = "Regime Type",
        ylab = "Composite Score (1-100)",
        col = c("salmon", "skyblue"))

legend("bottomright",                      # Position of the legend
       legend = c("Democracy", "Autocracy"),  # Group names
       fill = c("skyblue", "salmon"),        # Colors matching the boxplot
       border = "black",                     # Box border color (optional)
       bty = "n")   

dev.off()
```

##Summary Statistics and Quantile Frequency Tables 
```{r}
#New Vars - Quantile breaks 
quant_breaks_sdg <- quantile(merged$sdg_overall, probs = c(0, 0.25, 0.5, 0.75, 1), na.rm = TRUE)
quant_breaks_spi <- quantile(merged$spi_comp, probs = c(0, 0.25, 0.5, 0.75, 1), na.rm = TRUE)
quant_breaks_sci <- quantile(merged$sci_overall, probs = c(0, 0.25, 0.5, 0.75, 1), na.rm = TRUE)

# Inputting SDG quartiles in main DF 
merged$sdg_quant_bins <- cut(merged$sdg_overall, 
                      breaks = quant_breaks_sdg, 
                      include.lowest = TRUE, 
                      labels = c("Q1 (Lowest)", "Q2", "Q3", "Q4 (Highest)"))

#SDG IQR Summary Table 
sdg_summary <- merged %>%
  group_by(regime_type_2, sdg_quant_bins) %>%
  summarise(count = n(), .groups = "drop")
# Pivot to wide format
sdg_wide_table <- sdg_summary %>%
  pivot_wider(names_from = sdg_quant_bins, values_from = count, values_fill = 0)

# Inputting SPI quartiles in main DF 
merged$spi_quant_bins <- cut(merged$spi_comp, 
                      breaks = quant_breaks_spi, 
                      include.lowest = TRUE, 
                      labels = c("Q1 (Lowest)", "Q2", "Q3", "Q4 (Highest)"))

#SPI IQR Summary Table 
spi_summary <- merged %>%
  group_by(regime_type_2, spi_quant_bins) %>%
  summarise(count = n(), .groups = "drop")
# Pivot to wide format
spi_wide_table <- spi_summary %>%
  pivot_wider(names_from = spi_quant_bins, values_from = count, values_fill = 0)

# Inputting SCI quartiles in main DF 
merged$sci_quant_bins <- cut(merged$sci_overall, 
                      breaks = quant_breaks_sci, 
                      include.lowest = TRUE, 
                      labels = c("Q1 (Lowest)", "Q2", "Q3", "Q4 (Highest)"))

# SCI IQR Summary Table 
sci_summary <- merged %>%
  group_by(regime_type_2, sci_quant_bins) %>%
  summarise(count = n(), .groups = "drop")
# Pivot to wide format
sci_wide_table <- sci_summary %>%
  pivot_wider(names_from = sci_quant_bins, values_from = count, values_fill = 0)

#Summary tables by score: Min, 1st Qu. Median, Mean, 3rd Qu., Max, NA's 
summary(merged$sdg_overall)
summary(merged$spi_comp)
summary(merged$sci_overall)
summary(merged$di_score)

#write tables as CSV to directory 


# STACKED BAR PLOTS LATER 

```

summary(merged$sdg_overall)
   Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
  35.65   55.74   65.53   64.43   73.05   86.42 

summary(merged$spi_comp)
   Min. 1st Qu.  Median    Mean 3rd Qu.    Max.    NA's 
  11.77   54.82   67.09   66.71   81.48   95.26    2060 

summary(merged$sci_overall)
   Min. 1st Qu.  Median    Mean 3rd Qu.    Max.    NA's 
  16.67   57.78   70.00   67.96   80.00   98.89    1288 

summary(merged$di_score)
   Min. 1st Qu.  Median    Mean 3rd Qu.    Max.    NA's 
  0.260   3.520   5.780   5.509   7.280   9.930     800 

# Descriptive stats & Missing Data: Transitioning Countries Dataset 
```{r}
# new variable: experienced either kind of regime change event 
merged <- merged %>%
  group_by(country_code) %>%
  mutate(either_event = as.integer(any(dem_ep == 1 | aut_ep == 1))) %>%
  ungroup()

# subset of countries that experienced atleast 1 backsliding event  
regressed <- merged %>%
  group_by(country_code) %>%
  filter(aut_ep == 1) %>%
  ungroup()

# subset of countries that experienced atleast 1 democratization event 
democratized <- merged %>% 
  group_by(country_code) %>%
  filter(dem_ep == 1) %>%
  ungroup()

# number of countries that experienced a regime change event (autocratization or democratization)
n_countries_events <- length(unique(merged$country_code[merged$either_event == 1]))
n_countries_events

# number of countries that experienced autocratization
n_countries_aut <- length(unique(merged$country_code[merged$aut_ep == 1]))
n_countries_aut

# number of countries that experienced democratization
n_countries_dem <- length(unique(merged$country_code[merged$dem_ep == 1]))
n_countries_dem

#Frequency of countries that experienced democratization 2004-2024
hist(merged$year[merged$regch_event==1], breaks = 20)

#Frequency of countries that experienced autocratization 2004-2024
hist(merged$year[merged$regch_event==-1], breaks = 20)

#Frequency of countries that experienced no change 2004-2024
hist(merged$year[merged$regch_event==0], breaks = 20)

#frequency/proportion table: instances of backsliding, stability, democratization from 2004-2024
table(merged$regch_event) # frequency
prop.table(table(merged$regch_event)) # proportions 

```

# COMPONENT 1: COMPARING SPI & SCI X VARIABLES

**Aggregated SPI & SDG Scores**  
*H0: Null, there is no relationship*  
*H1: there is a statistically significant relationship between overall SPI and SDG composite scores*

```{r}
#correlation coefficients (r-squared), WITHOUT control variables 

#x-var 1 = spi
correlation_sdg_spi <- cor(merged$sdg_overall, merged$spi_comp, use = "complete.obs")^2

#x-var 2 = sci
correlation_sdg_sci <- cor(merged$sdg_overall, merged$sci_overall, use = "complete.obs")^2 

#x-var 3 = di
correlation_sdg_di <- cor(merged$sdg_overall, merged$di_score, use = "complete.obs")^2 

# pasting result 
string_corcoef <- "Correlation coefficient:" 
paste(string_corcoef, correlation_sdg_spi, "(SPI)", correlation_sdg_sci, "(SCI)", correlation_sdg_di, "(DI)") 
```

Correlation coefficient/R-sq (SPI): 0.616037202309322  
Correlation coefficient/R-sq (SCI): 0.410940651230861
Correlation coefficient/R-sq (DI): 0.452968121616442

## Comparing SPI & SCI w/o controls 
Finding estimated impact of variables on SDG status prior to adding controls or robust SEs
```{r}
ols_spi_naive <- lm(sdg_overall ~ spi_comp, data = merged)
summary(ols_spi_naive)

# 2. OLS for SCI and SDG - Overall 
ols_sci_naive <- lm(sdg_overall ~ sci_overall, data = merged)
summary(ols_sci_naive)

# 3. Multiple Regression with both SPI and SCI
ols_multiple_naive <- lm(sdg_overall ~ spi_comp + sci_overall, data = merged)
summary(ols_multiple_naive)
```
ols_spi_naive: 0.47806 (p-value < 0.001)  
ols_sci_naive: 0.39081 (p-value < 0.001)  
ols_multiple_naive: spi: 0.28779 (p-value < 0.001); sci: 0.15311 (p-value < 0.001)

The impact of SCI on SDG and SPI on SDG are statistically significant, in all models. SPI appears to have a greater impact on SDGs compared to that of SCI, regardless of the model. All of this is without controls. 

## TEST 1: Comparing SPI & SCI WITH controls AND Robust Standard Errors  
Applying controls and robust (Huber-White) standard errors  

*H0: Null, SCI model > SPI model*   
*H1: SPI model > SCI model*

```{r}
# 1. OLS for SPI and SDG - Overall 
ols_spi <- lm(sdg_overall ~ spi_comp + di_score + log_gdppc + population + gini + year_fct, data = merged)
summary(ols_spi)
# Robust SEs
coeftest(ols_spi, vcov = vcovHC(ols_spi, type = "HC1"))

# 2. OLS for SCI and SDG - Overall 
ols_sci <- lm(sdg_overall ~ sci_overall + log_gdppc + population + di_score + gini + year_fct, data = merged)
summary(ols_sci)
# Robust SEs
coeftest(ols_sci, vcov = vcovHC(ols_sci, type = "HC1"))

# 3. Multiple Regression with both SPI and SCI
ols_multiple <- lm(sdg_overall ~ spi_comp + sci_overall + log_gdppc + population + di_score + gini + year_fct, data = merged)
summary(ols_multiple)
# Robust SEs
coeftest(ols_multiple, vcov = vcovHC(ols_multiple, type = "HC1"))

#for right side pane & interactivity 
datatable(broom::tidy(ols_spi)) # SDG ~ SPI
datatable(broom::tidy(ols_sci)) # SDG ~ SCI
datatable(broom::tidy(ols_multiple)) # SDG ~ SPI & SCI 
```

We reject the null hypothesis that there is no relationship between SPI and SDG composite scores. Additionally, we reject the null hypothesis that there is no relationship between SCI and SDG composite scores. Holding all else constant (log GDP per capita, democracy score, population and GINI coefficient), SPI and SCI exhibit positive moderate and statistically significant relationships with SDG status.

ols_spi: 0.26573 (p-value < 0.001)  
ols_sci: 0.19833 (p-value < 0.001)  
ols_multiple: spi: 0.11928 (p-value < 0.001); sci: 0.13485 (p-value < 0.001)

When compared in separate models, SPI has a greater impact on SDG status (0.26573) than SCI (0.19833). This suggests that a one-unit increase in SPI is associated with a larger improvement in SDG outcomes compared to a one-unit increase in SCI, holding all controls constant. 

Interestingly, the opposite holds true in a multiple regression model containing both SPI and SCI. SPI's impact on SDG status (0.11928) (net of SPI) is less than that of SCI's (0.13485) (net of SCI), holding all controls constant. When together, the coefficients represent the unique impact of each predictor variable (measures of statistical capacity) on SDG status, net of all other variables.

Model 1 (ols_spi) does not control for SCI and model 2 (ols_sci) does not control for spi -- this is okay. SPI is the predecessor of the SCI, sharing/data overlap, and so it is expected to have significant statistical correlation (multicollinearity). This is likely what explains the significant reduction of both coefficients as seen in model 3: 0.26573 to 0.11928 for SPI (__% decrease); and from 0.19833 to 0.13485 for SCI (__% decrease). Such indicates that they're both capturing much of the same underlying relationship with SDG status.

However, the fact that both SPI and SCI remain significant when included together (model 3) with the highest adjusted R-sq (0.8414) suggests that they capture different dimensions of statistical capacity that independently contribute to SDG status.

### Multicolinearity & VIF: SPI & SCI 

```{r}
# Check correlation between SPI and SCI
cor(spi_comp, sci_overall, use = "complete.obs")

# Check VIF (Variance Inflation Factor) in Model 3
vif(ols_multiple)
```

**colinearity:** The correlation between SCI and SPI is about 0.8277. When placed within the same model, SCI inflated the standard error of SPI from 0.02010 to 0.02783. SCI had a similar reaction from the SPI with its standard error increasing from 0.01677 to 0.02822.

**VIF:** Such multicolinearity is reflected by the VIF test which accounts for all x variables in the model instead of just the two measures of statistical capacity (SCI & SPI).

GVIF -- continuous vars 
spi_comp: 2.996983 
sci_overall: 2.802763 
log_gdppc: 1.504559 
population: 1.055717 
di_score: 1.447762 
gini: 1.319256 

GVIF^(1/(2*Df)) -- categorical vars 
year: 1.038801 

Overall there reveals no severe multicollinearity (all GVIF < 5). There is mild correlation between your statistical capacity measures (spi_comp and sci_overall) with SPI slightly inflated by a factor of 2.996983 and SCI inflated by a factor of  2.802763. All other variables show minimal multicollinearity concerns, so it is acceptable to use in the same model as doing so will not severely impact estimates given both factors are less than 5.0.

### Missing Data Structure & Interpretations 
**Systematic, non-random missing data pattern:** SPI has near complete country data coverage, but with a stubborn temporal limitation (2016-2023). On the other hand, SCI has longer temporal coverage (2004-2020), but lacks reporting on high-income countries focusing primarily on the developing world. 

In model 1 (SDG ~ SPI), democracy score (di_score) is non-significant (-0.0243, p=0.8673). However, in model 2 (SDG ~ SCI) democracy score is highly significant (0.5556***, p < 0.001). In model 3, with both SCI and SPI, democracy score is marginally significant (0.0301, p=0.0484). This suggests that SCI's relationship with SDG outcomes may be closely linked to regime/democratic governance. Considering the non-random missing data structured previously mentioned, the difference in significance between the models for democracy score signals an even greater need to perform subgroup analysis of countries at different stages/levels of development. 


### AIC/BIC Checking Fit [FIX TEST- #N DIFFERS BTW MODELS]

```{r}
# Compare all three models with AIC
#AIC(ols_spi, ols_sci, ols_multiple)

# Compare all three models with BIC
#BIC(ols_spi, ols_sci, ols_multiple)
```

**Best fit:** ols_multiple (Adj Rsq: 0.7898) (AIC/BIC: ___)

Adj. R-sqs
SPI: 0.7983
SCI: 0.8019
Both: 0.8455

**Selecting model:** This analysis is specifically focused on overall statistical capacity rather than comparisons of measures. The SCI model reveals a better fit than the SPI model (Adj Rsq: 0.8019 > 0.7983). Model 3, containing both SPI and SCI, sustains a better fit compared to both model 1 with just SPI and model 2 with just SCI (Adj Rsq: 0.8455 > 0.7983; 0.8455 > 0.8019). Model 1 and 2 have significantly more country-year data points (n=457 and n=571, respectively) for regression analysis compared to model 3 (n=210). 
[STILL TO DECIDE]


### Combine to single index: principle component analysis (FOR FUTURE)

```{r}
# Standardize both measures
#spi_z <- scale(spi_comp)
#sci_z <- scale(sci_overall)

# Create composite (simple average)
#stat_capacity_index <- (spi_z + sci_z)/2

#extract common variance 
#pca_result <- prcomp(cbind(spi_comp, sci_overall), scale = TRUE)
#stat_capacity_pc1 <- pca_result$x[,1]  # First principal component
```

### Visual Analysis: SCI & SPI x SDG
```{r}
#visualize differences in fit
library(ggplot2)
library(plotly)

#define regression line colors
spi_line <- "steelblue4"
sci_line <- "darkgoldenrod"

# Creating scatterplot with both SPI and SCI on the same plot
Compare_fit <- ggplot(merged, aes(x = spi_comp, y = sdg_overall))+
  geom_smooth(aes(x = spi_comp, y = sdg_overall), 
              color = spi_line, 
              method = "lm", 
              linewidth = 0.75, 
              se = FALSE)+ # Regression line SPI
  geom_smooth(aes(x = sci_overall, y = sdg_overall), 
              color = sci_line, 
              method = "lm", 
              linewidth = 0.75, 
              se = FALSE)+ # Regression line for SCI
  geom_point(aes(color = "spi_comp"), alpha=0.50, size = 0.5)+ # Scatter plot for SPI
  geom_point(aes(x = sci_overall, y = sdg_overall, color = "sci_overall"), 
             alpha=0.5, size = 0.5)+ # Add SCI points w/different color
  scale_color_manual(values = c("spi_comp" = "steelblue1", 
                                "sci_overall" = "darkgoldenrod1")) +
  labs(title = "SDG vs. SPI and SCI",
       x = "SPI / SCI",
       y = "SDG",
       color = "Index") +  # Title for legend
  theme_bw() # Optional: adds a clean, black and white theme

Compare_fit

#make interactive
ggplotly(Compare_fit)

# Save to specific folder
# ggsave("~/Documents/GitHub/QMSS_Thesis_Sanchez/Output_CSVs/fd_plot.png", p, width = 10, height = 6)
```

# INTERACTIONS AND SUBGROUP ANALYSIS

## TEST 2: Checking for Interactions [REDO RESULTS FOR ROBUST SEs]: 
- Is there a need for subgroup analysis, and if so, by what kind of group? 
- Options: GNI Classification (income_level), regime_type_2, regime_type_4, di_score
```{r}
#interaction 1: does GNI Classification (income_level) affect the relationship between x (spi) & y (sdg)?
inc_lev_interaction <- lm(sdg_overall ~ spi_comp + spi_comp*income_level + log_gdppc + population + gini + year_fct,data = merged)
#summary(inc_lev_interaction)
coeftest(inc_lev_interaction, vcov = vcovHC(inc_lev_interaction, type = "HC1")) #Robust SE

#interaction 2: does regime_type_2 affect the relationship between x (spi) & y (sdg)?
reg_type2_interaction <- lm(sdg_overall ~ spi_comp + spi_comp*regime_type_2 + log_gdppc + population + gini + year_fct,data = merged)
#summary(reg_type2_interaction)
coeftest(reg_type2_interaction, vcov = vcovHC(reg_type2_interaction, type = "HC1")) #Robust SE

#interaction 3: does regime_type_4 affect the relationship between x (spi) & y (sdg)?
reg_type4_interaction <- lm(sdg_overall ~ spi_comp + spi_comp*regime_type_4 + log_gdppc + population + gini + year_fct,
                        data = merged)
summary(reg_type4_interaction)
coeftest(reg_type4_interaction, vcov = vcovHC(reg_type4_interaction, type = "HC1")) #Robust SE

#interaction 4: does di affect the relationship between x (spi) & y (sdg)?
reg_type_di_interaction <- lm(sdg_overall ~ spi_comp + spi_comp*di_score + log_gdppc + population + gini + year_fct,
                        data = merged)
summary(reg_type_di_interaction)
coeftest(reg_type_di_interaction, vcov = vcovHC(reg_type_di_interaction, type = "HC1")) #Robust SE
```
*GNI Income Classification: Yes* there are statistically significant interactions found from GNI classifications that affects the relationship between spi and sdgs

*Binary Regime Type: No* there is no statistically significant interactions found from regime type (autocracy vs democracy) that affects the relationship between spi and sdgs. 

*Categorical Regime type (4 options): No* there is no statistically significant interactions found from regime type (Closed autocracy, electoral autocracy, electoral democracy, liberal democracy) that affects the relationship between spi and sdgs. 

*Continuous di_score [0-1] Regime type: No* there is no statistically significant interactions found from regime type (infinate between 0-1) that affects the relationship between spi and sdgs. 

## TEST 3: WB GNI Classifications: income_level ("H", "UM", "LM", "L") [REDO RESULTS FOR ROBUST SEs]
**Disaggregated/Grouped by Development Status:** Make 4 regression models and then put them all together in a table to compare the slopes and R-sq values.
```{r}
# 1. Overall model (all countries)
overall_lm <- lm(sdg_overall ~ spi_comp + di_score + log_gdppc + population + gini + year_fct,
                        data = merged)
summary(overall_lm)
coeftest(overall_lm, vcov = vcovHC(overall_lm, type = "HC1")) #Robust SE

# 2. High income countries
high_inc_lm <- lm(sdg_overall ~ spi_comp + di_score + log_gdppc + population + gini + year_fct,
                        data = merged %>% 
                          filter(income_level == "H"))
summary(high_inc_lm)
coeftest(high_inc_lm, vcov = vcovHC(high_inc_lm, type = "HC1")) #Robust SE

# 3. Upper-middle income countries
upper_mid_lm <- lm(sdg_overall ~ spi_comp + di_score + log_gdppc + population + gini + year_fct,
                         data = merged %>% 
                           filter(income_level == "UM"))
summary(upper_mid_lm)
coeftest(upper_mid_lm, vcov = vcovHC(upper_mid_lm, type = "HC1")) #Robust SE

# 4. Lower-middle income countries
lower_mid_lm <- lm(sdg_overall ~ spi_comp + di_score + log_gdppc + population + gini + year_fct,
                         data = merged %>% 
                           filter(income_level == "LM"))
summary(lower_mid_lm)
coeftest(lower_mid_lm, vcov = vcovHC(lower_mid_lm, type = "HC1")) #Robust SE

# 5. Low income countries
low_inc_lm <- lm(sdg_overall ~ spi_comp + di_score + log_gdppc + population + gini + year_fct,
                       data = merged %>% 
                         filter(income_level == "L"))
summary(low_inc_lm)
coeftest(low_inc_lm, vcov = vcovHC(low_inc_lm, type = "HC1")) #Robust SE
```

### Extract coefficients and statistics of subgroups to comparison table
```{r}
# Creating mod_compare_tab - comparison table
mod_compare_tab <- data.frame(
  model = c("Overall", "High Income", "Upper-Middle Income", 
           "Lower-Middle Income", "Low Income"),
  coefficient = c(
    coef(overall_lm)["spi_comp"],
    coef(high_inc_lm)["spi_comp"],
    coef(upper_mid_lm)["spi_comp"],
    coef(lower_mid_lm)["spi_comp"],
    coef(low_inc_lm)["spi_comp"]
  ),
  
  intercept = c(
    coef(overall_lm)["(Intercept)"],
    coef(high_inc_lm)["(Intercept)"],
    coef(upper_mid_lm)["(Intercept)"],
    coef(lower_mid_lm)["(Intercept)"],
    coef(low_inc_lm)["(Intercept)"]
  ),
  
  Std_error = c(
  summary(overall_lm)$coefficients["spi_comp", 'Std. Error'], # just standard errors 
  summary(high_inc_lm)$coefficients["spi_comp", 'Std. Error'], # just standard errors 
  summary(upper_mid_lm)$coefficients["spi_comp", 'Std. Error'], # just standard errors 
  summary(lower_mid_lm)$coefficients["spi_comp", 'Std. Error'], # just standard errors 
  summary(low_inc_lm)$coefficients["spi_comp", 'Std. Error'] # just standard errors 
  ),

  t_value = c(
    summary(overall_lm)$coefficients["spi_comp", "t value"],
    summary(high_inc_lm)$coefficients["spi_comp", "t value"],
    summary(upper_mid_lm)$coefficients["spi_comp", "t value"],
    summary(lower_mid_lm)$coefficients["spi_comp", "t value"],
    summary(low_inc_lm)$coefficients["spi_comp", "t value"]
  ),
  
  p_value = c(
    summary(overall_lm)$coefficients["spi_comp", "Pr(>|t|)"],
    summary(high_inc_lm)$coefficients["spi_comp", "Pr(>|t|)"],
    summary(upper_mid_lm)$coefficients["spi_comp", "Pr(>|t|)"],
    summary(lower_mid_lm)$coefficients["spi_comp", "Pr(>|t|)"],
    summary(low_inc_lm)$coefficients["spi_comp", "Pr(>|t|)"]
  ),
  
  r_squared = c(
    summary(overall_lm)$r.squared,
    summary(high_inc_lm)$r.squared,
    summary(upper_mid_lm)$r.squared,
    summary(lower_mid_lm)$r.squared,
    summary(low_inc_lm)$r.squared
  ),
  
  adj_r_squared = c(
    summary(overall_lm)$adj.r.squared,
    summary(high_inc_lm)$adj.r.squared,
    summary(upper_mid_lm)$adj.r.squared,
    summary(lower_mid_lm)$adj.r.squared,
    summary(low_inc_lm)$adj.r.squared
  ),
  
  n_obs = c(
    nobs(overall_lm),
    nobs(high_inc_lm),
    nobs(upper_mid_lm),
    nobs(lower_mid_lm),
    nobs(low_inc_lm)
  )
)

# Defining function for significance stars based on p-values
sig_stars <- function(p_value) {
  if (p_value <= 0.001) {
    return("***")
  } else if (p_value <= 0.01) {
    return("**")
  } else if (p_value <= 0.05) {
    return("*")
  } else if (p_value <= 0.1) {
    return(".")
  } else {
    return("")
  }
}

# make table and round for better display
mod_compare_tab <- mod_compare_tab %>%
  mutate(
    coefficient = round(coefficient, 3),
    intercept = round(intercept, 3),
    Std_error = round(Std_error, 3),
    t_value = round(t_value, 3),
    p_value = p_value,
    significance = sapply(p_value, sig_stars), #significance stars 
    r_squared = round(r_squared, 3),
    adj_r_squared = round(adj_r_squared, 3)
  )

# Print the comparison table
print(mod_compare_tab)

# export mod_compare_tab 
# write.csv(mod_compare_tab, file = "ols_model_comparison.csv", row.names=F) 
```



##Visualizing Slopes: plotting multiple regression - by subgroup
```{r}
viz_gni_class <- ggplot(data = merged, aes(x = spi_comp, 
                                           y = sdg_overall, 
                                           color = income_level_lab)) +
  geom_point(alpha = 0.25, size = 0.75) + 
  # Overall regression line (black)
  geom_smooth(aes(group = 1), 
              method = "lm", 
              linewidth = 0.75, 
              se = FALSE, 
              color = "black") +
  # Group-specific regression lines
  geom_smooth(method = "lm", 
              linewidth = 0.65, 
              se = FALSE) +
  scale_color_manual(
    values = c("High Income Countries" = "#1D6A96", 
               "Upper-Middle Income Countries" = "#4CB5AE",
               "Lower-Middle Income Countries" = "#F3A738",
               "Low Income Countries" = "#C02942")
  ) +
  labs(title = "Relationship between SPI and SDG by World Bank Income Classification",
       x = "Statistical Performance Indicators (SPI)",
       y = "Sustainable Development Goals (SDG)",
       color = "Income Classification") +
  theme_bw() 

viz_gni_class
ggplotly(viz_gni_class)

# Save to specific folder
# ggsave("~/Documents/GitHub/QMSS_Thesis_Sanchez/Output_CSVs/fd_plot.png", p, width = 10, height = 6)
```

## HETERSCKEDISTACITY: Linearity vs non-linearity 
```{r}
plot(lm_spi, which = 1)  # residuals for SPI model  
plot(lm_sdg, which = 1)  # residuals for SDG model 
plot(lm_sdg_controlled, which = 1)  # residuals for SDG model controlled

library(lmtest)
resettest(lm_spi, power = 2:3, type = "fitted")
resettest(lm_sdg, power = 2:3, type = "fitted")
resettest(lm_sdg_controlled, power = 2:3, type = "fitted")

#validate with sensitivity test 
#sensmediation::sensmed(model_m, model_y, sims = 500)
```

## STILL LEFT TO INCLUDE 
- Include GNI Coefficient control in all models 
- Update results based on these changes

