---
title: "correlation_matrix_SDG_SPI"
author: "Sevastian Sanchez"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r cars}
#load libraries/packages
source("packages.R")

#load function 
source("df_years()_Function.R")
merged_matrix <- df_years(yr1 = 2015) 
```

#Partial correlation & package 
Partial correlation controls for confounding variables: Ensures correlations between SPI pillars and SDG goals are adjusted for external factors like log(GDP) and population (control1, control2).

Focuses on Direct Relationships: Removes indirect effects of controls on both SPI and SDG metrics.
```{r}
# Example: Filter relevant columns
matrix_data <- merged_matrix %>%
  dplyr::select(p1_use, p2_services, p3_products, p4_sources, p5_infra,
                goal1, goal2, goal3, goal4, goal5, goal6, goal7, goal8,
                goal9, goal10, goal11, goal12, goal13, goal14, goal15,
                goal16, goal17, gdppc, population)
```

# Initialize an empty matrix to store results
```{r}
spi_pillars <- c("p1_use", "p2_services", "p3_products", "p4_sources", "p5_infra")
sdg_goals <- paste0("goal", 1:17)
partial_corr_matrix <- matrix(NA, nrow = length(spi_pillars), ncol = length(sdg_goals),
                              dimnames = list(spi_pillars, sdg_goals))

# Loop through SPI pillars and SDG goals
for (spi in spi_pillars) {
  for (sdg in sdg_goals) {
    # Create temporary dataset with complete cases
    temp_data <- matrix_data[, c(spi, sdg, "gdppc", "population")] %>%  # Fixed: added quotes
      na.omit()
    
    # Only calculate if sufficient data remains
    if (nrow(temp_data) >= 4) {
      result <- ppcor::pcor(temp_data)  # Explicit package reference
      partial_corr_matrix[spi, sdg] <- result$estimate[1, 2]
    }
  }
}

# Convert to data frame
matrix_df <- as.data.frame(partial_corr_matrix)

```

#Visualize Correlation Matrix - option 1
```{r}
# Reshape the matrix into long format
melted_cormat <- melt(partial_corr_matrix)

# Visualize 
ggplot(data = melted_cormat, aes(x = Var2, y = Var1, fill = value)) +
  geom_tile(color = "white") +  # Add grid lines
  scale_fill_gradient2(low = "purple", mid = "white", high = "darkgreen", midpoint = 0,
                       limits = c(-1, 1), name = "Correlation") +  # Adjust color scale
  geom_tile(data = subset(melted_cormat, is.na(value)), aes(x = Var2, y = Var1), 
            fill = "grey", color = "white") +  # Grey tiles for NA values
  theme_minimal() +
  labs(title = "Correlation Matrix Heatmap",
       x = "SDG Goals",
       y = "SPI Pillars") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

#Visualize Correlation Matrix - option 2
```{r}
ggplot(data = melted_cormat, aes(x = Var2, y = Var1, fill = value)) +
  geom_tile(color = "white") +
  scale_fill_gradient2(low = "purple", mid = "white", high = "darkgreen", midpoint = 0,
                       limits = c(-1, 1), name = "Correlation") +
  # Add correlation values (2 decimal places)
  geom_text(aes(label = ifelse(is.na(value), "", round(value, 2))), 
            color = "black", size = 3) +
  # Grey tiles for NA values
  geom_tile(data = subset(melted_cormat, is.na(value)), 
            aes(x = Var2, y = Var1), fill = "grey90", color = "white") +
  theme_minimal() +
  labs(title = "Partial Correlation Matrix: SPI Pillars vs SDG Goals",
       x = "SDG Goals", 
       y = "SPI Pillars") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 9),
        axis.text.y = element_text(size = 9),
        legend.position = "right") +
  coord_fixed()  # Maintain square aspect ratio

geom_text(aes(color = abs(value) > 0.5), 
          label = ifelse(is.na(value), "", round(value, 2)), 
          size = 3) +
scale_color_manual(guide = "none", values = c("black", "white"))

```

