---
title: "correlation_matrix_SDG_SPI"
author: "Sevastian Sanchez"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Correlation Matrix: Stage 1 & 2 predictors 
create correlation matrix for SDG score, SPI score, DI score and log GDP per capita
```{r}
# Load necessary libraries
library(plm)
library(ggplot2)
library(reshape2)
library(dplyr)
library(car)  # For VIF calculation

# set working directory
setwd("~/Documents/GitHub/QMSS_Thesis_Sanchez")

#load libraries/packages
source("packages.R")

# load data
source("Comp2_panel_wrangling.R")

panel_data <- panel_data %>%
  dplyr::select(sdg_overall, spi_comp, di_score, log_gdppc, country_code, year)
# Impute missing data: log_gdppc
panel_data <- panel_data %>%
  mutate(
    log_gdppc = ifelse(is.na(log_gdppc), median(log_gdppc, na.rm = TRUE), log_gdppc)
  )

# Calculate the correlation matrix
correlation_matrix <- cor(panel_data[, c("sdg_overall", "spi_comp", "di_score", "log_gdppc")], use = "pairwise.complete.obs")

# Convert the correlation matrix to a data frame for ggplot
melted_correlation <- melt(correlation_matrix)

# Visualize the correlation matrix using ggplot2
correlation_plot <- ggplot(data = melted_correlation, aes(x = Var2, y = Var1, fill = value)) +
  geom_tile(color = "black") +
  scale_fill_gradient2(
    low = "red", mid = "white", high = "darkgreen", midpoint = 0,
    limits = c(-1, 1), name = "Correlation"
  ) +
  geom_text(aes(label = round(value, 2)), color = "black", size = 3) +
  theme_minimal() +
  labs(title = "Correlation Matrix: Main Variables",
       x = "", y = "") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Show the correlation plot
correlation_plot

# save the plot
ggsave("figures/correlation_matrix_main_vars.png", plot = correlation_plot, width = 5, height = 4, dpi = 300)

# VIF Test: Stage 1 & 2 predictors
# Load necessary libraries
library(plm)
library(car)  # For VIF calculation
# Calculate VIF for the predictors in the panel data
vif_data <- panel_data %>%
  dplyr::select(sdg_overall, spi_comp, di_score, log_gdppc)
# Fit a linear model to assess multicollinearity
vif_model <- lm(sdg_overall ~ spi_comp + di_score + log_gdppc, data = vif_data)
# Calculate VIF
vif_values <- vif(vif_model)
# Print VIF values
vif_values

```


# VIF Test: Stage 1 POLS Model 
```{r}
# BASELINE POLS MODEL
ols_spi_di_reg <- plm(formula = sdg_overall ~ spi_comp + di_score + log_gdppc + factor(year),
               model = "pooling", 
               index = c("country_code", "year"),
               data = panel_data)

# Assess multi-colinearity 
vif(ols_spi_di_reg)

# CENTERED POLS MODEL
ols_spi_di_cen <- plm(formula = sdg_overall ~ cen_spi_comp + cen_di_score + cen_log_gdppc + factor(year),
               model = "pooling", 
               index = c("country_code", "year"),
               data = panel_data)

# Assess multi-colinearity 
vif(ols_spi_di_cen)
```

# VIF Test: Stage 2 POLS Model
```{r}
# BASELINE POLS MODEL
ols_sdg_spi_reg <- plm(formula = sdg_overall ~ spi_comp + di_score + log_gdppc + factor(year),
               model = "pooling", 
               index = c("country_code", "year"),
               data = panel_data)

# Assess multi-colinearity 
vif(ols_sdg_spi_reg)

# CENTERED POLS MODEL
ols_sdg_spi_cen <- plm(formula = sdg_overall ~ cen_spi_comp + cen_di_score + cen_log_gdppc + factor(year),
               model = "pooling", 
               index = c("country_code", "year"),
               data = panel_data)

# Assess multi-colinearity 
vif(ols_sdg_spi_cen)
```




# Partial Correlation Matrix: Disaggregated SDG ~ SCI Scores 

```{r cars}
# set working directory 
setwd("~/Documents/GitHub/QMSS_Thesis_Sanchez")

#load libraries/packages
source("packages.R")

#load function 
source("df_years2.0_Function.R")
matrix_data <- df_years2.0(2015, 2023) %>% 
   dplyr::select(p1_use, p2_services, p3_products, p4_sources, p5_infra, spi_comp,
                goal1, goal2, goal3, goal4, goal5, goal6, goal7, goal8,
                goal9, goal10, goal11, goal12, goal13, goal14, goal15,
                goal16, goal17, sdg_overall, log_gdppc, population)

#Impute missing data: log_gdppc & population 
matrix_data <- matrix_data %>%
  mutate(
    log_gdppc = ifelse(is.na(log_gdppc), median(log_gdppc, na.rm = TRUE), log_gdppc),
    population = ifelse(is.na(population), median(population, na.rm = TRUE), population)
  )
```

#Partial correlation & package 
Partial correlation controls for confounding variables: Ensures correlations between SPI pillars and SDG goals are adjusted for external factors like log(GDP) and population.

Focuses on Direct Relationships: Removes indirect effects of controls on both SPI and SDG metrics.

# Initialize an empty matrix to store results
```{r}
spi_pillars <- c("spi_comp", "p1_use", "p2_services", "p3_products", "p4_sources", "p5_infra")
sdg_goals <- c("sdg_overall", paste0("goal", 1:17))

#devise matrix 
partial_corr_matrix <- matrix(NA, nrow = length(spi_pillars), 
                              ncol = length(sdg_goals),
                              dimnames = list(spi_pillars, sdg_goals))

for (spi in spi_pillars) {
  for (sdg in sdg_goals) {
    # Subset data with controls
    temp_data <- matrix_data[, c(spi, sdg, "log_gdppc", "population")] %>%
      na.omit()
    
    # Skip if insufficient data
    if (nrow(temp_data) < 10) {
      partial_corr_matrix[spi, sdg] <- NA
      next
    }
    
    # Check for near-constant variables (threshold)
    var_check <- sapply(temp_data[, c("log_gdppc", "population")], var)
    if (any(var_check < 0.01)) {  # Increased from 1e-3 to 0.01
      partial_corr_matrix[spi, sdg] <- NA
      next
    }
    
    # Attempt 1: Standard partial correlation
    tryCatch({
      result <- ppcor::pcor(temp_data)
      partial_corr_matrix[spi, sdg] <- result$estimate[1, 2]
    }, error = function(e) {
      # Attempt 2: Pseudoinverse-based correlation (more robust)
      tryCatch({
        cov_mat <- cov(temp_data)
        pcor_mat <- corpcor::pcor.shrink(temp_data, verbose = FALSE)
        partial_corr_matrix[spi, sdg] <- pcor_mat[1, 2]
      }, error = function(e) {
        # Fallback: Simple correlation without controls
        cor_value <- cor(temp_data[, spi], temp_data[, sdg])
        partial_corr_matrix[spi, sdg] <- cor_value
      })
    })
  }
}

# Convert to data frame
matrix_df <- as.data.frame(partial_corr_matrix)

# Reshape the matrix into long format
melted_cormat <- melt(partial_corr_matrix)
```

#visualizing partial correlation matrix
```{r}
matrix_spi_sdg <- ggplot(data = melted_cormat, aes(x = Var2, y = Var1, fill = value)) +
  geom_tile(color = "white") +
  scale_fill_gradient2(
    low = "red", mid = "white", high = "darkgreen", midpoint = 0,
    limits = c(-1, 1), name = "Correlation"
  ) +
  # Grey tiles for NA values (plot FIRST to avoid covering text)
  geom_tile(
    data = subset(melted_cormat, is.na(value)),
    aes(x = Var2, y = Var1), fill = "grey90", color = "white"
  ) +
  # Text with dynamic color based on |value|
  geom_text(
    aes(
      label = ifelse(is.na(value), "", round(value, 2)),
      color = abs(value) > 0.5  # White text for |cor| > 0.5
    ),
    size = 2.5
  ) +
  scale_color_manual(
    guide = "none",  # Hide the color legend
    values = c("black", "white")  # FALSE = black, TRUE = white
  ) +
  theme_minimal() +
  labs(
    title = "Partial Correlation Matrix: SPI Pillars vs SDG Goals",
    x = "SDG Goals", 
    y = "SPI Pillars"
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 9),
    axis.text.y = element_text(size = 9),
    legend.position = "right"
  ) +
  coord_fixed()  # Maintain square aspect ratio

#show
matrix_spi_sdg
```

