---
title: "Descriptive Stats: Variables and Datasets"
output: html_document
---

## ADJUST/WRANGLE DATA FIRST IN 'Comp1_CrossSec_OLS.Rmd' AS NEEDED

# Load Data/packages: merged_final_df.csv
```{r}
source("packages.R")

#load final merged df 
merged <- read_csv("data/Main CSV Outputs/merged_final_df.csv")
```

# NEW DF: counting the number of NAs (across all years) per country, by variables: SPI, SCI, SDG
```{r eval=FALSE, include=FALSE}
# List of columns to count NAs in
target_cols <- c("spi_comp", "sdg_overall", "sci_overall", "p1_use", "p2_services", 
                 "p3_products", "p4_sources", "p5_infra", paste0("goal", 1:17))

# Create new dataframe with NA counts per country
merged_na_counts <- merged %>%
  group_by(country_code) %>%
  summarise(across(
    .cols = all_of(target_cols),
    .fns = ~ sum(is.na(.x)),
    .names = "{.col}_na_count"
  )) %>%
  ungroup()

#saving merged_na_counts as CSV output 
#write.table(merged_na_counts, file = 'merged_na_counts.csv', row.names=F, sep = ",")

#load CSV 
merged_na_counts <- read_csv("data/Main CSV Outputs/merged_na_counts.csv")
```

# Descriptive stats & Missing Data: Full Dataset 
```{r}
# x & y variables 
df_sdg_statcap <- data.frame(merged$sdg_overall, merged$spi_comp, merged$sci_overall, merged$di_score, merged$log_gdppc, merged$population, merged$gini)

#N total observations: country-years
colSums(!is.na(df_sdg_statcap))
# N total NAs
colSums(is.na(df_sdg_statcap))

# N countries per variable? 
sapply(c("sdg_overall", "di_score", "spi_comp", "sci_overall", "log_gdppc", "population", "gini"), function(var) {
  length(unique(merged$country_code[!is.na(merged[[var]])]))
})

# N years per variable? 
sapply(c("sdg_overall", "di_score", "spi_comp", "sci_overall", "log_gdppc", "population", "gini"), function(var) {
  length(unique(merged$year[!is.na(merged[[var]])]))
})
```
**Years 2004-2023:**

Number of observations per variable:
sdg_overall  spi_comp sci_overall  di_score  log_gdppc  population  gini
   3360        1300      2072       2560       3333        3106     1404

Number of NAs per variable:
sdg_overall spi_comp  sci_overall  di_score  log_gdppc  population  gini
     0         2060      1288        800        27         254      1956
        
Number of countries per variable:
sdg_overall  spi_comp  sci_overall di_score log_gdppc   population  gini 
    168        165         123        160      168        164       151 

Number of years per variable:
sdg_overall  spi_comp  sci_overall di_score log_gdppc  population   gini 
     20         8          17        16        20          19        20 

# Histograms 
```{r}
# SPI & SCI HISTOGRAMS: side-by-side layout
png("figures/spi_sci_hist.png", width = 9, height = 5, unit = "in", res = 300) #SPI & SCI Side By Side 
par(mfrow = c(1, 2))

# SPI histogram
hist(merged$spi_comp, main = "Statistical Performance Index (2016-2023)", 
     xlab = "Composite Score (0-100)", 
     ylab = "Frequency (country-years)")

# SCI histogram
hist(merged$sci_overall, main = "Statistical Capacity Index (2004-2020)", 
     xlab = "Composite Score (0-100)", 
     ylab = "Frequency (country-years)")

par(mfrow = c(1, 1)) #side by side 
dev.off()

# SDG histogram 
png("figures/sdg_hist.png", width = 5, height = 4, unit = "in", res = 300) #Saving to figures folder 
sdg_hist <- hist(merged$sdg_overall, main = "Sustainable Development Goals (2004-2023)", 
                 xlab = "Composite Score (0-100)", 
                 ylab = "Frequency (country-years)")
dev.off()
```

*SDG VARIABLE NOTE: While all member states have a country profile, only those with less than 20% missing data are ranked in the SDG Index, resulting in 167 countries being ranked in the 2024 report.*

**Reason to remove GINI from study:**
while it is an important consideration, my study uses country-years as my primary unit of analysis, atleast in the corss-sectional component. Not all data points are available for every country, let alone every year. I suspect that this missingness is non-random, as not every country has the available statistical infrastructure to calculate a reliable estimate. With almost 2000 missing country-year values, including the GINI coefficient as a control would dramatically decrease my sample size and produce unreliable results. It is not good practice to impute list-wise values given the non-random nature of the missing values and especially to this extent.  

**Reason to remove Population from models**
The models will not be including total national population although is theoretically important for my research as it likely contributes the nations ability to capture high quality statistics on the population as surly this can challenge current structures in place that are not equiped to track a growing population. The reason is because doing so would introduce colinearity with national (logged) Gross Domestic Product per capita (population), which would bias results.

## Box plots
```{r}
#SPI & SCI BOX PLOTS: SIDE BY SIDE 
png("figures/spi_sci_boxplots.png", width = 9, height = 5, units = "in", res = 300)
par(mfrow = c(1, 2))

# SPI Distribution by regime type 
boxplot(spi_comp ~ regime_type_2, data = merged,
        main = "SPI Distribution by Regime Type",
        xlab = "Regime Type",
        ylab = "Composite Score (1-100)",
        col = c("salmon", "skyblue"))

legend("bottomright",                      # Position of the legend
       legend = c("Democracy", "Autocracy"),  # Group names
       fill = c("skyblue", "salmon"),        # Colors matching the boxplot
       border = "black",                     # Box border color (optional)
       bty = "n")   


# SCI Distribution by regime type (only considers lower income / developing countries)
boxplot(sci_overall ~ regime_type_2, data = merged,
        main = "SCI Distribution by Regime Type",
        xlab = "Regime Type",
        ylab = "Composite Score (1-100)",
        col = c("salmon", "skyblue"))

legend("bottomright",                      # Position of the legend
       legend = c("Democracy", "Autocracy"),  # Group names
       fill = c("skyblue", "salmon"),        # Colors matching the boxplot
       border = "black",                     # Box border color (optional)
       bty = "n")   

par(mfrow = c(1, 1)) #side by side 
dev.off()

# SDG Box plot 
png("figures/sdg_boxplots.png", width = 5, height = 4, units = "in", res = 300)
boxplot(sdg_overall ~ regime_type_2, data = merged,
        main = "SDG Distribution by Regime Type",
        xlab = "Regime Type",
        ylab = "Composite Score (1-100)",
        col = c("salmon", "skyblue"))

legend("bottomright",                      # Position of the legend
       legend = c("Democracy", "Autocracy"),  # Group names
       fill = c("skyblue", "salmon"),        # Colors matching the boxplot
       border = "black",                     # Box border color (optional)
       bty = "n")   

dev.off()
```

##Summary Statistics and Quantile Frequency Tables 
```{r}
#New Vars - Quantile breaks 
quant_breaks_sdg <- quantile(merged$sdg_overall, probs = c(0, 0.25, 0.5, 0.75, 1), na.rm = TRUE)
quant_breaks_spi <- quantile(merged$spi_comp, probs = c(0, 0.25, 0.5, 0.75, 1), na.rm = TRUE)
quant_breaks_sci <- quantile(merged$sci_overall, probs = c(0, 0.25, 0.5, 0.75, 1), na.rm = TRUE)

# Inputting SDG quantiles in main DF 
merged$sdg_quant_bins <- cut(merged$sdg_overall, 
                      breaks = quant_breaks_sdg, 
                      include.lowest = TRUE, 
                      labels = c("Q1 (Lowest)", "Q2", "Q3", "Q4 (Highest)"))

#SDG IQR Summary Table 
sdg_summary <- merged %>%
  group_by(regime_type_2, sdg_quant_bins) %>%
  summarise(count = n(), .groups = "drop")
# Pivot to wide format
sdg_wide_table <- sdg_summary %>%
  pivot_wider(names_from = sdg_quant_bins, values_from = count, values_fill = 0)

# Inputting SPI quartiles in main DF 
merged$spi_quant_bins <- cut(merged$spi_comp, 
                      breaks = quant_breaks_spi, 
                      include.lowest = TRUE, 
                      labels = c("Q1 (Lowest)", "Q2", "Q3", "Q4 (Highest)"))

#SPI IQR Summary Table 
spi_summary <- merged %>%
  group_by(regime_type_2, spi_quant_bins) %>%
  summarise(count = n(), .groups = "drop")
# Pivot to wide format
spi_wide_table <- spi_summary %>%
  pivot_wider(names_from = spi_quant_bins, values_from = count, values_fill = 0)

# Inputting SCI quartiles in main DF 
merged$sci_quant_bins <- cut(merged$sci_overall, 
                      breaks = quant_breaks_sci, 
                      include.lowest = TRUE, 
                      labels = c("Q1 (Lowest)", "Q2", "Q3", "Q4 (Highest)"))

# SCI IQR Summary Table 
sci_summary <- merged %>%
  group_by(regime_type_2, sci_quant_bins) %>%
  summarise(count = n(), .groups = "drop")
# Pivot to wide format
sci_wide_table <- sci_summary %>%
  pivot_wider(names_from = sci_quant_bins, values_from = count, values_fill = 0)

#Summary tables by score: Min, 1st Qu. Median, Mean, 3rd Qu., Max, NA's 
summary(merged$sdg_overall)
summary(merged$spi_comp)
summary(merged$sci_overall)
summary(merged$di_score)

#write tables as CSV to directory 


# STACKED BAR PLOTS LATER 

```

summary(merged$sdg_overall)
   Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
  35.65   55.74   65.53   64.43   73.05   86.42 

summary(merged$spi_comp)
   Min. 1st Qu.  Median    Mean 3rd Qu.    Max.    NA's 
  11.77   54.82   67.09   66.71   81.48   95.26    2060 

summary(merged$sci_overall)
   Min. 1st Qu.  Median    Mean 3rd Qu.    Max.    NA's 
  16.67   57.78   70.00   67.96   80.00   98.89    1288 

summary(merged$di_score)
   Min. 1st Qu.  Median    Mean 3rd Qu.    Max.    NA's 
  0.260   3.520   5.780   5.509   7.280   9.930     800 
