---
title: "Descriptive Statistics: Variables and Datasets"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: paper
    highlight: tango
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE,
                     fig.width = 10, fig.height = 6)
```

# 1. Load Data and Packages

```{r}
# set working directory 
setwd("~/Documents/GitHub/QMSS_Thesis_Sanchez")

# Load all packages from the packages.R file
source("packages.R")

# Load the merged (filtered, final) and merged_full (unfiltered, full) datasets
merged_full <- read_csv("data/Main CSV Outputs/merged_full_df.csv") # full unfiltered dataset
merged <- read_csv("data/Main CSV Outputs/merged_final_df.csv") # filtered dataset

# Define key variables for analysis
key_vars <- c("sdg_overall", "spi_comp", "sci_overall", "di_score", 
              "log_gdppc", "gdp_pc")
group_vars <- c("income_level", "regime_type_2", "regime_type_4")
```

*SDG VARIABLE NOTE: While all member states have a country profile, only those with less than 20% missing data are ranked in the SDG Index, resulting in 167 countries being ranked in the 2024 report.*

**Reason to remove GINI from study:**
This study uses country-years as my primary unit of analysis, at least in the cross-sectional segment of component 1. While inequalities in the distribution of wealth for a given country is an important consideration that may affect statistical capacity and sustainable development (source), not data points are available for every country, let alone every year, to produce a reliable estimate. I suspect that this missingness is non-random, as a significant share of countries, predominantly of emerging markets in the global south, face greater challenges in developing and consolidating the statistical infrastructure necessary for producing quality statistics (SOURCE). With almost 2000 missing country-year values, including the GINI coefficient as a control would dramatically decrease the sample size and yield unreliable results. To reiterate, this analysis refrains from using list-wise imputation, especially for variables like the GINI coefficient that are not available for all countries and years. Instead, the analysis focused on the available data, which is sufficient to draw meaningful conclusions about the relationship between statistical capacity and sustainable development.

**Reason to remove Population from models**
The models will not be including total national population although is theoretically important for my research as it likely contributes the nations ability to capture high quality statistics on the population as surly this can challenge current structures in place that are not equipped to track a growing population (source). The reason is because doing so would introduce colinearity with national (logged) Gross Domestic Product per capita (population), which would bias results.

# 2. Dataset Coverage

## 2.1 Overall Dataset Size

```{r}
# Basic dimensions of the dataset
cat("Total number of observations (rows):", nrow(merged_full), "\n")
cat("Total number of variables (columns):", ncol(merged_full), "\n")
cat("Year range:", min(merged_full$year, na.rm=TRUE), "to", max(merged_full$year, na.rm=TRUE), "\n")
cat("Total number of countries:", length(unique(merged_full$country_code)), "\n")
```

## 2.2 Coverage by Variable (full dataset)

```{r}
# Function to get coverage statistics for a variable
get_coverage <- function(data, var_name) {
  total_obs <- sum(!is.na(data[[var_name]]))
  total_countries <- length(unique(data$country_code[!is.na(data[[var_name]])]))
  total_years <- length(unique(data$year[!is.na(data[[var_name]])]))
  missing_obs <- sum(is.na(data[[var_name]]))
  missing_percent <- round(missing_obs / nrow(data) * 100, 2)
  
  return(data.frame(
    Variable = var_name,
    Observations = total_obs,
    Countries = total_countries,
    Years = total_years,
    Missing = missing_obs,
    Missing_Percent = missing_percent
  ))
}

# Get coverage for all key variables
coverage_stats <- bind_rows(lapply(key_vars, function(var) get_coverage(merged_full, var)))

# Create the table with APA styling and no caption number
coverage_table <- kable(coverage_stats, 
      caption = "Coverage Statistics for Key Variables (unfiltered)",
      col.names = c("Variable", "Observations", "Countries", "Years", "Missing Values", "Missing (%)"),
      align = c('l', rep('c', 5)),
      booktabs = TRUE) %>%
  kable_classic(full_width = FALSE, html_font = "Times New Roman", font_size = 12, "basic") %>%
  row_spec(0, bold = FALSE, align = "c") %>%
  column_spec(1, bold = FALSE)

# Save the table as HTML
save_kable(coverage_table, file = "figures/data coverage & summary stats/full_coverage_stats_table.html")

# Display the table in the document
coverage_table
```

## 2.3 Coverage by Year (full dataset)

```{r}
# Create a table showing variable coverage by year
year_coverage <- merged_full %>%
  group_by(year) %>%
  summarise(across(all_of(key_vars), ~sum(!is.na(.)), .names = "{.col}")) %>%
  arrange(year)

# Create the table with APA styling and no caption number
year_coverage_table <- kable(year_coverage, 
      caption = "Number of Countries Covered by Year for Key Variables (unfiltered)",
      booktabs = TRUE,
      align = c('l', rep('c', length(key_vars)))) %>%
  kable_classic(full_width = FALSE, html_font = "Times New Roman", font_size = 12, "basic") %>%
  row_spec(0, bold = FALSE, align = "c") %>%
  column_spec(1, bold = FALSE)

# Save the table as HTML
save_kable(year_coverage_table, file = "figures/data coverage & summary stats/full_year_coverage_table.html")

# Display the table in the document
year_coverage_table
```

## 2.4 Coverage by Variable (final dataset)

```{r}
# Get coverage for all key variables
coverage_stats <- bind_rows(lapply(key_vars, function(var) get_coverage(merged, var)))

# Create the table with APA styling and no caption number
coverage_table <- kable(coverage_stats, 
      caption = "Coverage Statistics for Key Variables (filtered)",
      col.names = c("Variable", "Observations", "Countries", "Years", "Missing Values", "Missing (%)"),
      align = c('l', rep('c', 5)),
      booktabs = TRUE) %>%
  kable_classic(full_width = FALSE, html_font = "Times New Roman", font_size = 12, "basic") %>%
  row_spec(0, bold = FALSE, align = "c") %>%
  column_spec(1, bold = FALSE)

# Save the table as HTML
save_kable(coverage_table, file = "figures/data coverage & summary stats/final_coverage_stats_table.html")

# Display the table in the document
coverage_table
```

## 2.5 Coverage by Year (final dataset)

```{r}
# Create a table showing variable coverage by year
year_coverage <- merged %>%
  group_by(year) %>%
  summarise(across(all_of(key_vars), ~sum(!is.na(.)), .names = "{.col}")) %>%
  arrange(year)

# Create the table with APA styling and no caption number
year_coverage_table <- kable(year_coverage, 
      caption = "Number of Countries Covered by Year for Key Variables (filtered)",
      booktabs = TRUE,
      align = c('l', rep('c', length(key_vars)))) %>%
  kable_classic(full_width = FALSE, html_font = "Times New Roman", font_size = 12, "basic") %>%
  row_spec(0, bold = FALSE, align = "c") %>%
  column_spec(1, bold = FALSE)

# Save the table as HTML
save_kable(year_coverage_table, file = "figures/data coverage & summary stats/final_year_coverage_table.html")

# Display the table in the document
year_coverage_table
```

# 3. Descriptive Statistics (Filtered Data)

## 3.1 Summary Statistics for Key Variables (All and subgrouped)
*Building function for sub-grouping in the future*
```{r}
# Define and list variables in desired order
key_vars <- c("sdg_overall", "spi_comp", "sci_overall", "di_score", "gdp_pc", "log_gdppc")

# Apply subgroup filter for GNI level
gni_filter <- "all" # specify desired GNI level (e.g., LIC, LMIC, UMIC, HIC, all)

# Apply subgroup filter level for regime type 
regime_filter <- "democracy" # specify desired regime type (e.g., all, Democracy, Autocracy)

# Generate summary statistics for key variables
merged_stats <- merged %>%
  filter(
    case_when(
      gni_filter == "HIC" ~ income_level == 'H', 
      gni_filter == "UMIC" ~ income_level == 'UM',
      gni_filter == "LMIC" ~ income_level == 'LM',
      gni_filter == "LIC" ~ income_level == 'L',
      gni_filter == "all" ~ TRUE,          # Keep all observations
      TRUE ~ FALSE                         # Exclude invalid filter options
    )
  ) %>% 
  filter(
    case_when(
      regime_filter == "autocracy" ~ regime_type_2 == 0,
      regime_filter == "democracy" ~ regime_type_2 == 1,
      regime_filter == "all" ~ TRUE,       # Keep all observations
      TRUE ~ FALSE # filter all else 
    )
  ) %>%
  select(all_of(key_vars)) %>%
  summarize(across(everything(), 
                   list(
                     min = ~min(., na.rm = TRUE),
                     q1 = ~quantile(., 0.25, na.rm = TRUE),
                     median = ~median(., na.rm = TRUE),
                     mean = ~mean(., na.rm = TRUE),
                     sd = ~sd(., na.rm = TRUE),
                     q3 = ~quantile(., 0.75, na.rm = TRUE),
                     max = ~max(., na.rm = TRUE),
                     n = ~sum(!is.na(.))
                     )))

# Reshape the data for better presentation
merged_stats_long <- merged_stats %>%
  pivot_longer(cols = everything(),
               names_to = c("variable", "stat"),
               names_pattern = "(.*)_(.*)",
               values_to = "value") %>%
  pivot_wider(names_from = stat, values_from = value) %>%
  arrange(match(variable, key_vars))

# Create the table with APA styling and no caption number
summary_stats_table <- kable(merged_stats_long, 
      caption = paste("Descriptive Statistics for Key Variables,", "Subgrouped to:", if(gni_filter == "all"){"All GNI Levels"} else {gni_filter}, "and", if(regime_filter == "all"){"All Regime Types"} else {regime_filter}),
      col.names = c("Variable", "Min", "1st Qu", "Median", "Mean", "SD", "3rd Qu", "Max", "N"),
      booktabs = TRUE,
      digits = 2,
      align = c('l', rep('c', 8))) %>%
  kable_classic(full_width = FALSE, html_font = "Times New Roman", font_size = 12) %>%
  row_spec(0, bold = FALSE, align = "c") %>%
  column_spec(1, bold = FALSE)

# Save the table as HTML
save_kable(summary_stats_table, 
file = paste("figures/data coverage & summary stats/summary", if(gni_filter == "all"){"_all_gni"} else {paste0("_", gni_filter)}, if(regime_filter == "all"){"_all_regime"} else {paste0("_", regime_filter)}, sep = "", ".html"))

# Display the table in the document
summary_stats_table
```

# 4. Visualizations

## 4.1 Distributions of Key Variables

```{r fig.height=8}
# Set line widths for mean and median lines
med_linewidth <- 0.6
mean_linewidth <- 0.5

# Create histograms for key index variables using ggplot2
p1 <- ggplot(merged, aes(x = spi_comp)) +
  geom_histogram(bins = 20, fill = "lightblue", color = "white", alpha = 0.7) +
  geom_vline(xintercept = median(merged$spi_comp, na.rm = TRUE), 
             color = "red", linewidth = med_linewidth, linetype = "dashed") +
  geom_vline(xintercept = mean(merged$spi_comp, na.rm = TRUE), 
             color = "black", linewidth = mean_linewidth) +
  labs(title = "Statistical Performance Index (2016-2023)",
       x = "Composite Score (0-100)",
       y = "Frequency (country-years)") +
  theme_minimal()
  
p2 <- ggplot(merged, aes(x = sci_overall)) +
  geom_histogram(bins = 20, fill = "lightgrey", color = "white", alpha = 0.7) +
  geom_vline(xintercept = median(merged$sci_overall, na.rm = TRUE), 
             color = "red", linewidth = med_linewidth, linetype = "dashed") +
  geom_vline(xintercept = mean(merged$sci_overall, na.rm = TRUE), 
             color = "black", linewidth = mean_linewidth) +
  labs(title = "Statistical Capacity Index (2004-2020)",
       x = "Composite Score (0-100)",
       y = "Frequency (country-years)") +
  theme_minimal()

p3 <- ggplot(merged, aes(x = di_score)) +
  geom_histogram(bins = 20, fill = "lightpink", color = "white", alpha = 0.7) +
  geom_vline(xintercept = median(merged$di_score, na.rm = TRUE), 
             color = "red", linewidth = med_linewidth, linetype = "dashed") +
  geom_vline(xintercept = mean(merged$di_score, na.rm = TRUE), 
             color = "black", linewidth = mean_linewidth) +
  labs(title = "Democracy Index (2006-2023)",
       x = "Score (0-10)",
       y = "Frequency (country-years)") +
  theme_minimal()

p4 <- ggplot(merged, aes(x = sdg_overall)) +
  geom_histogram(bins = 20, fill = "lightgreen", color = "white", alpha = 0.7) +
  geom_vline(xintercept = median(merged$sdg_overall, na.rm = TRUE), 
             color = "red", linewidth = med_linewidth, linetype = "dashed") +
  geom_vline(xintercept = mean(merged$sdg_overall, na.rm = TRUE), 
             color = "black", linewidth = mean_linewidth) +
  labs(title = "SDG Index (2004-2023)",
       x = "Composite Score (0-100)",
       y = "Frequency (country-years)") +
  theme_minimal()

# Combine plots using patchwork
(p1 + p2) / (p3 + p4)

# Save combined plot
ggsave("figures/data coverage & summary stats/hist_combined.png", (p1 + p2) / (p3 + p4), width = 10, height = 8, dpi = 300)
```

### GDP & Log(GDP) Distributions
*On-hand if needed*
```{r}
p5 <- ggplot(merged, aes(x = gdp_pc)) +
  geom_histogram(bins = 20, fill = "darkgreen", color = "white", alpha = 0.7) +
  geom_vline(xintercept = median(merged$di_score, na.rm = TRUE), 
             color = "red", linewidth = med_linewidth, linetype = "dashed") +
  geom_vline(xintercept = mean(merged$di_score, na.rm = TRUE), 
             color = "black", linewidth = mean_linewidth) +
  labs(title = "GDP Per Capita (2006-2023)",
       x = "Pre-transformed (USD)",
       y = "Frequency (country-years)") +
  theme_minimal()

p6 <- ggplot(merged, aes(x = log_gdppc)) +
  geom_histogram(bins = 20, fill = "darkgrey", color = "white", alpha = 0.7) +
  geom_vline(xintercept = median(merged$sdg_overall, na.rm = TRUE), 
             color = "red", linewidth = med_linewidth, linetype = "dashed") +
  geom_vline(xintercept = mean(merged$sdg_overall, na.rm = TRUE), 
             color = "black", linewidth = mean_linewidth) +
  labs(title = "Log of GDP Per Capita (2004-2023)",
       x = "Transformed (log(USD))",
       y = "Frequency (country-years)") +
  theme_minimal()

# Combine GDP plots using patchwork
gdp_combined <- (p5 + p6) 
#+ plot_layout(ncol = 2, guides = "collect") 
#+ plot_annotation(title = "GDP Per Capita Distributions", theme = theme(plot.title = element_text(hjust = 0.5)))
```


## 4.2 Boxplots by Regime Type

```{r fig.height=8}
# Create boxplots for variables split by regime type using ggplot2

# Add factor for regime type with proper labels
merged <- merged %>%
  mutate(regime_type_factor = factor(regime_type_2, 
                                    levels = c(0, 1),
                                    labels = c("Autocracy", "Democracy"))) %>% 
  # filter out NAs for regime type 
  filter(!is.na(regime_type_factor))

b1 <- ggplot(merged, aes(x = regime_type_factor, y = spi_comp, fill = regime_type_factor)) +
  geom_boxplot() +
  scale_fill_manual(values = c("salmon", "skyblue")) +
  labs(title = "SPI Distribution by Regime Type",
       x = NULL,
       y = "Composite Score (0-100)") +
  theme_minimal() +
  theme(legend.position = "none")

b2 <- ggplot(merged, aes(x = regime_type_factor, y = sci_overall, fill = regime_type_factor)) +
  geom_boxplot() +
  scale_fill_manual(values = c("salmon", "skyblue")) +
  labs(title = "SCI Distribution by Regime Type",
       x = NULL,
       y = "Composite Score (0-100)") +
  theme_minimal() +
  theme(legend.position = "none")

b3 <- ggplot(merged, aes(x = regime_type_factor, y = log_gdppc, fill = regime_type_factor)) +
  geom_boxplot() +
  scale_fill_manual(values = c("salmon", "skyblue")) +
  labs(title = "Log of GDP Per Capita by Regime Type",
       x = NULL,
       y = "Transformed Log(USD)") +
  theme_minimal() +
  theme(legend.position = "none")

b4 <- ggplot(merged, aes(x = regime_type_factor, y = sdg_overall, fill = regime_type_factor)) +
  geom_boxplot() +
  scale_fill_manual(values = c("salmon", "skyblue")) +
  labs(title = "SDG Index by Regime Type",
       x = NULL,
       y = "Composite Score (0-100)") +
  theme_minimal() +
  theme(legend.position = "none")

# Combine plots using patchwork
boxplot_combined <- (b1 + b2) / (b3 + b4)
boxplot_combined

# Save combined plot
ggsave("figures/data coverage & summary stats/boxplots_combined.png", boxplot_combined, width = 10, height = 8, dpi = 300)
```

# 5. Quantile Analysis # PICK UP HERE AND DOWN 

## 5.1 Quartile Breaks
*Just Quantiles, on-hand if needed*
```{r}
# Calculate quartile breaks for main indices
quant_breaks_sdg <- quantile(merged$sdg_overall, probs = c(0, 0.25, 0.5, 0.75, 1), na.rm = TRUE)
quant_breaks_spi <- quantile(merged$spi_comp, probs = c(0, 0.25, 0.5, 0.75, 1), na.rm = TRUE)
quant_breaks_sci <- quantile(merged$sci_overall, probs = c(0, 0.25, 0.5, 0.75, 1), na.rm = TRUE)
quant_breaks_di <- quantile(merged$di_score, probs = c(0, 0.25, 0.5, 0.75, 1), na.rm = TRUE)

# Display the quartile breaks
quartiles_df <- data.frame(
  Quartile = c("Minimum (0%)", "Q1 (25%)", "Median (50%)", "Q3 (75%)", "Maximum (100%)"),
  SDG_Index = quant_breaks_sdg,
  SPI = quant_breaks_spi,
  SCI = quant_breaks_sci,
  Democracy_Index = quant_breaks_di
)

# Create the table with APA styling and no caption number
quartiles_table <- kable(quartiles_df, 
      caption = "Quartile Breaks for Key Indices",
      booktabs = TRUE,
      digits = 2,
      align = c('l', rep('c', 4))) %>%
  kable_classic(full_width = FALSE, html_font = "Times New Roman", font_size = 12, "basic") %>%
  row_spec(0, bold = FALSE, align = "c") %>%
  column_spec(1, bold = FALSE)

# Save the table as HTML
#save_kable(quartiles_table, file = "figures/data coverage & summary stats/quartiles_table.html")

# Display the table in the document
quartiles_table
```

## 5.2 Quantile Frequency Tables

```{r}
# Create quartile bins for each index
merged <- merged %>%
  mutate(
    sdg_quant_bins = cut(sdg_overall, 
                      breaks = quant_breaks_sdg, 
                      include.lowest = TRUE, 
                      labels = c("Q1 (Lowest)", "Q2", "Q3", "Q4 (Highest)")),
    spi_quant_bins = cut(spi_comp, 
                      breaks = quant_breaks_spi, 
                      include.lowest = TRUE, 
                      labels = c("Q1 (Lowest)", "Q2", "Q3", "Q4 (Highest)")),
    sci_quant_bins = cut(sci_overall, 
                      breaks = quant_breaks_sci, 
                      include.lowest = TRUE, 
                      labels = c("Q1 (Lowest)", "Q2", "Q3", "Q4 (Highest)")),
    di_quant_bins = cut(di_score, 
                      breaks = quant_breaks_di, 
                      include.lowest = TRUE, 
                      labels = c("Q1 (Lowest)", "Q2", "Q3", "Q4 (Highest)"))
  )

# Create frequency tables by regime type for each index
# SDG Index
sdg_summary <- merged %>%
  filter(!is.na(sdg_quant_bins) & !is.na(regime_type_2)) %>%
  group_by(regime_type_2, sdg_quant_bins) %>%
  summarise(count = n(), .groups = "drop") %>%
  mutate(regime_type_name = ifelse(regime_type_2 == 1, "Democracy", "Autocracy"))

# Pivot to wide format
sdg_wide_table <- sdg_summary %>%
  select(-regime_type_2) %>%
  pivot_wider(names_from = sdg_quant_bins, values_from = count, values_fill = 0)

# Create the table with APA styling and no caption number
sdg_quartile_table <- kable(sdg_wide_table, 
      caption = "SDG Index Quartile Distribution by Regime Type",
      booktabs = TRUE,
      align = c('l', rep('c', 4))) %>%
  kable_classic(full_width = FALSE, html_font = "Times New Roman", font_size = 12, "basic") %>%
  row_spec(0, bold = FALSE, align = "c") %>%
  column_spec(1, bold = FALSE)

# Save the table as HTML
#save_kable(sdg_quartile_table, file = "figures/data coverage & summary stats/sdg_quartile_table.html")

# Display the table in the document
sdg_quartile_table

# SPI Index
spi_summary <- merged %>%
  filter(!is.na(spi_quant_bins) & !is.na(regime_type_2)) %>%
  group_by(regime_type_2, spi_quant_bins) %>%
  summarise(count = n(), .groups = "drop") %>%
  mutate(regime_type_name = ifelse(regime_type_2 == 1, "Democracy", "Autocracy"))

# Pivot to wide format
spi_wide_table <- spi_summary %>%
  select(-regime_type_2) %>%
  pivot_wider(names_from = spi_quant_bins, values_from = count, values_fill = 0)

# Create the table with APA styling and no caption number
spi_quartile_table <- kable(spi_wide_table, 
      caption = "SPI Quartile Distribution by Regime Type",
      booktabs = TRUE,
      align = c('l', rep('c', 4))) %>%
  kable_classic(full_width = FALSE, html_font = "Times New Roman", font_size = 12, "basic") %>%
  row_spec(0, bold = FALSE, align = "c") %>%
  column_spec(1, bold = FALSE)

# Save the table as HTML
#save_kable(spi_quartile_table, file = "figures/data coverage & summary stats/spi_quartile_table.html")

# Display the table in the document
spi_quartile_table

# SCI Index
sci_summary <- merged %>%
  filter(!is.na(sci_quant_bins) & !is.na(regime_type_2)) %>%
  group_by(regime_type_2, sci_quant_bins) %>%
  summarise(count = n(), .groups = "drop") %>%
  mutate(regime_type_name = ifelse(regime_type_2 == 1, "Democracy", "Autocracy"))

# Pivot to wide format
sci_wide_table <- sci_summary %>%
  select(-regime_type_2) %>%
  pivot_wider(names_from = sci_quant_bins, values_from = count, values_fill = 0)

# Create the table with APA styling and no caption number
sci_quartile_table <- kable(sci_wide_table, 
      caption = "SCI Quartile Distribution by Regime Type",
      booktabs = TRUE,
      align = c('l', rep('c', 4))) %>%
  kable_classic(full_width = FALSE, html_font = "Times New Roman", font_size = 12, "basic") %>%
  row_spec(0, bold = FALSE, align = "c") %>%
  column_spec(1, bold = FALSE)

# Save the table as HTML
#save_kable(sci_quartile_table, file = "figures/data coverage & summary stats/sci_quartile_table.html")

# Display the table in the document
sci_quartile_table
```

# 6. Summary Table of Findings

```{r}
# Create a summary table of key findings
summary_findings <- data.frame(
  Variable = c("SDG Overall", "SPI Composite", "SCI Overall", "Democracy Index"),
  Democracy_Mean = c(
    mean(merged$sdg_overall[merged$regime_type_2 == 1], na.rm = TRUE),
    mean(merged$spi_comp[merged$regime_type_2 == 1], na.rm = TRUE),
    mean(merged$sci_overall[merged$regime_type_2 == 1], na.rm = TRUE),
    mean(merged$di_score[merged$regime_type_2 == 1], na.rm = TRUE)
  ),
  Autocracy_Mean = c(
    mean(merged$sdg_overall[merged$regime_type_2 == 0], na.rm = TRUE),
    mean(merged$spi_comp[merged$regime_type_2 == 0], na.rm = TRUE),
    mean(merged$sci_overall[merged$regime_type_2 == 0], na.rm = TRUE),
    mean(merged$di_score[merged$regime_type_2 == 0], na.rm = TRUE)
  ),
  Difference = c(
    mean(merged$sdg_overall[merged$regime_type_2 == 1], na.rm = TRUE) - 
      mean(merged$sdg_overall[merged$regime_type_2 == 0], na.rm = TRUE),
    mean(merged$spi_comp[merged$regime_type_2 == 1], na.rm = TRUE) - 
      mean(merged$spi_comp[merged$regime_type_2 == 0], na.rm = TRUE),
    mean(merged$sci_overall[merged$regime_type_2 == 1], na.rm = TRUE) - 
      mean(merged$sci_overall[merged$regime_type_2 == 0], na.rm = TRUE),
    mean(merged$di_score[merged$regime_type_2 == 1], na.rm = TRUE) - 
      mean(merged$di_score[merged$regime_type_2 == 0], na.rm = TRUE)
  ),
  Percent_Difference = c(
    (mean(merged$sdg_overall[merged$regime_type_2 == 1], na.rm = TRUE) / 
      mean(merged$sdg_overall[merged$regime_type_2 == 0], na.rm = TRUE) - 1) * 100,
    (mean(merged$spi_comp[merged$regime_type_2 == 1], na.rm = TRUE) / 
      mean(merged$spi_comp[merged$regime_type_2 == 0], na.rm = TRUE) - 1) * 100,
    (mean(merged$sci_overall[merged$regime_type_2 == 1], na.rm = TRUE) / 
      mean(merged$sci_overall[merged$regime_type_2 == 0], na.rm = TRUE) - 1) * 100,
    (mean(merged$di_score[merged$regime_type_2 == 1], na.rm = TRUE) / 
      mean(merged$di_score[merged$regime_type_2 == 0], na.rm = TRUE) - 1) * 100
  )
)

# Format and display the summary table with APA styling and no caption number
summary_findings_table <- kable(summary_findings, 
      caption = "Summary of Differences Between Democracies and Autocracies",
      col.names = c("Variable", "Democracy Mean", "Autocracy Mean", "Absolute Difference", "% Difference"),
      booktabs = TRUE,
      digits = 2,
      align = c('l', rep('c', 4))) %>%
  kable_classic(full_width = FALSE, html_font = "Times New Roman", font_size = 12, "basic") %>%
  row_spec(0, bold = FALSE, align = "c") %>%
  column_spec(1, bold = FALSE) %>%
  add_footnote("Note: Positive differences indicate higher values in democracies compared to autocracies.", notation = "none") %>%
  footnote(general_title = "", 
           general = "Source: Author's calculations using merged dataset.")

# Save the table as HTML
save_kable(summary_findings_table, file = "figures/data coverage & summary stats/summary_findings_table.html")

# Display the table in the document
summary_findings_table
```