---
title: "First Tests OLS, FDs"
author: "Sevastian Sanchez"
date: "`r Sys.Date()`"
output: html_document
---
# All Countries, Preliminary Analysis (SPI x SDGs)
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Libraries 
```{r}
library(devtools)
library(tidyr)
library(dplyr)
library(vdemdata) # call vdem package 
library(readr) #loading data 
library(readxl)
library(stargazer)
library(broom)
```


##CLEANING / WRANGLING 
```{r}
#Vdem package from github 
vdem <- vdemdata::vdem  #loading data 

#spi csv from github 
url <- "https://raw.githubusercontent.com/worldbank/SPI/refs/heads/master/03_output_data/SPI_index.csv"
spi <- read_csv(url)

#sdg excel from directory 
sdg <- read_excel("~/Desktop/OneDrive/OneDrive/QMSS Thesis/Datasets/SDR2024-data.xlsx", sheet = "Backdated SDG Index")

#sci csv from directory 
sci_df <- read_csv("~/Desktop/OneDrive/OneDrive/QMSS Thesis/Datasets/SCI_All_Dim_TS.csv")

# set working directory 
setwd("~/Documents/GitHub/QMSS_Thesis_Sanchez")
```

##Vdem Wrangling / Cleaning
```{r}
#selecting democracy indices and controls 
vdem <- vdem %>%
  dplyr::select(country_name, country_text_id, year, v2x_regime, v2x_regime_amb, v2x_polyarchy, v2x_libdem, v2x_partipdem, v2x_delibdem, v2x_egaldem, v2xel_frefair, v2x_accountability, v2x_veracc, v2x_horacc, v2x_diagacc, v2xca_academ, v2x_freexp_altinf, e_gdp, e_gdppc, e_wb_pop, v3ststybcov, v3ststybpub, v3stcensus, v3ststatag) %>%
  filter(year > 1999) %>% #Only 1999 and up 
  rename(country_code = country_text_id, #renaming country code (new_name = old_name)
         regime_type = v2x_regime, 
         regime_type_amb = v2x_regime_amb,
         elect_dem = v2x_polyarchy,
         lib_dem = v2x_libdem,
         part_dem = v2x_partipdem,
         delib_dem = v2x_delibdem,
         egal_dem = v2x_egaldem,
         freefair = v2xel_frefair,
         #freefair_ord = e_v2xel_frefair,
         accountability = v2x_accountability,
         vt_account = v2x_veracc,
         hz_account = v2x_horacc,
         diag_account = v2x_diagacc,
         academ_free = v2xca_academ,
         freexp_altinfo = v2x_freexp_altinf,
         gdp = e_gdp,
         gdp_pc = e_gdppc,
         population = e_wb_pop,
         stat_agency = v3ststatag,
         stat_yb_cov = v3ststybcov,
         stat_yb_pub = v3ststybpub,
         census = v3stcensus,
         )
  
# To be added: e_v2xel_frefair,

#vdem$v2xca_academ <- vdem$v2xca_academ*100
#vdem$v2x_libdem <- vdem$v2x_libdem*100
```


## SPI wrangling / cleaning 
```{r}
#selecting and renaming spi vars 
spi <- spi %>% 
  select(country, iso3c, date, SPI.INDEX, SPI.INDEX.PIL1, SPI.INDEX.PIL2, SPI.INDEX.PIL3, SPI.INDEX.PIL4, SPI.INDEX.PIL5) %>% 
  rename(country_name = country) %>% 
  rename(country_code = iso3c) %>% 
  rename(year = date) %>% 
  rename(spi_comp = SPI.INDEX) %>% #SPI composite score, average of p1-5
  rename(p1_use = SPI.INDEX.PIL1) %>% #SPI pillar 1, Data use 
  rename(p2_services = SPI.INDEX.PIL2) %>% #SPI pillar, 2 Data services 
  rename(p3_products = SPI.INDEX.PIL3) %>% #SPI pillar, 3 Data products  
  rename(p4_sources = SPI.INDEX.PIL4) %>% #SPI pillar, 4 Data sources 
  rename(p5_infra = SPI.INDEX.PIL5) #SPI pillar 5, Data infrastructure
```

## SCI wrangling / cleaning 
```{r}
#full dataset 
sci <- sci_df %>% 
  dplyr::select(country_name, country_code, Year, IQ.SCI.OVRL, IQ.SCI.MTHD, IQ.SCI.PRDC, IQ.SCI.SRCE, everything()) %>% 
  rename(year = Year,
         sci_overall = IQ.SCI.OVRL, 
         sci_method = IQ.SCI.MTHD, 
         sci_periodicity = IQ.SCI.PRDC,
         sci_source = IQ.SCI.SRCE)

sci$year <- as.numeric(sci$year)

#for merging (only sci score variables selected)
sci_scores <- sci %>% 
  select(country_name, country_code, year, sci_overall, sci_method, sci_periodicity, sci_source)
```

##SDG wrangling / cleaning 
```{r}
#select sdg goals & composite 
sdg <- sdg %>% 
  dplyr::select("country_name", "country_code", "year", "sdg_overall", "goal1", "goal2", "goal3", "goal4", "goal5", "goal6", "goal7", "goal8", "goal9", "goal10", "goal11", "goal12", "goal3", "goal14", "goal15", "goal16", "goal17")

```

# Merging & subsetting (case study: Spain)
```{r}
#merge spi, sci, sdg and vdem data 
merged <- left_join(sdg, spi, by = c("year", "country_code")) 
merged <- left_join(merged, sci_scores, by = c("country_code", "year"))
merged <- left_join(merged, vdem, by = c("country_code", "year"))

#rm duplicate col names 
colnames(merged) <- make.unique(colnames(merged)) # Make column names unique

#data type changes 
merged$year <- as.integer(merged$year)
merged <- merged %>% 
  mutate_at(c("spi_comp", "p1_use", "p2_services", "p3_products", "p4_sources", "p5_infra", "sci_overall", "sci_method", "sci_periodicity", "sci_source"), as.numeric)

#exporting time series data AS: merged_ts.csv ; will save to current working directory 
#write.table(merged, file = "merged_ts.csv", row.names=F, sep = ",") 
```

## Wrangling for Spain - Test Subject 
```{r}
# subset for Spain 
merged_esp_ts <- merged[merged$country_code=="ESP",]

#exporting Spain time series AS: merged_esp_ts; will save to current working directory 
#write.table(merged_esp_ts, file = "merged_esp_ts.csv", row.names=F, sep = ",")
```

##Cross-Sectional Analysis - 2019 data only
```{r}
#subset for only 2019 
merged_cs_2019 <- merged[merged$year==2019,]

#exporting cross-sectional 2019 data AS: merged_cs2019; will save to current working directory 
#write.table(merged_cs_2019, file = "merged_cs2019.csv", row.names=F, sep = ",")
```

## NAIVE OLS (BI-VARIATE)
**Aggregated SPI & SDG Scores**
*H0: Null, there is no relationship*
*H1: there is a statistically significant correlation between overall SPI and SDG composite scores*
```{r}
# x & y variables 
y_var_sdg <- merged$sdg_overall 
x_var_spi <- merged$spi_comp
x_var_sci <- merged$sci_overall
controls <- merged %>% select(gdp, gdp_pc, population)
df_sdg_statcap <- data.frame(y_var_sdg, x_var_spi, x_var_sci)

# how many na's?
colSums(is.na(df_sdg_statcap))

# omit na's 
df_sdg_spi <- na.omit(data.frame(y_var_sdg, x_var_spi))
df_sdg_sci <- na.omit(data.frame(y_var_sdg, x_var_sci))
dim(df_sdg_spi) #spi observations, ommitted na's 
dim(df_sdg_sci) #sci observations, ommitted na's 

#correlation coefficient (x, y)
correlation_sdg_spi <- cor(df_sdg_spi$y_var_sdg, df_sdg_spi$x_var_spi)^2 #x-var 1 = spi
correlation_sdg_sci <- cor(df_sdg_sci$y_var_sdg, df_sdg_sci$x_var_sci)^2 #x-var 2 = sci
string_corcoef <- "Correlation coefficient:" 
paste(string_corcoef, correlation_sdg_spi) # pasting result 
paste(string_corcoef, correlation_sdg_sci) # pasting result 
```
Number of observations in SPI: 1300; number of NA's 3068 (not all years captured)
Number of observations in SCI: 2055; number of NA's 2313 (not all countries captured)

# Comparing SPI & SCI to identify best model 
```{r}
# 1. OLS for SPI and SDG - Overall 
ols_spi <- lm(y_var_sdg ~ x_var_spi + merged$gdp + merged$gdp_pc + merged$population)
summary(ols_spi)

# 2. OLS for SCI and SDG - Overall 
ols_sci <- lm(y_var_sdg ~ x_var_sci + merged$gdp + merged$gdp_pc + merged$population)
summary(ols_sci)

# 3. Multiple Regression with both SPI and SCI
ols_multiple <- lm(y_var_sdg ~ x_var_spi + x_var_sci + merged$gdp + merged$gdp_pc + merged$population)
summary(ols_multiple)

#visualize differences in fit
library(ggplot2)
library(plotly)

#define regression line colors
spi_line <- "steelblue4"
sci_line <- "darkgoldenrod"

# Creating scatterplot with both SPI and SCI on the same plot
Compare_fit <- ggplot(merged, aes(x = spi_comp, y = sdg_overall))+
  geom_smooth(aes(x = spi_comp, y = sdg_overall), color = spi_line, method = "lm", linewidth = 0.75, se = FALSE)+ # Regression line SPI
  geom_smooth(aes(x = sci_overall, y = sdg_overall), color = sci_line, method = "lm", linewidth = 0.75, se = FALSE)+ # Regression line for SCI
  geom_point(aes(color = "spi_comp"), alpha=0.50, size = 0.5)+  # Scatter plot for SPI, colored distinctly
  geom_point(aes(x = sci_overall, y = sdg_overall, color = "sci_overall"), alpha=0.5, size = 0.5)+ # Add SCI points w/different color
  scale_color_manual(values = c("spi_comp" = "steelblue1", "sci_overall" = "darkgoldenrod1")) +  # Define colors
  labs(title = "SDG vs. SPI and SCI",
       x = "SPI / SCI",
       y = "SDG",
       color = "Index") +  # Title for legend
  theme_bw() # Optional: adds a clean, black and white theme

#make interactive
ggplotly(Compare_fit)

# Save to specific folder
# ggsave("~/Documents/GitHub/QMSS_Thesis_Sanchez/Output_CSVs/fd_plot.png", p, width = 10, height = 6)
```
We reject the null hypothesis that there is no relationship between SPI and SDG composite scores. Holding all else constant, there is a positive moderate relationship between Statistical Performance (SPI) and SDG status. 


# MULTIPLE LINEAR REGRESSION 
**Disaggregated by Development Status:**  
*Make 4 regression models and then put them all together in a table to compare the slopes and R-sq values.* 
```{r}
#WB GNI income categories 
income_class <- read_csv("~/Documents/GitHub/QMSS_Thesis_Sanchez/Output_CSVs /world_bank_income_classifications.csv")

# New df with income classifications 
merged_inc <- merged %>% 
  left_join(income_class, by = c("country_code", "year"))

# Assuming your merged dataset contains:
# X = spi_comp (SPI values)
# Y = sdg_overall (SDG values)
# Subset: income_level (income classification: "H", "UM", "LM", "L")

# 1. Overall model (all countries)
overall_lm <- lm(sdg_overall ~ spi_comp + log(gdp_pc), #put log(GDP), check if pop is relevant, natural log (ln(GDP))
                        data = merged_inc)
summary(overall_lm)

# 2. High income countries
high_inc_lm <- lm(sdg_overall ~ spi_comp + gdp + gdp_pc + population,
                        data = merged_inc %>% 
                          filter(income_level == "H"))
summary(high_inc_lm)

# 3. Upper-middle income countries
upper_mid_lm <- lm(sdg_overall ~ spi_comp + gdp + gdp_pc + population,
                         data = merged_inc %>% 
                           filter(income_level == "UM"))
summary(upper_mid_lm)

# 4. Lower-middle income countries
lower_mid_lm <- lm(sdg_overall ~ spi_comp + gdp + gdp_pc + population,
                         data = merged_inc %>% 
                           filter(income_level == "LM"))
summary(lower_mid_lm)

# 5. Low income countries
low_inc_lm <- lm(sdg_overall ~ spi_comp + gdp + gdp_pc + population,
                       data = merged_inc %>% 
                         filter(income_level == "L"))
summary(low_inc_lm)


# take best regression in R squared and try with variables and ensure their relevance, 
# check to see if Open data is better, a proxy of SPI _ recommends for missing data proxy 
# also look into SCI 
# take few countries 
# send github and and document 
# comments of the limitations, specifically the time series \
# SPI has 1 pillar with 
# a good thesis is one that highlights the issues and being transparent 

```

# Extract coefficients and statistics from each model to create comparison table 
```{r}
# Creating mod_compare_tab - comparison table
mod_compare_tab <- data.frame(
  model = c("Overall", "High Income", "Upper-Middle Income", 
           "Lower-Middle Income", "Low Income"),
  coefficient = c(
    coef(overall_lm)["spi_comp"],
    coef(high_inc_lm)["spi_comp"],
    coef(upper_mid_lm)["spi_comp"],
    coef(lower_mid_lm)["spi_comp"],
    coef(low_inc_lm)["spi_comp"]
  ),
  intercept = c(
    coef(overall_lm)["(Intercept)"],
    coef(high_inc_lm)["(Intercept)"],
    coef(upper_mid_lm)["(Intercept)"],
    coef(lower_mid_lm)["(Intercept)"],
    coef(low_inc_lm)["(Intercept)"]
  ),
  t_value = c(
    summary(overall_lm)$coefficients["spi_comp", "t value"],
    summary(high_inc_lm)$coefficients["spi_comp", "t value"],
    summary(upper_mid_lm)$coefficients["spi_comp", "t value"],
    summary(lower_mid_lm)$coefficients["spi_comp", "t value"],
    summary(low_inc_lm)$coefficients["spi_comp", "t value"]
  ),
  p_value = c(
    summary(overall_lm)$coefficients["spi_comp", "Pr(>|t|)"],
    summary(high_inc_lm)$coefficients["spi_comp", "Pr(>|t|)"],
    summary(upper_mid_lm)$coefficients["spi_comp", "Pr(>|t|)"],
    summary(lower_mid_lm)$coefficients["spi_comp", "Pr(>|t|)"],
    summary(low_inc_lm)$coefficients["spi_comp", "Pr(>|t|)"]
  ),
  r_squared = c(
    summary(overall_lm)$r.squared,
    summary(high_inc_lm)$r.squared,
    summary(upper_mid_lm)$r.squared,
    summary(lower_mid_lm)$r.squared,
    summary(low_inc_lm)$r.squared
  ),
  adj_r_squared = c(
    summary(overall_lm)$adj.r.squared,
    summary(high_inc_lm)$adj.r.squared,
    summary(upper_mid_lm)$adj.r.squared,
    summary(lower_mid_lm)$adj.r.squared,
    summary(low_inc_lm)$adj.r.squared
  ),
  n_obs = c(
    nobs(overall_lm),
    nobs(high_inc_lm),
    nobs(upper_mid_lm),
    nobs(lower_mid_lm),
    nobs(low_inc_lm)
  )
)

# rounding for better display
mod_compare_tab <- mod_compare_tab %>%
  mutate(
    coefficient = round(coefficient, 3),
    intercept = round(intercept, 3),
    t_value = round(t_value, 3),
    p_value = p_value,
    r_squared = round(r_squared, 3),
    adj_r_squared = round(adj_r_squared, 3)
  )

# Print the comparison table
print(mod_compare_tab)

# export mod_compare_tab 
# write.csv(mod_compare_tab, file = "ols_model_comparison.csv", row.names=F) 
```

```{r}
# Convert income_level to factor with explicit labels
merged_inc$income_level <- factor(merged_inc$income_level,
                                 levels = c("H", "UM", "LM", "L"),
                                 labels = c("High Income Countries", 
                                           "Upper-Middle Income Countries", 
                                           "Lower-Middle Income Countries", 
                                           "Low Income Countries"))
```

**Visualize the slopes**
```{r}
# Create a visualization with different regression lines by income group
by_income_class <- ggplot(data = merged_inc, aes(x = spi_comp, y = sdg_overall, color = income_level)) +
  geom_point(alpha = 0.25, size = 0.75) + 
  geom_smooth(method = "lm", linewidth = 0.65, se = FALSE) + # 4 regressions by income classes
  scale_color_manual(values = c("High Income Countries" = "#1D6A96", 
                                "Upper-Middle Income Countries" = "#4CB5AE", 
                                "Lower-Middle Income Countries" = "#F3A738",
                                "Low Income Countries" = "#C02942"))

#Adding overall regression line 
by_income_class <- by_income_class + geom_smooth(data = merged_inc, aes(x = spi_comp, y = sdg_overall), method = "lm",
              linewidth = 0.75, se = FALSE, color = "black")

# Applying theme and labels 
by_income_class <- by_income_class + 
  labs(title = "Relationship between SPI and SDG by World Bank Income Classification",
       x = "Statistical Performance Indicators (SPI)",
       y = "Sustainable Development Goals (SDG)",
       color = "Income Classification" #legend title 
       ) +
  theme_bw()

ggplotly(by_income_class)

# Save to specific folder
# ggsave("~/Documents/GitHub/QMSS_Thesis_Sanchez/Output_CSVs/fd_plot.png", p, width = 10, height = 6)
```

# First Difference Models 
**PART 1: Create lagged variables and first differences**
```{r}
library(plm)
#library(dplyr)

# Create a simplified dataframe with essential variables
fd_data <- merged_inc %>%
  select(country_code, year, sdg_overall, spi_comp, income_level)

# PART 1: Create lagged variables and first differences with explicit dplyr reference
fd_data <- fd_data %>%
  arrange(country_code, year) %>% 
  group_by(country_code) %>%
  mutate(
    # Create lagged variables with explicit dplyr reference
    sdg_lag = dplyr::lag(sdg_overall, n = 1),
    spi_lag = dplyr::lag(spi_comp, n = 1),
    
    # Calculate first differences
    fd_sdg = sdg_overall - sdg_lag,
    fd_spi = spi_comp - spi_lag
  ) %>%
  ungroup()

# To verify the lag creation works correctly
# head(fd_data %>% select(country_code, year, sdg_overall, sdg_lag, fd_sdg, spi_comp, spi_lag, fd_spi) %>% arrange(country_code, year), 10)
```

**PART 2: Run FD models for each income group**
```{r}
# Overall model
fd_overall <- lm(fd_sdg ~ fd_spi, 
                data = fd_data %>% 
                  filter(!is.na(fd_sdg) & !is.na(fd_spi)))
summary(fd_overall)

# High income countries
fd_high <- lm(fd_sdg ~ fd_spi, 
             data = fd_data %>% 
               filter(!is.na(fd_sdg) & !is.na(fd_spi) & 
                     income_level == "High Income Countries"))
summary(fd_high)

# Upper-middle income countries
fd_upper <- lm(fd_sdg ~ fd_spi, 
              data = fd_data %>% 
                filter(!is.na(fd_sdg) & !is.na(fd_spi) & 
                      income_level == "Upper-Middle Income Countries"))
summary(fd_upper)

# Lower-middle income countries
fd_lower <- lm(fd_sdg ~ fd_spi, 
              data = fd_data %>% 
                filter(!is.na(fd_sdg) & !is.na(fd_spi) & 
                      income_level == "Lower-Middle Income Countries"))
summary(fd_lower)

# Low income countries
fd_low <- lm(fd_sdg ~ fd_spi, 
            data = fd_data %>% 
              filter(!is.na(fd_sdg) & !is.na(fd_spi) & 
                    income_level == "Low Income Countries"))
summary(fd_low)
```

**PART 3: Create a comparison table**
```{r}
# Extract coefficients and statistics from each model
fd_comparison <- data.frame(
  model = c("Overall", "High Income", "Upper-Middle Income", 
           "Lower-Middle Income", "Low Income"),
  coefficient = c(
    coef(fd_overall)["fd_spi"],
    coef(fd_high)["fd_spi"],
    coef(fd_upper)["fd_spi"],
    coef(fd_lower)["fd_spi"],
    coef(fd_low)["fd_spi"]
  ),
  intercept = c(
    coef(fd_overall)["(Intercept)"],
    coef(fd_high)["(Intercept)"],
    coef(fd_upper)["(Intercept)"],
    coef(fd_lower)["(Intercept)"],
    coef(fd_low)["(Intercept)"]
  ),
  t_value = c(
    summary(fd_overall)$coefficients["fd_spi", "t value"],
    summary(fd_high)$coefficients["fd_spi", "t value"],
    summary(fd_upper)$coefficients["fd_spi", "t value"],
    summary(fd_lower)$coefficients["fd_spi", "t value"],
    summary(fd_low)$coefficients["fd_spi", "t value"]
  ),
  p_value = c(
    summary(fd_overall)$coefficients["fd_spi", "Pr(>|t|)"],
    summary(fd_high)$coefficients["fd_spi", "Pr(>|t|)"],
    summary(fd_upper)$coefficients["fd_spi", "Pr(>|t|)"],
    summary(fd_lower)$coefficients["fd_spi", "Pr(>|t|)"],
    summary(fd_low)$coefficients["fd_spi", "Pr(>|t|)"]
  ),
  r_squared = c(
    summary(fd_overall)$r.squared,
    summary(fd_high)$r.squared,
    summary(fd_upper)$r.squared,
    summary(fd_lower)$r.squared,
    summary(fd_low)$r.squared
  ),
  adj_r_squared = c(
    summary(fd_overall)$adj.r.squared,
    summary(fd_high)$adj.r.squared,
    summary(fd_upper)$adj.r.squared,
    summary(fd_lower)$adj.r.squared,
    summary(fd_low)$adj.r.squared
  ),
  n_obs = c(
    nobs(fd_overall),
    nobs(fd_high),
    nobs(fd_upper),
    nobs(fd_lower),
    nobs(fd_low)
  )
)

# Format the numbers for better display
fd_comparison <- fd_comparison %>%
  mutate(
    coefficient = round(coefficient, 3),
    intercept = round(intercept, 3),
    t_value = round(t_value, 3),
    p_value = round(p_value, 4),
    r_squared = round(r_squared, 3),
    adj_r_squared = round(adj_r_squared, 3)
  )

# Print the comparison table
print(fd_comparison)

#export fd_comparison 
write.csv(fd_comparison, file = "fd_comparison.csv", row.names=F) 
```

**RESULTS:** 
Your first-difference (FD) model results reveal several important insights about the relationship between changes in Statistical Performance Indicators (SPI) and changes in Sustainable Development Goals (SDG) scores across different income groups:

[perplexity] The overall model shows a positive and statistically significant relationship (coefficient = 0.015, p = 0.0033) between improvements in statistical capacity and improvements in SDG performance. This confirms your hypothesis that better statistical systems contribute to better development outcomes.

However, the income group breakdown reveals fascinating variations:

1. **High Income Countries** show a negative coefficient (-0.012) that is not statistically significant (p = 0.2537). This suggests that for developed economies, improvements in statistical capacity don't necessarily translate to SDG improvements, possibly because they already have well-established statistical systems and other factors drive their SDG performance.

2. **Upper-Middle Income Countries** show a stronger positive effect (0.021) that is statistically significant (p = 0.0163). This indicates that for these emerging economies, investments in statistical capacity yield meaningful SDG improvements.

3. **Lower-Middle Income Countries** demonstrate an even stronger relationship (0.025) with high statistical significance (p = 0.0088). This is the most significant relationship among all income groups.

4. **Low Income Countries** show the strongest coefficient (0.026), suggesting the largest potential impact, though with slightly lower statistical significance (p = 0.0459) than lower-middle income countries.

The R-squared values are relatively low across all models (between 0.004 and 0.023), indicating that changes in statistical capacity explain only a small portion of the variation in SDG changes. This is common in first-difference models and suggests that many other factors also influence SDG progress.

For your project, these findings suggest that:

1. The relationship between statistical capacity and sustainable development follows an inverted U-shaped pattern across the development spectrum - minimal impact for high-income countries, strongest for middle-income countries, and slightly less strong (but still significant) for low-income countries.

2. Statistical capacity improvements appear most beneficial for developing and emerging economies, suggesting that international efforts to strengthen statistical systems should prioritize these countries.

3. The consistently positive intercepts (ranging from 0.224 to 0.364) indicate that SDG scores tend to improve over time regardless of changes in statistical capacity, though the rate of improvement varies by income level.

These results provide empirical support for the importance of statistical capacity building as a development tool, particularly for countries in the middle and lower ranges of economic development.

# Visual Analysis | FD Models 
**Error-bar**
```{r}
# Create dataframe for plotting 
plot_data <- fd_comparison %>%
  select(model, coefficient, p_value) %>%
  mutate(significant = p_value < 0.05)

# Create Horizontal error-bar graph 
fd_errorbar <- ggplot(plot_data, aes(x = coefficient, y = model, color = significant)) +
  geom_point(size = 3) +
  geom_errorbarh(aes(xmin = 0, xmax = coefficient), height = 0.2) + 
  geom_vline(xintercept = 0, linetype = "dashed") +
  scale_color_manual(values = c("gray", "blue")) +
  labs(title = "Effect of SPI Change on SDG Change by Income Group",
       x = "Coefficient (Effect Size)",
       y = "") +
  theme_bw() +
  theme(legend.position = "none")

fd_errorbar

#export fd_errorbar
# ggsave("fd_errorbar.png", width = 10, height = 6)
```

**Scatter plot of FDs**
```{r}
fd_scatter <- ggplot(fd_data %>% 
         filter(!is.na(fd_sdg) & !is.na(fd_spi)), 
       aes(x = fd_spi, y = fd_sdg, color = income_level)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", se = FALSE) +
  facet_wrap(~income_level) +
  labs(title = "Relationship between Changes in SPI and SDG by Income Group",
       x = "Change in Statistical Performance (SPI)",
       y = "Change in Sustainable Development (SDG)") +
  theme_bw()

fd_scatter

#export fd_scatter
# ggsave("iCloud_Drive/Documents/GitHub/QMSS_Thesis_Sanchez", width = 10, height = 6)
```








## NAIVE OLS W/Controls 
```{r}
# y & x vars + controls/
y_var <- merged$sdg_overall 
x1_var <- merged$spi_comp #insert x var 1
x2_var <- merged$p1_use #insert x var 2
x3_var <- merged$p2_services #insert x var 3
x4_var <- merged$p3_products #insert x var 3
x5_var <- merged$p4_sources #insert x var 3
x6_var <- merged$p5_infra #insert x var 3
df_2 <- data.frame(y_var, x1_var, x2_var, x3_var, x4_var, x5_var, x6_var)

#OLS lm & full summary table 
lm2 <- lm(y_var ~ x1_var + x2_var + x3_var + x4_var + x5_var + x6_var, data=df_2)
summary(lm2)
#plotting residuals
plot(lm2)
#plotting scatterplot 
plot(df$y_var, df$x1_var)
abline(lm2)
```

## Identify Regressing Countries 
```{r}
# 2015-present 






# 2000-present 


```




## First Difference Model 

```{r}


```


## Testing Index Reliability (Cronbach's Alpha)
```{r}

```




