---
title: "First Tests OLS, FDs"
author: "Sevastian Sanchez"
date: "`r Sys.Date()`"
output: html_document
---

# All Countries, Preliminary Analysis (SPI x SDGs)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#FIRST: Libraries

```{r}
#load libraries 
source("packages.R")

```

#SECOND: Load data

```{r}
# set working directory 
setwd("~/Documents/GitHub/QMSS_Thesis_Sanchez")

#load data 
source("data/data_sources.R")

```

#THIRD: run function in r-script: df_years_function.R

```{r}
#df_years: used for extracting data from specified year to now 

#load function
source("df_years()_Function.R")

#specify start year 
merged <- df_years(yr1 = 2000) 
```

##Vdem wrangling / cleaning (skip if df_years() used)
```{r eval=FALSE, include=FALSE}
#selecting democracy indices and controls 
vdem <- vdem %>% 
dplyr::select(country_name, country_text_id, year, v2x_regime, v2x_regime_amb, v2x_polyarchy, v2x_libdem, v2x_partipdem, v2x_delibdem, v2x_egaldem, v2xel_frefair, v2x_accountability, v2x_veracc, v2x_horacc, v2x_diagacc, v2xca_academ, v2x_freexp_altinf, e_gdp, e_gdppc, e_wb_pop, v3ststybcov, v3ststybpub, v3stcensus, v3ststatag) %>%
  filter(year > 1999) %>% #Only 1999 and up 
  rename(country_code = country_text_id, #renaming country code (new_name = old_name)
         regime_type = v2x_regime, 
         regime_type_amb = v2x_regime_amb,
         elect_dem = v2x_polyarchy,
         lib_dem = v2x_libdem,
         part_dem = v2x_partipdem,
         delib_dem = v2x_delibdem,
         egal_dem = v2x_egaldem,
         freefair = v2xel_frefair,
         #freefair_ord = e_v2xel_frefair,
         accountability = v2x_accountability,
         vt_account = v2x_veracc,
         hz_account = v2x_horacc,
         diag_account = v2x_diagacc,
         academ_free = v2xca_academ,
         freexp_altinfo = v2x_freexp_altinf,
         population = e_wb_pop,
         stat_agency = v3ststatag,
         stat_yb_cov = v3ststybcov,
         stat_yb_pub = v3ststybpub,
         census = v3stcensus,
         )
  
# To be added: e_v2xel_frefair
```


## SPI wrangling / cleaning (skip if df_years() used)

```{r eval=FALSE, include=FALSE}
#selecting and renaming spi vars 
spi <- spi %>% 
  select(country, iso3c, date, SPI.INDEX, SPI.INDEX.PIL1, SPI.INDEX.PIL2, SPI.INDEX.PIL3, SPI.INDEX.PIL4, SPI.INDEX.PIL5) %>% 
  rename(country_name = country) %>% 
  rename(country_code = iso3c) %>% 
  rename(year = date) %>% 
  rename(spi_comp = SPI.INDEX) %>% #SPI composite score, average of p1-5
  rename(p1_use = SPI.INDEX.PIL1) %>% #SPI pillar 1, Data use 
  rename(p2_services = SPI.INDEX.PIL2) %>% #SPI pillar, 2 Data services 
  rename(p3_products = SPI.INDEX.PIL3) %>% #SPI pillar, 3 Data products  
  rename(p4_sources = SPI.INDEX.PIL4) %>% #SPI pillar, 4 Data sources 
  rename(p5_infra = SPI.INDEX.PIL5) #SPI pillar 5, Data infrastructure
```


## SCI wrangling / cleaning (skip if df_years() used)

```{r eval=FALSE, include=FALSE}
#full dataset 
sci <- sci_df %>% 
  dplyr::select(country_name, country_code, Year, IQ.SCI.OVRL, IQ.SCI.MTHD, IQ.SCI.PRDC, IQ.SCI.SRCE, everything()) %>% 
  rename(year = Year,
         sci_overall = IQ.SCI.OVRL, 
         sci_method = IQ.SCI.MTHD, 
         sci_periodicity = IQ.SCI.PRDC,
         sci_source = IQ.SCI.SRCE)

#sci$year <- as.numeric(sci$year)

#for merging (only sci score variables selected)
sci_scores <- sci %>% 
  select(country_name, country_code, year, sci_overall, sci_method, sci_periodicity, sci_source)
```



##SDG wrangling / cleaning (skip if df_years() used)

```{r eval=FALSE, include=FALSE}
#select sdg goals & composite 
sdg <- sdg %>% 
  dplyr::select("country_name", "country_code", "year", "sdg_overall", "goal1", "goal2", "goal3", "goal4", "goal5", "goal6", "goal7", "goal8", "goal9", "goal10", "goal11", "goal12", "goal3", "goal14", "goal15", "goal16", "goal17")

```


##GDP_PC wrangling / cleaning (skip if df_years() used)
```{r eval=FALSE, include=FALSE}
gdppc_df <- gdppc_df %>%  #pivot back to long format
  pivot_longer(3:66, #use any way to select all columns (he used range)
               names_to = "year", # new var using cols 2:13 as values
               values_to = "gdp_pc") #new var containing all values within cols 2:13 as values 
View(gdppc_df)
```




## Merging & subsetting (skip if df_years() used)

```{r eval=FALSE, include=FALSE}
#merge spi, sci, sdg and vdem data 
merged <- left_join(sdg, spi, by = c("year", "country_code")) 
merged <- left_join(merged, sci_scores, by = c("country_code", "year"))
merged <- left_join(merged, vdem, by = c("country_code", "year"))

#rm duplicate col names 
colnames(merged) <- make.unique(colnames(merged)) # Make column names unique

#data type changes 
merged$year <- as.integer(merged$year)
merged <- merged %>% 
  mutate_at(c("spi_comp", "p1_use", "p2_services", "p3_products", "p4_sources", "p5_infra", "sci_overall", "sci_method", "sci_periodicity", "sci_source"), as.numeric)

#exporting time series data AS: merged_ts.csv ; will save to current working directory 
#write.table(merged, file = "merged_ts.csv", row.names=F, sep = ",") 
```


## All countries 2019 data only
```{r eval=FALSE, include=FALSE}
#subset for only 2019 
merged_cs_2019 <- merged[merged$year==2019,]

#exporting cross-sectional 2019 data AS: merged_cs2019; will save to current working directory 
#write.table(merged_cs_2019, file = "merged_cs2019.csv", row.names=F, sep = ",")
```





## Spain years 2000-present - Test Subject
```{r eval=FALSE, include=FALSE}
# subset for Brazil 
merged_bra_ts <- merged[merged$country_code=="BRA",]

# Subset for Turkey 
merged_tur_ts <- merged[merged$country_code=="TUR",]

# Subset for Russia 
merged_rus_ts <- merged[merged$country_code=="RUS",]

# Subset for Hungary
merged_hun_ts <- merged[merged$country_code=="HUN",]

# Subset for Spain 
merged_esp_ts <- merged[merged$country_code=="ESP",]

# Subset for USA 
merged_usa_ts <- merged[merged$country_code=="USA",]

#exporting Spain time series AS: merged_esp_ts; will save to current working directory 
#write.table(merged_esp_ts, file = "merged_esp_ts.csv", row.names=F, sep = ",")
```


##set up & missing data
```{r}
# x & y variables 
y_var_sdg <- merged$sdg_overall 
x_var_spi <- merged$spi_comp
x_var_sci <- merged$sci_overall
df_sdg_statcap <- data.frame(y_var_sdg, x_var_spi, x_var_sci)

# how many na's?
colSums(is.na(df_sdg_statcap))
#how many observations 
colSums(!is.na(df_sdg_statcap))
```

Number of observations in SPI: 1300; number of NA's 3068 (not all years captured) Number of observations in SCI: 2055; number of NA's 2313 (not all countries captured)

## NAIVE OLS (BI-VARIATE)

**Aggregated SPI & SDG Scores**  
*H0: Null, there is no relationship*  
*H1: there is a statistically significant relationship between overall SPI and SDG composite scores*

```{r}
#correlation coefficients (r-squared), WITHOUT control variables 

#x-var 1 = spi
correlation_sdg_spi <- cor(y_var_sdg, x_var_spi, use = "complete.obs")^2

#x-var 2 = sci
correlation_sdg_sci <- cor(y_var_sdg, x_var_sci, use = "complete.obs")^2 

# pasting result 
string_corcoef <- "Correlation coefficient:" 
paste(string_corcoef, correlation_sdg_spi, "(SPI)", correlation_sdg_sci, "(SCI)")
```

Correlation coefficient/R-sq (SPI): 0.616037202309322  
Correlation coefficient/R-sq (SCI): 0.417965563339242

# Comparing SPI & SCI to identify best model

*H0: Null, SCI model > SPI model*  
*H1: SPI model > SCI model*

```{r}
# 1. OLS for SPI and SDG - Overall 
ols_spi <- lm(y_var_sdg ~ x_var_spi + log(gdppc) + population, data = merged)
summary(ols_spi)

# 2. OLS for SCI and SDG - Overall 
ols_sci <- lm(y_var_sdg ~ x_var_sci + log(gdppc) + population, data = merged)
summary(ols_sci)

# 3. Multiple Regression with both SPI and SCI
ols_multiple <- lm(y_var_sdg ~ x_var_spi + x_var_sci + log(gdppc) + population, data = merged)
summary(ols_multiple)
```

We reject the null hypothesis that there is no relationship between SPI and SDG composite scores. Holding all else constant (log gdp per capita and population), there is a positive moderate relationship between Statistical Performance (SPI) and SDG status.

ols_spi: 0.280 (p-value < 0.001)  
ols_sci: 0.268 (p-value < 0.001)  
ols_multiple: spi: 0.1216 (p-value < 0.001); sci: 0.1288 (p-value < 0.001)

Comparing coefficients, SPI has a greater impact on SDG status (0.280) than SCI (0.268). However, in a multiple regression model containing both SPI and SCI, SCI has more of an impact on SDG status (0.1288) (net of SPI) than SPI (0.1216) (net of SCI). Here the coefficients represent the unique impact of each x variable on SDG status, net of all other variables.

Model 1 (ols_spi) does not control for SCI and model 2 (ols_sci) does not control for spi -- this is okay. SPI is the predecessor of the SCI, sharing/data overlap, and so it is expected to have significant statistical correlation (multicollinearity). This is likely what explains the significant reduction of coefficients (from 0.280 to 0.1216 for SPI, and from 0.268 to 0.1288 for SCI) indicating that they're both capturing much of the same underlying relationship with SDG status.

# AIC/BIC Checking Fit [FIX TEST- #N DIFFERS BTW MODELS]

```{r}
# Compare all three models with AIC
#AIC(ols_spi, ols_sci, ols_multiple)

# Compare all three models with BIC
#BIC(ols_spi, ols_sci, ols_multiple)
```

**Best fit:** ols_spi (Adj Rsq: 0.7696) (AIC/BIC: ___)

# SPI & SCI colinearity VIF

```{r}
# Check correlation between SPI and SCI
cor(x_var_spi, x_var_sci, use = "complete.obs")

# Check VIF (Variance Inflation Factor)
vif(ols_multiple)

```

**colinearity:** there is significant co-linearity between SCI and SPI, with a correlation of 0.8277. Upon integrating within the same model, SCI inflated the standard error of SPI from ___ to ___ SCI had a similar reaction fro the SPI with its standard error increasing from ___ to ___.

**VIF:** Such multicolinearity is reflected by the VIF test which accounts for all x variables in the model instead of just the two measures of statistical capacity (SCI & SPI).

x_var_spi: 3.34 
x_var_sci: 3.21 
log(gdppc): 1.27
population: 1.02

Unsurprisingly, the variance of the SPI coefficient is inflated by a factor of 3.34 due to correlation with other predictors. Similarly, the SCI coefficient’s variance is inflated by 3.21 times. However, it is acceptable to use in the same model as it will not severely impact estimates given both factors are less than 5.0.

**Selecting model:** My research question is about overall statistical capacity rather than comparing different measures. The SPI model reveals a better fit than the SCI model (Adj Rsq: 0.7696 > 0.7235). It is also a slightly better fit compared to the multiple OLS model containing both SPI and SCI (Adj Rsq: 0.7696 > 0.7624).

```{r}
#visualize differences in fit
library(ggplot2)
library(plotly)

#define regression line colors
spi_line <- "steelblue4"
sci_line <- "darkgoldenrod"

# Creating scatterplot with both SPI and SCI on the same plot
Compare_fit <- ggplot(merged, aes(x = spi_comp, y = sdg_overall))+
  geom_smooth(aes(x = spi_comp, y = sdg_overall), color = spi_line, method = "lm", linewidth = 0.75, se = FALSE)+ # Regression line SPI
  geom_smooth(aes(x = sci_overall, y = sdg_overall), color = sci_line, method = "lm", linewidth = 0.75, se = FALSE)+ # Regression line for SCI
  geom_point(aes(color = "spi_comp"), alpha=0.50, size = 0.5)+  # Scatter plot for SPI, colored distinctly
  geom_point(aes(x = sci_overall, y = sdg_overall, color = "sci_overall"), alpha=0.5, size = 0.5)+ # Add SCI points w/different color
  scale_color_manual(values = c("spi_comp" = "steelblue1", "sci_overall" = "darkgoldenrod1")) +  # Define colors
  labs(title = "SDG vs. SPI and SCI",
       x = "SPI / SCI",
       y = "SDG",
       color = "Index") +  # Title for legend
  theme_bw() # Optional: adds a clean, black and white theme

#make interactive
ggplotly(Compare_fit)

# Save to specific folder
# ggsave("~/Documents/GitHub/QMSS_Thesis_Sanchez/Output_CSVs/fd_plot.png", p, width = 10, height = 6)
```

#Combine to single index: principle component analysis (FOR FUTURE)

```{r}
# Standardize both measures
#spi_z <- scale(x_var_spi)
#sci_z <- scale(x_var_sci)

# Create composite (simple average)
#stat_capacity_index <- (spi_z + sci_z)/2

#extract common variance 
#pca_result <- prcomp(cbind(x_var_spi, x_var_sci), scale = TRUE)
#stat_capacity_pc1 <- pca_result$x[,1]  # First principal component
```

#Interactions 
```{r}
#interaction: does GNI Classification (income_level) affect the relationzhip between x (spi) & y (sdg)?
inc_lev_interaction <- lm(sdg_overall ~ spi_comp + spi_comp*income_level + log(gdppc) + population,
                        data = merged)
summary(inc_lev_interaction)

#interaction: does regime_type_2 affect the relationship between x (spi) & y (sdg)?
reg_type2_interaction <- lm(sdg_overall ~ spi_comp + spi_comp*regime_type_2 + log(gdppc) + population,
                        data = merged)
summary(reg_type2_interaction)

#interaction: does regime_type_4 affect the relationship between x (spi) & y (sdg)?
reg_type4_interaction <- lm(sdg_overall ~ spi_comp + spi_comp*regime_type_4 + log(gdppc) + population,
                        data = merged)
summary(reg_type4_interaction)

#interaction: does di affect the relationship between x (spi) & y (sdg)?
reg_type_di_interaction <- lm(sdg_overall ~ spi_comp + spi_comp*di_score + log(gdppc) + population,
                        data = merged)
summary(reg_type_di_interaction)
```
*GNI Income Classification: Yes* there are statistically significant interactions found from GNI classifications that affects the relationship between spi and sdgs

*Binary Regime Type: No* there is no statistically significant interactions found from regime type (autocracy vs democracy) that affects the relationship between spi and sdgs. 

*Categorical Regime type (4 options): No* there is no statistically significant interactions found from regime type (Closed autocracy, electoral autocracy, electoral democracy, liberal democracy) that affects the relationship between spi and sdgs. 

*Continuous di_score [0-1] Regime type: No* there is no statistically significant interactions found from regime type (infinate between 0-1) that affects the relationship between spi and sdgs. 


# SUBGROUP ANALYSIS BY WB GNI CLASSIFICATIONS 
**Disaggregated by Development Status:** Test models on low-income vs. middle-income countries to isolate resource-dependent mechanisms. *Make 4 regression models and then put them all together in a table to compare the slopes and R-sq values.*

##WB GNI income categories: income_level (income classification: "H", "UM", "LM", "L")
```{r}
# 1. Overall model (all countries)
overall_lm <- lm(sdg_overall ~ spi_comp + di_score + log(gdppc) + population,
                        data = merged)
summary(overall_lm)

# 2. High income countries
high_inc_lm <- lm(sdg_overall ~ spi_comp + di_score + log(gdppc) + population,
                        data = merged %>% 
                          filter(income_level == "H"))
summary(high_inc_lm)

# 3. Upper-middle income countries
upper_mid_lm <- lm(sdg_overall ~ spi_comp + di_score + log(gdppc) + population,
                         data = merged %>% 
                           filter(income_level == "UM"))
summary(upper_mid_lm)

# 4. Lower-middle income countries
lower_mid_lm <- lm(sdg_overall ~ spi_comp + di_score + log(gdppc) + population,
                         data = merged %>% 
                           filter(income_level == "LM"))
summary(lower_mid_lm)

# 5. Low income countries
low_inc_lm <- lm(sdg_overall ~ spi_comp + di_score + log(gdppc) + population,
                       data = merged %>% 
                         filter(income_level == "L"))
summary(low_inc_lm)
```

### Extract coefficients and statistics for Multiple Linear Regression Models to compare 
```{r}
# Creating mod_compare_tab - comparison table
mod_compare_tab <- data.frame(
  model = c("Overall", "High Income", "Upper-Middle Income", 
           "Lower-Middle Income", "Low Income"),
  coefficient = c(
    coef(overall_lm)["spi_comp"],
    coef(high_inc_lm)["spi_comp"],
    coef(upper_mid_lm)["spi_comp"],
    coef(lower_mid_lm)["spi_comp"],
    coef(low_inc_lm)["spi_comp"]
  ),
  
  intercept = c(
    coef(overall_lm)["(Intercept)"],
    coef(high_inc_lm)["(Intercept)"],
    coef(upper_mid_lm)["(Intercept)"],
    coef(lower_mid_lm)["(Intercept)"],
    coef(low_inc_lm)["(Intercept)"]
  ),
  
  Std_error = c(
  summary(overall_lm)$coefficients["spi_comp", 'Std. Error'], # just standard errors 
  summary(high_inc_lm)$coefficients["spi_comp", 'Std. Error'], # just standard errors 
  summary(upper_mid_lm)$coefficients["spi_comp", 'Std. Error'], # just standard errors 
  summary(lower_mid_lm)$coefficients["spi_comp", 'Std. Error'], # just standard errors 
  summary(low_inc_lm)$coefficients["spi_comp", 'Std. Error'] # just standard errors 
  ),

  t_value = c(
    summary(overall_lm)$coefficients["spi_comp", "t value"],
    summary(high_inc_lm)$coefficients["spi_comp", "t value"],
    summary(upper_mid_lm)$coefficients["spi_comp", "t value"],
    summary(lower_mid_lm)$coefficients["spi_comp", "t value"],
    summary(low_inc_lm)$coefficients["spi_comp", "t value"]
  ),
  
  p_value = c(
    summary(overall_lm)$coefficients["spi_comp", "Pr(>|t|)"],
    summary(high_inc_lm)$coefficients["spi_comp", "Pr(>|t|)"],
    summary(upper_mid_lm)$coefficients["spi_comp", "Pr(>|t|)"],
    summary(lower_mid_lm)$coefficients["spi_comp", "Pr(>|t|)"],
    summary(low_inc_lm)$coefficients["spi_comp", "Pr(>|t|)"]
  ),
  
  r_squared = c(
    summary(overall_lm)$r.squared,
    summary(high_inc_lm)$r.squared,
    summary(upper_mid_lm)$r.squared,
    summary(lower_mid_lm)$r.squared,
    summary(low_inc_lm)$r.squared
  ),
  
  adj_r_squared = c(
    summary(overall_lm)$adj.r.squared,
    summary(high_inc_lm)$adj.r.squared,
    summary(upper_mid_lm)$adj.r.squared,
    summary(lower_mid_lm)$adj.r.squared,
    summary(low_inc_lm)$adj.r.squared
  ),
  
  n_obs = c(
    nobs(overall_lm),
    nobs(high_inc_lm),
    nobs(upper_mid_lm),
    nobs(lower_mid_lm),
    nobs(low_inc_lm)
  )
)

# Defining function for significance stars based on p-values
sig_stars <- function(p_value) {
  if (p_value <= 0.001) {
    return("***")
  } else if (p_value <= 0.01) {
    return("**")
  } else if (p_value <= 0.05) {
    return("*")
  } else if (p_value <= 0.1) {
    return(".")
  } else {
    return("")
  }
}

# make table and round for better display
mod_compare_tab <- mod_compare_tab %>%
  mutate(
    coefficient = round(coefficient, 3),
    intercept = round(intercept, 3),
    Std_error = round(Std_error, 3),
    t_value = round(t_value, 3),
    p_value = p_value,
    significance = sapply(p_value, sig_stars), #significance stars 
    r_squared = round(r_squared, 3),
    adj_r_squared = round(adj_r_squared, 3)
  )

# Print the comparison table
print(mod_compare_tab)

# export mod_compare_tab 
#write.csv(mod_compare_tab, file = "ols_model_comparison.csv", row.names=F) 
```


##Visualizing Slopes: plotting
```{r}
viz_gni_class <- ggplot(data = merged, aes(x = spi_comp, y = sdg_overall, color = income_level_lab)) +
  geom_point(alpha = 0.25, size = 0.75) + 
  geom_smooth(method = "lm", linewidth = 0.65, se = FALSE) +
  scale_color_manual(
    values = c("High Income Countries" = "#1D6A96", 
               "Upper-Middle Income Countries" = "#4CB5AE",
               "Lower-Middle Income Countries" = "#F3A738",
               "Low Income Countries" = "#C02942")
  ) +
  labs(title = "Relationship between SPI and SDG by World Bank Income Classification",
       x = "Statistical Performance Indicators (SPI)",
       y = "Sustainable Development Goals (SDG)",
       color = "Income Classification") +
  theme_bw()

ggplotly(viz_gni_class)

# Save to specific folder
# ggsave("~/Documents/GitHub/QMSS_Thesis_Sanchez/Output_CSVs/fd_plot.png", p, width = 10, height = 6)
```


# Mediation analysis 
To test if SPI mediates the relationship between regime type and SDG outcomes: Democratic backsliding → reduces SPI → slows SDG progress  
- ACME (Average Causal Mediation Effect): SPI’s indirect effect.
- ADE (Average Direct Effect): Regime type’s direct effect, excluding SPI
```{r}
library(mediation)

#Total Effect: Check if regime type directly affects SDG scores (without SPI)
lm_sdg <- lm(sdg_overall ~ di_score + gdppc + population, data = merged)
summary(lm_sdg)

#Mediator model: Check if regime type affects SPI
lm_spi <- lm(spi_comp ~ di_score + gdppc + population, data = merged)
summary(lm_spi)

#outcome model: Check if SPI affects SDG scores while controlling for regime type
lm_sdg_controlled <- lm(sdg_overall ~ spi_comp + di_score + gdppc + population, data = merged)
summary(lm_sdg_controlled)

#Mediation test: Quantify how much of regime type’s effect on SDGs operates through SPI
med_model <- mediate(lm_spi, lm_sdg_controlled, treat = "di_score", mediator = "spi_comp")
summary(med_model)

```
ACME: SPI’s indirect effect. Because the ACME (indirect effect of SPI on sdg_overall) is significant, SPI mediates the regime-SDG relationship. 

ADE: Regime type’s direct effect, excluding SPI. Because the ADE (the direct effect between di_score on sdg_overall) is also significant, this suggests that SPI does not fully explain the connection between regime type and sdg status. 

# First Difference Models - Robustness check

**PART 1: Create lagged variables and first differences**

```{r}
library(plm)
#library(dplyr)

# Create a simplified data frame with essential variables
fd_fe_data <- merged %>%
  dplyr::select(country_code, year, year_fct, sdg_overall, spi_comp, di_score, income_level, income_level_lab)

# Create lagged variables and first differences with explicit dplyr reference
fd_data <- fd_data %>%
  arrange(country_code, year) %>% 
  group_by(country_code) %>%
  mutate(
    # Create lagged variables with explicit dplyr reference
    sdg_lag = dplyr::lag(sdg_overall, n = 1),
    spi_lag = dplyr::lag(spi_comp, n = 1),
    
    # Calculate first differences
    fd_sdg = sdg_overall - sdg_lag,
    fd_spi = spi_comp - spi_lag
  ) %>%
  ungroup()

# To verify the lag creation works correctly
# head(fd_data %>% select(country_code, year, sdg_overall, sdg_lag, fd_sdg, spi_comp, spi_lag, fd_spi) %>% arrange(country_code, year), 10)
```

**PART 2: Run FD models for each income group**
```{r}
# Overall model
fd_overall <- lm(fd_sdg ~ fd_spi, 
                data = fd_data %>% 
                  filter(!is.na(fd_sdg) & !is.na(fd_spi)))
summary(fd_overall)

# High income countries
fd_high <- lm(fd_sdg ~ fd_spi, 
             data = fd_data %>% 
               filter(!is.na(fd_sdg) & !is.na(fd_spi) & 
                     income_level_lab == "High Income Countries"))
summary(fd_high)

# Upper-middle income countries
fd_uppermid <- lm(fd_sdg ~ fd_spi, 
              data = fd_data %>% 
                filter(!is.na(fd_sdg) & !is.na(fd_spi) & 
                      income_level_lab == "Upper-Middle Income Countries"))
summary(fd_uppermid)

# Lower-middle income countries
fd_lowermid <- lm(fd_sdg ~ fd_spi, 
              data = fd_data %>% 
                filter(!is.na(fd_sdg) & !is.na(fd_spi) & 
                      income_level_lab == "Lower-Middle Income Countries"))
summary(fd_lowermid)

# Low income countries
fd_low <- lm(fd_sdg ~ fd_spi, 
            data = fd_data %>% 
              filter(!is.na(fd_sdg) & !is.na(fd_spi) & 
                    income_level_lab == "Low Income Countries"))
summary(fd_low)
```

**PART 3: Create a comparison table**

```{r}
# Extract coefficients and statistics from each model
fd_comparison <- data.frame(
  model = c("Overall", "High Income", "Upper-Middle Income", 
           "Lower-Middle Income", "Low Income"),
  coefficient = c(
    coef(fd_overall)["fd_spi"],
    coef(fd_high)["fd_spi"],
    coef(fd_upper)["fd_spi"],
    coef(fd_lower)["fd_spi"],
    coef(fd_low)["fd_spi"]
  ),
  intercept = c(
    coef(fd_overall)["(Intercept)"],
    coef(fd_high)["(Intercept)"],
    coef(fd_upper)["(Intercept)"],
    coef(fd_lower)["(Intercept)"],
    coef(fd_low)["(Intercept)"]
  ),
  t_value = c(
    summary(fd_overall)$coefficients["fd_spi", "t value"],
    summary(fd_high)$coefficients["fd_spi", "t value"],
    summary(fd_upper)$coefficients["fd_spi", "t value"],
    summary(fd_lower)$coefficients["fd_spi", "t value"],
    summary(fd_low)$coefficients["fd_spi", "t value"]
  ),
  p_value = c(
    summary(fd_overall)$coefficients["fd_spi", "Pr(>|t|)"],
    summary(fd_high)$coefficients["fd_spi", "Pr(>|t|)"],
    summary(fd_upper)$coefficients["fd_spi", "Pr(>|t|)"],
    summary(fd_lower)$coefficients["fd_spi", "Pr(>|t|)"],
    summary(fd_low)$coefficients["fd_spi", "Pr(>|t|)"]
  ),
  r_squared = c(
    summary(fd_overall)$r.squared,
    summary(fd_high)$r.squared,
    summary(fd_upper)$r.squared,
    summary(fd_lower)$r.squared,
    summary(fd_low)$r.squared
  ),
  adj_r_squared = c(
    summary(fd_overall)$adj.r.squared,
    summary(fd_high)$adj.r.squared,
    summary(fd_upper)$adj.r.squared,
    summary(fd_lower)$adj.r.squared,
    summary(fd_low)$adj.r.squared
  ),
  n_obs = c(
    nobs(fd_overall),
    nobs(fd_high),
    nobs(fd_upper),
    nobs(fd_lower),
    nobs(fd_low)
  )
)

# Format the numbers for better display
fd_comparison <- fd_comparison %>%
  mutate(
    coefficient = round(coefficient, 3),
    intercept = round(intercept, 3),
    t_value = round(t_value, 3),
    p_value = round(p_value, 4),
    r_squared = round(r_squared, 3),
    adj_r_squared = round(adj_r_squared, 3)
  )

# Print the comparison table
print(fd_comparison)

#export fd_comparison 
#write.csv(fd_comparison, file = "fd_comparison.csv", row.names=F) 
```

**RESULTS:**

The overall model shows a positive and statistically significant relationship (coefficient = 0.015, p = 0.0033) between improvements in statistical capacity and improvements in SDG performance. Although this confirms my hypothesis that better statistical systems contribute to better development outcomes, the intercept indicates that when SPI remains constant, SDG scores progress by 0.286 anyway.

The income group breakdown reveals fascinating variations:

1.  **High Income Countries** show a negative coefficient (-0.012) that is not statistically significant (p = 0.2537). This suggests that for developed economies, improvements in statistical capacity don't necessarily translate to SDG improvements, possibly because they already have well-established statistical systems and other factors drive their SDG performance.

2.  **Upper-Middle Income Countries** show a stronger positive effect (0.021) that is statistically significant (p = 0.0163). This indicates that for these emerging economies, investments in statistical capacity yield meaningful SDG improvements.

3.  **Lower-Middle Income Countries** demonstrate an even stronger relationship (0.025) with high statistical significance (p = 0.0088). This is the most significant relationship among all income groups.

4.  **Low Income Countries** show the strongest coefficient (0.026), suggesting the largest potential impact, though with slightly lower statistical significance (p = 0.0459) than lower-middle income countries.

The R-squared values are relatively low across all models (between 0.004 and 0.023), indicating that changes in statistical capacity explain only a small portion of the variation in SDG changes. This is common in first-difference models and suggests that many other factors also influence SDG progress.

1.  The relationship between statistical capacity and sustainable development follows an inverted U-shaped pattern across the development spectrum - minimal impact for high-income countries, strongest for middle-income countries, and slightly less strong (but still significant) for low-income countries.

2.  Statistical capacity improvements appear most beneficial for developing and emerging economies, suggesting that international efforts to strengthen statistical systems should prioritize these countries.

3.  The consistently positive intercepts (ranging from 0.224 to 0.364) indicate that SDG scores tend to improve over time regardless of changes in statistical capacity, though the rate of improvement varies by income level.


# Visual Analysis | FD Models

**Error-bar**

```{r}
# Create dataframe for plotting 
plot_data <- fd_comparison %>%
  select(model, coefficient, p_value) %>%
  mutate(significant = p_value < 0.05)

# Create Horizontal error-bar graph 
fd_errorbar <- ggplot(plot_data, aes(x = coefficient, y = model, color = significant)) +
  geom_point(size = 3) +
  geom_errorbarh(aes(xmin = 0, xmax = coefficient), height = 0.2) + 
  geom_vline(xintercept = 0, linetype = "dashed") +
  scale_color_manual(values = c("gray", "blue")) +
  labs(title = "Effect of SPI Change on SDG Change by Income Group",
       x = "Coefficient (Effect Size)",
       y = "") +
  theme_bw() +
  theme(legend.position = "none")

fd_errorbar

#export fd_errorbar
# ggsave("fd_errorbar.png", width = 10, height = 6)
```

**Scatter plot of FDs**

```{r}
fd_scatter <- ggplot(fd_data %>% 
         filter(!is.na(fd_sdg) & !is.na(fd_spi)), 
       aes(x = fd_spi, y = fd_sdg, color = income_level)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", se = FALSE) +
  facet_wrap(~income_level) +
  labs(title = "Relationship between Changes in SPI and SDG by Income Group",
       x = "Change in Statistical Performance (SPI)",
       y = "Change in Sustainable Development (SDG)") +
  theme_bw()

fd_scatter

#export fd_scatter
# ggsave("iCloud_Drive/Documents/GitHub/QMSS_Thesis_Sanchez", width = 10, height = 6)
```

## Multiple Regression: Xs = p1-5

```{r}
# y & x vars + controls
sdg_overall_score <- merged$sdg_overall 
p_1 <- merged$p1_use #insert x var 2
p_2 <- merged$p2_services #insert x var 3
p_3 <- merged$p3_products #insert x var 3
p_4 <- merged$p4_sources #insert x var 3
p_5 <- merged$p5_infra #insert x var 3
gdppc <- merged$gdppc

df_spi_p1to5 <- data.frame(p_1, p_2, p_3, p_4, p_5, gdppc)

#OLS lm & full summary table 
lm2 <- lm(sdg_overall_score ~ p_1 + p_2 + p_3 + p_4 + p_5 + log(gdppc), data=df_spi_p1to5)
summary(lm2)

#plotting residuals
plot(lm2)
#plotting scatterplot 
plot(df$y_var, df$x1_var)
abline(lm2)
```

##interaction term and control
```{r}
#incorperating regime type as control
ols_testing <- lm(sdg_overall ~ spi_comp + regime_type + log(gdppc) + population,
                        data = merged_inc)
summary(ols_testing)

#incorperating regime type as interaction term & control
ols_testing2 <- lm(sdg_overall ~ spi_comp + regime_type + regime_type*spi_comp + log(gdppc) + population,
                        data = merged_inc)
summary(ols_testing2)
```

## R Stan, GLM, Markove  
```{r}
library(rstanarm)

gml_testing <- stan_glm(sdg_overall ~ spi_comp + regime_type + regime_type*spi_comp + log(gdppc) + population, data = merged_inc)

# glm_testing
print(gml_testing)
summary(gml_testing)
```

## Fixed Effects Model
```{r}
library(plm)
fe_model <- plm(sdg_overall ~ spi_comp + regime_type + gdppc + population,
                data = merged_inc, index = c("country_code", "year"), model = "within")

summary(fe_model)
```


#Correlation matrix (look it up!)
```{r}
# 5 spi pillars by 17 and sdgs 

```


## Testing Index Reliability (Cronbach's Alpha)

```{r}

```
