---
title: "First Tests OLS, FDs"
author: "Sevastian Sanchez"
date: "`r Sys.Date()`"
output: html_document
---
# All Countries, Preliminary Analysis (SPI x SDGs)
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Libraries 
```{r}
library(devtools)
library(tidyr)
library(dplyr)
library(vdemdata) # call vdem package 
library(readr) #loading data 
library(readxl)
library(stargazer)
library(broom)
```


##CLEANING / WRANGLING 
```{r}
#Vdem package from github 
vdem <- vdemdata::vdem  #loading data 

#spi csv from github 
url <- "https://raw.githubusercontent.com/worldbank/SPI/refs/heads/master/03_output_data/SPI_index.csv"
spi <- read_csv(url)

#sdg excel from directory 
sdg <- read_excel("~/Desktop/OneDrive/OneDrive/QMSS Thesis/Datasets/SDR2024-data.xlsx", sheet = "Backdated SDG Index")

#sci csv from directory 
sci_df <- read_csv("~/Desktop/OneDrive/OneDrive/QMSS Thesis/Datasets/SCI_All_Dim_TS.csv")

#Lab 3 merged 
# merged_L3 <- read_csv("~/Documents/GitHub/Timeseries-SDGs-SPI-Vars/merged_L3.csv")
```

##Vdem Wrangling / Cleaning
```{r}
#selecting democracy indices and controls 
vdem <- vdem %>%
  dplyr::select(country_name, country_text_id, year, v2x_regime, v2x_regime_amb, v2x_polyarchy, v2x_libdem, v2x_partipdem, v2x_delibdem, v2x_egaldem, v2xel_frefair, v2x_accountability, v2x_veracc, v2x_horacc, v2x_diagacc, v2xca_academ, v2x_freexp_altinf, e_gdp, e_gdppc, e_wb_pop, v3ststybcov, v3ststybpub, v3stcensus, v3ststatag) %>%
  filter(year > 1999) %>% #Only 1999 and up 
  rename(country_code = country_text_id, #renaming country code (new_name = old_name)
         regime_type = v2x_regime, 
         regime_type_amb = v2x_regime_amb,
         elect_dem = v2x_polyarchy,
         lib_dem = v2x_libdem,
         part_dem = v2x_partipdem,
         delib_dem = v2x_delibdem,
         egal_dem = v2x_egaldem,
         freefair = v2xel_frefair,
         #freefair_ord = e_v2xel_frefair,
         accountability = v2x_accountability,
         vt_account = v2x_veracc,
         hz_account = v2x_horacc,
         diag_account = v2x_diagacc,
         academ_free = v2xca_academ,
         freexp_altinfo = v2x_freexp_altinf,
         gdp = e_gdp,
         gdp_pc = e_gdppc,
         population = e_wb_pop,
         stat_agency = v3ststatag,
         stat_yb_cov = v3ststybcov,
         stat_yb_pub = v3ststybpub,
         census = v3stcensus,
         )
  
# To be added: e_v2xel_frefair,

#vdem$v2xca_academ <- vdem$v2xca_academ*100
#vdem$v2x_libdem <- vdem$v2x_libdem*100
```


## SPI wrangling / cleaning 
```{r}
#selecting and renaming spi vars 
spi <- spi %>% 
  select(country, iso3c, date, SPI.INDEX, SPI.INDEX.PIL1, SPI.INDEX.PIL2, SPI.INDEX.PIL3, SPI.INDEX.PIL4, SPI.INDEX.PIL5) %>% 
  rename(country_name = country) %>% 
  rename(country_code = iso3c) %>% 
  rename(year = date) %>% 
  rename(spi_comp = SPI.INDEX) %>% #SPI composite score, average of p1-5
  rename(p1_use = SPI.INDEX.PIL1) %>% #SPI pillar 1, Data use 
  rename(p2_services = SPI.INDEX.PIL2) %>% #SPI pillar, 2 Data services 
  rename(p3_products = SPI.INDEX.PIL3) %>% #SPI pillar, 3 Data products  
  rename(p4_sources = SPI.INDEX.PIL4) %>% #SPI pillar, 4 Data sources 
  rename(p5_infra = SPI.INDEX.PIL5) #SPI pillar 5, Data infrastructure
```

## SCI wrangling / cleaning 
```{r}
#full dataset 
sci <- sci_df %>% 
  dplyr::select(country_name, country_code, Year, IQ.SCI.OVRL, IQ.SCI.MTHD, IQ.SCI.PRDC, IQ.SCI.SRCE, everything()) %>% 
  rename(year = Year,
         sci_overall = IQ.SCI.OVRL, 
         sci_method = IQ.SCI.MTHD, 
         sci_periodicity = IQ.SCI.PRDC,
         sci_source = IQ.SCI.SRCE)

sci$year <- as.numeric(sci$year)

#for merging (only sci score variables selected)
sci_scores <- sci %>% 
  select(country_name, country_code, year, sci_overall, sci_method, sci_periodicity, sci_source)
```

##SDG wrangling / cleaning 
```{r}
#select sdg goals & composite 
sdg <- sdg %>% 
  dplyr::select("country_name", "country_code", "year", "sdg_overall", "goal1", "goal2", "goal3", "goal4", "goal5", "goal6", "goal7", "goal8", "goal9", "goal10", "goal11", "goal12", "goal3", "goal14", "goal15", "goal16", "goal17")

```

# Merging & subsetting (case study: Spain)
```{r}
#merge spi, sci, sdg and vdem data 
merged <- left_join(sdg, spi, by = c("year", "country_code")) 
merged <- left_join(merged, sci_scores, by = c("country_code", "year"))
merged <- left_join(merged, vdem, by = c("country_code", "year"))

#rm duplicate col names 
colnames(merged) <- make.unique(colnames(merged)) # Make column names unique

#data type changes 
merged$year <- as.integer(merged$year)
merged <- merged %>% 
  mutate_at(c("spi_comp", "p1_use", "p2_services", "p3_products", "p4_sources", "p5_infra", "sci_overall", "sci_method", "sci_periodicity", "sci_source"), as.numeric)

#exporting time series data AS: merged_ts.csv ; will save to current working directory 
write.table(merged, file = "merged_ts.csv", row.names=F, sep = ",") 
```

## Wrangling for Spain - Test Subject 
```{r}
# subset for Spain 
merged_esp_ts <- merged[merged$country_code=="ESP",]

#exporting Spain time series AS: merged_esp_ts; will save to current working directory 
write.table(merged_esp_ts, file = "merged_esp_ts.csv", row.names=F, sep = ",")
```

##Cross-Sectional Analysis - 2019 data only
```{r}
#subset for only 2019 
merged_cs_2019 <- merged[merged$year==2019,]

#exporting cross-sectional 2019 data AS: merged_cs2019; will save to current working directory 
write.table(merged_cs_2019, file = "merged_cs2019.csv", row.names=F, sep = ",")
```

## NAIVE OLS (BI-VARIATE)
**Aggregated SPI & SDG Scores**
*H0: Null, there is no relationship*
*H1: there is a statistically significant correlation between overall SPI and SDG composite scores*
```{r}
# x & y variables 
y_var_sdg <- merged$sdg_overall 
x_var_spi <- merged$spi_comp
x_var_sci <- merged$sci_overall
controls <- merged %>% select(gdp, gdp_pc, population)
df_sdg_statcap <- data.frame(y_var_sdg, x_var_spi, x_var_sci)

# how many na's?
colSums(is.na(df_sdg_statcap))

# omit na's 
df_sdg_spi <- na.omit(data.frame(y_var_sdg, x_var_spi))
df_sdg_sci <- na.omit(data.frame(y_var_sdg, x_var_sci))
dim(df_sdg_spi) #spi observations, ommitted na's 
dim(df_sdg_sci) #sci observations, ommitted na's 

#correlation coefficient (x, y)
correlation_sdg_spi <- cor(df_sdg_spi$y_var_sdg, df_sdg_spi$x_var_spi)^2 #x-var 1 = spi
correlation_sdg_sci <- cor(df_sdg_sci$y_var_sdg, df_sdg_sci$x_var_sci)^2 #x-var 2 = sci
string_corcoef <- "Correlation coefficient:" 
paste(string_corcoef, correlation_sdg_spi) # pasting result 
paste(string_corcoef, correlation_sdg_sci) # pasting result 
```
Number of observations in SPI: 1300; number of NA's 3068 (not all years captured)
Number of observations in SCI: 2055; number of NA's 2313 (not all countries captured)

# Comparing SPI & SCI to identify best model 
```{r}
# 1. OLS for SPI and SDG - Overall 
ols_spi <- lm(y_var_sdg ~ x_var_spi + merged$gdp + merged$gdp_pc + merged$population)
summary(ols_spi)

# 2. OLS for SCI and SDG - Overall 
ols_sci <- lm(y_var_sdg ~ x_var_sci + merged$gdp + merged$gdp_pc + merged$population)
summary(ols_sci)

# 3. Multiple Regression with both SPI and SCI
ols_multiple <- lm(y_var_sdg ~ x_var_spi + x_var_sci + merged$gdp + merged$gdp_pc + merged$population)
summary(ols_multiple)

#visualize differences in fit
library(ggplot2)
library(plotly)

#define regression line colors
spi_line <- "steelblue4"
sci_line <- "darkgoldenrod"

# Creating scatterplot with both SPI and SCI on the same plot
Compare_fit <- ggplot(merged, aes(x = spi_comp, y = sdg_overall))+
  geom_smooth(aes(x = spi_comp, y = sdg_overall), color = spi_line, method = "lm", se = FALSE)+ # Regression line SPI
  geom_smooth(aes(x = sci_overall, y = sdg_overall), color = sci_line, method = "lm", se = FALSE)+ # Regression line for SCI
  geom_point(aes(color = "spi_comp"), alpha=0.50, size = 0.5)+  # Scatter plot for SPI, colored distinctly
  geom_point(aes(x = sci_overall, y = sdg_overall, color = "sci_overall"), alpha=0.5, size = 0.5)+ # Add SCI points w/different color
  scale_color_manual(values = c("spi_comp" = "steelblue1", "sci_overall" = "darkgoldenrod1")) +  # Define colors
  labs(title = "SDG vs. SPI and SCI",
       x = "SPI / SCI",
       y = "SDG",
       color = "Index") +  # Title for legend
  theme_bw() # Optional: adds a clean, black and white theme

#make interactive
ggplotly(Compare_fit)
```
We reject the null hypothesis that there is no relationship between SPI and SDG composite scores. Holding all else constant, there is a positive moderate relationship between Statistical Performance (SPI) and SDG status. 


## MULTIPLE LINEAR REGRESSION 
**Disaggregated by Development Status:**  
*Make 4 regression models and then put them all together in a table compare the slopes.* 
```{r}
#WB GNI income categories 
income_class <- read_csv("~/Documents/GitHub/QMSS_Thesis_Sanchez/Output_CSVs /world_bank_income_classifications.csv")

# New df with income classifications 
merged_inc <- merged %>% 
  left_join(income_class, by = c("country_code", "year"))

# Assuming your merged dataset contains:
# X = spi_comp (SPI values)
# Y = sdg_overall (SDG values)
# Subset: income_level (income classification: "H", "UM", "LM", "L")

# 1. Overall model (all countries)
overall_lm <- lm(sdg_overall ~ spi_comp + gdp + gdp_pc + population,
                        data = merged_inc)
summary(overall_lm)

# 2. High income countries
high_inc_lm <- lm(sdg_overall ~ spi_comp + gdp + gdp_pc + population,
                        data = merged_inc %>% 
                          filter(income_level == "H"))
summary(high_inc_lm)

# 3. Upper-middle income countries
upper_mid_lm <- lm(sdg_overall ~ spi_comp + gdp + gdp_pc + population,
                         data = merged_inc %>% 
                           filter(income_level == "UM"))
summary(upper_mid_lm)

# 4. Lower-middle income countries
lower_mid_lm <- lm(sdg_overall ~ spi_comp + gdp + gdp_pc + population,
                         data = merged_inc %>% 
                           filter(income_level == "LM"))
summary(lower_mid_lm)

# 5. Low income countries
low_inc_lm <- lm(sdg_overall ~ spi_comp + gdp + gdp_pc + population,
                       data = merged_inc %>% 
                         filter(income_level == "L"))
summary(low_inc_lm)

# Extract coefficients and create a comparison table
model_comparison <- bind_rows(
  tidy(overall_lm) %>% mutate(model = "Overall"),
  tidy(high_inc_lm) %>% mutate(model = "High Income"),
  tidy(upper_mid_lm) %>% mutate(model = "Upper-Middle Income"),
  tidy(lower_mid_lm) %>% mutate(model = "Lower-Middle Income"),
  tidy(low_inc_lm) %>% mutate(model = "Low Income")
) %>%
  select(model, term, estimate, std.error, p.value) %>% 
  filter(term=='spi_comp')

# Print the comparison table
print(model_comparison)
```

**Vizualize the slopes**
```{r}
# Create a visualization with different regression lines by income group
by_income_class <- ggplot(merged_inc, aes(x = spi_comp, y = sdg_overall, color = income_level)) +
  geom_point(alpha = 0.2) +
  geom_smooth(method = "lm", linewidth = 0.65, se = FALSE) +
  scale_color_manual(values = c("H" = "#1D6A96", 
                                "UM" = "#4CB5AE", 
                                "LM" = "#F3A738",
                                "L" = "#C02942"),
                    labels = c("High Income", 
                              "Upper-Middle Income", 
                              "Lower-Middle Income", 
                              "Low Income")) +
  labs(title = "Relationship between SPI and SDG by World Bank Income Classification",
       x = "Statistical Performance Indicators (SPI)",
       y = "Sustainable Development Goals (SDG)",
       color = "Income Classification") +
  theme_bw()

ggplotly(by_income_class)
```


## NAIVE OLS W/Controls 
```{r}
# y & x vars + controls/
y_var <- merged$sdg_overall 
x1_var <- merged$spi_comp #insert x var 1
x2_var <- merged$p1_use #insert x var 2
x3_var <- merged$p2_services #insert x var 3
x4_var <- merged$p3_products #insert x var 3
x5_var <- merged$p4_sources #insert x var 3
x6_var <- merged$p5_infra #insert x var 3
df_2 <- data.frame(y_var, x1_var, x2_var, x3_var, x4_var, x5_var, x6_var)

#OLS lm & full summary table 
lm2 <- lm(y_var ~ x1_var + x2_var + x3_var + x4_var + x5_var + x6_var, data=df_2)
summary(lm2)
#plotting residuals
plot(lm2)
#plotting scatterplot 
plot(df$y_var, df$x1_var)
abline(lm2)
```

## Identify Regressing Countries 
```{r}
# 2015-present 






# 2000-present 


```




## First Difference Model 

```{r}


```


## Testing Index Reliability (Cronbach's Alpha)
```{r}

```




