---
title: "First Tests OLS, FDs"
author: "Sevastian Sanchez"
date: "`r Sys.Date()`"
output: html_document
---
# All Countries, Preliminary Analysis (SPI x SDGs)
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Libraries 
```{r}
library(devtools)
library(tidyr)
library(dplyr)
library(vdemdata) # call vdem package 
library(readr) #loading data 
library(readxl)
library(stargazer)
```


##CLEANING / WRANGLING 
```{r}
#Vdem package from github 
vdem <- vdemdata::vdem  #loading data 

#spi csv from github 
url <- "https://raw.githubusercontent.com/worldbank/SPI/refs/heads/master/03_output_data/SPI_index.csv"
spi <- read_csv(url)

#sdg excel from directory 
sdg <- read_excel("~/Desktop/OneDrive/OneDrive/QMSS Thesis/Datasets/SDR2024-data.xlsx", sheet = "Backdated SDG Index")

#sci csv from directory 
sci_df <- read_csv("~/Desktop/OneDrive/OneDrive/QMSS Thesis/Datasets/SCI_All_Dim_TS.csv")

#Lab 3 merged 
# merged_L3 <- read_csv("~/Documents/GitHub/Timeseries-SDGs-SPI-Vars/merged_L3.csv")
```

##Vdem Wrangling / Cleaning
```{r}
#selecting democracy indices and controls 
vdem <- vdem %>%
  dplyr::select(country_name, country_text_id, year, v2x_regime, v2x_regime_amb, v2x_polyarchy, v2x_libdem, v2x_partipdem, v2x_delibdem, v2x_egaldem, v2xel_frefair, v2x_accountability, v2x_veracc, v2x_horacc, v2x_diagacc, v2xca_academ, v2x_freexp_altinf, e_gdp, e_gdppc, e_wb_pop, v3ststybcov, v3ststybpub, v3stcensus, v3ststatag) %>%
  filter(year > 1999) %>% #Only 1999 and up 
  rename(country_code = country_text_id, #renaming country code (new_name = old_name)
         regime_type = v2x_regime, 
         regime_type_amb = v2x_regime_amb,
         elect_dem = v2x_polyarchy,
         lib_dem = v2x_libdem,
         part_dem = v2x_partipdem,
         delib_dem = v2x_delibdem,
         egal_dem = v2x_egaldem,
         freefair = v2xel_frefair,
         #freefair_ord = e_v2xel_frefair,
         accountability = v2x_accountability,
         vt_account = v2x_veracc,
         hz_account = v2x_horacc,
         diag_account = v2x_diagacc,
         academ_free = v2xca_academ,
         freexp_altinfo = v2x_freexp_altinf,
         gdp = e_gdp,
         gdp_pc = e_gdppc,
         population = e_wb_pop,
         stat_agency = v3ststatag,
         stat_yb_cov = v3ststybcov,
         stat_yb_pub = v3ststybpub,
         census = v3stcensus,
         )
  
# To be added: e_v2xel_frefair,

#vdem$v2xca_academ <- vdem$v2xca_academ*100
#vdem$v2x_libdem <- vdem$v2x_libdem*100
```


## SPI wrangling / cleaning 
```{r}
#selecting and renaming spi vars 
spi <- spi %>% 
  select(country, iso3c, date, SPI.INDEX, SPI.INDEX.PIL1, SPI.INDEX.PIL2, SPI.INDEX.PIL3, SPI.INDEX.PIL4, SPI.INDEX.PIL5) %>% 
  rename(country_name = country) %>% 
  rename(country_code = iso3c) %>% 
  rename(year = date) %>% 
  rename(spi_comp = SPI.INDEX) %>% #SPI composite score, average of p1-5
  rename(p1_use = SPI.INDEX.PIL1) %>% #SPI pillar 1, Data use 
  rename(p2_services = SPI.INDEX.PIL2) %>% #SPI pillar, 2 Data services 
  rename(p3_products = SPI.INDEX.PIL3) %>% #SPI pillar, 3 Data products  
  rename(p4_sources = SPI.INDEX.PIL4) %>% #SPI pillar, 4 Data sources 
  rename(p5_infra = SPI.INDEX.PIL5) #SPI pillar 5, Data infrastructure
```

## SCI wrangling / cleaning 
```{r}
#full dataset 
sci <- sci_df %>% 
  dplyr::select(country_name, country_code, Year, IQ.SCI.OVRL, IQ.SCI.MTHD, IQ.SCI.PRDC, IQ.SCI.SRCE, everything()) %>% 
  rename(year = Year,
         sci_overall = IQ.SCI.OVRL, 
         sci_method = IQ.SCI.MTHD, 
         sci_periodicity = IQ.SCI.PRDC,
         sci_source = IQ.SCI.SRCE)

sci$year <- as.numeric(sci$year)

#for merging (only sci score variables selected)
sci_scores <- sci %>% 
  select(country_name, country_code, year, sci_overall, sci_method, sci_periodicity, sci_source)
```

##SDG wrangling / cleaning 
```{r}
#select sdg goals & composite 
sdg <- sdg %>% 
  dplyr::select("country_name", "country_code", "year", "sdg_overall", "goal1", "goal2", "goal3", "goal4", "goal5", "goal6", "goal7", "goal8", "goal9", "goal10", "goal11", "goal12", "goal3", "goal14", "goal15", "goal16", "goal17")

```

# Merging & subsetting (case study: Spain)
```{r}
#merge spi, sci, sdg and vdem data 
merged <- left_join(sdg, spi, by = c("year", "country_code")) 
merged <- left_join(merged, sci_scores, by = c("country_code", "year"))
merged <- left_join(merged, vdem, by = c("country_code", "year"))

#rm duplicate col names 
colnames(merged) <- make.unique(colnames(merged)) # Make column names unique

#data type changes 
merged$year <- as.integer(merged$year)
merged <- merged %>% 
  mutate_at(c("spi_comp", "p1_use", "p2_services", "p3_products", "p4_sources", "p5_infra", "sci_overall", "sci_method", "sci_periodicity", "sci_source"), as.numeric)

#exporting time series data AS: merged_ts.csv ; will save to current working directory 
write.table(merged, file = "merged_ts.csv", row.names=F, sep = ",") 
```

## Wrangling for Spain - Test Subject 
```{r}
# subset for Spain 
merged_esp_ts <- merged[merged$country_code=="ESP",]

#exporting Spain time series AS: merged_esp_ts; will save to current working directory 
write.table(merged_esp_ts, file = "merged_esp_ts.csv", row.names=F, sep = ",")
```

##Cross-Sectional Analysis - 2019 data only
```{r}
#subset for only 2019 
merged_cs_2019 <- merged[merged$year==2019,]

#exporting cross-sectional 2019 data AS: merged_cs2019; will save to current working directory 
write.table(merged_cs_2019, file = "merged_cs2019.csv", row.names=F, sep = ",")
```

## NAIVE OLS (BI-VARIATE)
**Aggregated SPI & SDG Scores**
*H0: Null, there is no relationship*
*H1: there is a statistically significant correlation between overall SPI and SDG composite scores*
```{r}
# x & y variables 
y_var_sdg <- merged$sdg_overall 
x_var_spi <- merged$spi_comp
x_var_sci <- merged$sci_overall
df_sdg_statcap <- data.frame(y_var_sdg, x_var_spi, x_var_sci)

# how many na's?
colSums(is.na(df_sdg_statcap))

# omit na's 
df_sdg_spi <- na.omit(data.frame(y_var_sdg, x_var_spi))
df_sdg_sci <- na.omit(data.frame(y_var_sdg, x_var_sci))
dim(df_sdg_spi) #spi observations, ommitted na's 
dim(df_sdg_sci) #sci observations, ommitted na's 

#correlation coefficient (x, y)
correlation_sdg_spi <- cor(df_sdg_spi$y_var_sdg, df_sdg_spi$x_var_spi)^2 #x-var 1 = spi
correlation_sdg_sci <- cor(df_sdg_sci$y_var_sdg, df_sdg_sci$x_var_sci)^2 #x-var 2 = sci
string_corcoef <- "Correlation coefficient:" 
paste(string_corcoef, correlation_sdg_spi) # pasting result 
paste(string_corcoef, correlation_sdg_sci) # pasting result 
```


```{r}
#OLS lm & full summary table 
lm1 <- lm(y_var ~ x_var, data=df)
summary(lm1)

#plotting scatter-plot 
plot(df$y_var, df$x_var)
abline(lm1)



# ...
#more possible predictors ...

#paste results 
paste(string_corcoef, sdg_spi_corr) # sdg x spi
paste(string_corcoef, sdg_sci_corr) # sdg x sci

```
We reject the null hypothesis that there is no relationship between SPI and SDG composite scores. There is a 


## ASSESSING MODEL FIT (R-Squared)
```{r}

```



## NAIVE OLS W/Controls 
```{r}
# y & x vars + controls/
y_var <- merged$sdg_overall 
x1_var <- merged$spi_comp #insert x var 1
x2_var <- merged$p1_use #insert x var 2
x3_var <- merged$p2_services #insert x var 3
x4_var <- merged$p3_products #insert x var 3
x5_var <- merged$p4_sources #insert x var 3
x6_var <- merged$p5_infra #insert x var 3
df_2 <- data.frame(y_var, x1_var, x2_var, x3_var, x4_var, x5_var, x6_var)

#OLS lm & full summary table 
lm2 <- lm(y_var ~ x1_var + x2_var + x3_var + x4_var + x5_var + x6_var, data=df_2)
summary(lm2)
#plotting residuals
plot(lm2)
#plotting scatterplot 
plot(df$y_var, df$x1_var)
abline(lm2)
```


## MULTIPLE LINEAR REGRESSION 
**Disaggregated by Development Status**
```{r}

```


## OLS W/INTERACTIONS 
**SPIxDemocracyScore & SDG 
```{r}

```


## Testing Index Reliability (Cronbach's Alpha)
```{r}

```


## First Difference Model 


```{r}


```




