---
title: "First Tests: OLS, Interactions, Subgroups"
author: "Sevastian Sanchez"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
  word_document: default
---

# All Countries, Preliminary Analysis (SPI x SDGs)

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  tidy.opts = list(width.cutoff = 60)
)
```

#FIRST: Libraries, Directory & Data 

```{r}
# set working directory 
setwd("~/Documents/GitHub/QMSS_Thesis_Sanchez")

#load libraries 
source("packages.R")

#load data 
source("data/data_sources.R")
```


#SECOND: run function in r-script: df_years_function.R

```{r}
#df_years: used for extracting data from specified year to now 

#load function
source("df_years()_Function.R")

#specify start year 
merged <- df_years(yr1 = 2000) 
```


##Vdem wrangling / cleaning (skip if df_years() used)
```{r eval=FALSE, include=FALSE}
#selecting democracy indices and controls 
vdem <- vdem %>% 
dplyr::select(country_name, country_text_id, year, v2x_regime, v2x_regime_amb, v2x_polyarchy, v2x_libdem, v2x_partipdem, v2x_delibdem, v2x_egaldem, v2xel_frefair, v2x_accountability, v2x_veracc, v2x_horacc, v2x_diagacc, v2xca_academ, v2x_freexp_altinf, e_gdp, e_gdppc, e_wb_pop, v3ststybcov, v3ststybpub, v3stcensus, v3ststatag) %>%
  filter(year > 1999) %>% #Only 1999 and up 
  rename(country_code = country_text_id, #renaming country code (new_name = old_name)
         regime_type = v2x_regime, 
         regime_type_amb = v2x_regime_amb,
         elect_dem = v2x_polyarchy,
         lib_dem = v2x_libdem,
         part_dem = v2x_partipdem,
         delib_dem = v2x_delibdem,
         egal_dem = v2x_egaldem,
         freefair = v2xel_frefair,
         #freefair_ord = e_v2xel_frefair,
         accountability = v2x_accountability,
         vt_account = v2x_veracc,
         hz_account = v2x_horacc,
         diag_account = v2x_diagacc,
         academ_free = v2xca_academ,
         freexp_altinfo = v2x_freexp_altinf,
         population = e_wb_pop,
         stat_agency = v3ststatag,
         stat_yb_cov = v3ststybcov,
         stat_yb_pub = v3ststybpub,
         census = v3stcensus,
         )
  
# To be added: e_v2xel_frefair
```


## SPI wrangling / cleaning (skip if df_years() used)
```{r eval=FALSE, include=FALSE}
#selecting and renaming spi vars 
spi <- spi %>% 
  select(country, iso3c, date, SPI.INDEX, SPI.INDEX.PIL1, SPI.INDEX.PIL2, SPI.INDEX.PIL3, SPI.INDEX.PIL4, SPI.INDEX.PIL5) %>% 
  rename(country_name = country) %>% 
  rename(country_code = iso3c) %>% 
  rename(year = date) %>% 
  rename(spi_comp = SPI.INDEX) %>% #SPI composite score, average of p1-5
  rename(p1_use = SPI.INDEX.PIL1) %>% #SPI pillar 1, Data use 
  rename(p2_services = SPI.INDEX.PIL2) %>% #SPI pillar, 2 Data services 
  rename(p3_products = SPI.INDEX.PIL3) %>% #SPI pillar, 3 Data products  
  rename(p4_sources = SPI.INDEX.PIL4) %>% #SPI pillar, 4 Data sources 
  rename(p5_infra = SPI.INDEX.PIL5) #SPI pillar 5, Data infrastructure
```



## SCI wrangling / cleaning (skip if df_years() used)

```{r eval=FALSE, include=FALSE}
#full dataset 
sci <- sci_df %>% 
  dplyr::select(country_name, country_code, Year, IQ.SCI.OVRL, IQ.SCI.MTHD, IQ.SCI.PRDC, IQ.SCI.SRCE, everything()) %>% 
  rename(year = Year,
         sci_overall = IQ.SCI.OVRL, 
         sci_method = IQ.SCI.MTHD, 
         sci_periodicity = IQ.SCI.PRDC,
         sci_source = IQ.SCI.SRCE)

#sci$year <- as.numeric(sci$year)

#for merging (only sci score variables selected)
sci_scores <- sci %>% 
  select(country_name, country_code, year, sci_overall, sci_method, sci_periodicity, sci_source)
```



##SDG wrangling / cleaning (skip if df_years() used)

```{r eval=FALSE, include=FALSE}
#select sdg goals & composite 
sdg <- sdg %>% 
  dplyr::select("country_name", "country_code", "year", "sdg_overall", "goal1", "goal2", "goal3", "goal4", "goal5", "goal6", "goal7", "goal8", "goal9", "goal10", "goal11", "goal12", "goal3", "goal14", "goal15", "goal16", "goal17")

```


##GDP_PC wrangling / cleaning (skip if df_years() used)
```{r eval=FALSE, include=FALSE}
gdppc_df <- gdppc_df %>%  #pivot back to long format
  pivot_longer(3:66, #use any way to select all columns (he used range)
               names_to = "year", # new var using cols 2:13 as values
               values_to = "gdp_pc") #new var containing all values within cols 2:13 as values 
View(gdppc_df)
```




## Merging & subsetting (skip if df_years() used)

```{r eval=FALSE, include=FALSE}
#merge spi, sci, sdg and vdem data 
merged <- left_join(sdg, spi, by = c("year", "country_code")) 
merged <- left_join(merged, sci_scores, by = c("country_code", "year"))
merged <- left_join(merged, vdem, by = c("country_code", "year"))

#rm duplicate col names 
colnames(merged) <- make.unique(colnames(merged)) # Make column names unique

#data type changes 
merged$year <- as.integer(merged$year)
merged <- merged %>% 
  mutate_at(c("spi_comp", "p1_use", "p2_services", "p3_products", "p4_sources", "p5_infra", "sci_overall", "sci_method", "sci_periodicity", "sci_source"), as.numeric)

#exporting time series data AS: merged_ts.csv ; will save to current working directory 
#write.table(merged, file = "merged_ts.csv", row.names=F, sep = ",") 
```


## All countries 2019 data only
```{r eval=FALSE, include=FALSE}
#subset for only 2019 
merged_cs_2019 <- merged[merged$year==2019,]

#exporting cross-sectional 2019 data AS: merged_cs2019; will save to current working directory 
#write.table(merged_cs_2019, file = "merged_cs2019.csv", row.names=F, sep = ",")
```






# Set up & missing data
```{r}
# x & y variables 
y_var_sdg <- merged$sdg_overall 
x_var_spi <- merged$spi_comp
x_var_sci <- merged$sci_overall
df_sdg_statcap <- data.frame(y_var_sdg, x_var_spi, x_var_sci)

# how many na's?
colSums(is.na(df_sdg_statcap))
#how many observations 
colSums(!is.na(df_sdg_statcap))

hist(x_var_spi)
hist(x_var_sci)
hist(y_var_sdg)
```

Number of observations in SPI: 1300; number of NA's 3068 (not all years captured) Number of observations in SCI: 2055; number of NA's 2313 (not all countries captured)

# COMPARING SPI & SCI X VARIABLES

**Aggregated SPI & SDG Scores**  
*H0: Null, there is no relationship*  
*H1: there is a statistically significant relationship between overall SPI and SDG composite scores*

```{r}
#correlation coefficients (r-squared), WITHOUT control variables 

#x-var 1 = spi
correlation_sdg_spi <- cor(y_var_sdg, x_var_spi, use = "complete.obs")^2

#x-var 2 = sci
correlation_sdg_sci <- cor(y_var_sdg, x_var_sci, use = "complete.obs")^2 

# pasting result 
string_corcoef <- "Correlation coefficient:" 
paste(string_corcoef, correlation_sdg_spi, "(SPI)", correlation_sdg_sci, "(SCI)")
```

Correlation coefficient/R-sq (SPI): 0.616037202309322  
Correlation coefficient/R-sq (SCI): 0.417965563339242

## Comparing SPI & SCI to identify best model, w/o controls 
Finding estimated impact of variables on SDG status prior to adding controls 
```{r}
ols_spi_naive <- lm(y_var_sdg ~ x_var_spi)
summary(ols_spi_naive)

# 2. OLS for SCI and SDG - Overall 
ols_sci_naive <- lm(y_var_sdg ~ x_var_sci)
summary(ols_sci_naive)

# 3. Multiple Regression with both SPI and SCI
ols_multiple_naive <- lm(y_var_sdg ~ x_var_spi + x_var_sci)
summary(ols_multiple_naive)
```
ols_spi_naive: 0.47806 (p-value < 0.001)  
ols_sci_naive: 0.39209 (p-value < 0.001)  
ols_multiple_naive: spi: 0.28779 (p-value < 0.001); sci: 0.15311 (p-value < 0.001)

The impact of SCI on SDG and SPI on SDG are statistically significant, in all models. SPI appears to have a greater impact on SDGs compared to that of SCI, regardless of the model. All of this is wothout controls. 

## Comparing SPI & SCI to identify best model, WITH controls 

*H0: Null, SCI model > SPI model*  
*H1: SPI model > SCI model*

```{r}
# 1. OLS for SPI and SDG - Overall 
ols_spi <- lm(y_var_sdg ~ x_var_spi + log_gdppc + population, data = merged)
summary(ols_spi)

# 2. OLS for SCI and SDG - Overall 
ols_sci <- lm(y_var_sdg ~ x_var_sci + log_gdppc + population, data = merged)
summary(ols_sci)

# 3. Multiple Regression with both SPI and SCI
ols_multiple <- lm(y_var_sdg ~ x_var_spi + x_var_sci + log_gdppc + population, data = merged)
summary(ols_multiple)
```

We reject the null hypothesis that there is no relationship between SPI and SDG composite scores. Holding all else constant (log gdp per capita and population), there is a positive moderate relationship between Statistical Performance (SPI) and SDG status.

ols_spi: 0.280 (p-value < 0.001)  
ols_sci: 0.268 (p-value < 0.001)  
ols_multiple: spi: 0.1216 (p-value < 0.001); sci: 0.1288 (p-value < 0.001)

Comparing coefficients, SPI has a greater impact on SDG status (0.280) than SCI (0.268). However, in a multiple regression model containing both SPI and SCI, SCI has more of an impact on SDG status (0.1288) (net of SPI) than SPI (0.1216) (net of SCI). Here the coefficients represent the unique impact of each x variable on SDG status, net of all other variables.

Model 1 (ols_spi) does not control for SCI and model 2 (ols_sci) does not control for spi -- this is okay. SPI is the predecessor of the SCI, sharing/data overlap, and so it is expected to have significant statistical correlation (multicollinearity). This is likely what explains the significant reduction of coefficients (from 0.280 to 0.1216 for SPI, and from 0.268 to 0.1288 for SCI) indicating that they're both capturing much of the same underlying relationship with SDG status.

### AIC/BIC Checking Fit [FIX TEST- #N DIFFERS BTW MODELS]

```{r}
# Compare all three models with AIC
#AIC(ols_spi, ols_sci, ols_multiple)

# Compare all three models with BIC
#BIC(ols_spi, ols_sci, ols_multiple)
```

**Best fit:** ols_spi (Adj Rsq: 0.7696) (AIC/BIC: ___)

### SPI & SCI colinearity VIF

```{r}
# Check correlation between SPI and SCI
cor(x_var_spi, x_var_sci, use = "complete.obs")

# Check VIF (Variance Inflation Factor)
vif(ols_multiple)
```

**colinearity:** there is significant co-linearity between SCI and SPI, with a correlation of 0.8277. Upon integrating within the same model, SCI inflated the standard error of SPI from 0.0115 to 0.0231. SCI had a similar reaction fro the SPI with its standard error increasing from 0.00796 to 0.0215.

**VIF:** Such multicolinearity is reflected by the VIF test which accounts for all x variables in the model instead of just the two measures of statistical capacity (SCI & SPI).

x_var_spi: 3.34 
x_var_sci: 3.21 
log_gdppc: 1.27
population: 1.02

Unsurprisingly, the variance of the SPI coefficient is inflated by a factor of 3.34 due to correlation with other predictors. Similarly, the SCI coefficient’s variance is inflated by 3.21 times. However, it is acceptable to use in the same model as it will not severely impact estimates given both factors are less than 5.0.

###Combine to single index: principle component analysis (FOR FUTURE)

```{r}
# Standardize both measures
#spi_z <- scale(x_var_spi)
#sci_z <- scale(x_var_sci)

# Create composite (simple average)
#stat_capacity_index <- (spi_z + sci_z)/2

#extract common variance 
#pca_result <- prcomp(cbind(x_var_spi, x_var_sci), scale = TRUE)
#stat_capacity_pc1 <- pca_result$x[,1]  # First principal component
```

**Selecting model:** My research question is about overall statistical capacity rather than comparing different measures. The SPI model reveals a better fit than the SCI model (Adj Rsq: 0.7696 > 0.7235). It is also a slightly better fit compared to the multiple OLS model containing both SPI and SCI (Adj Rsq: 0.7696 > 0.7624).

### Visual Analysis: SCI & SPI x SDG
```{r}
#visualize differences in fit
library(ggplot2)
library(plotly)

#define regression line colors
spi_line <- "steelblue4"
sci_line <- "darkgoldenrod"

# Creating scatterplot with both SPI and SCI on the same plot
Compare_fit <- ggplot(merged, aes(x = spi_comp, y = sdg_overall))+
  geom_smooth(aes(x = spi_comp, y = sdg_overall), 
              color = spi_line, 
              method = "lm", 
              linewidth = 0.75, 
              se = FALSE)+ # Regression line SPI
  geom_smooth(aes(x = sci_overall, y = sdg_overall), 
              color = sci_line, 
              method = "lm", 
              linewidth = 0.75, 
              se = FALSE)+ # Regression line for SCI
  geom_point(aes(color = "spi_comp"), alpha=0.50, size = 0.5)+ # Scatter plot for SPI
  geom_point(aes(x = sci_overall, y = sdg_overall, color = "sci_overall"), 
             alpha=0.5, size = 0.5)+ # Add SCI points w/different color
  scale_color_manual(values = c("spi_comp" = "steelblue1", 
                                "sci_overall" = "darkgoldenrod1")) +
  labs(title = "SDG vs. SPI and SCI",
       x = "SPI / SCI",
       y = "SDG",
       color = "Index") +  # Title for legend
  theme_bw() # Optional: adds a clean, black and white theme

Compare_fit

#make interactive
#ggplotly(Compare_fit)

# Save to specific folder
# ggsave("~/Documents/GitHub/QMSS_Thesis_Sanchez/Output_CSVs/fd_plot.png", p, width = 10, height = 6)
```

# INTERACTIONS AND SUBGROUP ANALYSIS

##Checking for Interactions: 
- Is there a need for subgroup analysis, and if so, by what kind of group? 
- Options: GNI Classification (income_level), regime_type_2, regime_type_4, di_score
```{r}
#interaction: does GNI Classification (income_level) affect the relationship between x (spi) & y (sdg)?
inc_lev_interaction <- lm(sdg_overall ~ spi_comp + spi_comp*income_level + log_gdppc + population,
                        data = merged)
summary(inc_lev_interaction)

#interaction: does regime_type_2 affect the relationship between x (spi) & y (sdg)?
reg_type2_interaction <- lm(sdg_overall ~ spi_comp + spi_comp*regime_type_2 + log_gdppc + population,
                        data = merged)
summary(reg_type2_interaction)

#interaction: does regime_type_4 affect the relationship between x (spi) & y (sdg)?
reg_type4_interaction <- lm(sdg_overall ~ spi_comp + spi_comp*regime_type_4 + log_gdppc + population,
                        data = merged)
summary(reg_type4_interaction)

#interaction: does di affect the relationship between x (spi) & y (sdg)?
reg_type_di_interaction <- lm(sdg_overall ~ spi_comp + spi_comp*di_score + log_gdppc + population,
                        data = merged)
summary(reg_type_di_interaction)
```
*GNI Income Classification: Yes* there are statistically significant interactions found from GNI classifications that affects the relationship between spi and sdgs

*Binary Regime Type: No* there is no statistically significant interactions found from regime type (autocracy vs democracy) that affects the relationship between spi and sdgs. 

*Categorical Regime type (4 options): No* there is no statistically significant interactions found from regime type (Closed autocracy, electoral autocracy, electoral democracy, liberal democracy) that affects the relationship between spi and sdgs. 

*Continuous di_score [0-1] Regime type: No* there is no statistically significant interactions found from regime type (infinate between 0-1) that affects the relationship between spi and sdgs. 


## WB GNI Classifications: income_level ("H", "UM", "LM", "L")
**Disaggregated/Grouped by Development Status:** Make 4 regression models and then put them all together in a table to compare the slopes and R-sq values.
```{r}
# 1. Overall model (all countries)
overall_lm <- lm(sdg_overall ~ spi_comp + di_score + log_gdppc + population + gini,
                        data = merged)
summary(overall_lm)

# 2. High income countries
high_inc_lm <- lm(sdg_overall ~ spi_comp + di_score + log_gdppc + population + gini,
                        data = merged %>% 
                          filter(income_level == "H"))
summary(high_inc_lm)

# 3. Upper-middle income countries
upper_mid_lm <- lm(sdg_overall ~ spi_comp + di_score + log_gdppc + population + gini,
                         data = merged %>% 
                           filter(income_level == "UM"))
summary(upper_mid_lm)

# 4. Lower-middle income countries
lower_mid_lm <- lm(sdg_overall ~ spi_comp + di_score + log_gdppc + population + gini,
                         data = merged %>% 
                           filter(income_level == "LM"))
summary(lower_mid_lm)

# 5. Low income countries
low_inc_lm <- lm(sdg_overall ~ spi_comp + di_score + log_gdppc + population + gini,
                       data = merged %>% 
                         filter(income_level == "L"))
summary(low_inc_lm)
```

### Extract coefficients and statistics of subgroups to comparison table
```{r}
# Creating mod_compare_tab - comparison table
mod_compare_tab <- data.frame(
  model = c("Overall", "High Income", "Upper-Middle Income", 
           "Lower-Middle Income", "Low Income"),
  coefficient = c(
    coef(overall_lm)["spi_comp"],
    coef(high_inc_lm)["spi_comp"],
    coef(upper_mid_lm)["spi_comp"],
    coef(lower_mid_lm)["spi_comp"],
    coef(low_inc_lm)["spi_comp"]
  ),
  
  intercept = c(
    coef(overall_lm)["(Intercept)"],
    coef(high_inc_lm)["(Intercept)"],
    coef(upper_mid_lm)["(Intercept)"],
    coef(lower_mid_lm)["(Intercept)"],
    coef(low_inc_lm)["(Intercept)"]
  ),
  
  Std_error = c(
  summary(overall_lm)$coefficients["spi_comp", 'Std. Error'], # just standard errors 
  summary(high_inc_lm)$coefficients["spi_comp", 'Std. Error'], # just standard errors 
  summary(upper_mid_lm)$coefficients["spi_comp", 'Std. Error'], # just standard errors 
  summary(lower_mid_lm)$coefficients["spi_comp", 'Std. Error'], # just standard errors 
  summary(low_inc_lm)$coefficients["spi_comp", 'Std. Error'] # just standard errors 
  ),

  t_value = c(
    summary(overall_lm)$coefficients["spi_comp", "t value"],
    summary(high_inc_lm)$coefficients["spi_comp", "t value"],
    summary(upper_mid_lm)$coefficients["spi_comp", "t value"],
    summary(lower_mid_lm)$coefficients["spi_comp", "t value"],
    summary(low_inc_lm)$coefficients["spi_comp", "t value"]
  ),
  
  p_value = c(
    summary(overall_lm)$coefficients["spi_comp", "Pr(>|t|)"],
    summary(high_inc_lm)$coefficients["spi_comp", "Pr(>|t|)"],
    summary(upper_mid_lm)$coefficients["spi_comp", "Pr(>|t|)"],
    summary(lower_mid_lm)$coefficients["spi_comp", "Pr(>|t|)"],
    summary(low_inc_lm)$coefficients["spi_comp", "Pr(>|t|)"]
  ),
  
  r_squared = c(
    summary(overall_lm)$r.squared,
    summary(high_inc_lm)$r.squared,
    summary(upper_mid_lm)$r.squared,
    summary(lower_mid_lm)$r.squared,
    summary(low_inc_lm)$r.squared
  ),
  
  adj_r_squared = c(
    summary(overall_lm)$adj.r.squared,
    summary(high_inc_lm)$adj.r.squared,
    summary(upper_mid_lm)$adj.r.squared,
    summary(lower_mid_lm)$adj.r.squared,
    summary(low_inc_lm)$adj.r.squared
  ),
  
  n_obs = c(
    nobs(overall_lm),
    nobs(high_inc_lm),
    nobs(upper_mid_lm),
    nobs(lower_mid_lm),
    nobs(low_inc_lm)
  )
)

# Defining function for significance stars based on p-values
sig_stars <- function(p_value) {
  if (p_value <= 0.001) {
    return("***")
  } else if (p_value <= 0.01) {
    return("**")
  } else if (p_value <= 0.05) {
    return("*")
  } else if (p_value <= 0.1) {
    return(".")
  } else {
    return("")
  }
}

# make table and round for better display
mod_compare_tab <- mod_compare_tab %>%
  mutate(
    coefficient = round(coefficient, 3),
    intercept = round(intercept, 3),
    Std_error = round(Std_error, 3),
    t_value = round(t_value, 3),
    p_value = p_value,
    significance = sapply(p_value, sig_stars), #significance stars 
    r_squared = round(r_squared, 3),
    adj_r_squared = round(adj_r_squared, 3)
  )

# Print the comparison table
print(mod_compare_tab)

# export mod_compare_tab 
# write.csv(mod_compare_tab, file = "ols_model_comparison.csv", row.names=F) 
```

##Visualizing Slopes: plotting multiple regression - by subgroup
```{r}
viz_gni_class <- ggplot(data = merged, aes(x = spi_comp, 
                                           y = sdg_overall, 
                                           color = income_level_lab)) +
  geom_point(alpha = 0.25, size = 0.75) + 
  # Overall regression line (black)
  geom_smooth(aes(group = 1), 
              method = "lm", 
              linewidth = 0.75, 
              se = FALSE, 
              color = "black") +
  # Group-specific regression lines
  geom_smooth(method = "lm", 
              linewidth = 0.65, 
              se = FALSE) +
  scale_color_manual(
    values = c("High Income Countries" = "#1D6A96", 
               "Upper-Middle Income Countries" = "#4CB5AE",
               "Lower-Middle Income Countries" = "#F3A738",
               "Low Income Countries" = "#C02942")
  ) +
  labs(title = "Relationship between SPI and SDG by World Bank Income Classification",
       x = "Statistical Performance Indicators (SPI)",
       y = "Sustainable Development Goals (SDG)",
       color = "Income Classification") +
  theme_bw() 

viz_gni_class
#ggplotly(viz_gni_class)

# Save to specific folder
# ggsave("~/Documents/GitHub/QMSS_Thesis_Sanchez/Output_CSVs/fd_plot.png", p, width = 10, height = 6)
```

# Mediation analysis 
To test if SPI mediates the relationship between regime type and SDG outcomes: 
Democratic backsliding → reduces SPI → slows SDG progress  

*H0: SPI DOES NOT mediate (indirectly effect) the relationship between regime type and SDG status* 
*H1: SPI mediates (indirectly effects) the relationship between regime type and SDG status* 

- ACME (Average Causal Mediation Effect): SPI’s indirect effect.
- ADE (Average Direct Effect): Regime type’s direct effect, excluding SPI
```{r}
library(mediation)

#Total Effect: Check if regime type directly affects SDG scores (without SPI)
lm_sdg <- lm(sdg_overall ~ di_score + log_gdppc + population, data = merged)
summary(lm_sdg)

#Mediator model: Check if regime type affects SPI
lm_spi <- lm(spi_comp ~ di_score + log_gdppc + population, data = merged)
summary(lm_spi)

#outcome model: Check if SPI affects SDG scores while controlling for regime type
lm_sdg_controlled <- lm(sdg_overall ~ spi_comp + di_score + log_gdppc + population, data = merged)
summary(lm_sdg_controlled)

#Mediation test: Quantify how much of regime type’s effect on SDGs operates through SPI
med_model <- mediate(lm_spi, lm_sdg_controlled, treat = "di_score", mediator = "spi_comp")
summary(med_model)

```
ACEM: SPI’s indirect effect = 0.917 units
ADE: Regime type’s direct effect, excluding SPI = 0.326 units
Total Effect: = 1.243 units
Proportion Mediated: = 74.1% of total units

Interpretation: A 1-unit DI increase boosts SDG scores by 1.243 total units, with 0.917 units (~74% of units) transmitted through SPI. The remaining 0.326 units reflect direct DI effects (e.g., governance reforms unrelated to statistics).

Because the ACME (indirect effect of SPI on sdg_overall) is highly significant (p < 0.001), SPI mediates the regime-SDG relationship, based on the model. 

Because the ADE (the direct effect between di_score on sdg_overall) is also significant (p = 0.002), although much less than the ACME estimate, SPI DOES NOT FULLY explain the connection between regime type and sdg status, based on the model. 

## Linearity vs non-linearity 
```{r}
plot(lm_spi, which = 1)  # residuals for SPI model  
plot(lm_sdg, which = 1)  # residuals for SDG model 
plot(lm_sdg_controlled, which = 1)  # residuals for SDG model controlled

library(lmtest)
resettest(lm_spi, power = 2:3, type = "fitted")
resettest(lm_sdg, power = 2:3, type = "fitted")
resettest(lm_sdg_controlled, power = 2:3, type = "fitted")

#validate with sensitivity test 
#sensmediation::sensmed(model_m, model_y, sims = 500)
```

## STILL LEFT TO INCLUDE 
- Use Robust Standard Errors across all models 
- Include GNI Coefficient control in all models 
- Update results based on these changes

