---
title: "Descriptive Statistics: Variables and Datasets"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: paper
    highlight: tango
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE,
                     fig.width = 10, fig.height = 6)
```

# 1. Load Data and Packages

```{r}
# set working directory 
setwd("~/Documents/GitHub/QMSS_Thesis_Sanchez")

# Load all packages from the packages.R file
source("packages.R")

# Load the merged (filtered, final) and merged_full (unfiltered, full) datasets
merged_full <- read_csv("data/Main CSV Outputs/merged_full_df.csv") # full unfiltered dataset
merged <- read_csv("data/Main CSV Outputs/merged_final_df.csv") # filtered dataset

# Define key variables for analysis
key_vars <- c("sdg_overall", "spi_comp", "sci_overall", "di_score", 
              "log_gdppc", "gdp_pc")
group_vars <- c("income_level", "regime_type_2", "regime_type_4")
```

# 2. Dataset Coverage

## 2.1 Overall Dataset Size

```{r}
# Basic dimensions of the dataset
cat("Total number of observations (rows):", nrow(merged_full), "\n")
cat("Total number of variables (columns):", ncol(merged_full), "\n")
cat("Year range:", min(merged_full$year, na.rm=TRUE), "to", max(merged_full$year, na.rm=TRUE), "\n")
cat("Total number of countries:", length(unique(merged_full$country_code)), "\n")
```

## 2.2 Coverage by Variable (full dataset)

```{r}
# Function to get coverage statistics for a variable
get_coverage <- function(data, var_name) {
  total_obs <- sum(!is.na(data[[var_name]]))
  total_countries <- length(unique(data$country_code[!is.na(data[[var_name]])]))
  total_years <- length(unique(data$year[!is.na(data[[var_name]])]))
  missing_obs <- sum(is.na(data[[var_name]]))
  missing_percent <- round(missing_obs / nrow(data) * 100, 2)
  
  return(data.frame(
    Variable = var_name,
    Observations = total_obs,
    Countries = total_countries,
    Years = total_years,
    Missing = missing_obs,
    Missing_Percent = missing_percent
  ))
}

# Get coverage for all key variables
coverage_stats <- bind_rows(lapply(key_vars, function(var) get_coverage(merged_full, var)))

# Create the table with APA styling and no caption number
coverage_table <- kable(coverage_stats, 
      caption = "Coverage Statistics for Key Variables",
      col.names = c("Variable", "Observations", "Countries", "Years", "Missing Values", "Missing (%)"),
      align = c('l', rep('c', 5)),
      booktabs = TRUE) %>%
  kable_classic(full_width = FALSE, html_font = "Times New Roman", font_size = 12, "basic") %>%
  row_spec(0, bold = FALSE, align = "c") %>%
  column_spec(1, bold = FALSE)

# Save the table as HTML
#save_kable(coverage_table, file = "figures/descriptive_stats/full_coverage_stats_table.html")

# Display the table in the document
coverage_table
```

## 2.3 Coverage by Year (full dataset)

```{r}
# Create a table showing variable coverage by year
year_coverage <- merged_full %>%
  group_by(year) %>%
  summarise(across(all_of(key_vars), ~sum(!is.na(.)), .names = "{.col}_count")) %>%
  arrange(year)

# Create the table with APA styling and no caption number
year_coverage_table <- kable(year_coverage, 
      caption = "Number of Non-Missing Observations by Year for Key Variables",
      booktabs = TRUE,
      align = c('l', rep('c', length(key_vars)))) %>%
  kable_classic(full_width = FALSE, html_font = "Times New Roman", font_size = 12, "basic") %>%
  row_spec(0, bold = FALSE, align = "c") %>%
  column_spec(1, bold = FALSE)

# Save the table as HTML
save_kable(year_coverage_table, file = "figures/descriptive_stats/full_year_coverage_table.html")

# Display the table in the document
year_coverage_table
```

## 2.4 Coverage by Variable (final dataset)

```{r}
# Function to get coverage statistics for a variable
get_coverage <- function(data, var_name) {
  total_obs <- sum(!is.na(data[[var_name]]))
  total_countries <- length(unique(data$country_code[!is.na(data[[var_name]])]))
  total_years <- length(unique(data$year[!is.na(data[[var_name]])]))
  missing_obs <- sum(is.na(data[[var_name]]))
  missing_percent <- round(missing_obs / nrow(data) * 100, 2)
  
  return(data.frame(
    Variable = var_name,
    Observations = total_obs,
    Countries = total_countries,
    Years = total_years,
    Missing = missing_obs,
    Missing_Percent = missing_percent
  ))
}

# Get coverage for all key variables
coverage_stats <- bind_rows(lapply(key_vars, function(var) get_coverage(merged, var)))

# Create the table with APA styling and no caption number
coverage_table <- kable(coverage_stats, 
      caption = "Coverage Statistics for Key Variables",
      col.names = c("Variable", "Observations", "Countries", "Years", "Missing Values", "Missing (%)"),
      align = c('l', rep('c', 5)),
      booktabs = TRUE) %>%
  kable_classic(full_width = FALSE, html_font = "Times New Roman", font_size = 12, "basic") %>%
  row_spec(0, bold = FALSE, align = "c") %>%
  column_spec(1, bold = FALSE)

# Save the table as HTML
save_kable(coverage_table, file = "figures/descriptive_stats/final_coverage_stats_table.html")

# Display the table in the document
coverage_table
```

## 2.5 Coverage by Year (final dataset)

```{r}
# Create a table showing variable coverage by year
year_coverage <- merged %>%
  group_by(year) %>%
  summarise(across(all_of(key_vars), ~sum(!is.na(.)), .names = "{.col}_count")) %>%
  arrange(year)

# Create the table with APA styling and no caption number
year_coverage_table <- kable(year_coverage, 
      caption = "Number of Non-Missing Observations by Year for Key Variables",
      booktabs = TRUE,
      align = c('l', rep('c', length(key_vars)))) %>%
  kable_classic(full_width = FALSE, html_font = "Times New Roman", font_size = 12, "basic") %>%
  row_spec(0, bold = FALSE, align = "c") %>%
  column_spec(1, bold = FALSE)

# Save the table as HTML
save_kable(year_coverage_table, file = "figures/descriptive_stats/final_year_coverage_table.html")

# Display the table in the document
year_coverage_table
```

# 3. Descriptive Statistics

## 3.1 Summary Statistics for Key Variables

```{r}
# Generate summary statistics for key variables
merged_stats <- merged %>%
  select(all_of(key_vars)) %>%
  summarize(across(everything(), 
                   list(
                     mean = ~mean(., na.rm = TRUE),
                     sd = ~sd(., na.rm = TRUE),
                     min = ~min(., na.rm = TRUE),
                     max = ~max(., na.rm = TRUE)
                   )))

# Reshape the data for better presentation
merged_stats_long <- merged_stats %>%
  pivot_longer(cols = everything(),
               names_to = c("variable", "stat"),
               names_pattern = "(.*)_(.*)",
               values_to = "value") %>%
  pivot_wider(names_from = stat, values_from = value) %>%
  arrange(variable)

# Create the table with APA styling and no caption number
summary_stats_table <- kable(merged_stats_long, 
      caption = "Descriptive Statistics for Key Variables",
      col.names = c("Variable", "Mean", "SD", "Minimum", "Maximum"),
      booktabs = TRUE,
      digits = 2,
      align = c('l', rep('c', 4))) %>%
  kable_classic(full_width = FALSE, html_font = "Times New Roman", font_size = 12, "basic") %>%
  row_spec(0, bold = FALSE, align = "c") %>%
  column_spec(1, bold = FALSE)

# Save the table as HTML
save_kable(summary_stats_table, file = "figures/descriptive_stats/summary_stats_table.html")

# Display the table in the document
summary_stats_table
```

## 3.2 Summary Statistics by Income Level

```{r}
# Calculate mean values by income level
income_level_stats <- merged %>%
  group_by(income_level) %>%
  summarise(across(all_of(key_vars), 
                  ~mean(.x, na.rm = TRUE), 
                  .names = "mean_{.col}"),
            n = n()) %>%
  arrange(factor(income_level, levels = c("L", "LM", "UM", "H")))

# Add income level labels for better readability
income_level_stats <- income_level_stats %>%
  mutate(income_level_name = case_when(
    income_level == "L" ~ "Low Income",
    income_level == "LM" ~ "Lower-Middle Income",
    income_level == "UM" ~ "Upper-Middle Income",
    income_level == "H" ~ "High Income",
    TRUE ~ income_level
  ))

# Create the table with APA styling and no caption number
income_level_table <- kable(income_level_stats %>% 
        select(income_level_name, n, starts_with("mean_")), 
      caption = "Mean Values by Income Level",
      col.names = c("Income Level", "N", "SDG Overall", "SPI Composite", 
                   "SCI Overall", "Democracy Index", "Log GDP PC", "GDP PC"),
      booktabs = TRUE,
      digits = 2,
      align = c('l', rep('c', 7))) %>%
  kable_classic(full_width = FALSE, html_font = "Times New Roman", font_size = 12, "basic") %>%
  row_spec(0, bold = FALSE, align = "c") %>%
  column_spec(1, bold = FALSE)

# Save the table as HTML
save_kable(income_level_table, file = "figures/descriptive_stats/income_level_table.html")

# Display the table in the document
income_level_table
```

## 3.3 Summary Statistics by Regime Type

```{r}
# Calculate mean values by regime type
regime_type_stats <- merged %>%
  group_by(regime_type_2) %>%
  summarise(across(all_of(key_vars), 
                  ~mean(.x, na.rm = TRUE), 
                  .names = "mean_{.col}"),
            across(all_of(key_vars), 
                  ~sd(.x, na.rm = TRUE), 
                  .names = "sd_{.col}"),
            n = n()) %>%
  mutate(regime_type_name = ifelse(regime_type_2 == 1, "Democracy", "Autocracy"))

# Create the table with APA styling and no caption number
regime_type_table <- kable(regime_type_stats %>% 
        select(regime_type_name, n, starts_with("mean_")), 
      caption = "Mean Values by Regime Type",
      col.names = c("Regime Type", "N", "SDG Overall", "SPI Composite", 
                   "SCI Overall", "Democracy Index", "Log GDP PC", "GDP PC"),
      booktabs = TRUE,
      digits = 2,
      align = c('l', rep('c', 7))) %>%
  kable_classic(full_width = FALSE, html_font = "Times New Roman", font_size = 12, "basic") %>%
  row_spec(0, bold = FALSE, align = "c") %>%
  column_spec(1, bold = FALSE) %>%
  add_footnote("Note: Values represent the mean score for each variable by regime type.", notation = "none") %>%
  footnote(general_title = "", 
           general = "Source: Author's calculations using merged dataset.")

# Save the table as HTML
save_kable(regime_type_table, file = "figures/descriptive_stats/regime_type_table.html")

# Display the table in the document
regime_type_table
```

# 4. Visualizations

## 4.1 Distributions of Key Variables

```{r fig.height=8}
# Create histograms for key index variables using ggplot2
p1 <- ggplot(merged, aes(x = sdg_overall)) +
  geom_histogram(bins = 20, fill = "lightblue", color = "black", alpha = 0.7) +
  geom_vline(xintercept = median(merged$sdg_overall, na.rm = TRUE), 
             color = "red", linewidth = 1, linetype = "dashed") +
  labs(title = "Sustainable Development Goals Index (2004-2023)",
       x = "Composite Score (0-100)",
       y = "Frequency (country-years)") +
  theme_minimal()

p2 <- ggplot(merged, aes(x = spi_comp)) +
  geom_histogram(bins = 20, fill = "lightgreen", color = "black", alpha = 0.7) +
  geom_vline(xintercept = median(merged$spi_comp, na.rm = TRUE), 
             color = "red", linewidth = 1, linetype = "dashed") +
  labs(title = "Statistical Performance Index (2016-2023)",
       x = "Composite Score (0-100)",
       y = "Frequency (country-years)") +
  theme_minimal()

p3 <- ggplot(merged, aes(x = sci_overall)) +
  geom_histogram(bins = 20, fill = "lightyellow", color = "black", alpha = 0.7) +
  geom_vline(xintercept = median(merged$sci_overall, na.rm = TRUE), 
             color = "red", linewidth = 1, linetype = "dashed") +
  labs(title = "Statistical Capacity Index (2004-2020)",
       x = "Composite Score (0-100)",
       y = "Frequency (country-years)") +
  theme_minimal()

p4 <- ggplot(merged, aes(x = di_score)) +
  geom_histogram(bins = 20, fill = "lightpink", color = "black", alpha = 0.7) +
  geom_vline(xintercept = median(merged$di_score, na.rm = TRUE), 
             color = "red", linewidth = 1, linetype = "dashed") +
  labs(title = "Democracy Index",
       x = "Score (0-10)",
       y = "Frequency (country-years)") +
  theme_minimal()

# Combine plots using patchwork
(p1 + p2) / (p3 + p4)

# Save combined plot
ggsave("figures/descriptive_stats/histograms_combined.png", (p1 + p2) / (p3 + p4), width = 10, height = 8, dpi = 300)
```

## 4.2 Boxplots by Regime Type

```{r fig.height=8}
# Create boxplots for variables split by regime type using ggplot2
# Add factor for regime type with proper labels
merged <- merged %>%
  mutate(regime_type_factor = factor(regime_type_2, 
                                    levels = c(0, 1),
                                    labels = c("Autocracy", "Democracy")))

b1 <- ggplot(merged, aes(x = regime_type_factor, y = sdg_overall, fill = regime_type_factor)) +
  geom_boxplot() +
  scale_fill_manual(values = c("salmon", "skyblue")) +
  labs(title = "SDG Index by Regime Type",
       x = "Regime Type",
       y = "Composite Score (0-100)") +
  theme_minimal() +
  theme(legend.position = "none")

b2 <- ggplot(merged, aes(x = regime_type_factor, y = spi_comp, fill = regime_type_factor)) +
  geom_boxplot() +
  scale_fill_manual(values = c("salmon", "skyblue")) +
  labs(title = "SPI Distribution by Regime Type",
       x = "Regime Type",
       y = "Composite Score (0-100)") +
  theme_minimal() +
  theme(legend.position = "none")

b3 <- ggplot(merged, aes(x = regime_type_factor, y = sci_overall, fill = regime_type_factor)) +
  geom_boxplot() +
  scale_fill_manual(values = c("salmon", "skyblue")) +
  labs(title = "SCI Distribution by Regime Type",
       x = "Regime Type",
       y = "Composite Score (0-100)") +
  theme_minimal() +
  theme(legend.position = "none")

b4 <- ggplot(merged, aes(x = regime_type_factor, y = di_score, fill = regime_type_factor)) +
  geom_boxplot() +
  scale_fill_manual(values = c("salmon", "skyblue")) +
  labs(title = "Democracy Index by Regime Type",
       x = "Regime Type",
       y = "Score (0-10)") +
  theme_minimal() +
  theme(legend.position = "none")

# Combine plots using patchwork
boxplot_combined <- (b1 + b2) / (b3 + b4)
boxplot_combined

# Save combined plot
ggsave("figures/descriptive_stats/boxplots_combined.png", boxplot_combined, width = 10, height = 8, dpi = 300)
```

# 5. Quantile Analysis

## 5.1 Quartile Breaks

```{r}
# Calculate quartile breaks for main indices
quant_breaks_sdg <- quantile(merged$sdg_overall, probs = c(0, 0.25, 0.5, 0.75, 1), na.rm = TRUE)
quant_breaks_spi <- quantile(merged$spi_comp, probs = c(0, 0.25, 0.5, 0.75, 1), na.rm = TRUE)
quant_breaks_sci <- quantile(merged$sci_overall, probs = c(0, 0.25, 0.5, 0.75, 1), na.rm = TRUE)
quant_breaks_di <- quantile(merged$di_score, probs = c(0, 0.25, 0.5, 0.75, 1), na.rm = TRUE)

# Display the quartile breaks
quartiles_df <- data.frame(
  Quartile = c("Minimum (0%)", "Q1 (25%)", "Median (50%)", "Q3 (75%)", "Maximum (100%)"),
  SDG_Index = quant_breaks_sdg,
  SPI = quant_breaks_spi,
  SCI = quant_breaks_sci,
  Democracy_Index = quant_breaks_di
)

# Create the table with APA styling and no caption number
quartiles_table <- kable(quartiles_df, 
      caption = "Quartile Breaks for Key Indices",
      booktabs = TRUE,
      digits = 2,
      align = c('l', rep('c', 4))) %>%
  kable_classic(full_width = FALSE, html_font = "Times New Roman", font_size = 12, "basic") %>%
  row_spec(0, bold = FALSE, align = "c") %>%
  column_spec(1, bold = FALSE)

# Save the table as HTML
save_kable(quartiles_table, file = "figures/descriptive_stats/quartiles_table.html")

# Display the table in the document
quartiles_table
```

## 5.2 Quantile Frequency Tables

```{r}
# Create quartile bins for each index
merged <- merged %>%
  mutate(
    sdg_quant_bins = cut(sdg_overall, 
                      breaks = quant_breaks_sdg, 
                      include.lowest = TRUE, 
                      labels = c("Q1 (Lowest)", "Q2", "Q3", "Q4 (Highest)")),
    spi_quant_bins = cut(spi_comp, 
                      breaks = quant_breaks_spi, 
                      include.lowest = TRUE, 
                      labels = c("Q1 (Lowest)", "Q2", "Q3", "Q4 (Highest)")),
    sci_quant_bins = cut(sci_overall, 
                      breaks = quant_breaks_sci, 
                      include.lowest = TRUE, 
                      labels = c("Q1 (Lowest)", "Q2", "Q3", "Q4 (Highest)")),
    di_quant_bins = cut(di_score, 
                      breaks = quant_breaks_di, 
                      include.lowest = TRUE, 
                      labels = c("Q1 (Lowest)", "Q2", "Q3", "Q4 (Highest)"))
  )

# Create frequency tables by regime type for each index
# SDG Index
sdg_summary <- merged %>%
  filter(!is.na(sdg_quant_bins) & !is.na(regime_type_2)) %>%
  group_by(regime_type_2, sdg_quant_bins) %>%
  summarise(count = n(), .groups = "drop") %>%
  mutate(regime_type_name = ifelse(regime_type_2 == 1, "Democracy", "Autocracy"))

# Pivot to wide format
sdg_wide_table <- sdg_summary %>%
  select(-regime_type_2) %>%
  pivot_wider(names_from = sdg_quant_bins, values_from = count, values_fill = 0)

# Create the table with APA styling and no caption number
sdg_quartile_table <- kable(sdg_wide_table, 
      caption = "SDG Index Quartile Distribution by Regime Type",
      booktabs = TRUE,
      align = c('l', rep('c', 4))) %>%
  kable_classic(full_width = FALSE, html_font = "Times New Roman", font_size = 12, "basic") %>%
  row_spec(0, bold = FALSE, align = "c") %>%
  column_spec(1, bold = FALSE)

# Save the table as HTML
save_kable(sdg_quartile_table, file = "figures/descriptive_stats/sdg_quartile_table.html")

# Display the table in the document
sdg_quartile_table

# SPI Index
spi_summary <- merged %>%
  filter(!is.na(spi_quant_bins) & !is.na(regime_type_2)) %>%
  group_by(regime_type_2, spi_quant_bins) %>%
  summarise(count = n(), .groups = "drop") %>%
  mutate(regime_type_name = ifelse(regime_type_2 == 1, "Democracy", "Autocracy"))

# Pivot to wide format
spi_wide_table <- spi_summary %>%
  select(-regime_type_2) %>%
  pivot_wider(names_from = spi_quant_bins, values_from = count, values_fill = 0)

# Create the table with APA styling and no caption number
spi_quartile_table <- kable(spi_wide_table, 
      caption = "SPI Quartile Distribution by Regime Type",
      booktabs = TRUE,
      align = c('l', rep('c', 4))) %>%
  kable_classic(full_width = FALSE, html_font = "Times New Roman", font_size = 12, "basic") %>%
  row_spec(0, bold = FALSE, align = "c") %>%
  column_spec(1, bold = FALSE)

# Save the table as HTML
save_kable(spi_quartile_table, file = "figures/descriptive_stats/spi_quartile_table.html")

# Display the table in the document
spi_quartile_table

# SCI Index
sci_summary <- merged %>%
  filter(!is.na(sci_quant_bins) & !is.na(regime_type_2)) %>%
  group_by(regime_type_2, sci_quant_bins) %>%
  summarise(count = n(), .groups = "drop") %>%
  mutate(regime_type_name = ifelse(regime_type_2 == 1, "Democracy", "Autocracy"))

# Pivot to wide format
sci_wide_table <- sci_summary %>%
  select(-regime_type_2) %>%
  pivot_wider(names_from = sci_quant_bins, values_from = count, values_fill = 0)

# Create the table with APA styling and no caption number
sci_quartile_table <- kable(sci_wide_table, 
      caption = "SCI Quartile Distribution by Regime Type",
      booktabs = TRUE,
      align = c('l', rep('c', 4))) %>%
  kable_classic(full_width = FALSE, html_font = "Times New Roman", font_size = 12, "basic") %>%
  row_spec(0, bold = FALSE, align = "c") %>%
  column_spec(1, bold = FALSE)

# Save the table as HTML
save_kable(sci_quartile_table, file = "figures/descriptive_stats/sci_quartile_table.html")

# Display the table in the document
sci_quartile_table
```

# 6. Summary Table of Findings

```{r}
# Create a summary table of key findings
summary_findings <- data.frame(
  Variable = c("SDG Overall", "SPI Composite", "SCI Overall", "Democracy Index"),
  Democracy_Mean = c(
    mean(merged$sdg_overall[merged$regime_type_2 == 1], na.rm = TRUE),
    mean(merged$spi_comp[merged$regime_type_2 == 1], na.rm = TRUE),
    mean(merged$sci_overall[merged$regime_type_2 == 1], na.rm = TRUE),
    mean(merged$di_score[merged$regime_type_2 == 1], na.rm = TRUE)
  ),
  Autocracy_Mean = c(
    mean(merged$sdg_overall[merged$regime_type_2 == 0], na.rm = TRUE),
    mean(merged$spi_comp[merged$regime_type_2 == 0], na.rm = TRUE),
    mean(merged$sci_overall[merged$regime_type_2 == 0], na.rm = TRUE),
    mean(merged$di_score[merged$regime_type_2 == 0], na.rm = TRUE)
  ),
  Difference = c(
    mean(merged$sdg_overall[merged$regime_type_2 == 1], na.rm = TRUE) - 
      mean(merged$sdg_overall[merged$regime_type_2 == 0], na.rm = TRUE),
    mean(merged$spi_comp[merged$regime_type_2 == 1], na.rm = TRUE) - 
      mean(merged$spi_comp[merged$regime_type_2 == 0], na.rm = TRUE),
    mean(merged$sci_overall[merged$regime_type_2 == 1], na.rm = TRUE) - 
      mean(merged$sci_overall[merged$regime_type_2 == 0], na.rm = TRUE),
    mean(merged$di_score[merged$regime_type_2 == 1], na.rm = TRUE) - 
      mean(merged$di_score[merged$regime_type_2 == 0], na.rm = TRUE)
  ),
  Percent_Difference = c(
    (mean(merged$sdg_overall[merged$regime_type_2 == 1], na.rm = TRUE) / 
      mean(merged$sdg_overall[merged$regime_type_2 == 0], na.rm = TRUE) - 1) * 100,
    (mean(merged$spi_comp[merged$regime_type_2 == 1], na.rm = TRUE) / 
      mean(merged$spi_comp[merged$regime_type_2 == 0], na.rm = TRUE) - 1) * 100,
    (mean(merged$sci_overall[merged$regime_type_2 == 1], na.rm = TRUE) / 
      mean(merged$sci_overall[merged$regime_type_2 == 0], na.rm = TRUE) - 1) * 100,
    (mean(merged$di_score[merged$regime_type_2 == 1], na.rm = TRUE) / 
      mean(merged$di_score[merged$regime_type_2 == 0], na.rm = TRUE) - 1) * 100
  )
)

# Format and display the summary table with APA styling and no caption number
summary_findings_table <- kable(summary_findings, 
      caption = "Summary of Differences Between Democracies and Autocracies",
      col.names = c("Variable", "Democracy Mean", "Autocracy Mean", "Absolute Difference", "% Difference"),
      booktabs = TRUE,
      digits = 2,
      align = c('l', rep('c', 4))) %>%
  kable_classic(full_width = FALSE, html_font = "Times New Roman", font_size = 12, "basic") %>%
  row_spec(0, bold = FALSE, align = "c") %>%
  column_spec(1, bold = FALSE) %>%
  add_footnote("Note: Positive differences indicate higher values in democracies compared to autocracies.", notation = "none") %>%
  footnote(general_title = "", 
           general = "Source: Author's calculations using merged dataset.")

# Save the table as HTML
save_kable(summary_findings_table, file = "figures/descriptive_stats/summary_findings_table.html")

# Display the table in the document
summary_findings_table
```