---
title: "Comp2_residual_diagn"
author: "Sevastian Sanchez"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loading Data & Packages
```{r}
setwd("~/Documents/GitHub/QMSS_Thesis_Sanchez")

#load libraries 
source("packages.R")

#load function 
#source("df_years2.0_Function.R")

#load df_years() function: 2015-present
#all_data <- df_years2.0(2004, 2023)

#read merged_df csv
all_data <- read_csv("data/Main CSV Outputs/merged_final_df.csv")

# Adding has_aut_ep and has_dem_ep for regressing regimes (experienced atleast 1 aut_ep/dem_ep)
panel_data <- all_data %>%
  # Group by country to check for any event
  group_by(country_code) %>%
  mutate(has_aut_ep = any(aut_ep == 1, na.rm = TRUE)) %>%
  mutate(has_dem_ep = any(dem_ep == 1, na.rm = TRUE)) %>% 
  ungroup() 

# centralized variables for higher-order multicollinearity 
panel_data <- panel_data %>%
  mutate(
    cen_sdg_overall = sdg_overall - mean(sdg_overall, na.rm = TRUE),
    cen_spi_comp = spi_comp - mean(spi_comp, na.rm = TRUE),
    cen_di_score = di_score - mean(di_score, na.rm = TRUE),
    cen_log_gdppc = log_gdppc - mean(log_gdppc, na.rm = TRUE)
  )

# year-year lags for FD Models - DF
fd_data <- panel_data %>%
  select(country_code, year, sdg_overall, di_score, spi_comp, log_gdppc, income_level, aut_ep, dem_ep, regch_event) %>%
  filter(!is.na(di_score) & !is.na(spi_comp) | !is.na(spi_comp) & !is.na(sdg_overall)) %>% 
  arrange(country_code, year) %>%
  group_by(country_code) %>%
  mutate(
    # first differences for selected variables
    sdg_diff = sdg_overall - dplyr::lag(sdg_overall, n=1),
    di_diff = di_score - dplyr::lag(di_score, n=1),
    spi_diff = spi_comp - dplyr::lag(spi_comp,n=1),
    log_gdppc_diff = log_gdppc - dplyr::lag(log_gdppc, n=1),
    # lagged first differences 
    di_diff_lag1 = dplyr::lag(di_diff, n=1),
    di_diff_lag2 = dplyr::lag(di_diff, n=2),
    spi_diff_lag1 = dplyr::lag(spi_diff, n=1),
    spi_diff_lag2 = dplyr::lag(spi_diff, n=2),
    log_gdppc_diff_lag1 = dplyr::lag(log_gdppc_diff, n=1),
    log_gdppc_diff_lag1 = dplyr::lag(log_gdppc_diff, n=2)
  ) %>%
  ungroup()
```

# STAGE 1 MODELS RESIGUAL DIAGNOSTICS
- checking residuals ONLY for Stage 1 and 2 Models, ommitting all Comp 1 models
Stage 1: 
- ols_spi_di
- fd_spi_di
- fe_spi_di
Stage 2: 
- ols_base_lm
- fd_base_lm 
- fe_base_lm 

## PLOS Residuals for mediator model: SPI ~ DI + Controls [STAGE 1]
```{r}
# Extract the data actually used in the model
ols_spi_di <- lm(spi_comp ~ dplyr::lag(di_score, 1) + log_gdppc + factor(year), data = panel_data) # regular linear model
model_data1 <- model.frame(ols_spi_di)

# saving as png 
#png("figures/residual_plots/ols1_residual_plots.png", width = 8, height = 3, units = "in", res = 300)
#par(mfrow = c(1, 3))

# switch to all stage-one plots 
png("figures/residual_plots/all_s1_residual_plots.png", width = 8, height = 6.5, units = "in", res = 300)
par(mfrow = c(3, 3))

# Residuals vs Fitted Values plot for ols_spi_di
plot(ols_spi_di, which = 1,
     main = "Residuals vs Fitted Model", caption = "ols_spi_di [Stage 1]", pch = 1, cex = 0.35, col = "#595959")
abline(h = 0, col = "black", lty = 2, lwd = 1)
fitted_vals <- fitted(ols_spi_di)
lines(lowess(fitted_vals, resid(ols_spi_di)), col = "red", lwd =  1.5)

# Residuals vs di_score plot
plot(model_data1$di_score, resid(ols_spi_di),
     xlab = "di_score",  ylab = "", yaxt = "n", main = "Residuals vs di_score", pch = 1, cex = 0.35, col = "darkred")
abline(h = 0, col = "black", lty = 2, lwd = 1)
lines(lowess(model_data1$di_score, resid(ols_spi_di)), col = "red", lwd = 1.5)

# Residuals vs log_gdppc plot
plot(model_data1$log_gdppc, resid(ols_spi_di),
     xlab = "log_gdppc",  ylab = "", yaxt = "n", main = "Residuals vs log_gdppc", pch = 1, cex = 0.35, col = "darkgreen")
abline(h = 0, col = "black", lty = 2, lwd = 1)
lines(lowess(model_data1$log_gdppc, resid(ols_spi_di)), col = "red", lwd =  1.5)

#dev.off() # to save 

#par(mfrow = c(1, 1))
```

## First Differences Residuals for mediator model: SPI ~ DI + Controls [STAGE 1]
```{r}
# Extract the data actually used in the model
fd_spi_di <- lm(spi_diff ~ dplyr::lag(di_diff, 1) + log_gdppc_diff, data = fd_data)
model_data_fd1 <- model.frame(fd_spi_di)

# saving as png
#png("figures/residual_plots/fd1_residual_plots.png", width = 8, height = 3, units = "in", res = 300)
#par(mfrow = c(1, 3))

# Residuals vs Fitted Values plot for fd_spi_di
plot(fd_spi_di, which = 1,
     main = "FD Residuals vs Fitted Model", caption = "fd_spi_di [Stage 1]", pch = 1, cex = 0.35, col = "#595959")
abline(h = 0, col = "black", lty = 2, lwd = 1)
fitted_vals_fd <- fitted(fd_spi_di)
lines(lowess(fitted_vals_fd, resid(fd_spi_di)), col = "red", lwd =  1.5)

# Residuals vs di_diff plot
plot(model_data_fd1$di_diff, resid(fd_spi_di),
     xlab = "di_diff",  ylab = "", yaxt = "n", main = "FD Residuals vs di_diff", pch = 1, cex = 0.35, col = "darkred")
abline(h = 0, col = "black", lty = 2, lwd = 1)
lines(lowess(model_data_fd1$di_diff, resid(fd_spi_di)), col = "red", lwd = 1.5)

# Residuals vs log_gdppc_diff plot
plot(model_data_fd1$log_gdppc_diff, resid(fd_spi_di),
     xlab = "log_gdppc_diff",  ylab = "", yaxt = "n", main = "FD Residuals vs log_gdppc_diff", pch = 1, cex = 0.35, col = "darkgreen")
abline(h = 0, col = "black", lty = 2, lwd = 1)
lines(lowess(model_data_fd1$log_gdppc_diff, resid(fd_spi_di)), col = "red", lwd =  1.5)

#dev.off() # to save
#par(mfrow = c(1, 1))
```

## FE Residuals for mediator model: SPI ~ DI + Controls [STAGE 1]
```{r}
# Extract the data actually used in the model
fe_spi_di <- lm(spi_comp ~ dplyr::lag(di_score, 1) + log_gdppc + factor(year) + factor(country_code), data = panel_data)
model_data_fe1 <- model.frame(fe_spi_di)

# saving as png
#png("figures/residual_plots/fe1_residual_plots.png", width = 8, height = 3, units = "in", res = 300)
#par(mfrow = c(1, 3))

# Residuals vs Fitted Values plot for fe_spi_di
plot(fe_spi_di, which = 1,
     main = "FE Residuals vs Fitted Model", caption = "fe_spi_di [Stage 1]", pch = 1, cex = 0.35, col = "#595959")
abline(h = 0, col = "black", lty = 2, lwd = 1)
fitted_vals_fe <- fitted(fe_spi_di)
lines(lowess(fitted_vals_fe, resid(fe_spi_di)), col = "red", lwd =  1.5)

# Residuals vs di_score plot
plot(model_data_fe1$di_score, resid(fe_spi_di),
     xlab = "di_score",  ylab = "", yaxt = "n", main = "FE Residuals vs di_score", pch = 1, cex = 0.35, col = "darkred")
abline(h = 0, col = "black", lty = 2, lwd = 1)
lines(lowess(model_data_fe1$di_score, resid(fe_spi_di)), col = "red", lwd = 1.5)

# Residuals vs log_gdppc plot
plot(model_data_fe1$log_gdppc, resid(fe_spi_di),
     xlab = "log_gdppc",  ylab = "", yaxt = "n", main = "FE Residuals vs log_gdppc", pch = 1, cex = 0.35, col = "darkgreen")
abline(h = 0, col = "black", lty = 2, lwd = 1)
lines(lowess(model_data_fe1$log_gdppc, resid(fe_spi_di)), col = "red", lwd =  1.5)

dev.off() # to save
par(mfrow = c(1, 1))
```

## POLS Residuals for base/outcome model: SDG ~ SPI + DI + Controls [STAGE 2]
```{r}
# Extract the data actually used in the base model
ols_base_lm <- lm(sdg_overall ~ dplyr::lag(spi_comp, 1) + dplyr::lag(di_score, 1) + log_gdppc + factor(year), data = panel_data) # regular linear model
model_data2 <- model.frame(ols_base_lm)

# saving as png 
#png("figures/residual_plots/ols2_residual_plots.png", width = 8, height = 2.5, units = "in", res = 300)
#par(mfrow = c(1, 4))

# switch to - saving all stage-two plots
png("figures/residual_plots/all_s2_residual_plots.png", width = 8.5, height = 6.5, units = "in", res = 300)
par(mfrow = c(3, 4))

# Residuals vs Fitted Values plot for ols_base_lm
plot(ols_base_lm, which = 1,
     main = "Residuals vs Fitted Model", caption = "ols_base [Stage 2]", pch = 1, cex = 0.35, col = "#595959")
abline(h = 0, col = "black", lty = 2, lwd = 1)
fitted_vals <- fitted(ols_base_lm)
lines(lowess(fitted_vals, resid(ols_base_lm)), col = "red", lwd =  1.5)

# Residuals vs spi_comp plots
plot(model_data2$spi_comp, resid(ols_base_lm),
     xlab = "spi_comp", ylab = "", yaxt = "n", main = "Residuals vs spi_comp", pch = 1, cex = 0.35, col = "darkblue")
abline(h = 0, col = "black", lty = 2, lwd = 1)
lines(lowess(model_data2$spi_comp, resid(ols_base_lm)), col = "red", lwd =  1.5)

# Residuals vs di_score plot
plot(model_data2$di_score, resid(ols_base_lm),
     xlab = "di_score", ylab = "", yaxt = "n", main = "Residuals vs di_score", pch = 1, cex = 0.35, col = "darkred")
abline(h = 0, col = "black", lty = 2, lwd = 1)
lines(lowess(model_data2$di_score, resid(ols_base_lm)), col = "red", lwd = 1.5)

# Residuals vs log_gdppc plot
plot(model_data2$log_gdppc, resid(ols_base_lm),
     xlab = "log_gdppc", ylab = "", yaxt = "n", main = "Residuals vs log_gdppc", pch = 1, cex = 0.35, col = "darkgreen")
abline(h = 0, col = "black", lty = 2, lwd = 1)
lines(lowess(model_data2$log_gdppc, resid(ols_base_lm)), col = "red", lwd =  1.5)

#dev.off() # to save 
#par(mfrow = c(1, 1))
```
The residuals vs fitted values plot shows a clear pattern, indicating non-linearity in the relationship between the predictors and the SDG performance. The residuals are not randomly distributed around zero, suggesting that the linear model does not adequately capture the underlying relationship.

### Log Transformation: log(gdppc) & log(gdppc)^2 Residuals Plots [Extra - Do later]
```{r}
# Residuals vs log(gdppc) plot
plot(log(model_data$gdp_pc), resid(ols_base_lm),
     xlab = "log_gdppc", ylab = "Residuals", main = "Residuals vs gdp_pc", pch = 1, col = "darkgreen")
abline(h = 0, col = "red", lty = 2, lwd = 1)
lines(lowess(log(model_data$gdp_pc), resid(ols_base_lm)), col = "black", lwd = 2)

# Residuals vs log(gdppc)^2 plot
plot(I(log(model_data$gdp_pc)^2), resid(ols_base_lm),
     xlab = "log_gdppc", ylab = "Residuals", main = "Residuals vs gdp_pc", pch = 16, col = "darkgreen")
abline(h = 0, col = "red", lty = 2, lwd = 1)
lines(lowess(I(log(model_data$gdp_pc)^2), resid(ols_base_lm)), col = "black", lwd = 2)

# Set up plotting area for side-by-side comparison
par(mfrow = c(1, 3))  # 1 row, 3 columns for the 3 continuous variables
```


### Experimental - Quad term for log_gdppc in base model
```{r eval=FALSE, include=FALSE}
quad2_log_gdppc <- I(panel_data$log_gdppc^2)

ols_base_lm_quad <- lm(sdg_overall ~ spi_comp + di_score + log_gdppc + quad2_log_gdppc + factor(year), data = panel_data) # regular linear model
model_data3 <- model.frame(ols_base_lm_quad)

par(mfrow = c(2, 2))

# Residuals vs Fitted Values plot for ols_base_lm_quad
plot(ols_base_lm_quad, which = 1,
     main = "Residuals vs Fitted Model", caption = "ols_base", pch = 1, cex = 0.35, col = "#595959")
abline(h = 0, col = "black", lty = 2, lwd = 1)
fitted_vals <- fitted(ols_base_lm_quad)
lines(lowess(fitted_vals, resid(ols_base_lm_quad)), col = "red", lwd =  1.5)

# Residuals vs spi_comp plots
plot(model_data3$spi_comp, resid(ols_base_lm_quad),
     xlab = "spi_comp", ylab = "Residuals", main = "Residuals vs spi_comp", pch = 1, cex = 0.35, col = "darkblue")
abline(h = 0, col = "black", lty = 2, lwd = 1)
lines(lowess(model_data3$spi_comp, resid(ols_base_lm_quad)), col = "red", lwd =  1.5)

# Residuals vs di_score plot
plot(model_data3$di_score, resid(ols_base_lm_quad),
     xlab = "di_score", ylab = "Residuals", main = "Residuals vs di_score", pch = 1, cex = 0.35, col = "darkred")
abline(h = 0, col = "black", lty = 2, lwd = 1)
lines(lowess(model_data3$di_score, resid(ols_base_lm_quad)), col = "red", lwd = 1.5)

# Residuals vs log_gdppc plot
plot(model_data3$quad2_log_gdppc, resid(ols_base_lm_quad),
     xlab = "quad2_log_gdppc", ylab = "Residuals", main = "Residuals vs log_gdppc", pch = 1, cex = 0.35, col = "darkgreen")
abline(h = 0, col = "black", lty = 2, lwd = 1)
lines(lowess(model_data3$quad2_log_gdppc, resid(ols_base_lm_quad)), col = "red", lwd =  1.5)

par(mfrow = c(1, 1))
```



## First Differences Residuals for base/outcome model: SDG ~ SPI + DI + Controls [STAGE 2]
```{r}
# Extract the data actually used in the base model
fd_sdg_spi <- lm(sdg_diff ~ dplyr::lag(spi_diff, 1) + dplyr::lag(di_diff, 1) + log_gdppc_diff, data = fd_data)
model_data_fd2 <- model.frame(fd_base_lm)

# saving as png
#png("figures/residual_plots/fd2_residual_plots.png", width = 8, height = 2.5, units = "in", res = 300)
#par(mfrow = c(1, 4))

# Residuals vs Fitted Values plot for fd_base_lm
plot(fd_base_lm, which = 1,
     main = "FD Residuals vs Fitted Model", caption = "fd_base [Stage 2]", pch = 1, cex = 0.35, col = "#595959")
abline(h = 0, col = "black", lty = 2, lwd = 1)
fitted_vals_fd2 <- fitted(fd_base_lm)
lines(lowess(fitted_vals_fd2, resid(fd_base_lm)), col = "red", lwd =  1.5)

# Residuals vs spi_comp plots
plot(model_data_fd2$spi_diff, resid(fd_base_lm),
     xlab = "spi_diff", ylab = "", yaxt = "n", main = "FD Residuals vs spi_comp", pch = 1, cex = 0.35, col = "darkblue")
abline(h = 0, col = "black", lty = 2, lwd = 1)
lines(lowess(model_data_fd2$spi_diff, resid(fd_base_lm)), col = "red", lwd =  1.5)

# Residuals vs di_score plot
plot(model_data_fd2$di_diff, resid(fd_base_lm),
     xlab = "di_diff",  ylab = "", yaxt = "n", main = "FD Residuals vs di_score", pch = 1, cex = 0.35, col = "darkred")
abline(h = 0, col = "black", lty = 2, lwd = 1)
lines(lowess(model_data_fd2$di_diff, resid(fd_base_lm)), col = "red", lwd = 1.5)

# Residuals vs log_gdppc_diff plot
plot(model_data_fd2$log_gdppc_diff, resid(fd_base_lm),
     xlab = "log_gdppc_diff",  ylab = "", yaxt = "n", main = "FD Residuals vs log_gdppc", pch = 1, cex = 0.35, col = "darkgreen")
abline(h = 0, col = "black", lty = 2, lwd = 1)
lines(lowess(model_data_fd2$log_gdppc_diff, resid(fd_base_lm)), col = "red", lwd =  1.5)

#dev.off() # to save
#par(mfrow = c(1, 1))
```

## FE Residuals for base/outcome model: SDG ~ SPI + DI + Controls [STAGE 2]
```{r}
# Extract the data actually used in the base model
fe_sdg_spi <- lm(sdg_overall ~ dplyr::lag(spi_comp, 1) + dplyr::lag(di_score, 1) + log_gdppc + factor(year) + factor(country_code), data = panel_data)
model_data_fe2 <- model.frame(fe_base_lm)

# saving as png
#png("figures/residual_plots/fe2_residual_plots.png", width = 8, height = 2.5, units = "in", res = 300)
#par(mfrow = c(1, 4))

# Residuals vs Fitted Values plot for fe_base_lm
plot(fe_base_lm, which = 1,
     main = "FE Residuals vs Fitted Model", caption = "fe_base [Stage 2]", pch = 1, cex = 0.35, col = "#595959")
abline(h = 0, col = "black", lty = 2, lwd = 1)
fitted_vals_fe2 <- fitted(fe_base_lm)
lines(lowess(fitted_vals_fe2, resid(fe_base_lm)), col = "red", lwd =  1.5)

# Residuals vs spi_comp plots
plot(model_data_fe2$spi_comp, resid(fe_base_lm),
     xlab = "spi_comp", ylab = "", yaxt = "n", main = "FE Residuals vs spi_comp", pch = 1, cex = 0.35, col = "darkblue")
abline(h = 0, col = "black", lty = 2, lwd = 1)
lines(lowess(model_data_fe2$spi_comp, resid(fe_base_lm)), col = "red", lwd =  1.5)

# Residuals vs di_score plot
plot(model_data_fe2$di_score, resid(fe_base_lm),
     xlab = "di_score", ylab = "", yaxt = "n", main = "FE Residuals vs di_score", pch = 1, cex = 0.35, col = "darkred")
abline(h = 0, col = "black", lty = 2, lwd = 1)
lines(lowess(model_data_fe2$di_score, resid(fe_base_lm)), col = "red", lwd = 1.5)

# Residuals vs log_gdppc plot
plot(model_data_fe2$log_gdppc, resid(fe_base_lm),
     xlab = "log_gdppc",  ylab = "", yaxt = "n", main = "FE Residuals vs log_gdppc", pch = 1, cex = 0.35, col = "darkgreen")
abline(h = 0, col = "black", lty = 2, lwd = 1)
lines(lowess(model_data_fe2$log_gdppc, resid(fe_base_lm)), col = "red", lwd =  1.5)

dev.off() # to save
par(mfrow = c(1, 1))
```
- No unsusual curvatures or patterns. 
- Homoscedasticity seems reasonable.
- No obvious outliers.
- Residuals appear roughly normally distributed.
- Overall, the residual diagnostics suggest that the linear regression assumptions are reasonably met for both models.

