---
title: "Comp3_interactions"
output: html_document
---

# Set up and Wrangling 
```{r}
# set working directory
setwd("~/Documents/GitHub/QMSS_Thesis_Sanchez")

#load libraries/packages
source("packages.R")

# load data
source("Comp2_panel_wrangling.R")
```

**Quick Counts**
```{r}
# EIU DI binary regime type 
table(panel_data$di_reg_type_2)

# RoW V-Dem binary regime type
table(panel_data$regime_type_binary)

# income_level_recoded 
table(panel_data$income_level_recoded)
```


## Stepwise Check: Moderator/Interaction Terms [Stage 1]

## EIU DI binary regime type, Subgrouping 
```{r}
panel_sub <- panel_data %>%
  filter(di_reg_type_2 == 1) # Democracies only

# Load the panel data
FE_testing <- plm(formula = spi_comp ~ di_score + dplyr::lag(di_score, 1) + dplyr::lag(di_score, 2) + log_gdppc + factor(income_level_recoded),
                  model = "within", 
                  index = c("country_code", "year"),
                  data = panel_data, 
                  #subset = di_reg_type_2 == 1, # modifying based on regime type
                  effect = "twoways")
summary(FE_testing)
summary(FE_testing, vcov = vcovHC(FE_testing, cluster = "group", type = "HC1"))
```

## WB GNI Class, Subgrouping 
```{r}
# GNI Class 
FE_testing <- plm(formula = spi_comp ~ di_score + dplyr::lag(di_score, 1) + dplyr::lag(di_score, 2) + log_gdppc,
                  model = "within", 
                  index = c("country_code", "year"),
                  data = panel_data, 
                  #subset = income_level_recoded == 3, # modifying based on income level
                  effect = "twoways")
summary(FE_testing)
summary(FE_testing, vcov = vcovHC(FE_testing, cluster = "group", type = "HC1"))
```

## Regime Change Episodes (has_aut_ep), Subgrouping
```{r}
FE_testing <- plm(formula = spi_comp ~ di_score + dplyr::lag(di_score, 1) + dplyr::lag(di_score, 2) + log_gdppc,
                  model = "within", 
                  index = c("country_code", "year"),
                  data = panel_data, 
                  subset = has_aut_ep == TRUE, # modifying based regime stability
                  effect = "twoways")
summary(FE_testing)
summary(FE_testing, vcov = vcovHC(FE_testing, cluster = "group", type = "HC1"))
```

## Regime Change Episodes (has_dem_ep), Subgrouping
```{r}
FE_testing <- plm(formula = spi_comp ~ di_score + dplyr::lag(di_score, 1) + dplyr::lag(di_score, 2) + log_gdppc,
                  model = "within", 
                  index = c("country_code", "year"),
                  data = panel_data, 
                  subset = has_dem_ep == TRUE, # modifying based regime stability
                  effect = "twoways")
summary(FE_testing)
summary(FE_testing, vcov = vcovHC(FE_testing, cluster = "group", type = "HC1"))
```


## Complete Regime Change (autocratized), Subgrouping [FOUND SOMETHING!!!]
Of countries (15 countries) that experienced complete change from democracy to autocracy, how did the decrease in DI score impact their SPI performance? Was the impact lagged or immediate?
```{r}
# autocratized variable 
FE_testing <- plm(formula = spi_comp ~ di_score + dplyr::lag(di_score, 1) + dplyr::lag(di_score, 2) + log_gdppc,
                  model = "within", 
                  index = c("country_code", "year"),
                  data = panel_data, 
                  subset = autocratized == TRUE, # modifying based regime stability
                  effect = "twoways")
summary(FE_testing)
summary(FE_testing, vcov = vcovHC(FE_testing, cluster = "group", type = "HC1"))
```

## Complete Regime Change (democratized), Subgrouping [FOUND SOMETHING!!!]
Of countries (15 countries) that experienced complete change from autocracy to democracy, how did the decrease in DI score impact their SPI performance? Was the impact lagged or immediate?
```{r}
# autocratized variable 
FE_testing <- plm(formula = spi_comp ~ di_score + dplyr::lag(di_score, 1) + dplyr::lag(di_score, 2) + log_gdppc,
                  model = "within", 
                  index = c("country_code", "year"),
                  data = panel_data, 
                  subset = democratized == TRUE, # modifying based on regime stability
                  effect = "twoways")
summary(FE_testing)
summary(FE_testing, vcov = vcovHC(FE_testing, cluster = "group", type = "HC1"))
```



## Stepwise Check: Moderator/Interaction Terms [Stage 1]
*H0:* NUll, the impact of DI on SPI Performance does NOT vary depending on Z [income_level_recoded, regime_type_binary, di_score]
*H1:* The impact of DI on SPI Performance varies depending on Z [income_level_recoded, regime_type_binary,  di_score]

**Recoded Moderator Variables:**
- income_level_recoded: 1 = Low; 2 = Lower-Middle; 3 = Upper-Middle; 4 = High
- regime_type_binary: 0 = Autocracy; 1 = Democracy
- aut_ep: 0 = No Autocratization; 1 = Autocratization Event
- dem_ep: 0 = No Democratization; 1 = Democratization Event
**NOTE**: Based on the results of the RESET test, I will not include polynomial terms in the interaction models, as the linear model is appropriate.
```{r}
### HYPOTHESES ###

# Interaction 1: DI x GNI Class 
# HA: The impact of SPI on SDG performance varies depending on GNI Classification 
di_x_inc_lvl <- plm(formula = spi_comp ~ dplyr::lag(cen_di_score, 1)*dplyr::lag(income_level_recoded, 1) + cen_log_gdppc,
                     model = "within", 
                     index = c("country_code", "year"),
                     data = panel_data,
                     effect = "twoways")
summary(di_x_inc_lvl, vcov = vcovHC(di_x_inc_lvl, cluster = "group", type = "HC1"))

# Interaction 2: DI x Regime Type (Binary)
# HA: The impact of SPI on SDG performance varies depending on Regime Type (Binary)
di_x_reg_binary <- plm(formula = spi_comp ~ dplyr::lag(cen_di_score, 1)*dplyr::lag(autocracy, 1) + cen_log_gdppc, 
                       model = "within", 
                       index = c("country_code", "year"),
                       effect = "twoways",
                       data = panel_data)
summary(di_x_reg_binary, vcov = vcovHC(di_x_reg_binary, cluster = "group", type = "HC1"))

# Interaction 3 DI x Regime Change Episode/Direction
# HA: The impact of SPI on SDG performance varies depending on Regime Change Status/Direction
di_x_reg_eps <- plm(formula = spi_comp ~ dplyr::lag(cen_di_score, 1)*dplyr::lag(aut_ep, 1) + dplyr::lag(cen_di_score, 1)*dplyr::lag(dem_ep, 1) + cen_log_gdppc,
                   model = "within", 
                   index = c("country_code", "year"),
                   effect = "twoways",
                   data = panel_data)
summary(di_x_reg_eps, vcov = vcovHC(di_x_reg_eps, cluster = "group", type = "HC1"))

# Enhanced helper function
tidy_plm_robust <- function(model, model_name) {
  tidy_coef <- tidy(model, 
                    conf.int = TRUE, 
                    vcov. = vcovHC(model, cluster = "group", type = "HC1")) %>%
    mutate(model = model_name)
  
  # Extract model statistics
  model_summary <- summary(model)
  tidy_coef %>%
    mutate(
      adj_r_squared = model_summary$r.squared["adjrsq"]
    )
}

# Apply tidy_plm_robust() function to all models
tidy_di_x_inc_lvl    <- tidy_plm_robust(di_x_inc_lvl, "Moderator: DI x Income Level")
tidy_di_x_reg_binary <- tidy_plm_robust(di_x_reg_binary, "Moderator: DI x Regime Type (Binary)")
tidy_di_x_reg_eps  <- tidy_plm_robust(di_x_reg_eps, "Moderator: DI x Regime Change Episode (aut_ep & dem_ep)")

model_n_obs <- tibble(
  model = c("Moderator: DI x Income Level", "Moderator: DI x Regime Type Binary", 
            "Moderator: DI x Regime Change Episode (aut_ep & dem_ep)"),
  n_obs = c(nobs(di_x_inc_lvl), nobs(di_x_reg_binary),
            nobs(di_x_reg_eps))
)

# Combine all tidy data frames into one
all_models_tidy <- bind_rows(tidy_di_x_inc_lvl, 
                             tidy_di_x_reg_binary, 
                             tidy_di_x_reg_eps) %>%
  left_join(model_n_obs, by = "model") %>% 
  select(model, term, estimate, std.error, p.value, adj_r_squared, n_obs, everything()) %>%
  arrange(model, term)

# Save the tidy data frame to a CSV file
write_csv(all_models_tidy, "results_csv/fe_s1_interactions.csv")


# extracting robust standard errors for each interaction model
rob_se_di_x_inc_lvl <- coeftest(di_x_inc_lvl, vcov = vcovHC(di_x_inc_lvl, cluster = "group", type = "HC1"))
rob_se_di_x_reg_binary <- coeftest(di_x_reg_binary, vcov = vcovHC(di_x_reg_binary, cluster = "group", type = "HC1"))
rob_se_di_x_reg_eps <- coeftest(di_x_reg_eps, vcov = vcovHC(di_x_reg_eps, cluster = "group", type = "HC1"))

# Create a list of robust standard errors for stargazer
robust_se_list <- list(
  rob_se_di_x_inc_lvl[, 2], 
  rob_se_di_x_reg_binary[, 2], 
  rob_se_di_x_reg_eps[, 2]
  )
          
stargazer(
  di_x_inc_lvl, di_x_reg_binary, di_x_reg_eps, 
  se = robust_se_list,
  type = "html",
  out = "figures/interact_s1_models.html",
  report = "vcs*",
  title = "Stage 1: Applying Moderators to Two-Way Fixed Effects Models",
  align = TRUE,
  dep.var.labels = "SDG Overall",
  column.labels = c("DI x Income Level", "DI x Regime Type (Binary: Autocracy)", "DI x Regime Change Episode"),
  omit = "factor\\(year\\)",  # Omits year fixed effects from display
  model.names = FALSE
)
```
*Interaction 1: GNI Income Classification:* 
*HA: The impact of SPI on SDG performance varies depending on GNI Classification (income_level_recoded: 1 = Low; 2 = Lower-Middle; 3 = Upper-Middle; 4 = High)*
Not significant 

*Interaction 2: Regime Type (Binary):*
*HA: The impact of SPI on SDG performance varies depending on Regime Type (Binary: 0 = Autocracy; 1 = Democracy)*
Not significant 

*Interaction 3: Regime Change Status/Direction:*
*HA: The impact of SPI on SDG performance varies depending on Regime Change Status/Direction*
Not significant

### Partial F Test for Interaction Terms
```{r}
# Partial F Test for Interaction Terms
# restricted model without interaction terms 
restricted_model <- plm(formula = spi_comp ~ cen_di_score + cen_log_gdppc + factor(income_level_recoded),
                        model = "within", 
                        index = c("country_code", "year"),
                        data = panel_data,
                        effect = "twoways")

# Perform the partial F test[FIGURE OUT WHY IS NOT WORKINGGG]
## This is the right approach according to the lecture!!!
partial_f_test <- pFtest(restricted_model, di_x_inc_lvl)
partial_f_test


```



# Interaction Effects [Stage 2]
*H0:* NUll, the impact of SPI on SDG Performance does NOT vary depending on Z [income_level_recoded, regime_type_binary, regime_type_categ, di_score]
*H1:* The impact of SPI on SDG Performance varies depending on Z [income_level_recoded, regime_type_binary, regime_type_categ, di_score]

**Recoded Moderator Variables:**
- income_level_recoded: 1 = Low; 2 = Lower-Middle; 3 = Upper-Middle; 4 = High
- regime_type_binary: 0 = Autocracy; 1 = Democracy
- di_score: 0-1, continuous (Democracy Index)
- aut_ep: 0 = No Autocratization; 1 = Autocratization Event
- dem_ep: 0 = No Democratization; 1 = Democratization Event
```{r}
# Interaction 1: SPI x GNI Class 
# HA: The impact of SPI on SDG performance varies depending on GNI Classification 
spi_x_inc_lvl <- plm(formula = sdg_overall ~ cen_spi_comp*dplyr::lag(income_level_recoded, 1) + cen_di_score +
                       cen_log_gdppc + I(cen_log_gdppc^2),
                     model = "within", 
                     index = c("country_code", "year"),
                     data = panel_data,
                     effect = "twoways")
summary(spi_x_inc_lvl, vcov = vcovHC(spi_x_inc_lvl, cluster = "group", type = "HC1"))

# Interaction 2: SPI x DI
# HA: The impact of SPI on SDG performance varies depending on Democracy Score (cen_di_score: 0-1, continuous)
spi_x_di_score <- plm(formula = sdg_overall ~ cen_spi_comp*cen_di_score + cen_log_gdppc + I(cen_log_gdppc^2) + factor(income_level_recoded),
                      model = "within", 
                      index = c("country_code", "year"),
                      data = panel_data,
                      effect = "twoways")
summary(spi_x_di_score, vcov = vcovHC(spi_x_di_score, cluster = "group", type = "HC1"))

# Interaction 3: SPI x Regime Type (binary)
# HA: The impact of SPI on SDG performance varies depending on Regime Type (Binary: 0 = Autocracy; 1 = Democracy)
spi_x_reg_binary <- plm(formula = sdg_overall ~ cen_spi_comp*factor(regime_type_binary) + cen_di_score +
                          cen_log_gdppc + I(cen_log_gdppc^2) + factor(income_level_recoded),
                        model = "within", 
                        index = c("country_code", "year"),
                        data = panel_data,
                        effect = "twoways")
summary(spi_x_reg_binary, vcov = vcovHC(spi_x_reg_binary, cluster = "group", type = "HC1"))

# Interaction 4: SPI x Regime Change Event (Autocratization or Democratization)
# HA: The impact of SPI on SDG performance varies depending on Regime Change Status/Direction
spi_x_reg_eps <- plm(formula = sdg_overall ~ cen_spi_comp*aut_ep + cen_spi_comp*dem_ep +
                     cen_log_gdppc + I(cen_log_gdppc^2) + factor(income_level_recoded),
                   model = "within", 
                   index = c("country_code", "year"),
                   data = panel_data,
                   effect = "twoways")
summary(spi_x_reg_eps, vcov = vcovHC(spi_x_reg_eps, cluster = "group", type = "HC1"))


```


```{r}
# Final Model [TO BE UPDATED BASED ON STAGE 2 RESULTS]
final_model <- plm(formula = sdg_overall ~ cen_spi_comp*factor(income_level_recoded) + cen_spi_comp*cen_di_score + cen_log_gdppc + I(cen_log_gdppc^2), 
                   model = "within", 
                   index = c("country_code", "year"),
                   data = panel_data,
                   effect = "twoways")
summary(final_model, vcov = vcovHC(final_model, cluster = "group", type = "HC1"))
```


