---
title: "H2_SPI~DI"
output: html_document
---


# Set up and Wrangling 
```{r}
# set working directory
setwd("~/Documents/GitHub/QMSS_Thesis_Sanchez")

#load libraries/packages
source("packages.R")

# load data
source("Comp2_panel_wrangling.R")
```

# SPI and DI Relationship
## Baseline Model [FOUND SOMETHING!!!]
Using fixed effects regression, how does a change in DI score impact SPI performance? Is the impact lagged or immediate?
```{r}
# baseline model
FE_testing <- plm(formula = spi_comp ~ di_score + dplyr::lag(di_score, 1) + dplyr::lag(di_score, 2) + log_gdppc,
                  model = "within", 
                  index = c("country_code", "year"),
                  data = panel_data, 
                  effect = "twoways")
summary(FE_testing)
summary(FE_testing, vcov = vcovHC(FE_testing, cluster = "group", type = "HC1"))
```


## Complete Regime Change (autocratized), Subgrouping [FOUND SOMETHING!!!]
Of countries (15 countries) that experienced complete change from democracy to autocracy, how did the decrease in DI score impact their SPI performance? Was the impact lagged or immediate?
```{r}
# autocratized variable 
FE_testing <- plm(formula = spi_comp ~ di_score + dplyr::lag(di_score, 1) + dplyr::lag(di_score, 2) + log_gdppc,
                  model = "within", 
                  index = c("country_code", "year"),
                  data = panel_data, 
                  subset = autocratized == 1, # modifying based regime stability
                  effect = "twoways")
summary(FE_testing)
summary(FE_testing, vcov = vcovHC(FE_testing, cluster = "group", type = "HC1"))
```


## Complete Regime Change (democratized), Subgrouping [FOUND SOMETHING!!!]
Of countries (15 countries) that experienced complete change from autocracy to democracy, how did the decrease in DI score impact their SPI performance? Was the impact lagged or immediate?
```{r}
# autocratized variable 
FE_testing <- plm(formula = spi_comp ~ di_score + dplyr::lag(di_score, 1) + dplyr::lag(di_score, 2) + log_gdppc,
                  model = "within", 
                  index = c("country_code", "year"),
                  data = panel_data, 
                  subset = democratized == 1, # modifying based on regime stability
                  effect = "twoways")
summary(FE_testing)
summary(FE_testing, vcov = vcovHC(FE_testing, cluster = "group", type = "HC1"))
```
