---
title: "H2_SPI~DI"
output: html_document
---


# Set up and Wrangling 
```{r}
# set working directory
setwd("~/Documents/GitHub/QMSS_Thesis_Sanchez")

#load libraries/packages
source("packages.R")

# load data
source("Comp2_panel_wrangling.R")
```

# SPI and DI Relationship
## Baseline Model [FOUND SOMETHING!!!]
Using fixed effects regression, how does a change in DI score impact SPI performance? Is the impact lagged or immediate?
```{r}
# baseline model
FE_all <- plm(formula = spi_comp ~ di_score + dplyr::lag(di_score, 1) + dplyr::lag(di_score, 2) + log_gdppc,
                  model = "within", 
                  index = c("country_code", "year"),
                  data = panel_data, 
                  effect = "twoways")
summary(FE_all)
summary(FE_all, vcov = vcovHC(FE_all, cluster = "group", type = "HC1"))
```


## Complete Regime Change (autocratized), Subgrouping [FOUND SOMETHING!!!]
Of countries (15 countries) that experienced complete change from democracy to autocracy, how did the decrease in DI score impact their SPI performance? Was the impact lagged or immediate?
```{r}
# autocratized variable 
FE_aut <- plm(formula = spi_comp ~ di_score + dplyr::lag(di_score, 1) + dplyr::lag(di_score, 2) + log_gdppc,
                  model = "within", 
                  index = c("country_code", "year"),
                  data = panel_data, 
                  subset = autocratized == 1, # modifying based regime stability
                  effect = "twoways")
summary(FE_aut)
summary(FE_aut, vcov = vcovHC(FE_aut, cluster = "group", type = "HC1"))
```


## Complete Regime Change (democratized), Subgrouping [FOUND SOMETHING!!!]
Of countries (15 countries) that experienced complete change from autocracy to democracy, how did the decrease in DI score impact their SPI performance? Was the impact lagged or immediate?
```{r}
# autocratized variable 
FE_dem <- plm(formula = spi_comp ~ di_score + dplyr::lag(di_score, 1) + dplyr::lag(di_score, 2) + log_gdppc,
                  model = "within", 
                  index = c("country_code", "year"),
                  data = panel_data, 
                  subset = democratized == 1, # modifying based on regime stability
                  effect = "twoways")
summary(FE_dem)
summary(FE_dem, vcov = vcovHC(FE_dem, cluster = "group", type = "HC1"))
```

## Put all three models in a table
```{r}
# Extracting robust standard errors for the three FE models
rob_se_fe <- list(
  coeftest(FE_all, vcov = vcovHC(FE_all, cluster = "group", type = "HC1"))[, 2],
  coeftest(FE_aut, vcov = vcovHC(FE_aut, cluster = "group", type = "HC1"))[, 2],
  coeftest(FE_dem, vcov = vcovHC(FE_dem, cluster = "group", type = "HC1"))[, 2]
)
# stargazer table for FE models
stargazer(
  FE_all, FE_aut, FE_dem,
  se = rob_se_fe,
  type = "html",
  title = "H4: SPI and DI Fixed Effects Models (Regime Change)",
  align = TRUE,
  dep.var.labels = "SPI Overall Score",
  column.labels = c("All", "Autocratized", "Democratized"),
  #covariate.labels = c("DI Composite", "Lagged DI (t-1)", "Lagged DI (t-2)",  "Log(GDPper-capita)", "Lower-Middle Income (LMIC)", "Upper-Middle Income (UMIC)", "High Income (HIC)", "Intercept"),
  report = "vcs*",
  notes = "Robust standard errors clustered by country in parentheses",
  omit = "factor\\(year\\)",
  model.names = FALSE,
  out = "compiled_materials/results/h4_spi_di_FE_models.html"
)

```